### 2017-04-12 thesis/brainstorm.Rmd
adding brainstorm questions from week 6

---
title: "Brainstorm Questions for Melbourne Pedestrian Data"
output:
  html_document: default
  html_notebook: default
---
***

###Can we classify groups of sensor locations into types?###  
* Cluster analysis?  
* Types of areas? (restaurants and bars, clubs, shopping, train stations, etc..)  
    + peak traffic periods different?

###How does the pedestrian flow change with weather?###  
* Does rain result in less people in the city? Only on weekends and holidays?  
* Modelling based on past weather  
* Forecasting using weather forecast  
    + differences between business/office districts vs. leisure districts (shopping, cultural/tourist) - maybe lcassify these areas manually and see if we can confirm by cluster analysis?

###Is there a “maximum capacity” to how many people can move through the city in a given time frame? And if so, what affects this capacity?###

* At quantiles for given locations, say 95th percentile  
* Look at features of those locations to see what may affect flow (linked to clustering? maybe factor analysis?)  
    + Does this change with weather conditions? (similar to road traffic being affected by adverse weather potentially?)  
    + Any difference for areas with limited road traffic (Swanston and Bourke)

###Can we predict whether a day is a public holiday?###

* We expect that business areas to be quiet, and leisure areas to be just as busy/busier
    + maybe need to compare ratios between areas

### 2017-04-13 impute_missing.r
impute code

@@ -1,0 +1,98 @@
+# Make missing values explict
+df_time <- sub_ped %>% 
+  split(.$Sensor_ID) %>% 
+  map(extract2, "Date_Time") %>% 
+  map(full_seq, period = 3600)
+sensors_ls <- sub_ped %>% 
+  distinct(Sensor_ID, Sensor_Name) %>% 
+  arrange(Sensor_ID) %>% 
+  split(.$Sensor_ID)
+df_time <- map2(df_time, sensors_ls, data.frame) %>% 
+  map(set_colnames, c("Date_Time", "Sensor_ID", "Sensor_Name"))
+
+ped_full <- sub_ped %>% 
+  split(.$Sensor_ID) %>% 
+  map2(df_time, right_join, 
+       by = c("Date_Time", "Sensor_ID", "Sensor_Name")) %>% 
+  bind_rows()
+
+ped_full <- ped_full %>% 
+  mutate(
+    Year = year(Date_Time),
+    Month = as.character(month(Date_Time, label = TRUE)),
+    Mdate = day(Date_Time),
+    Day = as.character(wday(Date_Time, label = TRUE)),
+    Time = hour(Date_Time),
+    Date = as_date(Date_Time),
+    Holiday = if_else(Date %in% pub_holiday$date, "Y", "N"),
+    HDay = if_else(Holiday == "Y", "Holiday", Day)
+  )
+
+# Fit a glm model using dummies
+mlevels <- unique(ped_full$Month)
+dlevels <- unique(ped_full$Day)[c(6:7, 1:5)]
+hdlevels <- unique(ped_full$HDay)[c(7:8, 2:6, 1)]
+ped_dummy <- ped_full %>% 
+  mutate(
+    Year = factor(Year, levels = 2014:2016),
+    Month = factor(Month, levels = mlevels),
+    Day = factor(Day, levels = dlevels),
+    HDay = factor(HDay, levels = hdlevels),
+    Time = factor(Time, levels = 0:23)
+  )
+
+ped_list <- ped_dummy %>% 
+  split(.$Sensor_ID)
+
+ped_fit <- ped_list %>% 
+  map(arrange, Date_Time) %>% 
+  map(~ glm(Hourly_Counts ~ Year * HDay * Time + Month, data = .,
+    family = poisson())
+  )
+
+ped_resid <- sub_ped %>% 
[…]

### 2017-04-26 thesis/proposal_draft.Rmd
thesis proposal draft
still need to add lit review regarding modelling pedestrian traffic and forecasting with multiple seasonalities

---
title: "Draft Proposal"
author: "Gavin Chin"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(ggmap)
library(grid)
library(png)
library(knitr)
```

```{r, echo = FALSE, messages = FALSE, include=FALSE, cache=TRUE}
  ped_df <- read_csv("data/Pedestrian_volume__updated_monthly_.csv")
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])

  ped_loc <- read_csv("data/Pedestrian_sensor_locations.csv")

  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
                               mean(range(ped_loc$Latitude))),
                                zoom=14)
```

## Topic
The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives" $^1$. This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD. In particular, we want to be able to model the pedestrian traffic at different locations in the CBD, and forecast the traffic. In addition, we aim to have an interactive, dynamic data visualisation of the model to help

--------------------------------------------------------------------------------------------------------------------------------------

## Background

#### The Data: City of Melbourne Pedestrian Data

The dataset being investigated can be obtained from the official City of Melbourne's open data platform. The pedestrian data is in the form of hourly pedestrian counts for `r nrow(ped_loc)` sensor locations. In total, there is `r nrow(ped_df)` observations in the dataset. It is available in `.csv` format, allowing for easy data import into R.

```{r, echo = F, fig.align= "center", fig.width=3.5, fig.height=3.5}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(), axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```
\begin{center}
Figure 1:  Locations of all pedestrian sensors around Melbourne's CBD as at April 2017
\end{center}

These sensors are placed above flowing traffic either under an awning or on a street pole, recording movements from multiple directions. The pedestrian counting system was introduced by the council with the objectives to help inform decisions about urban planning and management, identify opportunities to improve city walkability and transport, to measure the impact of events and specific marketing campaings, as well as to monitor retail activity and aiding the business community with marketing and resource management.

Each observation in the data contains three main variables: the specific hour (date and time of day), location (sensor ID) and the pedestrian count for the hour. Other variables were also included such as the month, date, hour of the day, etc.. to help with the analysis. These variables do not being in any additional information however.

```{r, message= FALSE, warning= FALSE, fig.height=4, echo = FALSE}
  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
    geom_smooth(aes(x = Timeplot, y = Hourly_Counts, colour = Month), se = F) + 
    xlab(NULL) + ylab("Pedestrian Count") + 
    ggtitle(ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"]) +
    facet_wrap(~ Day) + theme(plot.title = element_text(hjust = 0.5))
```

\begin{center}
Figure 2: Smoothed pedestrian counts for Melbourne Central by day of the week and month of the year
\end{center}

We are faced with two large limitations due to the nature of the collection of the data. Firstly, pedestrian counts are aggregated hourly. This does not allow for highly accurate modelling. Secondly, we note that there is a lack of sensors, and hence pedestrian data, for the area between Spencer Street and Elizabeth Street. The majority of the sensors have been installed by the City of Melbourne along Swanston Street. The consequence of this is the potential to make misleading conclusions regarding the pedestrian activity in that area. Plotting pedestrian traffic of the entire CBD may give the inaccurate perception of very low/no pedestrian activity.  
It is hoped that pedestrian data based on WiFi points around the CBD, which has greater spatial coverage and a higher degree of resolution (counts aggregated every 15 minutes), will be available. The analysis based on the sensor data should remain valid for the WiFi data.

#### Current Pedestrian Data Visualisation

The City of Melbourne currently has an online visualisation tool which is publicly accessible at [City of Melbourne Pedestrian Counting System](http://www.pedestrian.melbourne.vic.gov.au) (`http://www.pedestrian.melbourne.vic.gov.au`). This visualisation features an interactive map of Melbourne's CBD with markers indicating the the location and the pedestrian activity at the location for a given time and date.

```{r, out.width= '70%', echo = FALSE, fig.align= "center"}
include_graphics("img/ped_vis_melb.PNG")
```

\begin{center}
Figure 3: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au
\end{center}

Pedestrian activity is represented by two aesthetic features - size and colour. The markers, in the shape of thermometers, change in height based on the absolute pedestrian count, with a colour scale from blue to white to yellow, analogous to temperature, to map the relative diference in activity. The colour changes when the hourly counts deviate by more than 10% from the average count over the previous four weeks for the same day of the week and time of the day.  In addition to these aesthetics, the plot itself is both interactive and animated.

The main advantage of an animated/dynamic plot is that the changes over time can be visualised effectively for multiple sensor locations. The UI (user interface) allows the end user to change the date visualisation, and either view pedestrian traffic of a particular hour of the day, or to run an animation of the whole day's pedestrian traffic.

## Methodology

  The first step in performing the analysis is to clean the data. In particular, we need to impute the missing values at certain sensor locations. This is cause by two primary reasons: 1) new sensors being installed over time, and 2) sensors not operational for periods of time. A possible model which has been proposed is a GLM (generalized linear model) to estimate a Poisson regression: $$\text{HourlyCounts} \sim \text{Year} \times \text{HDay} \times \text{Time} + \text{Month} \label{poisson_model} \tag{1}$$

This estimation is computed for each sensor individually, as each location will have unique characteristics. This model assumes that the hourly pedestrian counts for a given location is dependent on the time of day, `Time`, where the `Year` and day of the week has as interaction effect. `Month` has an additive effect in this model, rather than an interactive effect. This would imply that the month of the year does not affect the pattern of activity for a given day of the week and time of day.

In equation \ref{poisson_model}, the variable `HDay` refers to a factor variable with 8 factor levels: the 7 days of the week and a factor for a public holiday. This is generated such that, if the day is a public holiday, then `HDay = "Holiday"`, regardless of the day of the week. Otherwise, `HDay` is the same as the factor for the day of the week.  
Alternatively, it would be possible to allow for the effect of a public holiday by including a separate dummy variable. However, the likelihood of overfitting is greater, as we have a lack of public holidays as it is, let alone the lack of holidays falling on each day of the week. Similarly, the inclusion of the `Year` variable may introduce similar problems, especially at newer sensor locations with a lack of historical data.

Another decision which needs to be made is the starting date to model from. With multiple sensors being installed years later after the initial 28 sensors were installed and capturing data, we do not have a complete set of observations over the total observation period at each sensor location. These locations will need to have pedestrian activity modelled based on neighbouring sensors to predict missing values for the time period prior to the sensor's installation and operation.

After the imputation of the missing values, we want to model the data (under the assumption that it is now complete for a given time period). In particular, we want to see if we can model pedestrian activity with weather data from the Bureau of Meteorology. The motivation for this is the availability of weather forecasts. Without looking at the data, we would expect that the effects of weather to have a greater effect on weekends and public holidays than on weekdays. The intuition behind this hypothesis is that activity in the CBD during weekends and public holidays is generally for leisure, whereas weekday activity is primarily for business and commercial activities.

Weather may also have a casual effect by month of the year, due to seasonal weather. This could help explain differences in activity levels in summer compared to in winter in certain locations. This effect may also only have an effect in outdoor locations, and not on indoor locations. For example, Melbourne Central, a retail shopping area, is affected by `Month`, but most likely not weather. From _Figure 2_, we see that the average peak actvity on Fridays and Saturdays in January is significantly lower than other months of the year, while weather conditions are similar to December and February. This is most likely to be explained by typical seasonal consumer behaviour: peak consumer spending occurs in December due to Christmas gift shopping and Boxing Day sales. As a result, guilt ridden consumers tend to wind back retail spending during January. It can also be explained by people going on vacations out into regional/rural areas.

## References  
  1. City of Melbourne, _'About Melbourne Data'_, https://data.melbourne.vic.gov.au/about, _accessed 18 April 2017_  
  2. City of Melbourne, _'Pedestrian Counting System'_,  
  http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf 4 May 2015, _accessed 18 April 2017_

### 2017-05-02 thesis/proposal_draft.Rmd
update wednesday morning: still needs lit review for modelling time series and visualisation

@@
 library(readr)
+library(dplyr)
+library(purrr)
 library(ggplot2)
@@
 library(knitr)
+library(lubridate)
 ```
@@
 ```{r, echo = FALSE, messages = FALSE, include=FALSE, cache=TRUE}
-  ped_df <- read_csv("data/Pedestrian_volume__updated_monthly_.csv")
+  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
       ped_df$Year <- as.factor(ped_df$Year)
@@
 
-  ped_loc <- read_csv("data/Pedestrian_sensor_locations.csv")
+  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
 
@@
 In particular, we want to be able to model the pedestrian traffic at different locations in the CBD, and forecast the traffic.
-In addition, we aim to have an interactive, dynamic data visualisation of the model to help
+In addition, we aim to have an interactive, dynamic data visualisation of the model to help us to understand how Melbourne operates.
 
@@
 
-```{r, message= FALSE, warning= FALSE, fig.height=4, echo = FALSE}
+```{r, message= FALSE, warning= FALSE, fig.height=4, fig.width=6, echo = FALSE}
   ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
     geom_smooth(aes(x = Timeplot, y = Hourly_Counts, colour = Month), se = F) + 
-    xlab(NULL) + ylab("Pedestrian Count") + 
+    xlab("Time of the Day") + ylab("Pedestrian Count") + 
     ggtitle(ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"]) +
-    facet_wrap(~ Day) + theme(plot.title = element_text(hjust = 0.5))
+    facet_wrap(~ Day) + theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
+    scale_x_continuous(limits = c(0,24))
 ```
@@
 ```{r, out.width= '70%', echo = FALSE, fig.align= "center"}
-include_graphics("img/ped_vis_melb.PNG")
+include_graphics("../img/ped_vis_melb.PNG")
 ```
@@
 The UI (user interface) allows the end user to change the date visualisation, and either view pedestrian traffic of a particular hour of the day, or to run an animation of the whole day's pedestrian traffic.
+
+#### Modelling a time series with multiple seasonal patterns
+
+The pedestrian data is presented as a time series.
+However, we face a major obstacle with pedestrian data: multiple seasonal patterns.
+Looking at the data on a month to month basis at a single sensor location, we have two main cycles: a daily sub-cycle within a weekly cycle.
+```{r, message= FALSE, warning= FALSE, fig.height=2, echo = FALSE}
+bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
+ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) +
+      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
+```
+
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE}
+ggplot(filter(ped_df, Sensor_ID == "13")) + geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day)) +
+      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
+
+```
 
@@
 ## References  
-  1. City of Melbourne, _'About Melbourne Data'_, https://data.melbourne.vic.gov.au/about, _accessed 18 April 2017_  
+  1. City of Melbourne, _'About Melbourne Data'_, https://data.melbourne.vic.gov.au/about,  
+  _accessed 18 April 2017_  
   2. City of Melbourne, _'Pedestrian Counting System'_,  
-  http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf 4 May 2015, _accessed 18 April 2017_
+  http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf 
+  4 May 2015, _accessed 18 April 2017_

### 2017-05-02 thesis/proposal_draft.Rmd
wednesday morning, update 2: added comments to plot in figure 3

@@
   pdf_document: default
+  html_document: default
 ---
@@
 
-```{r, message= FALSE, warning= FALSE, fig.height=4, fig.width=6, echo = FALSE}
+```{r, message= FALSE, warning= FALSE, fig.height=4, fig.width=6, echo = FALSE, fig.align= "center"}
   ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
@@
 Looking at the data on a month to month basis at a single sensor location, we have two main cycles: a daily sub-cycle within a weekly cycle.
-```{r, message= FALSE, warning= FALSE, fig.height=2, echo = FALSE}
+```{r, message= FALSE, warning= FALSE, fig.height=2, echo = FALSE, fig.align= "center"}
 bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
@@
 ```
-
-```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE}
+\begin{center}
+Figure 3:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016)
+\end{center}  
+From _Figure 3_, we can observe multiple characteristics of the time series.
+Firstly, we see the weekly cycle, where weekdays (Monday to Friday) have lots of activity and weekends (Saturday and Sunday) are relatively quiet.
+This would be highly suggestive that Flagstaff Station is primarily used by city office workers.
+Another phenomenon we observe is the effect of a public holiday.
+Looking near the middle of the month (Monday, 13 June 2016), we have the Queen's Birthday public holiday.
+We must also note that, prior to the start of 2016, Flagstaff Station was not operating on weekends, nor public holidays.
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
 ggplot(filter(ped_df, Sensor_ID == "13")) + geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day)) +
@@
 ```
+\begin{center}
+Figure 4:  Estimated hourly counts at Flagstaff Station by Day using `geom_smooth()`
+\end{center}
+
+In _Figure 4_, we see that, within the weekly cycle, each day has a daily sub-cycle.
+The plot in _Figure 4_ is generated using `geom_smooth` in `ggplot2`, which fitted a line to estimate hourly counts over time for each day using a GAM (generalized additive model) for smoothing.
 

### 2017-05-03 thesis/proposal_draft.Rmd
wednesday update 3

@@
 ---
-title: "Draft Proposal"
-author: "Gavin Chin"
+title: "What can we learn about life in the city by mining the Melbourne’s Open Data Platform?"
+subtitle: "Thesis Proposal"
+date: April 2017
+author: 
+- "Gavin Chin"
+- "supervised by Dianne Cook"
 output:
@@
 \begin{center}
-Figure 4:  Estimated hourly counts at Flagstaff Station by Day using `geom_smooth()`
+Figure 4:  Estimated hourly counts at Flagstaff Station by Day 
 \end{center}
@@
 The plot in _Figure 4_ is generated using `geom_smooth` in `ggplot2`, which fitted a line to estimate hourly counts over time for each day using a GAM (generalized additive model) for smoothing.
+
+#### R and Shiny Apps
+
+R is an open source programming language used for statistical computing.
 
@@
 This is cause by two primary reasons: 1) new sensors being installed over time, and 2) sensors not operational for periods of time.
-A possible model which has been proposed is a GLM (generalized linear model) to estimate a Poisson regression: $$\text{HourlyCounts} \sim \text{Year} \times \text{HDay} \times \text{Time} + \text{Month} \label{poisson_model} \tag{1}$$
+A possible model which has been proposed is a GLM (generalized linear model) to estimate a Poisson regression:
+\begin{align*}
+\text{HourlyCounts} & \sim\text{Year}\times\text{HDay}\times\text{Time}+\text{Month}\tag{1}\\
+\Rightarrow\qquad\lambda_{i} & =\beta_{0}+\beta_{1}\text{Year}\times\text{HDay}\times\text{Time}+\beta_{2}\text{Month}\\
+\text{HourlyCounts}_{i} & =\text{Poisson}\left(\lambda_{i}\right)
+\end{align*}
 

### 2017-05-03 thesis/proposal_draft.Rmd
update

@@
 The main advantage of an animated/dynamic plot is that the changes over time can be visualised effectively for multiple sensor locations.
+This is particularly useful in reducing change blindness in the peception of the data visualisation, as it is easier for people viewing the plot to see relative changes from one hour to the next.
 The UI (user interface) allows the end user to change the date visualisation, and either view pedestrian traffic of a particular hour of the day, or to run an animation of the whole day's pedestrian traffic.
@@
 The plot in _Figure 4_ is generated using `geom_smooth` in `ggplot2`, which fitted a line to estimate hourly counts over time for each day using a GAM (generalized additive model) for smoothing.
+We can see the patterns we observe on a weekly cycle here when we seperately estimate by the day of the week.
+What we can see more clearly is also the difference in the pattern of pedestrian activity for each hour of the day, where there are two main peaks per weekday.
+This is consistent with the peak travel times for people commuting to and from work.
+
+The consequence of having multiple seasonalities in the time series is that typical time series
 
@@
 R is an open source programming language used for statistical computing.
+Shiny, made by RStudio, is a web application framework for R.
+It allows us to easily develop interactive web applications which interface with R in the background.
+Combining these with open source packages in R, such as `ggplot2` or `plotly`, we can develop dynamic, interactive visualisations for data.
 

### 2017-05-03 thesis/proposal_draft.Rmd
Update proposal_draft.Rmd

@@
 ## Topic
-The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives" $^1$.
+The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives"[^1].
 This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.
@@
 
-The dataset being investigated can be obtained from the official City of Melbourne's open data platform.
+The dataset being investigated can be obtained from the official City of Melbourne's open data[^2].
 The pedestrian data is in the form of hourly pedestrian counts for `r nrow(ped_loc)` sensor locations.
@@
 
-```{r, echo = F, fig.align= "center", fig.width=3.5, fig.height=3.5}
+[^1]: from _'About Melbourne Data'_ page at https://data.melbourne.vic.gov.au/about
+
+[^2]: the specific pedestrian data is available at   https://data.melbourne.vic.gov.au/Transport-Movement/Pedestrian-volume-updated-monthly-/b2ak-trbp
+
+```{r, echo = F, fig.align= "center", fig.width=3.5, fig.height=3.5, cache=TRUE}
 ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
@@
 
-```{r, message= FALSE, warning= FALSE, fig.height=4, fig.width=6, echo = FALSE, fig.align= "center"}
+```{r, message= FALSE, warning= FALSE, fig.height=4, fig.width=6, echo = FALSE, fig.align= "center", cache=TRUE}
   ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
@@
 We must also note that, prior to the start of 2016, Flagstaff Station was not operating on weekends, nor public holidays.
-```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center", cache=TRUE}
 ggplot(filter(ped_df, Sensor_ID == "13")) + geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day)) +
@@
 
+In particular, the objective of this project is to have the modelled pedestrian data presented as an interactive web app using Shiny.
+A key part which makes the visualisations interactive is the ability to have user inputs.
+This may be in the form of selecting a time period for the data, or changing how a certain plot is generated (size, colour, etc..).
+Another key
+
 ## Methodology
@@
 A possible model which has been proposed is a GLM (generalized linear model) to estimate a Poisson regression:
+
 \begin{align*}

### 2017-05-03 thesis/proposal_draft.Rmd
update test wednesday evening

@@
 library(ggmap)
-library(grid)
-library(png)
 library(knitr)

### 2017-05-03 thesis/proposal_draft.Rmd
update evening 2

@@
 The main advantage of an animated/dynamic plot is that the changes over time can be visualised effectively for multiple sensor locations.
-This is particularly useful in reducing change blindness in the peception of the data visualisation, as it is easier for people viewing the plot to see relative changes from one hour to the next.
-The UI (user interface) allows the end user to change the date visualisation, and either view pedestrian traffic of a particular hour of the day, or to run an animation of the whole day's pedestrian traffic.
+This is particularly useful in reducing change blindness[^3] in the peception of the data visualisation, as it is easier for people viewing the plot to see relative changes from one hour to the next at a certain location.
+However, a particular aspect which needs to be improved on is the ability to visualise the pedestrian activity levels around the whole city at once.
+The UI (user interface) allows some interaction by the end user.
+It allows the user to change the date visualisation, and either view pedestrian traffic of a particular hour of the day, or to run an animation of the whole day's pedestrian traffic.
+Clicking on a sensor location brings up a line graph plotting hourly counts for the day.
+
+Another
+
+[^3]: change blindness is the inability to notice a change in visual stimulus introduced
 
@@
 
-The consequence of having multiple seasonalities in the time series is that typical time series
+The consequence of having multiple seasonalities in the time series is that typical time series do not deal with multiple seasonalities well.
 
@@
   4 May 2015, _accessed 18 April 2017_  
+  3. Christopher G. Healey, _'Attention and Visual Memory in Visualization and Computer Graphics'_, IEEE Transactions on Visualization and Computer Graphics,  
+  https://www.csc2.ncsu.edu/faculty/healey/download/tvcg.12a.pdf

### 2017-05-03 thesis/proposal_draft.Rmd
update night 1

@@
 
-Another
+Another interactive visualisation featured on the Open Data Platform also presents pedestrian counts as average daily volume on a map of Melbourne's CBD.
+Daily counts are mapped to a size aesthetic.
+However, this visualisation is very limited.
+While it useful to see the overall average activity at a sensor location relative to other locations in the CBD, it fails to present key characteristics of the time series (discussed later).
+
+```{r, out.width= '70%', echo = FALSE, fig.align= "center"}
+include_graphics("../img/ped_vis_melb_2.PNG")
+```
 

### 2017-05-03 thesis/proposal_draft.Rmd
update night 2

@@
 
-Another interactive visualisation featured on the Open Data Platform also presents pedestrian counts as average daily volume on a map of Melbourne's CBD.
+Another interactive visualisation featured on the Open Data Platform ( _Figure 4_ )also presents pedestrian counts as average daily volume on a map of Melbourne's CBD.
 Daily counts are mapped to a size aesthetic.
@@
 While it useful to see the overall average activity at a sensor location relative to other locations in the CBD, it fails to present key characteristics of the time series (discussed later).
+It also lacks basic features such as a legend for what colour represents.
 
@@
 ```
+
+\begin{center}
+Figure 4: Interactive data visualisation created using Tableau
+\end{center}
 
@@
 \begin{center}
-Figure 3:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016)
+Figure 5:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016)
 \end{center}  
@@
 \begin{center}
-Figure 4:  Estimated hourly counts at Flagstaff Station by Day 
+Figure 6:  Estimated hourly counts at Flagstaff Station by Day 
 \end{center}
@@
 
+### Data Cleaning
+
   The first step in performing the analysis is to clean the data.
@@
 
+The purpose of this estimation is purely for the imputation of missing values, so prediction within this period is the highest priority at this point.
+While it would be desirable, this model is not intended to forecast future values.
+
 Another decision which needs to be made is the starting date to model from.
@@
 These locations will need to have pedestrian activity modelled based on neighbouring sensors to predict missing values for the time period prior to the sensor's installation and operation.
+
+### Data Modelling
 
@@
 
+### Data Visualisation
+
 ## References  

### 2017-05-03 thesis/proposal_draft.Rmd
update night 3

@@
 
+The use of weather data to improve forecasting performance would be ideal, due to the availability of weather forecasts compared to many other forms of data.
+Machine learning techniques will also be explored.
+In particular, clustering of sensor locations may be useful in identifying patterns at certain types of location, such as tourist attractions and shopping districts.
+
 ### Data Visualisation
+
+Once a model for forecasting the data has been established, an interactive web application to present the model will be developed.
+A key aspect which we want to include is the ability for the end user to forecast pedestrian activity in the future.
 
@@
   4 May 2015, _accessed 18 April 2017_  
-  3. Christopher G. Healey, _'Attention and Visual Memory in Visualization and Computer Graphics'_, IEEE Transactions on Visualization and Computer Graphics,  
+  3. Healey, Christopher G.,  _'Attention and Visual Memory in Visualization and Computer Graphics'_, IEEE Transactions on Visualization and Computer Graphics,  
   https://www.csc2.ncsu.edu/faculty/healey/download/tvcg.12a.pdf

### 2017-05-03 thesis/proposal_draft.Rmd
thurs morning update

@@
 \begin{center}
-Figure 4: Interactive data visualisation created using Tableau
+Figure 4: An interactive data visualisation created using Tableau
 \end{center}
@@
 In particular, clustering of sensor locations may be useful in identifying patterns at certain types of location, such as tourist attractions and shopping districts.
+To evaluate the performance of the model, we look to minimise the RMSE (root mean square error):
+$$ \text{RMSE} = \sqrt{\frac{\sum_{t=1}^{T}(\hat{y_t} - y_t)^2}{T}}$$
 
@@
 Once a model for forecasting the data has been established, an interactive web application to present the model will be developed.
-A key aspect which we want to include is the ability for the end user to forecast pedestrian activity in the future.
+A key aspect which we want to implement is the ability for the end user to forecast pedestrian activity in the future.
+Improving on the currently available visualisations, the aesthetics of the visualisation will be designed with human perception theory taken into consideration.
+Preattentive processing theory will be considered to ensure certain elements of the visualisation allow for quick, preattentive identification of busy and quiet areas.
+Change blindness theory should also be reviewed in order to avoid misperception of relative changes in pedestrian activity.
 

### 2017-05-03 presentation/proposal_pres.Rmd
quick presentation draft

---
title: "Modelling the City of Melbourne Pedestrian Data"
author: "Gavin Chin, supervised by Di Cook"
date: "04/05/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(dplyr)
library(purrr)
library(ggplot2)
library(ggmap)
library(knitr)
library(lubridate)

  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])

  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")

  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
                               mean(range(ped_loc$Latitude))),
                                zoom=14)
```

## Introduction

The City of Melbourne provides an open data platform to access council datasets through its Open Data Platform  
This project focuses primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD

## Pedestrian Sensors

```{r, echo = F, fig.align= "center", cache=TRUE}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(), axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

## Time Series of Pedestrian Activity

```{r, message= FALSE, warning= FALSE, fig.align= "center"}
ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13") %>%
  ggplot() + geom_line(aes(x = MTime, y = Hourly_Counts)) +
      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

### 2017-05-03 code/work.R
initial code for imputing missing values

@@ -1,0 +1,31 @@
+
+
+library(readr)
+library(dplyr)
+library(purrr)
+library(ggplot2)
+library(knitr)
+library(lubridate)
+
+ped_df <- read_csv("data/Pedestrian_volume__updated_monthly_.csv")
+  ped_df$Year <- as.factor(ped_df$Year)
+  ped_df$Month <- as.factor(ped_df$Month)
+  ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
+  ped_df$Time <- as.factor(ped_df$Time)
+  ped_df$Timeplot <- as.integer(ped_df$Time)
+  ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
+  ped_df$Day <- as.factor(ped_df$Day)
+  ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])
+
+ped_loc <- read_csv("data/Pedestrian_sensor_locations.csv")
+
+pub_hday14 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
+pub_hday15 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
+pub_hday16 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
+
+  pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)
+  
+  pub_hdays$Date <- ymd(pub_hdays$Date)
+  pub_hdays$Month <- pub_hdays$Date %>% month(label = TRUE, abbr = FALSE)
+  
+  

### 2017-05-04 code/work.R
update thurs 2

@@ -6,6 +6,7 @@ library(purrr)
 library(ggplot2)
 library(knitr)
 library(lubridate)
+library(tidytext)
 
 ped_df <- read_csv("data/Pedestrian_volume__updated_monthly_.csv")
   ped_df$Year <- as.factor(ped_df$Year)
@@ -27,5 +28,15 @@ pub_hday16 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab
   
   pub_hdays$Date <- ymd(pub_hdays$Date)
   pub_hdays$Month <- pub_hdays$Date %>% month(label = TRUE, abbr = FALSE)
-  
+  pub_hdays$VIC <- 0
+  pub_hdays$VIC[grep(glob2rx("*VIC*"), pub_hdays$`Applicable To`)] <- 1
+  pub_hdays$VIC[grep("NAT", pub_hdays$`Applicable To`)] <- 1
+
+  ped_df$HDay <- 0
+for (i in 1:nrow(ped_df)){
+  if (ped_df$Date_Time[i] %in% pub_hdays$Date[pub_hdays$VIC == 1]) {
+      ped_df$HDay[i] <- "Holiday"
+  }
+  else {ped_df$HDay[i] <- ped_df$Day[i]}
+}
   

### 2017-05-04 ped_app/app.R

@@ -7,9 +7,6 @@ library(ggplot2)
 library(readr)
 library(ggmap)
 ##desktop
-  setwd("C:/Users/Gavin/Documents/honours2017/honours2017/")
-  Sys.setenv('MAPBOX_TOKEN' = 'pk.eyJ1IjoiYm9iYnl6MHIiLCJhIjoiMTk5MDczNWVmY2MzZjc2MmY0ZDI0YmZiYzEyNDc1MDYifQ.EFS7w-Ni47OoMpwy-zT_5A')
-  
   ped_df <- read_csv("data/Pedestrian_volume__updated_monthly_.csv")
   ped_df$Year <- as.factor(ped_df$Year)
   ped_df$Month <- as.factor(ped_df$Month)

### 2017-05-05 thesis/proposal_draft.Rmd
update thurs 2
added plots of time series in appendix

@@
 
+For testing the proof of concept, a simple web application for viewing average hourly counts for a given month (for a given year) on a map was developed. 
+```{r, fig.align = 'center', echo = FALSE, out.width = "66%"}
+include_graphics("../img/ped_vis_melb_me.png")
+```
+
+### Limitations
+
+#### Model  
+The model itself is currently limited to being treating the data as static panel data. A major reason for this is the nature of the data, where the patterns from year to year do not change.
+
+#### Visualisation
+
+### Research Plan
+- Imputation that incorporates time and space
+- Predicitive model for each sensor
+- Visual methods for space-time counts
+- Shiny app - interface design, computation
+- dynamic
+
 ## References  

### 2017-05-05 thesis/proposal_draft.Rmd
update friday - still to do: more detail in research plan, more detail in limitations, more detail in methodology of visualisation and app development.

@@
 knitr::opts_chunk$set(echo = TRUE)
+options(scipen = 100)
 library(readr)
@@
 library(lubridate)
+library(reshape2)
+
 ```
@@
 
---------------------------------------------------------------------------------------------------------------------------------------
+---------------------------------------------------------------------------------------
 
@@
 
-```{r, out.width= '70%', echo = FALSE, fig.align= "center"}
+```{r, out.width= '60%', echo = FALSE, fig.align= "center"}
 include_graphics("../img/ped_vis_melb_2.PNG")
@@
 While it would be desirable, this model is not intended to forecast future values.
+Evaluating the model's goodness of fit, we perform cross validation.
+Splitting the data randomly 60/40 into a training set (data for model estimation) and a test set (data set to test the goodness of fit), we can get a quick evaluation of the model's performance.
+LOOCV (leave one out cross validation) is not computationally feasible due to the large number of observations.
 
@@
 With multiple sensors being installed years later after the initial 28 sensors were installed and capturing data, we do not have a complete set of observations over the total observation period at each sensor location.
-These locations will need to have pedestrian activity modelled based on neighbouring sensors to predict missing values for the time period prior to the sensor's installation and operation.
+Consequently, we will only use data from 2014 onwards, as this appears to be the most complete and comprehensive time period where the obeservations provide useful information.
 
@@
 In particular, clustering of sensor locations may be useful in identifying patterns at certain types of location, such as tourist attractions and shopping districts.
-To evaluate the performance of the model, we look to minimise the RMSE (root mean square error):
-$$ \text{RMSE} = \sqrt{\frac{\sum_{t=1}^{T}(\hat{y_t} - y_t)^2}{T}}$$
+Different models for different types of locations may be used, and clustering may give insights as to which type of model to use.
+For example, the major city loop train stations, Flinders St, Southern Cross and Melbourne Central may have similar patterns.
+Then, we would expect to be able to use the same parameterisation of the model to estimate.
+After a model has been established, we need to evaluate the performance of the model.
+For this, we have decided to minimise the RMSFE (root mean square forecast error):
+$$ \text{RMSFE} = \sqrt{\frac{\sum_{t=1}^{T}(\hat{y_t} - y_t)^2}{T}}$$
 
@@
 
-Once a model for forecasting the data has been established, an interactive web application to present the model will be developed.
-A key aspect which we want to implement is the ability for the end user to forecast pedestrian activity in the future.
+Once a model for forecasting the data has been established, visual methods for illustrating the model need to be developed.
+
 Improving on the currently available visualisations, the aesthetics of the visualisation will be designed with human perception theory taken into consideration.
@@
 
+### Shiny App Development
+
+To implement the visual methods, an interactive web application to present the model will be developed.
+
 For testing the proof of concept, a simple web application for viewing average hourly counts for a given month (for a given year) on a map was developed.
+A key aspect which we want to implement is the ability for the end user to forecast pedestrian activity in the future.
 ```{r, fig.align = 'center', echo = FALSE, out.width = "66%"}
@@
 
-### Limitations
+## Limitations
+
+#### Data  
+The pedestrian sensor dataset is very limited in its spatial coverage. There is a possibility of data collected from WiFi devices being made available
 
@@
 #### Visualisation  
+Ideally, having a dynamic spatial model would allow us to visualise the pedestrian activity in areas not covered by the currently installed sensors.
+However, due to time constraints, the visualisation will most likely be based on a model of predictions of pedestrian counts at each sensor location.
 
-### Research Plan
+## Research Plan
+The plan to approach this research project is as follows:
+
 - Imputation that incorporates time and space
-- Predicitive model for each sensor
-- Visual methods for space-time counts
-- Shiny app - interface design, computation
-- dynamic
+- Creating a predicitive model for each sensor
+- Investiagtion of visual methods for space-time counts
+- Development of a Shiny app: interface design, computation and data backend
+- Generation of a dynamic spatial model (if time permits)
 
@@
   1. City of Melbourne, _'About Melbourne Data'_, https://data.melbourne.vic.gov.au/about,  
-  _accessed 18 April 2017_  
+   $\quad$ _accessed 18 April 2017_
+
   2. City of Melbourne, _'Pedestrian Counting System'_,  
-  http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf 
+   $\quad$ http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf  
   4 May 2015, _accessed 18 April 2017_
-  3. Healey, Christopher G.,  _'Attention and Visual Memory in Visualization and Computer Graphics'_, IEEE Transactions on Visualization and Computer Graphics,  
+
+  3. Gould, Phillip G., Koehler, Anne B., Ord, J Keith., Snyder, Ralph D., Hyndman, Rob J.,  
+   $\quad$ Vahid-Araghi, Farshid, (2008) _'Forecasting Time Series with Multiple Seasonal Patterns'_, European Journal of Operational Research 191(1), 207–220
+
+  4. Healey, Christopher G., (2012) _'Attention and Visual Memory in Visualization and Computer Graphics'_, IEEE Transactions on Visualization and Computer Graphics,  
   https://www.csc2.ncsu.edu/faculty/healey/download/tvcg.12a.pdf
+\newpage
+
+# Appendix
+
+### Plots of time series at each sensor location over entire time span from 2009:  
+```{r, echo = FALSE, cache=TRUE}
+ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)
+
+ped_plot$counts <- ped_plot$x
+
+ped_plot$Time <- parse_date(paste(ped_plot$MDate, ped_plot$Month, ped_plot$Year, sep = "-"), format = "%D-%B-%Y")
+
+ped_plot$ID <- as.integer(ped_plot$Sensor_ID)
+
+ggplot(filter(ped_plot, ID < 10)) + geom_line(aes(x = Time, y = counts)) + facet_wrap(~ Sensor_ID)
+
+ggplot(filter(ped_plot, ID > 10, ID < 20)) + geom_line(aes(x = Time, y = counts)) + facet_wrap(~ Sensor_ID)
+
+ggplot(filter(ped_plot, ID > 20, ID < 30)) + geom_line(aes(x = Time, y = counts)) + facet_wrap(~ Sensor_ID)
+
+ggplot(filter(ped_plot, ID > 30, ID < 40)) + geom_line(aes(x = Time, y = counts)) + facet_wrap(~ Sensor_ID)
+
+ggplot(filter(ped_plot, ID > 40)) + geom_line(aes(x = Time, y = counts)) + facet_wrap(~ Sensor_ID)
+
+```
+
+Looking at the plots of the monthly aggregates, we can inspect the data quickly for obvious outliers.
+After detrending the time series for both the deterministic trends (non-zero mean) and stochastic trends (weekly and daily cycles), we expect to have stationary series.
+
+Note that the Sensor ID #41 is missing from the data for unknown reasons.

### 2017-05-05 thesis/proposal_draft.Rmd
update appendix

@@
 Looking at the plots of the monthly aggregates, we can inspect the data quickly for obvious outliers.
-After detrending the time series for both the deterministic trends (non-zero mean) and stochastic trends (weekly and daily cycles), we expect to have stationary series.
+After detrending the time series for both the deterministic trends (non-zero mean) and stochastic trends (weekly and daily cycles), we expect to have a stationary series.
+Anything outside these "expected" trends from the cycles could indicate an unobserved variable (such as weather) having a casual effect on the pedestrian count.
+Large "dips" in the series, or odd values over a long time span may be indicative of missing values which need to be imputed.
 

### 2017-05-05 thesis/proposal_draft.Rmd
changed figure 6 from geom_smooth to geom_path

@@
 library(dplyr)
-library(purrr)
 library(ggplot2)
@@
 \end{center}  
-From _Figure 3_, we can observe multiple characteristics of the time series.
+From _Figure 5_, we can observe multiple characteristics of the time series.
 Firstly, we see the weekly cycle, where weekdays (Monday to Friday) have lots of activity and weekends (Saturday and Sunday) are relatively quiet.
@@
 ```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center", cache=TRUE}
-ggplot(filter(ped_df, Sensor_ID == "13")) + geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day)) +
+ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
       xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
@@
 
-In _Figure 4_, we see that, within the weekly cycle, each day has a daily sub-cycle.
-The plot in _Figure 4_ is generated using `geom_smooth` in `ggplot2`, which fitted a line to estimate hourly counts over time for each day using a GAM (generalized additive model) for smoothing.
+In _Figure 6_, we see that, within the weekly cycle, each day has a daily sub-cycle.
 We can see the patterns we observe on a weekly cycle here when we seperately estimate by the day of the week.
@@
 Splitting the data randomly 60/40 into a training set (data for model estimation) and a test set (data set to test the goodness of fit), we can get a quick evaluation of the model's performance.
+Where total number of observations is $T$, we split the data into $N \approx 0.6 \times T$ and $M \approx 0.4 \times T$ so that $T=N+M$.
+Then, we have:
+$$\text{MSE}_\text{Tr} = \frac{1}{M} \sum_{j=1}^{M} (\hat{y_t}-y_t)^2$$
 LOOCV (leave one out cross validation) is not computationally feasible due to the large number of observations.
@@
 
+## Acknowledgements
+
 ## References  
@@
   https://www.csc2.ncsu.edu/faculty/healey/download/tvcg.12a.pdf
+
+  5. Wickham, Hadley, Chang, Winston  _'ggplot2'_ , http://ggplot2.tidyverse.org
+
+  6. Wickham, Hadley, Francois, Romain  _'dplyr'_ , http://dplyr.tidyverse.org
+
 \newpage

### 2017-05-06 thesis/proposal_draft.Rmd
added acknowledgements (still WIP)

@@
 library(knitr)
-library(lubridate)
-library(reshape2)
 
@@
 \begin{center}
-Figure 1:  Locations of all pedestrian sensors around Melbourne's CBD as at April 2017
+Figure 1:  Locations of all pedestrian sensors around Melbourne's CBD as at April 2017 - Plotted using `ggmap`, the red points map the locations of the sensors.
+The majority of the pedestrian sensors are placed in the eastern side of the CBD, particularly along Swanston St
 \end{center}
@@
 \begin{center}
-Figure 3: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au
+Figure 3: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au - The UI allows the user to choose specific dates and hours of the day to visualise on a 3D perspective map
+
 \end{center}
@@
 \begin{center}
-Figure 4: An interactive data visualisation created using Tableau
+Figure 4: An interactive data visualisation created using Tableau available from 
 \end{center}
@@
 We must also note that, prior to the start of 2016, Flagstaff Station was not operating on weekends, nor public holidays.
-```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center", cache=TRUE}
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
 ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
@@
 \begin{center}
-Figure 6:  Estimated hourly counts at Flagstaff Station by Day 
+ **Figure 6**:  Hourly counts at Flagstaff Station, with colour mapped to day of the week - There are two distinctly different daily patterns.
+The typical weekday traffic pattern has two peaks at 9am and 5pm, while it remains relatively quiet over the weekend as seen by the flatter lines.
+Also note that the second peak is lower than the first peak on average.
 \end{center}
@@
 
+I would like to thank my supervisor, Dianne Cook, for her help and support throughout this project.
+I would also like to thank Earo Wang for coming along to our meetings to help out, in particular for her code for imputing missing values.
+
+This proposal is written in R Markdown (`.Rmd`) using RStudio and the code used to produce the results outputted are embedded in this document.
+The packages used for this proposal are: `readr, dplyr, ggplot2, ggmap, knitr`
 ## References  
@@
 
-  5. Wickham, Hadley, Chang, Winston  _'ggplot2'_ , http://ggplot2.tidyverse.org
+  5. Wickham, Hadley, Chang, Winston,  _'ggplot2'_ , http://ggplot2.tidyverse.org
 
-  6. Wickham, Hadley, Francois, Romain  _'dplyr'_ , http://dplyr.tidyverse.org
+  6. Wickham, Hadley, Francois, Romain,  _'dplyr'_ , http://dplyr.tidyverse.org
+
+  7. Wickham, Hadley, Kahle, David J., _'ggmap'_ , https://github.com/dkahle/ggmap
 

### 2017-05-06 thesis/proposal_draft.Rmd
edited formatting, corrected RMSFE equation

@@
 \begin{center}
-Figure 4: An interactive data visualisation created using Tableau available from 
+Figure 4: An interactive data visualisation map created using Tableau by Boreak Silk.
+Average daily volumes are mapped to size, and colours map a positive or negative change in the average for the period compared to a specified reference period.
+Note the lack of a legend to help users understand how to interpret different aesthetic mapping of the visualisation.
 \end{center}
+_Source:https://public.tableau.com/profile/boreak.silk#!/vizhome/PedestrianactivityinMelbourne/Dailyvolumesandcomparison _
 
@@
 
+To plot values on a map of the CBD, the R package `ggmap` is being used currently.
+The plots are created using the package `ggplot2`, which has the capability to generate many different types of plots.
+Choosing an appropriate combinations of plots will be important here to present the model accurately.
+
 ### Shiny App Development
@@
 ```
+
+Methods for making the app run efficiently need to be investigated, as each plot needs to be rendered after any changes to the input values in the UI (eg.
+changing time of day requires the plot to be rendered again).
+This is highly inefficient and not feasible for use on a public website.
+This problem will be exacerbated when a predictive model is added, as the model will need to predict values before the plots are rendered.
+
+Decisions regarding the design of the user interface also need to be made, specifically with the inputs which will be made available to users.
 
@@
 ### Plots of time series at each sensor location over entire time span from 2009:  
-```{r, echo = FALSE, cache=TRUE}
+```{r, echo = FALSE, cache=TRUE, out.height = "70%"}
 ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)

### 2017-05-06 thesis/proposal_draft.Rmd
updated

@@
 output:
-  pdf_document: default
+  pdf_document: 
+    keep_tex: yes
   html_document: default
@@
 
----------------------------------------------------------------------------------------
-
 ## Background
@@
 \begin{center}
-Figure 1:  Locations of all pedestrian sensors around Melbourne's CBD as at April 2017 - Plotted using `ggmap`, the red points map the locations of the sensors.
+\textbf{Figure 1}:  Locations of all pedestrian sensors around Melbourne's CBD as at April 2017 - Plotted using `ggmap`, the red points map the locations of the sensors.
 The majority of the pedestrian sensors are placed in the eastern side of the CBD, particularly along Swanston St
@@
 \begin{center}
-Figure 2: Smoothed pedestrian counts for Melbourne Central by day of the week and month of the year
+\textbf{Figure 2}: Smoothed pedestrian counts for Melbourne Central by day of the week and month of the year
 \end{center}
@@
 \begin{center}
-Figure 3: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au - The UI allows the user to choose specific dates and hours of the day to visualise on a 3D perspective map
+\textbf{Figure 3}: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au - The UI allows the user to choose specific dates and hours of the day to visualise on a 3D perspective map
 
@@
 \begin{center}
-Figure 4: An interactive data visualisation map created using Tableau by Boreak Silk.
+\textbf{Figure 4}: An interactive data visualisation map created using Tableau by Boreak Silk.
 Average daily volumes are mapped to size, and colours map a positive or negative change in the average for the period compared to a specified reference period.
@@
 \begin{center}
-Figure 5:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016)
+\textbf{Figure 5}:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016)
 \end{center}  
@@
 \begin{center}
- **Figure 6**:  Hourly counts at Flagstaff Station, with colour mapped to day of the week - There are two distinctly different daily patterns.
+ \textbf{Figure 6}:  Hourly counts at Flagstaff Station, with colour mapped to day of the week - There are two distinctly different daily patterns.
 The typical weekday traffic pattern has two peaks at 9am and 5pm, while it remains relatively quiet over the weekend as seen by the flatter lines.
@@
 After a model has been established, we need to evaluate the performance of the model.
-For this, we have decided to minimise the RMSFE (root mean square forecast error):
-$$ \text{RMSFE} = \sqrt{\frac{\sum_{t=1}^{T}(\hat{y_t} - y_t)^2}{T}}$$
+For this, we have decided to minimise the RMSFE (root mean square forecast error).
+Assuming that we estimate the model with $T$ observations, and we have $k$ points to forecast (so we need to have $T+k$ observations available), then our RMSFE is:
+$$ \text{RMSFE} = \sqrt{\frac{\sum_{t=T+1}^{T+k}(\hat{y_t} - y_t)^2}{k}}$$
 
@@
 The packages used for this proposal are: `readr, dplyr, ggplot2, ggmap, knitr`
+
 ## References  
@@
 ### Plots of time series at each sensor location over entire time span from 2009:
-```{r, echo = FALSE, cache=TRUE, out.height = "70%"}
+
+```{r, echo = FALSE, cache = TRUE, fig.height = 2.5}
 ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)
@@
 ```
-
 Looking at the plots of the monthly aggregates, we can inspect the data quickly for obvious outliers.

### 2017-05-06 thesis/proposal_draft.Rmd
update

@@
 \begin{center}
-\textbf{Figure 5}:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016)
+\textbf{Figure 5}:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016).
+Two repeating cycles can be seen, with a public holiday on 13 July showing atypical pedestrian counts for a weekday (Monday).
 \end{center}  
@@
 This proposal is written in R Markdown (`.Rmd`) using RStudio and the code used to produce the results outputted are embedded in this document.
-The packages used for this proposal are: `readr, dplyr, ggplot2, ggmap, knitr`
+The packages used for this proposal are: `readr, dplyr, ggplot2, ggmap, knitr`, while packages `lubridate, shiny` are also being used for the project iteself.
 
@@
 
-  5. Wickham, Hadley, Chang, Winston,  _'ggplot2'_ , http://ggplot2.tidyverse.org
+  5. Wickham, Hadley, Francois, Romain,  _'dplyr'_ , http://dplyr.tidyverse.org
 
-  6. Wickham, Hadley, Francois, Romain,  _'dplyr'_ , http://dplyr.tidyverse.org
+  6. Wickham, Hadley, Chang, Winston,  _'ggplot2'_ , http://ggplot2.tidyverse.org
 
   7. Wickham, Hadley, Kahle, David J., _'ggmap'_ , https://github.com/dkahle/ggmap
+
+  8. Wickham, Hadley, Hester, Jim, Francois, Romain, _'readr'_ , http://readr.tidyverse.org
+
+  9. Xie, Yihui, _'knitr'_ , https://yihui.name/knitr/
 
@@
 
-```{r, echo = FALSE, cache = TRUE, fig.height = 2.5}
+```{r, echo = FALSE}
 ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)

### 2017-05-06 thesis/proposal_draft.Rmd
update- close to final

@@
 \begin{center}
-\textbf{Figure 3}: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au - The UI allows the user to choose specific dates and hours of the day to visualise on a 3D perspective map
+\textbf{Figure 3}: Pedestrian Counting System Visualisation from www.pedestrian.melbourne.vic.gov.au - The UI allows the user to choose specific dates and hours of the day to visualise on a 3D perspective map.
+The visualisation can also animate pedestrian activity, allowing for changes over time to be easily seen.
 
@@
 However, due to time constraints, the visualisation will most likely be based on a model of predictions of pedestrian counts at each sensor location.
+Improvements to the quality of the map can be made with utilisation of the R package `leaflet`.
+Other packages being explored for visualisations also include `plotly`.
 

### 2017-05-06 presentation/proposal_pres.Rmd
initial code for imputing missing values
working draft (very much a WIP)

@@
 title: "Modelling the City of Melbourne Pedestrian Data"
+subtitle: "Thesis Proposal"
 author: "Gavin Chin, supervised by Di Cook"
-date: "04/05/2017"
-output: ioslides_presentation
+date: "8 May 2016"
+output:
+  xaringan::moon_reader:
+    lib_dir: libs
+    nature:
+      highlightStyle: github
+      highlightLines: true
+      countIncrementalSlides: false
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
+options(htmltools.dir.version = FALSE)
 library(readr)
 library(dplyr)
-library(purrr)
 library(ggplot2)
@@
 library(knitr)
+library(pander)
 library(lubridate)
+  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
 
-  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
+  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
+
+  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
+                               mean(range(ped_loc$Latitude))),
+                                zoom=14)
+```
+
+class: center
+
+# Introduction
+
+The City of Melbourne has an Open Data Platform which makes council data available to the public
+
+This project will focus on the pedestrian sensor data
+
+```{r, echo = FALSE, fig.height=5}
+ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
+              labs(title = "Melbourne Pedestrian Sensor Locations") +
+              xlab(NULL) + ylab(NULL) +
+              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(), axis.text.y = element_blank(),
+                    axis.ticks = element_blank())
+```
+
+---
+
+# How is the data collected?
+
+The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
+These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+
+---
+
+# What does the data look like?
+
+```{r}
+glimpse(ped_df)
+```
+
+```{r, echo = FALSE}
       ped_df$Year <- as.factor(ped_df$Year)
@@
 
-  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
-
-  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
-                               mean(range(ped_loc$Latitude))),
-                                zoom=14)
 ```
 
-## Introduction
+---
 
-The City of Melbourne provides an open data platform to access council datasets through its Open Data Platform  
-This project focuses primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD
+class: center, middle
 
-## Pedestrian Sensors
+## Time series at each sensor location
 
-```{r, echo = F, fig.align= "center", cache=TRUE}
-ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
-              labs(title = "Melbourne Pedestrian Sensor Locations") +
-              xlab(NULL) + ylab(NULL) +
-              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
-                    axis.text.x = element_blank(), axis.text.y = element_blank(),
-                    axis.ticks = element_blank())
-```
-
-## Time Series of Pedestrian Activity
-
-```{r, message= FALSE, warning= FALSE, fig.align= "center"}
-ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13") %>%
-  ggplot() + geom_line(aes(x = MTime, y = Hourly_Counts)) +
+```{r, message= FALSE, warning= FALSE, fig.height = 5, fig.width = 12, echo = FALSE, fig.align= "center"}
+bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
+ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) + ggtitle("Hourly Pedestrian Count at Flagstaff Station", subtitle = "during June 2016") +
       xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
@@
 
-## Slide with Plot
+---
+# Why do we want to model pedestrian traffic?
 
-```{r pressure}
-plot(pressure)
+Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector
+
+Different ways it can be used:
+
+  Government example uses:  
+    - Infrastructure planning
+    - Security planning
+  Private example uses:
+    - Marketing campaign planning
+    - Resource management
+    - Investment planning
+
+  and many other various ways..
+
+---
+class: center, inverse
+
+# Methodology
+
+The project has been split into multiple steps:  
+  - Data Cleaning
+  - Modelling
+  - Visualisation
+  - Shiny app development
+
+---
+
+# Data Cleaning
+
+While the data variables themselves are relatively tidy, there is a large number of observations we need to impute
+
+--
+
+```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12}
+ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)
+
+ped_plot$counts <- ped_plot$x
+
+ped_plot$Time <- parse_date(paste(ped_plot$MDate, ped_plot$Month, ped_plot$Year, sep = "-"), format = "%D-%B-%Y")
+
+ped_plot$ID <- as.integer(ped_plot$Sensor_ID)
+
+ped_plot$Sensor_ID <- as.factor(ped_plot$Sensor_ID)
+
+ggplot(filter(ped_plot)) + geom_line(aes(x = Time, y = counts, colour = Sensor_ID)) + theme_light() + theme(legend.position = "none") + labs(x = "Year", y = "Pedestrian Counts") + ggtitle("Monthly Total Counts by Sensor")
+
 ```
+
+---
+
+# R Code
+
+```{r comment='#'}
+# a boring regression
+fit = lm(dist ~ 1 + speed, data = cars)
+coef(summary(fit))
+dojutsu = c('地爆天星', '天照', '加具土命', '神威', '須佐能乎', '無限月読')
+grep('天', dojutsu, value = TRUE)
+```
+
+---
+
+# R Plots
+
+```{r cars, fig.height=4, dev='svg'}
+par(mar = c(4, 4, 1, .1))
+plot(cars, pch = 19, col = 'darkgray', las = 1)
+abline(fit, lwd = 2)
+```
+
+---
+
+# HTML Widgets
+
+I have not thoroughly tested HTML widgets against **xaringan**. Some may work well, and some may not. It is a little tricky.
+
+Similarly, the Shiny mode (`runtime: shiny`) does not work.
+I might get these issues fixed in the future, but these are not of high priority to me.
+I never turn my presentation into a Shiny app.
+When I need to demonstrate more complicated examples, I just launch them separately.
+It is convenient to share slides with other people when they are plain HTML/JS applications.
+
+See the next page for two HTML widgets.
+
+---
+
+```{r out.width='100%', fig.height=6, eval=require('leaflet')}
+library(leaflet)
+leaflet() %>% addTiles() %>% setView(mean(ped_loc$Longitude), mean(ped_loc$Latitude), zoom = 14)
+```
+
+---
+
+```{r eval=require('DT'), tidy=FALSE}
+DT::datatable(
+  head(iris, 10),
+  fillContainer = FALSE, options = list(pageLength = 5)
+)
+```
+
+---
+
+# Some Tips
+
+- When you use the "Infinite Moon Reader" addin in RStudio, your R session will be blocked by default.
+You can click the red button on the right of the console to stop serving the slides, or use the _daemonized_ mode so that it does not block your R session.
+To do the latter, you can set the option
+
+    ```r
+    options(servr.daemon = TRUE)
+    ```
+
+    in your current R session, or in `~/.Rprofile` so that it is applied to all future R sessions. I do the latter by myself.
+
+    To know more about the web server, see the [**servr**](https://github.com/yihui/servr) package.
+
+--
+
+- Do not forget to try the `yolo` option of `xaringan::moon_reader`.
+
+    ```yaml
+    output:
+      xaringan::moon_reader:
+        yolo: true
+    ```
+
+---
+
+# Some Tips
+
+- Slides can be automatically played if you set the `autoplay` option under `nature`, e.g. go to the next slide every 30 seconds (this can be useful for lightning talks):
+
+    ```yaml
+    output:
+      xaringan::moon_reader:
+        nature:
+          autoplay: 30000
+    ```
+
+--
+
+- The option `highlightLines: true` of `nature` will highlight code lines that start with `*` or are wrapped in `{{ }}`;
+
+    ```yaml
+    output:
+      xaringan::moon_reader:
+        nature:
+          highlightLines: true
+    ```
+
+    See examples on the next page.
+
+---
+
+# Some Tips
+
+.pull-left[
+An example using a leading `*`:
+
+    ```r
+    if (TRUE) {
+    ** message("Very important!")
+    }
+    ```
+Output:
+```r
+if (TRUE) {
+* message("Very important!")
+}
+```
+
+This is invalid R code, so it is a plain fenced code block that is not executed.
+]
+
+.pull-right[
+An example using `{{}}`:
+
+    `r ''````{r tidy=FALSE}
+    if (TRUE) {
+    *{{ message("Very important!") }}
+    }
+    ```
+Output:
+```{r tidy=FALSE}
+if (TRUE) {
+{{ message("Very important!") }}
+}
+```
+
+It is valid R code so you can run it.
+]
+
+---
+
+# Some Tips
+
+- To make slides work offline, you need to download a copy of remark.js in advance, because **xaringan** uses the online version by default (see the help page `?xaringan::moon_reader`).
+
+- You can use `xaringan::summon_remark()` to download the latest or a specified version of remark.js. By default, it is downloaded to `libs/remark-latest.min.js`.
+
+- Then change the `chakra` option in YAML to point to this file, e.g.
+
+    ```yaml
+    output:
+      xaringan::moon_reader:
+        chakra: libs/remark-latest.min.js
+    ```
+
+- If you used Google fonts in slides (the default theme uses _Yanone Kaffeesatz_, _Droid Serif_, and _Source Code Pro_), they won't work offline unless you download or install them locally.
+The Heroku app [google-webfonts-helper](https://google-webfonts-helper.herokuapp.com/fonts) can help you download fonts and generate the necessary CSS.
+
+---
+
+# CSS
+
+Among all options in `xaringan::moon_reader`, the most challenging but perhaps also the most rewarding one is `css`, because it allows you to customize the appearance of your slides using any CSS rules or hacks you know.
+
+You can see the default CSS file [here](https://github.com/yihui/xaringan/blob/master/inst/rmarkdown/templates/xaringan/resources/example.css).
+You can completely replace it with your own CSS files, or define new rules to override the default.
+
+---
+
+# CSS
+
+For example, suppose you want to change the font for code from the default "Source Code Pro" to "Ubuntu Mono". You can create a CSS file named, say, `ubuntu-mono.css`:
+
+```css
+@import url(https://fonts.googleapis.com/css?family=Ubuntu+Mono:400,700,400italic);
+
+.remark-code, .remark-inline-code { font-family: 'Ubuntu Mono'; }
+```
+
+Then set the `css` option in the YAML metadata:
+
+```yaml
+output:
+  xaringan::moon_reader:
+    css: ["default", "ubuntu-mono.css"]
+```
+
+Here I assume `ubuntu-mono.css` is under the same directory as your Rmd.
+
+---
+class: inverse, middle, center
+background-image: url(https://upload.wikimedia.org/wikipedia/commons/3/39/Naruto_Shiki_Fujin.svg)
+background-size: contain
+
+# Naruto
+
+---
+background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)
+background-size: 100px
+background-position: 90% 8%
+
+# Sharingan
+
+The R package name **xaringan** was derived<sup>1</sup> from **Sharingan**, a dōjutsu in the Japanese anime _Naruto_ with two abilities:
+
+- the "Eye of Insight"
+
+- the "Eye of Hypnotism"
+
+I think a presentation is basically a way to communicate insights to the audience, and a great presentation may even "hypnotize" the audience.<sup>2,3</sup>
+
+.footnote[
+[1] In Chinese, the pronounciation of _X_ is _Sh_ /ʃ/ (as in _shrimp_). Now you should have a better idea of how to pronounce my last name _Xie_.
+
+[2] By comparison, bad presentations only put the audience to sleep.
+
+[3] Personally I find that setting background images for slides is a killer feature of remark.js. It is an effective way to bring visual impact into your presentations.
+]
+
+---
+
+# Naruto terminology
+
+The **xaringan** package borrowed a few terms from Naruto, such as
+
+- [Sharingan](http://naruto.wikia.com/wiki/Sharingan) (写輪眼; the package name)
+
+- The [moon reader](http://naruto.wikia.com/wiki/Moon_Reader) (月読; an attractive R Markdown output format)
+
+- [Chakra](http://naruto.wikia.com/wiki/Chakra) (查克拉; the path to the remark.js library, which is the power to drive the presentation)
+
+- [Nature transformation](http://naruto.wikia.com/wiki/Nature_Transformation) (性質変化; transform the chakra by setting different options)
+
+- The [infinite moon reader](http://naruto.wikia.com/wiki/Infinite_Tsukuyomi) (無限月読; start a local web server to continuously serve your slides)
+
+- The [summoning technique](http://naruto.wikia.com/wiki/Summoning_Technique) (download remark.js from the web)
+
+You can click the links to know more about them if you want. The jutsu "Moon Reader" may seem a little evil, but that does not mean your slides are evil.
+
+---
+
+class: center
+
+# Hand seals (印)
+
+Press `h` or `?` to see the possible ninjutsu you can use in remark.js.
+
+![](https://upload.wikimedia.org/wikipedia/commons/7/7e/Mudra-Naruto-KageBunshin.svg)
+
+---
+
+class: center, middle
+
+# Thanks!
+
+Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).
+
+The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).

### 2017-05-06 thesis/proposal_draft.Rmd
fixed issues: shiny/r not ending properly, removed equation reference + corrected equation (missing log(lambda) ), finished sentence for limitations of data

@@
 This may be in the form of selecting a time period for the data, or changing how a certain plot is generated (size, colour, etc..).
-Another key
+A feature which we can implement is an interactive forecast, allowing the user to change some of the input variables in the predicitive model.
 
@@
 \begin{align*}
-\text{HourlyCounts} & \sim\text{Year}\times\text{HDay}\times\text{Time}+\text{Month}\tag{1}\\
-\Rightarrow\qquad\lambda_{i} & =\beta_{0}+\beta_{1}\text{Year}\times\text{HDay}\times\text{Time}+\beta_{2}\text{Month}\\
+\text{HourlyCounts} & \sim\text{Year}\times\text{HDay}\times\text{Time}+\text{Month}\\
+\Rightarrow\qquad \log{\lambda_{i}} & =\beta_{0}+\beta_{1}\text{Year}\times\text{HDay}\times\text{Time}+\beta_{2}\text{Month}\\
 \text{HourlyCounts}_{i} & =\text{Poisson}\left(\lambda_{i}\right)
@@
 #### Data  
-The pedestrian sensor dataset is very limited in its spatial coverage. There is a possibility of data collected from WiFi devices being made available
+The pedestrian sensor dataset is very limited in its spatial coverage.
+There is a possibility of data collected from WiFi devices being made available, which would increase both the spatial coverage and the resolution (counts aggregated every 15 minutes instead of hourly).
 

### 2017-05-06 presentation/proposal_pres.Rmd
slides update

@@
 ```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
-options(htmltools.dir.version = FALSE)
+options(htmltools.dir.version = FALSE, scipen = 100)
 library(readr)
@@
 ---
-class: center, inverse
+class: inverse
 
@@
 
-The project has been split into multiple steps:  
+### The project has been split into multiple steps:
+
   - Data Cleaning
@@
 
-While the data variables themselves are relatively tidy, there is a large number of observations we need to impute
+### While the data variables themselves are relatively tidy, there is a large number of observations we need to impute:
 
@@
 ```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12}
-ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)
+ped_plot <- aggregate(x = ped_df$Hourly_Counts,
+                      by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month),
+                      FUN = sum)
 
@@
 
-ped_plot$Time <- parse_date(paste(ped_plot$MDate, ped_plot$Month, ped_plot$Year, sep = "-"), format = "%D-%B-%Y")
+ped_plot$Time <- parse_date(paste(ped_plot$MDate, ped_plot$Month, ped_plot$Year,
+                                  sep = "-"), format = "%D-%B-%Y")
 
@@
 
-ggplot(filter(ped_plot)) + geom_line(aes(x = Time, y = counts, colour = Sensor_ID)) + theme_light() + theme(legend.position = "none") + labs(x = "Year", y = "Pedestrian Counts") + ggtitle("Monthly Total Counts by Sensor")
+ggplot(filter(ped_plot)) + geom_line(aes(x = Time, y = counts, colour = Sensor_ID)) + theme_light() +
+  theme(legend.position = "none") + labs(x = "Year", y = "Pedestrian Counts") + ggtitle("Monthly Total Counts by Sensor")
 

### 2017-05-06 presentation/proposal_pres.Rmd
removed template content, WIP - going through methodology

--- 2017-05-03 presentation/proposal_pres.Rmd
@@
 title: "Modelling the City of Melbourne Pedestrian Data"
+subtitle: "Thesis Proposal"
 author: "Gavin Chin, supervised by Di Cook"
-date: "04/05/2017"
-output: ioslides_presentation
+date: "8 May 2016"
+output:
+  xaringan::moon_reader:
+    lib_dir: libs
+    nature:
+      highlightStyle: github
+      highlightLines: true
+      countIncrementalSlides: false
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
+options(htmltools.dir.version = FALSE, scipen = 100)
 library(readr)
 library(dplyr)
-library(purrr)
 library(ggplot2)
@@
 library(knitr)
+library(pander)
 library(lubridate)
+  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
 
-  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
+  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
+
+  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
+                               mean(range(ped_loc$Latitude))),
+                                zoom=14)
+```
+
+class: center
+
+# Introduction
+
+The City of Melbourne has an Open Data Platform which makes council data available to the public
+
+This project will focus on the pedestrian sensor data
+
+```{r, echo = FALSE, fig.height=5}
+ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
+              labs(title = "Melbourne Pedestrian Sensor Locations") +
+              xlab(NULL) + ylab(NULL) +
+              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(), axis.text.y = element_blank(),
+                    axis.ticks = element_blank())
+```
+
+---
+
+# How is the data collected?
+
+The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
+These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+
+---
+
+# What does the data look like?
+
+```{r}
+glimpse(ped_df)
+```
+
+--
+
+#### Basic variables available are: time and date, sensor ID and hourly counts
+
+```{r, echo = FALSE}
       ped_df$Year <- as.factor(ped_df$Year)
@@
 
-## Slide with Plot
+---
 
-```{r pressure}
-plot(pressure)
+# Why do we want to model pedestrian traffic?
+
+### Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector
+
+#### Different ways it can be used:
+
+  Government example uses:  
+    - Infrastructure planning
+    - Security planning
+  Private example uses:
+    - Marketing campaign planning
+    - Resource management
+    - Investment planning
+
+  and many other various ways..
+
+---
+class: inverse, middle, center
+
+# Methodology
+
+### The project has been split into multiple steps:  
+#### Data Cleaning
+#### Modelling
+#### Visualisation
+#### Shiny app development
+
+---
+
+# Data Cleaning: Problems in the data
+
+### While the data variables themselves are relatively tidy, there is a large number of observations we need to impute:
+
+--
+
+```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12}
+ped_plot <- aggregate(x = ped_df$Hourly_Counts,
+                      by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month),
+                      FUN = sum)
+
+ped_plot$counts <- ped_plot$x
+
+ped_plot$Time <- parse_date(paste(ped_plot$MDate, ped_plot$Month, ped_plot$Year,
+                                  sep = "-"), format = "%D-%B-%Y")
+
+ped_plot$ID <- as.integer(ped_plot$Sensor_ID)
+
+ped_plot$Sensor_ID <- as.factor(ped_plot$Sensor_ID)
+
+ggplot(filter(ped_plot)) + geom_line(aes(x = Time, y = counts, colour = Sensor_ID)) + theme_light() +
+  theme(legend.position = "none") + labs(x = "Year", y = "Pedestrian Counts") + ggtitle("Monthly Total Counts by Sensor")
+
 ```
+
+---
+
+# Data Cleaning: Imputing the missings
+
+#### A simple way to impute these missings is estimating a Poisson regression using GLM in R:
+
+```{r, eval = FALSE, tidy = TRUE}
+missing_glm <- glm(Hourly_Counts ~ Year * HDay * Time + Month)
+```
+
+This can be expressed as:
+
+$$\text{HourlyCounts} \sim\text{Year}\times\text{HDay}\times\text{Time}+\text{Month}$$
+$$\Rightarrow\qquad \log{\lambda_{i}}  =\beta_{0}+\beta_{1}\text{Year}\times\text{HDay}\times\text{Time}+\beta_{2}\text{Month}$$
+$$\text{HourlyCounts}_{i}  =\text{Poisson}\left(\lambda_{i}\right)$$

### 2017-05-06 presentation/proposal_pres.Rmd
reordered some slides, added more plots

@@
 
+# Why do we want to model pedestrian traffic?
+
+### Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector
+
+#### Different ways it can be used:
+
+  Government example uses:  
+    - Infrastructure planning
+    - Security planning
+  Private example uses:
+    - Marketing campaign planning
+    - Resource management
+    - Investment planning
+
+  and many other various ways..
+
+---
+
 # What does the data look like?
@@
 
-# Why do we want to model pedestrian traffic?
+## Multiple seasonal patterns
 
-### Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector
-
-#### Different ways it can be used:
-
-  Government example uses:  
-    - Infrastructure planning
-    - Security planning
-  Private example uses:
-    - Marketing campaign planning
-    - Resource management
-    - Investment planning
-
-  and many other various ways..
+#### For example, at Flagstaff Station, we notice a couple of key patterns:
+- Weekly cycle
+```{r, message= FALSE, warning= FALSE, fig.height = 4, fig.width = 12, echo = FALSE, fig.align= "center"}
+bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
+ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) + ggtitle("Hourly Pedestrian Count at Flagstaff Station", subtitle = "during June 2016") +
+      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
+```
+- Daily sub-cycle
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
+ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
+      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
+```
 
@@
 $$\text{HourlyCounts}_{i}  =\text{Poisson}\left(\lambda_{i}\right)$$
+
+--
+
+This needs to be estimated for each location, due to the different characteristics
+
+---

### 2017-05-06 presentation/proposal_pres.Rmd
corrected date, added explanation to imputation model

--- 2017-05-06 presentation/proposal_pres.Rmd
@@
 author: "Gavin Chin, supervised by Di Cook"
-date: "8 May 2016"
+date: "8 May 2017"
 output:
@@
 
-# Why do we want to model pedestrian traffic?
+## Multiple seasonal patterns
 
-### Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector
-
-#### Different ways it can be used:
-
-  Government example uses:  
-    - Infrastructure planning
-    - Security planning
-  Private example uses:
-    - Marketing campaign planning
-    - Resource management
-    - Investment planning
-
-  and many other various ways..
+#### For example, at Flagstaff Station, we notice a couple of key patterns:
+- Weekly cycle
+```{r, message= FALSE, warning= FALSE, fig.height = 3, fig.width = 12, echo = FALSE, fig.align= "center"}
+bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
+ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) + ggtitle("Hourly Pedestrian Count at Flagstaff Station", subtitle = "during June 2016") +
+      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
+```
+- Daily sub-cycle
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
+ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
+      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
+```
 
@@
 $$\text{HourlyCounts}_{i}  =\text{Poisson}\left(\lambda_{i}\right)$$
+
+--
+
+- This model specification implies an interaction effect between the year, time of the day, and day of the week (or if the day is a public holiday)
+  + Month of the year is modelled as an additive effect
+
+- The model needs to be estimated for each location, due to the different characteristics between locations
+
+---

### 2017-05-07 presentation/proposal_pres.Rmd
improved plot resolution using fig.retina

@@
 
-```{r, echo = FALSE, fig.height=5}
+```{r, echo = FALSE, fig.height=5, fig.retina = 2}
 ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
@@
 
-```{r, message= FALSE, warning= FALSE, fig.height = 5, fig.width = 12, echo = FALSE, fig.align= "center"}
-bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
-ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) + ggtitle("Hourly Pedestrian Count at Flagstaff Station", subtitle = "during June 2016") +
+```{r, message= FALSE, warning= FALSE, fig.height = 5, fig.width = 12, echo = FALSE, fig.retina = 2, fig.align= "center"}
+bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "40")
+ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts), colour = "black") + ggtitle(paste("Hourly Pedestrian Count at", unique(bsplot$Sensor_Name), sep = " "), subtitle = "during June 2016") +
       xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
@@
 - Weekly cycle
-```{r, message= FALSE, warning= FALSE, fig.height = 3, fig.width = 12, echo = FALSE, fig.align= "center"}
+```{r, message= FALSE, warning= FALSE, fig.height = 3, fig.width = 12, echo = FALSE, fig.retina = 2, fig.align= "center"}
 bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
@@
 - Daily sub-cycle
-```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
+```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center", fig.retina = 2}
 ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
@@
 
-```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12}
+```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12, fig.retina = 2}
 ped_plot <- aggregate(x = ped_df$Hourly_Counts,

### 2017-05-07 presentation/proposal_pres.Rmd
updates

@@
 The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD
+
 These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+
+```{r echo = FALSE}
+include_graphics("../img/ped_sensors.png")
+```
 

### 2017-05-07 thesis/proposal_draft.Rmd
updates

@@
       xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
-
 ```

### 2017-05-07 code/work.R

@@ -7,6 +7,7 @@ library(ggplot2)
 library(knitr)
 library(lubridate)
 library(tidytext)
+library(speedglm)
 
 ped_df <- read_csv("data/Pedestrian_volume__updated_monthly_.csv")
   ped_df$Year <- as.factor(ped_df$Year)

### 2017-05-07 presentation/proposal_pres.Rmd
update: added diagram from city of melb. pdf

--- 2017-05-07 presentation/proposal_pres.Rmd
@@
 ---
+class: center
 
@@
 The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
-These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+- These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+
+![figure of pedestrian sensors](../img/ped_sensors.png)
+
+_source: http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf _
 

### 2017-05-07 thesis/proposal_draft.Rmd
set alpha = 1 in plot legend for figure 6

--- 2017-05-06 thesis/proposal_draft.Rmd
@@
 ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
-      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
-
+      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light() + guides(colour = guide_legend(override.aes = list(alpha = 1)))
 ```

### 2017-05-07 presentation/proposal_pres.Rmd
fixed diagram of sensors, alpha = 1 in geom_path plot of daily patterns

@@
 
-# How is the data collected?
+## How is the data collected?
 
@@
 
-_source: http://www.pedestrian.melbourne.vic.gov.au/pdf/CityofMelbournePedestrianCountingSystem.pdf _
+_source: http://www.pedestrian.melbourne.vic.gov.au/ _
 
@@
 ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
-      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()
+      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light() +
+      guides(colour = guide_legend(override.aes = list(alpha = 1)))
 ```

### 2017-05-07 thesis/proposal_draft.Rmd
spell checked

@@
 
-These sensors are placed above flowing traffic either under an awning or on a street pole, recording movements from multiple directions.
-The pedestrian counting system was introduced by the council with the objectives to help inform decisions about urban planning and management, identify opportunities to improve city walkability and transport, to measure the impact of events and specific marketing campaings, as well as to monitor retail activity and aiding the business community with marketing and resource management.
+These sensors are placed above flowing traffic, either under an awning or on a street pole, recording movements from multiple directions.
+The pedestrian counting system was introduced by the council with the objectives: to help inform decisions about urban planning and management, identify opportunities to improve city walk-ability and transport, to measure the impact of events and specific marketing campaigns, as well as to monitor retail activity and aiding the business community with marketing and resource management.
 
@@
 Pedestrian activity is represented by two aesthetic features - size and colour.
-The markers, in the shape of thermometers, change in height based on the absolute pedestrian count, with a colour scale from blue to white to yellow, analogous to temperature, to map the relative diference in activity.
+The markers, in the shape of thermometers, change in height based on the absolute pedestrian count, with a colour scale from blue to white to yellow, analogous to temperature, to map the relative difference in activity.
 The colour changes when the hourly counts deviate by more than 10% from the average count over the previous four weeks for the same day of the week and time of the day.
@@
 The main advantage of an animated/dynamic plot is that the changes over time can be visualised effectively for multiple sensor locations.
-This is particularly useful in reducing change blindness[^3] in the peception of the data visualisation, as it is easier for people viewing the plot to see relative changes from one hour to the next at a certain location.
+This is particularly useful in reducing change blindness[^3] in the perception of the data visualisation, as it is easier for people viewing the plot to see relative changes from one hour to the next at a certain location.
 However, a particular aspect which needs to be improved on is the ability to visualise the pedestrian activity levels around the whole city at once.
@@
 In _Figure 6_, we see that, within the weekly cycle, each day has a daily sub-cycle.
-We can see the patterns we observe on a weekly cycle here when we seperately estimate by the day of the week.
+We can see the patterns we observe on a weekly cycle here when we separately estimate by the day of the week.
 What we can see more clearly is also the difference in the pattern of pedestrian activity for each hour of the day, where there are two main peaks per weekday.
@@
 
-In particular, the objective of this project is to have the modelled pedestrian data presented as an interactive web app using Shiny.
+In particular, the objective of this project is to have the modeled pedestrian data presented as an interactive web app using Shiny.
 A key part which makes the visualisations interactive is the ability to have user inputs.
 This may be in the form of selecting a time period for the data, or changing how a certain plot is generated (size, colour, etc..).
-A feature which we can implement is an interactive forecast, allowing the user to change some of the input variables in the predicitive model.
+A feature which we can implement is an interactive forecast, allowing the user to change some of the input variables in the predictive model.
 
@@
 
-In equation \ref{poisson_model}, the variable `HDay` refers to a factor variable with 8 factor levels: the 7 days of the week and a factor for a public holiday.
+In the equation above, the variable `HDay` refers to a factor variable with 8 factor levels: the 7 days of the week and a factor for a public holiday.
 This is generated such that, if the day is a public holiday, then `HDay = "Holiday"`, regardless of the day of the week.
@@
 With multiple sensors being installed years later after the initial 28 sensors were installed and capturing data, we do not have a complete set of observations over the total observation period at each sensor location.
-Consequently, we will only use data from 2014 onwards, as this appears to be the most complete and comprehensive time period where the obeservations provide useful information.
+Consequently, we will only use data from 2014 onwards, as this appears to be the most complete and comprehensive time period where the observations provide useful information.
 
@@
 For example, Melbourne Central, a retail shopping area, is affected by `Month`, but most likely not weather.
-From _Figure 2_, we see that the average peak actvity on Fridays and Saturdays in January is significantly lower than other months of the year, while weather conditions are similar to December and February.
+From _Figure 2_, we see that the average peak activity on Fridays and Saturdays in January is significantly lower than other months of the year, while weather conditions are similar to December and February.
 This is most likely to be explained by typical seasonal consumer behaviour: peak consumer spending occurs in December due to Christmas gift shopping and Boxing Day sales.
@@
 
+Other models, such as the MS (Multiple Seasonal) process, need to be investigated for its use.
+
 #### Visualisation  
@@
 - Imputation that incorporates time and space
-- Creating a predicitive model for each sensor
-- Investiagtion of visual methods for space-time counts
+- Creating a predictive model for each sensor
+- Investigation of visual methods for space-time counts
 - Development of a Shiny app: interface design, computation and data backend
@@
 This proposal is written in R Markdown (`.Rmd`) using RStudio and the code used to produce the results outputted are embedded in this document.
-The packages used for this proposal are: `readr, dplyr, ggplot2, ggmap, knitr`, while packages `lubridate, shiny` are also being used for the project iteself.
+The packages used for this proposal are: `readr, dplyr, ggplot2, ggmap, knitr`, while packages `lubridate, shiny` are also being used for the project itself.
 

### 2017-05-07 presentation/proposal_pres.Rmd
more slides - still WIP

@@
 
-```{r, echo = FALSE, fig.height=5, fig.retina = 2}
+```{r, echo = FALSE, fig.height=5, fig.retina = 3}
 ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
@@
 - These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+- The data is available as a `.csv` at https://data.melbourne.vic.gov.au/, and is updated monthly
 
@@
   and many other various ways..
-
 ---
 
-# What does the data look like?
+class: center, middle
 
-```{r}
-glimpse(ped_df)
-```
-
---
-
-#### Basic variables available are: time and date, sensor ID and hourly counts
+## Time series at each sensor location
 
@@
       ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])
+      for (i in 1:dim(ped_loc)[1])
+      {
+      ped_df$Lat[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Latitude[i]
+      ped_df$Lon[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Longitude[i]
+      }
 
 ```
-
----
-
-class: center, middle
-
-## Time series at each sensor location
 
@@
       xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
+```
+
+#### Basic variables available are: time and date, sensor ID and hourly counts
+
+```{r, echo = FALSE}
+head(ped_df[, -c(11,12)], 1)
 ```
@@
 ---
+
+# Modelling
+
+- Main purpose of the model is to forecast pedestrian counts
+- Data used for estimatation (training dataset) will be the period between 2014-2016
+  + The data before 2014 is not stable, and also lacks data from newly installed sensors
+- Data used to evaluate the forecasting performance will be the 2017 data available
+
+- Due to the multiple seasonal patterns in the time series, typical models like AR, ARMA and ARIMA do not work very well
+  + Modelling using other sources of data is being explored, in particular weather data
+  + Alternative models, such as a Multiple Seasonal (MS) process available through the `forecast` R package, should also be investigated
+
+---
+
+class: center, middle
+
+## Visualisation: Currently available
+#### Official City of Melbourne data visualisation
+
+```{r, echo = FALSE, out.height= "400px", out.width = "540px"}
+include_graphics("../img/ped_vis_melb.PNG")
+```
+
+---
+class: center, middle
+
+## Visualisation
+
+With this project, we want to improve the way we visualise Melbourne's pedestrian activity, and use a predicitive model to get a better understanding of how Melbourne operates
+
+```{r, echo = FALSE}
+ggmap(melb) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), alpha = 0.02, data = filter(ped_df, Year == "2016", Month == "March", Time == "9"), colour = "violetred") + scale_size_area(max_size = 10) + guides(colour = guide_legend(override.aes = list(alpha = 1)))
+```

### 2017-05-07 presentation/proposal_pres.Rmd
still need to add: shiny app info, limitations, q&a slide, acknowledgements, more visualisation

@@
                                 zoom=14)
+  melbbw <- get_map(location=c(mean(range(ped_loc$Longitude)),
+                               mean(range(ped_loc$Latitude))),
+                                zoom=14,
+                    color = "bw")
 ```
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, message=FALSE, warning=FALSE}
       ped_df$Year <- as.factor(ped_df$Year)
@@
       ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])
+
+      ped_df$Lat <- 0
+      ped_df$Lon <- 0
+
       for (i in 1:dim(ped_loc)[1])
@@
 ```{r, eval = FALSE, tidy = TRUE}
-missing_glm <- glm(Hourly_Counts ~ Year * HDay * Time + Month)
+missing_glm <- glm(Hourly_Counts ~ Year * HDay * Time + Month,
+                   data = ped_df, family = "poisson")
 ```
@@
 ---
+class: middle, center
+
+## Modelling: evaluating model performance
+The criteria we will use to evaluate the model's forecasting performance is the root mean square forecast error:
+$$ \text{RMSFE} = \sqrt{\frac{\sum_{t=T+1}^{T+k}(\hat{y_t} - y_t)^2}{k}}$$
+where the model is estimated with $T$ observations, and we have $k$ points to forecast (so we need to have $T+k$ observations available)
+- evaluating performance of the model on its forecasting accuracy is important, as it will reveal overfitting of data
+
+---
 
@@
 
-With this project, we want to improve the way we visualise Melbourne's pedestrian activity, and use a predicitive model to get a better understanding of how Melbourne operates
+With this project, we want to improve the way we visualise Melbourne's pedestrian activity
+```{r, echo = FALSE, fig.retina = 3, out.height="500px", fig.align = 'center'}
+ggmap(melbbw) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), colour = "red", alpha = 0.03, data = filter(ped_df, Year == "2016", Month == "March", Time == "9")) + scale_size_area(max_size = 10) + scale_color_distiller(palette = "PuOr") + guides(colour = guide_legend(override.aes = list(alpha = 1))) + theme_light() + labs(x = NULL, y = NULL)
+```
 
-```{r, echo = FALSE}
-ggmap(melb) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), alpha = 0.02, data = filter(ped_df, Year == "2016", Month == "March", Time == "9"), colour = "violetred") + scale_size_area(max_size = 10) + guides(colour = guide_legend(override.aes = list(alpha = 1)))
+---
+class: center
+
+## Visualisation: different methods
+
+```{r, echo = FALSE, message=FALSE, out.height="500px", fig.retina=2}
+  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
+    geom_smooth(aes(x = Timeplot, y = Hourly_Counts, colour = Month), se = F) + 
+    xlab("Time of the Day") + ylab("Pedestrian Count") + 
+    ggtitle(paste("Monthly Average Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +
+    facet_wrap(~ Day, nrow = 2) + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
+    scale_x_continuous(limits = c(0,24))
 ```

### 2017-05-07 presentation/proposal_pres.Rmd
majority of content complete - just needs some polishing up and potentially some trimming

@@
 library(knitr)
-library(pander)
 library(lubridate)
@@
 
---
-
 ```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12, fig.retina = 2}
@@
   + Alternative models, such as a Multiple Seasonal (MS) process available through the `forecast` R package, should also be investigated
+  + Machine learning techniques, such as clustering and other non-parametric estimation methods, will also be considered
 
@@
 
-```{r, echo = FALSE, message=FALSE, out.height="500px", fig.retina=2}
+```{r, echo = FALSE, message=FALSE, out.height="480px", fig.retina=2, fig.width = 12}
   ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
@@
 ```
+
+---
+
+class: center
+
+## Visualisation: different methods
+
+```{r, echo = FALSE, message=FALSE, out.height="480px", fig.retina=2, fig.width = 12}
+  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
+    geom_density(aes(x = Hourly_Counts, colour = Day), adjust = 2, alpha = 0.2) + 
+    xlab("Time of the Day") + ylab("Density") + 
+    ggtitle(paste("Kernel Density Estimates of Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +
+    theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")
+```
+
+---
+class: middle
+
+## Visualisation: importance of design
+
+As we can see, there are many different ways to present the same data
+
+When designing visualisations for pedestrian data, we need to take into consideration:
+- Preattentive processing
+- Change blindness
+
+Need to communicate the intended message to the person viewing the plots
+
+---
+class: center
+
+## Interactive Visualisation: Shiny apps
+
+The package `shiny` allows us to build web apps using R's capabilities in the background
+We can add interactive inputs to visualisations.
+
+An example of what we can make:
+
+---
+
+class: inverse, middle
+
+# Research Plan
+
+### The plan to approach this research project is as follows:
+
+- Imputation that incorporates time and space
+- Creating a predictive model for each sensor
+- Investigation of visual methods for space-time counts
+- Development of a Shiny app: interface design, computation and data backend
+- Generation of a dynamic spatial model (if time permits)
+
+---
+
+class: inverse, middle, center
+
+# Questions or comments?
+
+---
+
+# Acknowledgements
+
+##### Thanks to Di Cook and Earo Wang for helping me thus far in the project!
+
+This presentation was written in R Markdown using the template `xaringan`.
+Other packages used for this presentation are: `readr, dplyr, ggplot2, ggmap, knitr, lubridate`

### 2017-05-07 presentation/proposal_pres.Rmd
corrections made to some errors

@@
 
-```{r, eval = FALSE, tidy = TRUE}
+```{r, eval = FALSE}
 missing_glm <- glm(Hourly_Counts ~ Year * HDay * Time + Month,
@@
 ```{r, echo = FALSE, fig.retina = 3, out.height="500px", fig.align = 'center'}
-ggmap(melbbw) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), colour = "red", alpha = 0.03, data = filter(ped_df, Year == "2016", Month == "March", Time == "9")) + scale_size_area(max_size = 10) + scale_color_distiller(palette = "PuOr") + guides(colour = guide_legend(override.aes = list(alpha = 1))) + theme_light() + labs(x = NULL, y = NULL)
+ggmap(melbbw) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), colour = "red", alpha = 0.03, data = filter(ped_df, Year == "2016", Month == "January", Time == "9")) + scale_size_area(max_size = 10) + scale_color_distiller(palette = "PuOr") + guides(colour = guide_legend(override.aes = list(alpha = 1))) + theme_light() + labs(x = NULL, y = NULL)
 ```
@@
     geom_density(aes(x = Hourly_Counts, colour = Day), adjust = 2, alpha = 0.2) + 
-    xlab("Time of the Day") + ylab("Density") + 
+    xlab("Pedestrian Counts") + ylab("Density") + 
     ggtitle(paste("Kernel Density Estimates of Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +

### 2017-05-07 presentation/proposal_pres.Rmd
changed research plan to numbering

@@
 
-- Imputation that incorporates time and space
-- Creating a predictive model for each sensor
-- Investigation of visual methods for space-time counts
-- Development of a Shiny app: interface design, computation and data backend
-- Generation of a dynamic spatial model (if time permits)
+1. Imputation that incorporates time and space
+2. Creating a predictive model for each sensor
+3. Investigation of visual methods for space-time counts
+4. Development of a Shiny app: interface design, computation and data backend
+5. Generation of a dynamic spatial model (if time permits)
 

### 2017-05-07 thesis/proposal_draft.Rmd
added more detailed caption to figure 2

@@
 \begin{center}
-\textbf{Figure 2}: Smoothed pedestrian counts for Melbourne Central by day of the week and month of the year
+\textbf{Figure 2}: Smoothed pedestrian counts for Melbourne Central by day of the week and month of the year.
+Month of the year is mapped to colour.
+We see that the daily patterns are quite consistent between days of the week, although we note that the amount of pedestrians at late hours of the day are higher during Friday and Saturday compared to other days.
 \end{center}

### 2017-05-07 presentation/proposal_pres.Rmd
last minute updates to slides and proposal

@@
 The package `shiny` allows us to build web apps using R's capabilities in the background
-We can add interactive inputs to visualisations.
 
-An example of what we can make:
+We can add interactive inputs to visualisations
+
+#### An example of what we can make:
 

### 2017-05-07 thesis/proposal_draft.Rmd
last minute updates to slides and proposal

@@
 subtitle: "Thesis Proposal"
-date: April 2017
+date: 7 May 2017
 author: 

### 2017-08-18 thesis/missing_impute.Rmd
working missing values imputation with walk_melb() data; need to get raw data working due to issues with making missings explicit

---
title: "Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset"
output: html_notebook
---

```{r, echo = FALSE}
library(plyr)
library(tidyverse)
library(lubridate)
library(magrittr)
## devtools::install_github("earowang/rwalkr")
library(rwalkr)
# 
# ped_2014 <- run_melb(year = 2014)
# ped_2015 <- run_melb(year = 2015)
# ped_2016 <- run_melb(year = 2016)
# ped_2017 <- run_melb(year = 2017)
# 
# ped_data <- bind_rows(ped_2014, ped_2015, ped_2016, ped_2017)

ped_data <- walk_melb(from = "2014-01-01")
# ped_data <- read_csv("../data/ped_df.csv")
# ped_data$X1 <- NULL
ped_data$Sensor_ID <- ped_data$Sensor
ped_data$Hourly_Counts <- ped_data$Count
ped_data$Date_Time <- ymd_h(paste(ped_data$Date, ped_data$Time, sep = " "))
ped_data$Day <- wday(ped_data$Date_Time, label = T, abbr = F)
ped_data$Month <- month(ped_data$Date_Time, label = T, abbr = F)
ped_data$Time <- as.factor(ped_data$Time)

# for offline use
# ped_dat <- read_csv(file = "../data/ped_df_2.csv")
# ped_data <- ped_dat %>%
#  mutate(
#    Date_Time = as.POSIXct(strptime(Date_Time, format = "%d-%B-%Y %H:%M"),
#      tz = ""),
#    ID = NULL,
#    Mdate = as.integer(Mdate),
#    Hourly_Counts = as.integer(Hourly_Counts),
#    Sensor_ID = as.integer(Sensor_ID),
#    Time = as.integer(Time),
#    Year = as.integer(Year)
#  ) %>% as.data.frame() %>%
#  filter(Date_Time > "2013-12-31 23:00:01") %>%
#  arrange(Date_Time)

# rm(ped_dat)

# ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
```

```{r}
## making missings explicit
# dates11 <- full_seq(ped_data$Date_Time, period = 3600)
# id_seq <- rep(1:43, no_T)
# 
# ped_full_seq <- data.frame(Date_Time = dates11, Sensor_ID = id_seq)
# 
# ped_full <- merge(ped_data, ped_full_seq, all.y = T)
# 
# unique(ped_data$Date_Time) %>% length()
# 
#   df_t <- list()
#   no_T <- nrow(as.data.frame(dates11))
#     for (i in 1:43) {
#     i_id <- rep(i, no_T)
#     df_t[[i]] <- data.frame(Date_Time = dates11, Sensor_ID = i_id)
#     }
#   
#   
#   
#   ped_full <- list()
#   ped_list_d <- ped_data %>% split(.$Sensor_ID)
#     for (i in 1:43) {
#     df <- ped_list_d[[i]]
#     ped_full[[i]] <- right_join(df, df_t[[i]], by = "Date_Time")
#     }
#   ped_data <- bind_rows(ped_full) %>% arrange(Date_Time) %>%
#     mutate(Mdate = mday(Date_Time),
#            Day = wday(Date_Time, label = T, abbr = F),
#            Time = hour(Date_Time),
#            Year = year(Date_Time),
#            Month = month(Date_Time, label = T, abbr = F),
#            ID = NULL,
#            Sensor_Name = NULL)
#   
#   ped_names <- ped_loc %>% select(`Sensor ID`, `Sensor Description`)
#   colnames(ped_names) <- c("Sensor_ID", "Sensor_Name")
#   
#   ped_data <- ped_data %>% left_join(ped_names)
#   
#   ped_a <- ped_data %>% split(.$Sensor_ID)
#   ped_n <- numeric()
#   for (i in 1:43) {
#     ped_n[i] <- nrow(df_t[[i]])
#   }
#   
#   sum(ped_n)
#   
#   ad <- setdiff(as.data.frame(ped_a[[1]]$Date_Time), as.data.frame(dates1))
```

## Public Holiday Code
```{r pubhdays}
pub_hday14 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
pub_hday15 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
pub_hday16 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")

pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)

pub_hdays$Date <- ymd(pub_hdays$Date)
pub_hdays$Month <- pub_hdays$Date %>% month(label = TRUE, abbr = FALSE)
pub_hdays$VIC <- 0
pub_hdays$VIC[grep(glob2rx("*VIC*"), pub_hdays$`Applicable To`)] <- 1
pub_hdays$VIC[grep("NAT", pub_hdays$`Applicable To`)] <- 1

dates1 <- full_seq(ped_data$Date_Time, period = 3600)
hdayvals <- rep(0, length(dates1))
hdayvals[as.Date(dates1) %in% pub_hdays$Date[pub_hdays$VIC == 1]] <- 1
dates2 <- as.Date(rep(dates1, each = 43))
hdayvals2 <- as.numeric(rep(hdayvals, each = 43))

ped_data$IsHDay <- hdayvals2
ped_data$HDay <- ped_data$Day
levels(ped_data$HDay) <- c(levels(ped_data$Day), "Holiday")
ped_data$HDay[ped_data$IsHDay == 1] <- "Holiday"

ped_data$DayType <- ifelse(ped_data$Day == "Sunday", "Sunday",
                  ifelse(ped_data$Day == "Saturday", "Saturday",
                  ifelse(ped_data$Day == "Monday", "Monday",
                  ifelse(ped_data$Day == "Friday", "Friday",
                                                 "Midweek"))))
ped_data$DayType[ped_data$IsHDay == 1] <- "Holiday"

```

## GLM model for imputing missings
```{r}
ped_list <- ped_data %>% filter(Date < "2017-01-01") %>%
            split(.$Sensor_ID)

ped_coeffs <- list()
ped_model_fitted <- list()
ped_model_summary <- list()
p <- progress_estimated(43)
options(na.action = 'na.omit')
for (i in 1:43) {
  p$tick()$print()
  train <- ped_list[[i]]
  ped_model_i <- glm(Hourly_Counts ~ Month + DayType*Time,
                     family = quasipoisson(), data = train)
  ped_coeffs[[i]] <- ped_model_i$coefficients
  ped_model_summary[[i]] <- summary(ped_model_i)
}
rm(ped_model_i)

```

```{r}
test_ped <- ped_data %>% filter(Date > "2016-12-31") %>%
            split(.$Sensor_ID)
## required for model.matrix to allow NAs
options(na.action = 'na.pass')

for (i in 1:43) {
  test_df <- model.matrix(Hourly_Counts ~ Month + DayType*Time,
                          data = test_ped[[i]])
  test_coeff <- ped_coeffs[[i]]
  if (length(test_coeff) < ncol(test_df)) {
    missing_cols <- setdiff(colnames(test_df), rownames(as.data.frame(test_coeff)))
    test_df <- test_df[, !(colnames(test_df) %in% missing_cols)]
  }

  test_preds <- exp(test_df %*% test_coeff)
  test_ped[[i]]$Fitted <- test_preds
}

test_fitted <- bind_rows(test_ped)

```

```{r}
ggplot(filter(test_fitted, Date > "2017-07-31")) + geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
                         geom_line(aes(x = Date_Time, y = Fitted), colour = "darkred", alpha = 0.75) +
                         facet_wrap(~ Sensor)
```

### 2017-08-19 thesis/missing_impute.Rmd
update

@@
 ped_data$HDay[ped_data$IsHDay == 1] <- "Holiday"
+```
+This code uses official public holiday data from `data.gov.au` to check if dates fall on public holidays. The code ensures only public holidays which apply to Victoria are used.
 
+## Code for Day Type
+```{r daytypecode}
 ped_data$DayType <- ifelse(ped_data$Day == "Sunday", "Sunday",
@@
 ped_data$DayType[ped_data$IsHDay == 1] <- "Holiday"
-
 ```
+Compacts the `HDay` factor into 6 levels, rather than 8 levels, by coding Tuesday, Wednesday and Thursday as factor `"Midweek"`. This is because at most locations, these days have similar patterns.
 
 ## GLM model for imputing missings
-```{r}
+```{r glm_models}
 ped_list <- ped_data %>% filter(Date < "2017-01-01") %>%
@@
 ```
+The code above fits a GLM model for each sensor location.
+This model uses a quasi-poisson distribution to allow of overdispersion.
+Using a standard poisson distribution would give the restriction of mean = variance.
+The model is specified as `Hourly_Counts ~ Month + DayType * Time`, where `Month` is given an additive effect, and `DayType * Time` gives an interaction term between `DayType` and `Time`, as well as any other indirect additive effects.
+To improve efficiency, only the coefficients of the model are retained, as well as model summarries for model evaluation.  
+We train the model only on the data from 2014 to 2016. The test data set is the data available from 2017.
 
+### Test data fit
 ```{r}
@@
 ```
+Using the coefficients saved for each location, we can easily fit the test data.  
+With the function `model.matrix()`, `R` will generate a matrix, $X$, with the factor levels converted into dummy variables for use with the GLM.
+Then, we can obtain our estimates of $\text{Hourly_Counts}_ij$, $\widehat{\text{Hourly_Counts}_i} = \exp \left( X \hat{\beta_j} \right)$ where $\hat{\beta_j}$ is a column vector of the GLM estimated coefficients at the $j^{th}$ sensor location.
+
+# Goodness of Fit
+### Visually
+```{r}
+ggplot(filter(test_fitted, Date > "2017-07-23")) + geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
+                         geom_line(aes(x = Date_Time, y = Fitted), colour = "violetred2", alpha = 0.75) +
+                         facet_wrap(~ Sensor, scales = "free_y", ncol = 5) + theme_minimal() + 
+                    theme(axis.text.x = element_text(size = 3),
+                          strip.text = element_text(size = 5),
+                          axis.text.y = element_text(size = 0)) +
+                    xlab("Date") + ylab("Hourly Counts")
+```
+Looking at plots of actuals (black) against fitted (red) of data from 24 July 2017 to 6 August 2017, we see that some sensors appear to be not operational.
+We see that there is a mix of very good fitting models and not so good ones.
+An example of of a reasonably good fitting model is the _Collins Place (North)_ sensor location:
+```{r}
+ggplot(test_fitted %>%
+         filter(Date > "2017-07-23",
+                Sensor == "Collins Place (North)")) +
+      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
+      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7)
+```
+```{r}
+col_pl_test_mse <- test_fitted %>% filter(Date > "2017-07-23", Sensor == "Collins Place (North)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean()
+col_pl_test_rmse <- sqrt(col_pl_test_mse)
+```
+
+Using RMSE (root mean square error) as a goodness of fit measure, we find that the RMSE of the prediction represented in the plot above is `r col_pl_test_rmse`.
 
 ```{r}
-ggplot(filter(test_fitted, Date > "2017-07-31")) + geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
-                         geom_line(aes(x = Date_Time, y = Fitted), colour = "darkred", alpha = 0.75) +
-                         facet_wrap(~ Sensor)
+ggplot(filter(test_fitted, Sensor_ID == "Melbourne Central")) +
+  geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.1) +
+  geom_path(aes(x = Time, y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
+  facet_wrap(~ DayType)
 ```
+
+```{r}
+ggplot(filter(ped_data, Sensor == "Collins Place (South)")) + geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day))
+```
+
+```{r}
+
+```

### 2017-08-20 thesis/missing_impute.Rmd
formatting fix

--- 2017-08-18 thesis/missing_impute.Rmd
@@
 title: "Imputation of Missing Values in the Melbourne CBD Pedestrian Dataset"
-output: html_notebook
+output:
+  html_notebook: default
+  html_document: default
+  pdf_document: default
+  word_document: default
 ---
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, message=F, warning=F, results='hide'}
 library(plyr)
@@
 library(magrittr)
+
 ## devtools::install_github("earowang/rwalkr")
-library(rwalkr)
+# library(rwalkr)
 # 
@@
 
-ped_data <- walk_melb(from = "2014-01-01")
-# ped_data <- read_csv("../data/ped_df.csv")
-# ped_data$X1 <- NULL
+# ped_data <- walk_melb(from = "2014-01-01")
+ped_data <- read_csv("../data/ped_df.csv")
+ped_data$X1 <- NULL
 ped_data$Sensor_ID <- ped_data$Sensor
@@
 
-```{r}
+```{r missing_explicit, echo = F}
 ## making missings explicit
@@
 ## Public Holiday Code
-```{r pubhdays}
-pub_hday14 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
-pub_hday15 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
-pub_hday16 <- read_csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
 
+```{r pubhdays, message=F, warning=F}
+pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv", )
+pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
+pub_hday16 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
 pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)
@@
 ped_data$HDay[ped_data$IsHDay == 1] <- "Holiday"
+```
+This code uses official public holiday data from `data.gov.au` to check if dates fall on public holidays. The code ensures only public holidays which apply to Victoria are used.
 
+## Code for Day Type
+```{r daytypecode, results='hide'}
 ped_data$DayType <- ifelse(ped_data$Day == "Sunday", "Sunday",
@@
 ped_data$DayType[ped_data$IsHDay == 1] <- "Holiday"
-
 ```
+Compacts the `HDay` factor into 6 levels, rather than 8 levels, by coding Tuesday, Wednesday and Thursday as factor `"Midweek"`. This is because at most locations, these days have similar patterns.
 
 ## GLM model for imputing missings
-```{r}
+```{r glm_models, results='hide'}
 ped_list <- ped_data %>% filter(Date < "2017-01-01") %>%
@@
 ```
+The code above fits a GLM model for each sensor location.
+This model uses a quasi-poisson distribution to allow of overdispersion.
+Using a standard poisson distribution would give the restriction of mean = variance.
+The model is specified as `Hourly_Counts ~ Month + DayType * Time`, where `Month` is given an additive effect, and `DayType * Time` gives an interaction term between `DayType` and `Time`, as well as any other indirect additive effects.
+To improve efficiency, only the coefficients of the model are retained, as well as model summarries for model evaluation.  
+We train the model only on the data from 2014 to 2016. The test data set is the data available from 2017.
 
-```{r}
+### Test data fit
+```{r, results='hide'}
 test_ped <- ped_data %>% filter(Date > "2016-12-31") %>%
@@
 ```
+Using the coefficients saved for each location, we can easily fit the test data.  
+With the function `model.matrix()`, `R` will generate a matrix, $X$, with the factor levels converted into dummy variables for use with the GLM.
+Then, we can obtain our estimates of $\text{Hourly_Counts}_{ij}$, $\widehat{\text{Hourly_Counts}_i} = \exp \left( X \hat{\beta_j} \right)$ where $\hat{\beta_j}$ is a column vector of the GLM estimated coefficients at the $j^{th}$ sensor location.
+
+# Goodness of Fit
+### Visually
+```{r, fig.align= 'center', fig.height=24, fig.retina=2, dev='png', dev.args=list(type="cairo"), dpi = 128}
+ggplot(filter(test_fitted, Date > "2017-07-23")) +
+  geom_line(aes(x = Date_Time, y = Hourly_Counts), size = 2) +
+  geom_line(aes(x = Date_Time, y = Fitted),
+            colour = "violetred2", alpha = 0.75, size = 2) +
+                         facet_wrap(~ Sensor, scales = "free_y", ncol = 4) + theme_minimal() + 
+                    theme(axis.text.x = element_text(size = 3),
+                          strip.text = element_text(size = 32),
+                          axis.text.y = element_text(size = 0)) +
+                    xlab("Date") + ylab("Hourly Counts")
+```
+Looking at plots of actuals (black) against fitted (red) of data from 24 July 2017 to 6 August 2017, we see that some sensors appear to be not operational.
+We see that there is a mix of very good fitting models and not so good ones.
+An example of of a reasonably good fitting model is the _Collins Place (North)_ sensor location:
+```{r, fig.align= 'center', fig.retina=4, dev='png', dev.args=list(type="cairo"), dpi = 128}
+ggplot(test_fitted %>%
+         filter(Date > "2017-07-23",
+                Sensor == "Collins Place (North)")) +
+      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
+      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7)
+```
 
 ```{r}
-ggplot(filter(test_fitted, Date > "2017-07-31")) + geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
-                         geom_line(aes(x = Date_Time, y = Fitted), colour = "darkred", alpha = 0.75) +
-                         facet_wrap(~ Sensor)
+col_pl_test_mse <- test_fitted %>% filter(Date > "2017-07-23", Sensor == "Collins Place (North)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean()
+col_pl_test_rmse <- sqrt(col_pl_test_mse)
 ```
+
+Using RMSE (root mean square error) as a goodness of fit measure, we find that the RMSE of the prediction represented in the plot above is `r round(col_pl_test_rmse, 4)`.
+
+```{r, fig.align= 'center'}
+ggplot(filter(test_fitted, Sensor_ID == "Melbourne Central")) +
+  geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.1) +
+  geom_path(aes(x = Time, y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
+  facet_wrap(~ DayType)
+```
+
+```{r, message = F, warning = F, fig.align= 'center', dev='png', dev.args=list(type="cairo"), dpi = 128}
+ggplot(filter(ped_data, Sensor == "Melbourne Central")) + 
+  geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), se = FALSE) +
+  xlab("Time of Day") + ylab("Hourly Count") + theme_minimal() +
+  ggtitle("Average pedestrian count at Melbourne Central by day of the week")
+```
+
+```{r}
+
+```

### 2017-08-20 thesis/missing_impute.Rmd
update with more evaluation

@@
 output:
+  html_document: default
   html_notebook: default
-  html_document: default
-  pdf_document: default
-  word_document: default
 ---
-
-```{r, echo = FALSE, message=F, warning=F, results='hide'}
+```{r, include=FALSE}
 library(plyr)
@@
 library(magrittr)
+library(viridis)
+library(gridExtra)
+set.seed(12345)
+```
 
+```{r data_read, message=F, warning=F, include=FALSE, results='hide', cache=TRUE}
 ## devtools::install_github("earowang/rwalkr")
@@
 
-```{r pubhdays, message=F, warning=F}
-pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv", )
+```{r pubhdays, message=F, warning=F, cache=TRUE}
+## read data from data.gov.au
+pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
 pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
@@
 ## Code for Day Type
-```{r daytypecode, results='hide'}
+```{r daytypecode, results='hide', cache=TRUE, echo=FALSE}
 ped_data$DayType <- ifelse(ped_data$Day == "Sunday", "Sunday",
@@
 ```
-Compacts the `HDay` factor into 6 levels, rather than 8 levels, by coding Tuesday, Wednesday and Thursday as factor `"Midweek"`. This is because at most locations, these days have similar patterns.
+Compacts the `HDay` factor into 6 levels, rather than 8 levels, by coding Tuesday, Wednesday and Thursday as factor `"Midweek"`.
+This is because at most locations, these days have similar patterns.
+For instance, we see this at Melbourne Central:
+
+```{r, message = F, warning = F, fig.align= 'center', cache=TRUE, echo=FALSE}
+ggplot(filter(ped_data, Sensor == "Melbourne Central")) + 
+  geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), se = FALSE) +
+  xlab("Time of Day") + ylab("Hourly Count") + theme_minimal() +
+  ggtitle("Average pedestrian count at Melbourne Central by day of the week")
+```
 
 ## GLM model for imputing missings
-```{r glm_models, results='hide'}
+```{r glm_models, results='hide', cache=TRUE}
 ped_list <- ped_data %>% filter(Date < "2017-01-01") %>%
@@
 ### Test data fit
-```{r, results='hide'}
+```{r, results='hide', cache=TRUE}
 test_ped <- ped_data %>% filter(Date > "2016-12-31") %>%
@@
 ### Visually
-```{r, fig.align= 'center', fig.height=24, fig.retina=2, dev='png', dev.args=list(type="cairo"), dpi = 128}
+```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=4, cache=TRUE, echo=FALSE}
 ggplot(filter(test_fitted, Date > "2017-07-23")) +
-  geom_line(aes(x = Date_Time, y = Hourly_Counts), size = 2) +
+  geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
   geom_line(aes(x = Date_Time, y = Fitted),
-            colour = "violetred2", alpha = 0.75, size = 2) +
+            colour = "violetred2", alpha = 0.75) +
                          facet_wrap(~ Sensor, scales = "free_y", ncol = 4) + theme_minimal() + 
                     theme(axis.text.x = element_text(size = 3),
-                          strip.text = element_text(size = 32),
+                          strip.text = element_text(size = 10),
                           axis.text.y = element_text(size = 0)) +
@@
 An example of of a reasonably good fitting model is the _Collins Place (North)_ sensor location:
-```{r, fig.align= 'center', fig.retina=4, dev='png', dev.args=list(type="cairo"), dpi = 128}
+```{r, fig.align= 'center', fig.retina = 4, cache=TRUE}
 ggplot(test_fitted %>%
@@
       geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
-      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7)
+      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7) +
+      theme_minimal()
 ```
 
-```{r}
-col_pl_test_mse <- test_fitted %>% filter(Date > "2017-07-23", Sensor == "Collins Place (North)") %>%
-               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean()
-col_pl_test_rmse <- sqrt(col_pl_test_mse)
+```{r, cache=TRUE}
+col_pl_test_rmse <- test_fitted %>% filter(Date > "2017-07-23", Sensor == "Collins Place (North)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
 ```
@@
 Using RMSE (root mean square error) as a goodness of fit measure, we find that the RMSE of the prediction represented in the plot above is `r round(col_pl_test_rmse, 4)`.
+In comparison, the model predicted for the sensor located on _Grattan St-Swanston St (West)_ fits very poorly:
 
-```{r, fig.align= 'center'}
-ggplot(filter(test_fitted, Sensor_ID == "Melbourne Central")) +
+```{r, echo=FALSE, fig.align='center', fig.retina=4, cache=TRUE}
+ggplot(test_fitted %>%
+         filter(Date > "2017-07-23",
+                Sensor == "Grattan St-Swanston St (West)")) +
+      geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
+      geom_line(aes(x = Date_Time, y = Fitted), colour = 'violetred2', alpha = 0.7) +
+      theme_minimal()
+```
+
+```{r, echo=FALSE, cache=TRUE}
+gratt_test_rmse <- test_fitted %>% filter(Date > "2017-07-23", Sensor == "Grattan St-Swanston St (West)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
+
+gratt_test_rmse_july <- test_fitted %>% filter(Date > "2017-07-23", Date < "2017-08-01", Sensor == "Grattan St-Swanston St (West)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
+
+gratt_test_rmse_aug <- test_fitted %>% filter(Date > "2017-07-31", Sensor == "Grattan St-Swanston St (West)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
+```
+
+Here, the poor fit results in a RMSE of `r round(gratt_test_rmse, 4)` for the period between Jul 24 to Aug 6.
+It appears as though the additive effect of `Month` is exaggerated, where we see that predictions in July are significantly underestimated, compared to August predictions which have adequate fit.
+If we compare the RMSE between July predictions and August predictions, we find that the RMSE of the predictions of Aug 1 to Aug 6 is much lower than that of the predictions for Jul 24 to Jul 31 (`r round(gratt_test_rmse_aug, 4)` compared to `r round(gratt_test_rmse_july, 4)`)
+
+```{r, fig.align= 'center', fig.retina=3, echo=FALSE, cache=TRUE}
+ggplot(filter(test_fitted, Sensor_ID == "Grattan St-Swanston St (West)")) +
   geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.1) +
   geom_path(aes(x = Time, y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
-  facet_wrap(~ DayType)
+  facet_wrap(~ DayType) +
+      theme_minimal()
 ```
 
-```{r, message = F, warning = F, fig.align= 'center', dev='png', dev.args=list(type="cairo"), dpi = 128}
-ggplot(filter(ped_data, Sensor == "Melbourne Central")) + 
-  geom_smooth(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), se = FALSE) +
-  xlab("Time of Day") + ylab("Hourly Count") + theme_minimal() +
-  ggtitle("Average pedestrian count at Melbourne Central by day of the week")
+Looking at the counts for each day, we see that the variance during weekdays is quite large.
+This would suggest that the day of the week and time of the day do not explain a lot of the variance in the data.
+
+```{r, fig.align= 'center', fig.retina=3, echo=FALSE, cache=TRUE}
+ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) +
+  geom_path(aes(x = Time, y = Hourly_Counts, group = Date, colour = Month), alpha = 0.1) +
+  facet_wrap(~ DayType) +
+      theme_minimal()
+
+ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) + geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
+      theme_minimal()
 ```
+
+However, we see that the variance within a month is quite low comparatively. It is worth noting that the data available at this location is very poor.
 

### 2017-08-20 thesis/missing_impute.Rmd
updated with knitted html

@@
 output:
-  html_document: default
+  html_document:
+    theme: united
   html_notebook: default
+  pdf_document: default
 ---
@@
 Using the coefficients saved for each location, we can easily fit the test data.  
-With the function `model.matrix()`, `R` will generate a matrix, $X$, with the factor levels converted into dummy variables for use with the GLM.
-Then, we can obtain our estimates of $\text{Hourly_Counts}_{ij}$, $\widehat{\text{Hourly_Counts}_i} = \exp \left( X \hat{\beta_j} \right)$ where $\hat{\beta_j}$ is a column vector of the GLM estimated coefficients at the $j^{th}$ sensor location.
+With the function `model.matrix()`, `R` will generate a matrix, $X_j = \{ X_{1j}, X_{2j}, \dots , X_{Tj} \}$, with the factor levels converted into dummy variables for use with the GLM.
+Then, we can obtain our estimates of $\text{Hourly_Counts}_{j}$, $\widehat{\text{Hourly} \_ \text{Counts}}_{j} = \exp \left( X_j\hat{\beta_j} \right)$ where $\hat{\beta_j}$ is a column vector of the GLM estimated coefficients at the $j^{th}$ sensor location.
 
@@
 ### Visually
-```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=4, cache=TRUE, echo=FALSE}
+```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=1, cache=TRUE, echo=FALSE}
 ggplot(filter(test_fitted, Date > "2017-07-23")) +
@@
 An example of of a reasonably good fitting model is the _Collins Place (North)_ sensor location:
-```{r, fig.align= 'center', fig.retina = 4, cache=TRUE}
+```{r, fig.align= 'center', fig.retina = 1, cache=TRUE}
 ggplot(test_fitted %>%
@@
 
-```{r, echo=FALSE, fig.align='center', fig.retina=4, cache=TRUE}
+```{r, echo=FALSE, fig.align='center', fig.retina=1, cache=TRUE}
 ggplot(test_fitted %>%
@@
 
-```{r, fig.align= 'center', fig.retina=3, echo=FALSE, cache=TRUE}
+```{r, fig.align= 'center', fig.retina=1, echo=FALSE, cache=TRUE, warning=F}
 ggplot(filter(test_fitted, Sensor_ID == "Grattan St-Swanston St (West)")) +
@@
 
-```{r, fig.align= 'center', fig.retina=3, echo=FALSE, cache=TRUE}
+```{r, fig.align= 'center', fig.retina=1, echo=FALSE, cache=TRUE, warning = F}
 ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) +
@@
 However, we see that the variance within a month is quite low comparatively. It is worth noting that the data available at this location is very poor.  
-
-```{r}
-
-```
+Plotting all available observations, we see we only have data starting from `r head(na.omit(ped_data[ped_data$Sensor_ID == "Grattan St-Swanston St (West)", ])$Date, 1)`.

### 2017-08-20 thesis/missing_impute.Rmd
added rmse of test fit for all models

@@
 library(gridExtra)
+library(pander)
 set.seed(12345)
@@
           axis.text.y = element_text(size = 0)) +
-                    xlab("Date") + ylab("Hourly Counts")
+    xlab("Date") + ylab("Hourly Counts") +
+    ggtitle("Plot of Actual (black) against Fitted for period 24 July 2017 to 6 August 2017")        
 ```
-Looking at plots of actuals (black) against fitted (red) of data from 24 July 2017 to 6 August 2017, we see that some sensors appear to be not operational.
+Looking at plots of actuals (black) against fitted of data from 24 July 2017 to 6 August 2017, we see that some sensors appear to be not operational.
 We see that there is a mix of very good fitting models and not so good ones.
@@
 gratt_test_rmse <- test_fitted %>% filter(Date > "2017-07-23", Sensor == "Grattan St-Swanston St (West)") %>%
-               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+              .$residsq %>% mean() %>% sqrt()
 
-gratt_test_rmse_july <- test_fitted %>% filter(Date > "2017-07-23", Date < "2017-08-01", Sensor == "Grattan St-Swanston St (West)") %>%
-               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
+gratt_test_rmse_july <- test_fitted %>% filter(Date > "2017-07-23", Date < "2017-08-01",
+                                               Sensor == "Grattan St-Swanston St (West)") %>%
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+              .$residsq %>% mean() %>% sqrt()
 
 gratt_test_rmse_aug <- test_fitted %>% filter(Date > "2017-07-31", Sensor == "Grattan St-Swanston St (West)") %>%
-               mutate(residsq = (Fitted - Hourly_Counts)^2) %>% .$residsq %>% mean() %>% sqrt()
+               mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+              .$residsq %>% mean() %>% sqrt()
 ```
@@
 
-ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) + geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
+ggplot(filter(ped_data, Sensor_ID == "Grattan St-Swanston St (West)")) +
+  geom_line(aes(x = Date_Time, y = Hourly_Counts)) +
       theme_minimal()
@@
 Plotting all available observations, we see we only have data starting from `r head(na.omit(ped_data[ped_data$Sensor_ID == "Grattan St-Swanston St (West)", ])$Date, 1)`.
+
+Computing the RMSE for all of the test set (2017 data) for each sensor:
+
+```{r}
+test_df <- test_fitted %>% mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+            split(.$Sensor_ID)
+test_df_rmse <- data.frame(Sensor_ID = rep(0,43), RMSE = rep(0, 43))
+  for (i in 1:43) {
+      test_df_rmse$Sensor_ID[i] <- names(test_df)[i]
+      test_df_rmse$RMSE[i] <- mean(test_df[[i]]$residsq) %>% sqrt()
+  }
+
+test_df_plot <- test_fitted %>% filter(Date > "2017-07-23") %>%
+            mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+            split(.$Sensor_ID)
+test_df_rmse_plot <- data.frame(Sensor_ID = rep(0,43), RMSE = rep(0, 43))
+  for (i in 1:43) {
+      test_df_rmse_plot$Sensor_ID[i] <- names(test_df_plot)[i]
+      test_df_rmse_plot$RMSE[i] <- mean(test_df_plot[[i]]$residsq) %>% sqrt()
+  }
+pander(arrange(test_df_rmse, by= RMSE))
+```

### 2017-09-03 thesis/missing_impute.Rmd
model for flinders st - swanston st (west) using neighbouring sensors

--- 2017-08-20 thesis/missing_impute.Rmd
@@
 library(lubridate)
-library(magrittr)
-library(viridis)
+library(pander)
 library(gridExtra)
+library(fortify)
 set.seed(12345)
@@
 
-ped_coeffs <- list()
-ped_model_fitted <- list()
-ped_model_summary <- list()
-p <- progress_estimated(43)
-options(na.action = 'na.omit')
-for (i in 1:43) {
-  p$tick()$print()
-  train <- ped_list[[i]]
-  ped_model_i <- glm(Hourly_Counts ~ Month + DayType*Time,
-                     family = quasipoisson(), data = train)
-  ped_coeffs[[i]] <- ped_model_i$coefficients
-  ped_model_summary[[i]] <- summary(ped_model_i)
+# ped_coeffs <- list()
+# ped_model_fitted <- list()
+# ped_model_summary <- list()
+# p <- progress_estimated(43)
+# options(na.action = 'na.omit')
+# for (i in 1:43) {
+#   p$tick()$print()
+#   train <- ped_list[[i]]
+#   ped_model_i <- glm(Hourly_Counts ~ Month + DayType*Time,
+#                      family = quasipoisson(), data = train)
+#   ped_coeffs[[i]] <- ped_model_i$coefficients
+#   ped_model_summary[[i]] <- summary(ped_model_i)
+# }
+# rm(ped_model_i)
+```
+
+```{r, cache = T}
+## Alternative code for GLM
+library(foreach)
+library(doSNOW)
+### code from:
+## http://www.win-vector.com/blog/2014/05/trimming-the-fat-from-glm-models-in-r/
+stripGlmLR = function(cm) {
+  cm$y = c()
+  cm$model = c()
+
+  cm$residuals = c()
+  cm$fitted.values = c()
+  cm$effects = c()
+  cm$qr$qr = c()  
+  cm$linear.predictors = c()
+  cm$weights = c()
+  cm$prior.weights = c()
+  cm$data = c()
+
+  cm$family$variance = c()
+  cm$family$dev.resids = c()
+  cm$family$aic = c()
+  cm$family$validmu = c()
+  cm$family$simulate = c()
+  attr(cm$terms,".Environment") = c()
+  attr(cm$formula,".Environment") = c()
+
+  cm
 }
-rm(ped_model_i)
+#######################################
+
+cl <- makeCluster(2)
+registerDoSNOW(cl)
+ped_models <- foreach(i = 1:43) %dopar% {
+  model <- glm(Hourly_Counts ~ Month + DayType*Time, data = ped_list[[i]], family = quasipoisson())
+  model <- stripGlmLR(model)
+  return(model)
+}
+
+ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels
 
 ```
+
 The code above fits a GLM model for each sensor location.
@@
 for (i in 1:43) {
-  test_df <- model.matrix(Hourly_Counts ~ Month + DayType*Time,
-                          data = test_ped[[i]])
-  test_coeff <- ped_coeffs[[i]]
-  if (length(test_coeff) < ncol(test_df)) {
-    missing_cols <- setdiff(colnames(test_df), rownames(as.data.frame(test_coeff)))
-    test_df <- test_df[, !(colnames(test_df) %in% missing_cols)]
+  test_ped[[i]]$Fitted <- predict.glm(ped_models[[i]], newdata = test_ped[[i]], type = "response")
 }
 
-  test_preds <- exp(test_df %*% test_coeff)
-  test_ped[[i]]$Fitted <- test_preds
-}
+# 
+# for (i in 1:43) {
+#   test_df <- model.matrix(Hourly_Counts ~ Month + DayType*Time,
+#                           data = test_ped[[i]])
+#   test_coeff <- ped_coeffs[[i]]
+#   if (length(test_coeff) < ncol(test_df)) {
+#     missing_cols <- setdiff(colnames(test_df), rownames(as.data.frame(test_coeff)))
+#     test_df <- test_df[, !(colnames(test_df) %in% missing_cols)]
+#   }
+#   
+#   test_preds <- exp(test_df %*% test_coeff)
+#   test_ped[[i]]$Fitted <- test_preds
+# }
 
@@
 ### Visually
-```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=1, cache=TRUE, echo=FALSE}
+```{r, fig.align= 'center', fig.height=24, fig.width=10, fig.retina=2, cache=TRUE, echo=FALSE}
 ggplot(filter(test_fitted, Date > "2017-07-23")) +
@@
 ggplot(filter(test_fitted, Sensor_ID == "Grattan St-Swanston St (West)")) +
-  geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.1) +
-  geom_path(aes(x = Time, y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
+  geom_path(aes(x = as.integer(Time), y = Hourly_Counts, group = Date), alpha = 0.1) +
+  geom_path(aes(x = as.integer(Time), y = Fitted, group = Date), colour = "violetred2", alpha = 0.1) +
   facet_wrap(~ DayType) +
@@
 Plotting all available observations, we see we only have data starting from `r head(na.omit(ped_data[ped_data$Sensor_ID == "Grattan St-Swanston St (West)", ])$Date, 1)`.
+
+Computing the RMSE for all of the test set (2017 data) for each sensor:
+
+```{r}
+test_df <- test_fitted %>% mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+            split(.$Sensor_ID)
+test_df_rmse <- data.frame(Sensor_ID = rep(0,43), RMSE = rep(0, 43), StdRMSE = rep(0, 43))
+  for (i in 1:43) {
+      test_df_rmse$Sensor_ID[i] <- names(test_df)[i]
+      test_df_rmse$RMSE[i] <- mean(test_df[[i]]$residsq) %>% sqrt()
+      test_df_rmse$StdRMSE[i] <- test_df_rmse$RMSE[i] / mean(test_df[[i]]$Hourly_Counts)
+
+  }
+
+test_df_plot <- test_fitted %>% filter(Date > "2017-07-23") %>%
+            mutate(residsq = (Fitted - Hourly_Counts)^2) %>%
+            split(.$Sensor_ID)
+test_df_rmse_plot <- data.frame(Sensor_ID = rep(0,43), RMSE = rep(0, 43), StdRMSE = rep(0, 43))
+  for (i in 1:43) {
+      test_df_rmse_plot$Sensor_ID[i] <- names(test_df_plot)[i]
+      test_df_rmse_plot$RMSE[i] <- mean(test_df_plot[[i]]$residsq) %>% sqrt()
+      test_df_rmse_plot$StdRMSE[i] <- test_df_rmse_plot$RMSE[i] / mean(test_df_plot[[i]]$Hourly_Counts)
+  }
+pander(arrange(test_df_rmse, by= StdRMSE))
+```
+
+Here, the `StdRMSE` is the RMSE standardised by the mean.
+
+```{r}
+library(ggmap)
+  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
+```
+
+Looking at some problematic sensor locations:
+```{r}
+pander(test_df_rmse[test_df_rmse$StdRMSE > 0.65, ] %>% arrange(by = StdRMSE))
+test_df_rmse[test_df_rmse$StdRMSE > 0.65, ] %>% .$Sensor_ID -> bad_sensors
+```
+
+```{r, warning=F}
+p1 <- ggplot(test_ped$`Birrarung Marr`) + geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.05) +
+                                  geom_path(aes(x = Time, y = Fitted, group = Date), alpha = 0.05,
+                                            colour = "violetred2") + ggtitle("Birrarung Marr")
+p2 <- ggplot(test_ped$`Birrarung Marr`) + geom_line(aes(x = Date_Time, y = Hourly_Counts), alpha = 0.5) +
+                                  geom_line(aes(x = Date_Time, y = Fitted), alpha = 0.5, colour = "violetred2")
+
+grid.arrange(p1, p2, nrow = 1)
+```
+
+```{r}
+p1 <- ggplot(test_ped$`Flagstaff Station`) + geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.05) +
+                                  geom_path(aes(x = Time, y = Fitted, group = Date), alpha = 0.05,
+                                            colour = "violetred2") + ggtitle("Flagstaff Station")
+p2 <- ggplot(test_ped$`Flagstaff Station`) + geom_line(aes(x = Date_Time, y = Hourly_Counts), alpha = 0.5) +
+                                  geom_line(aes(x = Date_Time, y = Fitted), alpha = 0.5, colour = "violetred2")
+grid.arrange(p1, p2, nrow = 1)
+
+```
+
+```{r}
+p1 <- ggplot(test_ped$`Flinders St-Swanston St (West)`) + geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.05) +
+                                  geom_path(aes(x = Time, y = Fitted, group = Date), alpha = 0.05,
+                                            colour = "violetred2") + ggtitle("Flinders St-Swanston St (West)")
+p2 <- ggplot(test_ped$`Flinders St-Swanston St (West)`) + geom_line(aes(x = Date_Time, y = Hourly_Counts), alpha = 0.5) +
+                                  geom_line(aes(x = Date_Time, y = Fitted), alpha = 0.5, colour = "violetred2")
+grid.arrange(p1, p2, nrow = 1)
+
+```
+
+```{r}
+p1 <- ggplot(test_ped$`New Quay`) + geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.05) +
+                                  geom_path(aes(x = Time, y = Fitted, group = Date), alpha = 0.05,
+                                  colour = "violetred2") + ggtitle("New Quay")
+p2 <- ggplot(test_ped$`New Quay`) + geom_line(aes(x = Date_Time, y = Hourly_Counts), alpha = 0.5) +
+                                  geom_line(aes(x = Date_Time, y = Fitted), alpha = 0.5, colour = "violetred2")
+grid.arrange(p1, p2, nrow = 1)
+
+```
+
+```{r}
+p1 <- ggplot(test_ped$`Princes Bridge`) + geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.05) +
+                                  geom_path(aes(x = Time, y = Fitted, group = Date), alpha = 0.05,
+                                  colour = "violetred2") + ggtitle("Princes Bridge")
+p2 <- ggplot(test_ped$`Princes Bridge`) + geom_line(aes(x = Date_Time, y = Hourly_Counts), alpha = 0.5) +
+                                  geom_line(aes(x = Date_Time, y = Fitted), alpha = 0.5, colour = "violetred2")
+grid.arrange(p1, p2, nrow = 1)
+```
+
+```{r}
+p1 <- ggplot(test_ped$`Southbank`) + geom_path(aes(x = Time, y = Hourly_Counts, group = Date), alpha = 0.05) +
+                                  geom_path(aes(x = Time, y = Fitted, group = Date), alpha = 0.05,
+                                  colour = "violetred2") + ggtitle("Southbank")
+p2 <- ggplot(test_ped$`Southbank`) + geom_line(aes(x = Date_Time, y = Hourly_Counts), alpha = 0.5) +
+                                  geom_line(aes(x = Date_Time, y = Fitted), alpha = 0.5, colour = "violetred2")
+grid.arrange(p1, p2, nrow = 1)
+```
+
+### Standardise Counts
+
+$$HourlyCounts_{i,j} \sim N(0,1)$$
+
+```{r}
+for (i in 1:43) {
+  ped_list[[i]]$scaled_count <- scale(ped_list[[i]]$Hourly_Counts)
+}
+ped_df <- bind_rows(ped_list)
+ped_df %>% mutate(Timep = as.integer(Time)) -> ped_df
+```
+
+```{r}
+ggplot(ped_df[ped_df$Date > "2015-01-01", ]) + 
+  geom_path(aes(x = Timep, y = scaled_count, colour = Sensor_ID, group = Date), alpha = 0.01) +
+      theme(legend.position = 'none') + xlab("Time") + ylab("Standardised Counts")
+```
+
+```{r}
+ggplot(ped_df[ped_df$Sensor_ID == "Flinders St-Swanston St (West)" | 
+              ped_df$Sensor_ID == "Flinders St-Elizabeth St (East)" |
+              ped_df$Sensor_ID == "Flinders St Station Underpass" , ]) +
+      geom_smooth(aes(x = Timep, y = scaled_count, colour = Sensor_ID))
+
+ggplot(ped_df[ped_df$Sensor_ID == "Flinders St-Swanston St (West)" | 
+              ped_df$Sensor_ID == "Flinders St-Elizabeth St (East)" |
+              ped_df$Sensor_ID == "Flinders St Station Underpass" , ]) +
+      geom_path(aes(x = Timep, y = scaled_count, group = Date, colour = Sensor_ID),
+                alpha = 0.05) + facet_wrap(~ DayType) + theme(legend.position = 'none')
+```
+
+```{r}
+ped_df[ped_df$Sensor_ID == "Flinders St-Swanston St (West)", ]$scaled_count -> swan
+ped_df[ped_df$Sensor_ID == "Flinders St-Elizabeth St (East)", ]$scaled_count -> liz
+ped_df[ped_df$Sensor_ID == "Flinders St Station Underpass", ]$scaled_count -> underpass
+
+flin_swan_df <- data.frame(swan = swan, liz = liz, underpass = underpass,
+                           date = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Date,
+                           time = as.integer(ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Time),
+                           timefact = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Time,
+                           month = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Month,
+                           daytype = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$DayType,
+                           dtm = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Date_Time,
+                           hday = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$IsHDay)
+
+flin_swan <- lm(swan ~ liz + underpass, data = flin_swan_df) 
+summary(flin_swan)
+flin_swan_df$swan_fit <- predict.lm(flin_swan, newdata = flin_swan_df)
+flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
+plot(flin_swan)
+ggplot(flin_swan_df) + 
+  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
+  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
+  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')
+
+```
+
+```{r}
+flin_swan_df %>% mutate(midday = ifelse(time > 10 & time < 17, 1, 0)) -> flin_swan_df
+
+flin_swan <- lm(swan ~ daytype*liz*timefact, data = flin_swan_df) 
+summary(flin_swan)
+flin_swan_df$swan_fit <- predict.lm(flin_swan, newdata = flin_swan_df)
+flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
+plot(flin_swan)
+ggplot(flin_swan_df) + 
+  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
+  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
+  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')
+
+```
+
+```{r}
+flin_swan_2 <- lm(swan ~ daytype*timefact*liz + daytype*timefact*underpass, data = flin_swan_df) 
+summary(flin_swan_2)
+flin_swan_df$swan_fit <- predict.lm(flin_swan_2, newdata = flin_swan_df)
+flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
+plot(flin_swan_2)
+
+ggplot(flin_swan_df) + 
+  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
+  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
+  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')
+
+```
+
+Issues with potentially `NA` having `Hourly_Counts = 0`:
+
+```{r}
+ggplot(flin_swan_df) + geom_line(aes(x = dtm, y = swan))
+ggplot(ped_data %>% filter(Sensor_ID == "Flinders St-Swanston St (West)")) + geom_line(aes(x = Date_Time, y = Hourly_Counts))
+```
+
+Need to filter out pre-2015 data from training set:
+
+```{r}
+flin_swan_df_train <- flin_swan_df %>% filter(date > "2014-12-31")
+
+flin_swan_2 <- lm(swan ~ timefact*liz + timefact*underpass, data = flin_swan_df_train) 
+summary(flin_swan_2)
+flin_swan_df$swan_fit <- predict.lm(flin_swan_2, newdata = flin_swan_df)
+flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
+plot(flin_swan_2)
+
+ggplot(flin_swan_df) + 
+  geom_path(aes(x = time, y = swan, group = date), alpha = 0.05) + 
+  geom_path(aes(x = time, y = liz, group = date), alpha = 0.02, colour = 'limegreen') + 
+  geom_path(aes(x = time, y = swan_fit, group = date), alpha = 0.02, colour = 'red')
+
+```
+Also removed `DayType` due to high leverage observations.
+
+```{r}
+flin_swan_df[8763:8764, ]
+```
+
+High leverage points here are the New Year's Day celebrations.

### 2017-09-03 code/trimmedGLM.R

@@ -34,15 +34,9 @@ stripGlmLR = function(cm) {
 
 cl <- makeCluster(2)
 registerDoSNOW(cl)
-pb <- txtProgressBar(max=43, style=3)
-progress <- function(n) setTxtProgressBar(pb, n)
-opts <- list(progress=progress)
 
-ped_models <- foreach(i = 1:43,  .options.snow=opts) %dopar% {
+ped_models <- foreach(i = 1:43) %dopar% {
   model <- glm(Hourly_Counts ~ Month + DayType*Time, data = ped_list[[i]], family = quasipoisson())
   model <- stripGlmLR(model)
   return(model)
 }
-close(pb)
-
-predictions <- predict.glm(ped_models[[1]], newdata = test_ped[[1]], type = "response")      

### 2017-09-03 thesis/missing_impute.Rmd
southbank model

@@
 library(gridExtra)
-library(fortify)
+library(ggfortify)
 set.seed(12345)
@@
 ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels
+
+for (i in 1:43) {
+  ped_list[[i]]$glm_fitted <- predict.glm(ped_models[[i]], ped_list[[i]], type = "response")
+}
 
@@
 
-### Standardise Counts
+### Standardising Counts
 
@@
 ```
+
+#### Flinders St. - Swanston St. (West)
 
@@
 flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
-plot(flin_swan)
+autoplot(flin_swan)
 ggplot(flin_swan_df) + 
@@
 flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
-plot(flin_swan_2)
-
+autoplot(flin_swan_2)
 ggplot(flin_swan_df) + 
@@
 
-flin_swan_2 <- lm(swan ~ timefact*liz + timefact*underpass, data = flin_swan_df_train) 
+flin_swan_2 <- lm(swan ~ timefact + timefact:liz + timefact:underpass, data = flin_swan_df_train) 
 summary(flin_swan_2)
@@
 flin_swan_df$resid <- flin_swan_df$swan_fit - flin_swan_df$swan
-plot(flin_swan_2)
+autoplot(flin_swan_2)
 
@@
 ```{r}
-flin_swan_df[8763:8764, ]
+flin_swan_df[c(8763:8764, 1248), ]
 ```
@@
 High leverage points here are the New Year's Day celebrations.  
+Comparing this fit to that Quasipoisson GLM:
+```{r}
+flin_swan_var <- ped_list$`Flinders St-Swanston St (West)` %>%
+  .$Hourly_Counts %>% var(na.rm = TRUE)
+flin_swan_mu <- ped_list$`Flinders St-Swanston St (West)` %>%
+  .$Hourly_Counts %>% mean(na.rm = TRUE)
+flin_swan_df$glmfits <- ped_list$`Flinders St-Swanston St (West)` %>% .$glm_fitted
+flin_swan_df$fits <- flin_swan_df$swan_fit*sqrt(flin_swan_var) + flin_swan_mu
+flin_swan_df$swan_unscaled <- flin_swan_df$swan*sqrt(flin_swan_var) + flin_swan_mu
+
+ggplot(flin_swan_df) + 
+  geom_path(aes(x = time, y = swan_unscaled, group = date), alpha = 0.05) + 
+  geom_path(aes(x = time, y = glmfits, group = date), alpha = 0.02, colour = 'blue') + 
+  geom_path(aes(x = time, y = fits, group = date), alpha = 0.02, colour = 'red')
+
+```
+
+#### Southbank
+```{r}
+ped_df[ped_df$Sensor_ID == "Southbank", ]$scaled_count -> sbank
+ped_df[ped_df$Sensor_ID == "Waterfront City", ]$scaled_count -> wfront
+ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$scaled_count -> wbridge
+ped_df[ped_df$Sensor_ID == "Vic Arts Centre", ]$scaled_count -> vicarts
+
+sbank_df <- data.frame(sbank_y = sbank, wfront, wbridge, vicarts,
+                       date = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Date,
+                       time = as.integer(ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Time),
+                       timefact = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Time,
+                       month = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Month,
+                       daytype = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$DayType,
+                       dtm = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$Date_Time,
+                       hday = ped_df[ped_df$Sensor_ID == "Webb Bridge", ]$IsHDay)
+
+qplot(x = dtm, y = sbank, data = sbank_df)
+```
+
+```{r}
+ggplot(sbank_df) + geom_point(aes(x = sbank, y = wbridge))
+
+sbank_fit <- lm(sbank_y ~ wfront + wbridge + vicarts, data = sbank_df)
+summary(sbank_fit)
+autoplot(sbank_fit)
+sbank_df$sbank_fit <- predict(sbank_fit, sbank_df)
+ggplot(sbank_df) + geom_path(aes(x = time, y = sbank_y, group = date), alpha = 0.06) + 
+                   geom_path(aes(x = time, y = sbank_fit, group = date), alpha = 0.02, colour = 'red')
+```
+
+```{r}
+sbank_fit <- lm(sbank ~ timefact:wbridge + timefact:vicarts + timefact:wfront, data = sbank_df)
+summary(sbank_fit)
+autoplot(sbank_fit)
+sbank_df$sbank_fit <- predict(sbank_fit, sbank_df)
+ggplot(sbank_df) + geom_path(aes(x = time, y = sbank_y, group = date), alpha = 0.06) + 
+                   geom_path(aes(x = time, y = sbank_fit, group = date), alpha = 0.02, colour = 'red')
+```
+
+```{r}
+sbank_df[26302:26304, ]
+```
+
+We also have a huge potential outlier:
+```{r}
+sbank_df %>% arrange(desc(fits)) %>% head(5) %>%
+    select(dtm, daytype, sbank_y, wfront, vicarts,
+            wbridge, sbank_fit, fits, glmfits) %>%
+    pander()
+```
+
+Comparing to the Quasipoisson GLM:
+```{r}
+sbank_var <- ped_list$`Southbank` %>%
+  .$Hourly_Counts %>% var(na.rm = TRUE)
+sbank_mu <- ped_list$`Southbank` %>%
+  .$Hourly_Counts %>% mean(na.rm = TRUE)
+sbank_df$glmfits <- ped_list$`Southbank` %>% .$glm_fitted
+sbank_df$fits <- sbank_df$sbank_fit*sqrt(sbank_var) + sbank_mu
+sbank_df$sbank_unscaled <- sbank_df$sbank_y * sqrt(sbank_var) + sbank_mu
+
+ggplot(sbank_df[sbank_df$date != "2014-04-13", ]) + 
+  geom_path(aes(x = time, y = sbank_unscaled, group = date), alpha = 0.15) + 
+  geom_path(aes(x = time, y = glmfits, group = date), alpha = 0.007, colour = 'blue') +
+  geom_path(aes(x = time, y = fits, group = date), alpha = 0.0088, colour = 'red')
+```

### 2017-09-26 thesis/draft.Rmd
update with short thesis draft. much to add

---
title: "Thesis Draft"
author: "Gavin Chin"
date: "21 September 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
library(tidyverse)
library(lubridate)
library(foreach)
library(doSNOW)
library(rwalkr)
library(pander)
```

```{r data_loading, warning=F, message=F}
ped_data <- NULL
if(file.exists("../data/ped_df.csv")){
  ped_data <- read_csv("../data/ped_df.csv")
  ped_data$X1 <- NULL
  }

## update data
last_df_date <- ifelse(file.exists("../data/ped_df.csv"),
                        ped_data$Date %>% sort() %>%
                        tail(1),
                        as.Date("2013-12-31")) + 1 %>% as.Date(origin = "1970-01-01")

if(last_df_date != today(tzone = "Australia/Melbourne")){
  new_data <- rwalkr::walk_melb(from = last_df_date, to = today()-1)
  ped_data <- rbind(ped_data, new_data)
  write_csv(ped_data, "../data/ped_df.csv")
  }

ped_data$Sensor_ID <- ped_data$Sensor
ped_data$Hourly_Counts <- ped_data$Count
ped_data$Date_Time <- ymd_h(paste(ped_data$Date, ped_data$Time, sep = " "))
ped_data$Day <- lubridate::wday(ped_data$Date_Time, label = TRUE, abbr = F)
ped_data$Month <- lubridate::month(ped_data$Date_Time, label = TRUE, abbr = F)
ped_data$Time <- as.factor(ped_data$Time)

## long to wide
ped_data %>% select(Sensor_ID, Count, Date_Time, Time, Month, Day) %>% 
  spread(Sensor_ID, Count) -> dfa

pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
pub_hday16 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)

pub_hdays$Date <- ymd(pub_hdays$Date)
pub_hdays$Month <- lubridate::month(pub_hdays$Date, label = TRUE, abbr = F)
pub_hdays$VIC <- 0
pub_hdays$VIC[grep(glob2rx("*VIC*"), pub_hdays$`Applicable.To`)] <- 1
pub_hdays$VIC[grep("NAT", pub_hdays$`Applicable.To`)] <- 1

dfa <- dfa %>% mutate(IsHDay = (date(Date_Time) %in% pub_hdays$Date[pub_hdays$VIC == 1]))

dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
                     ifelse(dfa$Day == "Saturday", "Saturday",
                    ifelse(dfa$Day == "Monday", "Monday",
                     ifelse(dfa$Day == "Friday", "Friday",
                                  "Midweek"))))
dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
```
## Background

#### The Data: City of Melbourne Pedestrian Data

The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives"[^1]. This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD. In particular, we want to be able to model the pedestrian traffic at different locations in the CBD, and forecast the traffic. In addition, we aim to have an interactive, dynamic data visualisation of the model to help us to understand how Melbourne operates.

The dataset being investigated can be obtained from the official City of Melbourne's open data[^2]. The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations. In total, there is `r nrow(ped_data)` observations in the dataset, representing `r nrow(dfa)` hours of data. It is available in `.csv` format, allowing for easy data import into R. We can also access the pedestrian data using the the `rwalkr` package by Earo Wang[^3]. This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.

```{r}
ped_data %>% select(Sensor, Hourly_Counts) %>% split(.$Sensor)  %>%
            map(is.na) %>% map(sum) %>% unlist -> sensor_missvals
```

Not even an single sensor location has complete data for the period between between 1/1/2014 and `r paste(day(today()), month(today()), year(today()), sep = "/")`.

[^1]: from _'About Melbourne Data'_ page at https://data.melbourne.vic.gov.au/about

[^2]: the specific pedestrian data is available at   https://data.melbourne.vic.gov.au/Transport-Movement/Pedestrian-volume-updated-monthly-/b2ak-trbp

[^3]: https://github.com/earowang/rwalkr/

### Imputation Literature Review Notes
<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->

<!-- http://ir.lib.uwo.ca/cgi/viewcontent.cgi?article=1250&context=csce2016   -->
<!-- - traffic counters with 40% to 60% missing values -->
<!-- - heuristic methods -->
<!--   - using "good historical values" or historical average values resulted -->
<!--     in MARE (mean absolute relative error) reaching up to 80% -->
<!--   - using moving average values worked better than other heuristic methods -->
<!--   - all heuristic methods inhenrently cannot reflect sudden fluctuations during abnormal periods -->
<!-- - pattern matching methods -->
<!--   - find "candidate" volume pattern to compare to "study curve" and use the candidate which best matches the study curve -->
<!-- - ARIMA -->
<!--   - did not work well with multiple seasonalities -->
<!-- - AI methods -->
<!--   - genetic algorithms (GAs) and artificial neural networks (ANNs) -->

<!-- Method used in the study was kNN with an interaction using a weather model, which they found to work best with large variations in traffic during winter -->
<!-- (severe weather conditions affecting traffic volume)   -->

Reviewing current research on imputation of missing values in traffic data, a paper on imputation of missing classified traffic data during winter season (Imputation of Missing Classified Traffic Data During Winter Season - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_) stated some methods which were found to be poor for imputation. The data they were attempting imputation for were traffic counters located on the highway network in Alberta, Canada with between 40% and 60% missing data.

Replacement with "good" historical values, or using historical average values was found to have resulted in very poor fit with high MARE (mean absolute relative error). Thet did find that using a moving average worked the best out of all the heuristic methods used. The largest problem with heuristic methods, however, is the inherent inability to reflect sudden fluctuations/shocks during abnormal periods. With such a large proportion of missing values, pattern matching methods were investigated. This involves comparing the _study curve_, the pattern at the counter which is to be imputed, to candidate patterns (patterns at other locations).

For the purposes of that paper, it was found to be inappropriate with the data being analysed.  However, unlike _Roh et al._ where patterns were compared to traffic volumes in different juristictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller. Another difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data. They proposed using non-parametric estimation methods with k-Nearest Neighbours for their imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.

<!-- #### Flexible and Robust Method for Missing Loop Detector Data Imputation - _K. C. Henrickson, Y. Zou, and Y. Wang_   -->

<!-- http://docs.trb.org/prp/15-5804.pdf   -->
<!-- - also use kNN -->

<!-- #### Key Factors of K-nearest Neighbors Nonparametric Regression in Short-time Traffic Flow Forecasting -->
<!-- http://www.springer.com/cda/content/document/cda_downloaddocument/9789462391017-c2.pdf?SGWID=0-0-45-1492476-p177147492 -->

<!-- #### K-NEAREST NEIGHBOR ALGORITHM FOR UNIVARIATE TIME SERIES PREDICTION -->

<!-- http://webbut.unitbv.ro/BU2012/Series%20III/BULETIN%20III/Sasu%20A.pdf -->

## Methodology: Imputation of Missing Values

### Basic GLM Approach

The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
$$E \left[\text{Hourly_Counts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
Both these variables are categorical/factors. We cannot treat time of the day as an integer or numeric because it does not have a linear relationship with counts.s

We use the Quasi-Poisson error distribution as opposed to normal/Gaussian errors due to the response variables being count data. The Poisson distribution allows for only non-negative integer values, while estimating a Quasi-Poisson model estimates an additional parameter for overdispersion. This allows more flexibility in the model by allowing the assumption that $E[Y]=Var[Y]=\mu$ to be relaxed. Instead, the Quasi-Poisson distribution allows for $Var[Y] = \theta \mu$.

```{r}
## Base GLM models

### code from:
## http://www.win-vector.com/blog/2014/05/trimming-the-fat-from-glm-models-in-r/
stripGlmLR = function(cm) {
  cm$y = c()
  cm$model = c()

  cm$residuals = c()
  cm$fitted.values = c()
  cm$effects = c()
  cm$qr$qr = c()  
  cm$linear.predictors = c()
  cm$weights = c()
  cm$prior.weights = c()
  cm$data = c()

  cm$family$variance = c()
  cm$family$dev.resids = c()
  cm$family$aic = c()
  cm$family$validmu = c()
  cm$family$simulate = c()
  attr(cm$terms,".Environment") = c()
  attr(cm$formula,".Environment") = c()

  cm
}
#######################################
```

```{r}
cl <- makeCluster(2)
registerDoSNOW(cl)
models <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
  colnames(tr_dat)[1] <- "Hourly_Counts"
  model <- glm(Hourly_Counts ~ HDay*Time, data = tr_dat, family = quasipoisson())
  model <- stripGlmLR(model)
  return(model)
}

# ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels

```

While this works well with a small proportion of missing values, it also requires a large number of observations. Specifically, with this paramaterisation, we have $23 + 7 + (23 \times 7) + 1=192$ coefficients to estimate. Another disadvantage of using this model is the lack of robustness to outliers due to the sensitivity of some parameters. This becomes particularly problematic at sensor locations which have been recently installed, where there is a lack of historical data.

### Small-Large Split Approach

The proposed alternative approach to imputing the missing values in the data focuses on improving the models used at locations with a large proportion of missing values. A potential threshold value which could be used to class a sensor as having a "large" proportion of missing values vs a "small" proportion of missing values is 10%. At locations with a "small" proportion of missings, we can continue to use the GLM quasi-possion regression specified in $(1)$. At the locations with a large proportion of missing values, we need to use information from neighbouring sensors.

The first issue which needs to be addressed is the potential of values which are actually missing due to sensor failure or malfunction having a zero count. This issue will cause bias in the estimates, as well as affect the classification of a sensor. Of course, we cannot simply classify all zero counts as `NA`, as true zero values are possible.

```{r}
n_obs <- nrow(dfa)
threshold <- 0.1
threshold_miss_hrs <- 50

run_lengths <- foreach(i = unique(ped_data$Sensor)) %do%
  {
    dfa[is.na(dfa[, i]), i] <- 0
    na_length_check <- rle(as.numeric(unlist(dfa[,i])))
    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
    rl <- dfa[, i] %>% unlist %>% rle()
    dfa[maybeNA, i] <- NA
    return(rl)

    }
run_lengths %>% lapply(., `[[`, 1) %>%
    unlist -> sequences

pander(summary(sequences[sequences > 1]))

# ggplot() + geom_boxplot(aes(y = sequences, x = "A"), alpha = 0.2) +
#     xlab(NULL) + ylab("Length of same counts value") +
#     theme(axis.text.x = element_blank(),
#           axis.ticks.x = element_blank(),
#           axis.title.x = element_blank()) +
#     theme_minimal()
```

Instead, a simple check we can implement is to look for long sequences of zero counts. For example, we may want to classify any sequence of `Hourly_Counts = 0` running for longer than 24 hours as being considered `NA`. From the summary of lengths of repeated values (which are repeated at least once), we see that the distribution of the lengths are very positively skewed. With a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data. Classifying any sequence of repeated values, typically 0, which has a length greater than 24 as `NA` should allow true zero counts to avoid being misclassified as `NA`.

After we have replaced questionable zero values with `NA`, we can calculate the proportion of missing values for each sensor location. From these calculated proportion, we can classify sensors as either having a "large" or "small" proportion of missing values, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.05`.

```{r}
na_prop <- numeric()

for (i in unique(ped_data$Sensor)){
  na_prop[i] <- sum(is.na(dfa[, i])) / n_obs
}

large_miss_sensors <- names(na_prop[na_prop > threshold])
small_miss_sensors <- setdiff(names(na_prop), large_miss_sensors)
```

Because the models for the sensors with a large proportion of missing values rely on having neighbouring sensors having complete cases (no missing values), we want to impute these values first. This will then allow us to have as much information to be available for the models which use neighbouring sensors to to train on.

For sensors with a small proportion of missing values, we use a GLM quasi-poisson model again. To improve on the previous model used, we use the model:
$$E \left[\text{Hourly_Counts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$

$$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$

```{r alg_glm}
models <- list()
fitted_data <- list()
used_sensors <- list()
p <- progress_estimated(length(unique(ped_data$Sensor)))
cl <- makeCluster(2)
registerDoSNOW(cl)

models <- foreach(i = small_miss_sensors) %dopar%
{
  p$tick()$print()
  options(na.action = na.omit)
  # if(max(na_length_check$lengths) > max_miss_hrs) 
  # {
  #   model <- NULL
  # }

  sensor_df <- unlist(dfa[, i])

  model <- glm(sensor_df ~  Month + DayType*Time, family = 'quasipoisson', data = dfa)

  options(na.action = na.pass)
  fitted_data[[i]] <- predict(model, newdata = dfa, type = "response")
  options(na.action = na.omit)
  dfa[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
  return(stripGlmLR(model))
}
```

Instead of `HDay`, we use `DayType` which classifies Tuesday, Wednesday and Thursday as a single factor level, Midweek. We do this as the daily patterns on these weekdays are similar. This reduces the number of parameters to estimate by 46. A result of this is we also have more robust estimates for the midweek days, being less sensitive to outliers. Because this model will be estimated for sensor locations with few missing values,  we have enough data to add in month as an additive effect. This will help capture the annual seasonality.

Using these estimated models, we impute the missing values to produce complete data at all these sensor locations. For the locations with high proportions of missing values can be predicted using the now complete data from all the locations which had a small proportion of missing values. Using a threshold of 10%, this gives us `r length(small_miss_sensors)` sensor locations to use.

For each of the sensor locations with high proportions of missings, we need to decide which sensor to use for prediction. A simple method of finding a geographical closest neighbour is to take the haversine/great-circle distance between the sensor to be imputed and all the possible candidates.
```{r, message=F}
# Calculates the geodesic distance between two points specified by radian latitude/longitude using the
# Spherical Law of Cosines (slc)
# source: https://www.r-bloggers.com/great-circle-distance-calculations-in-r/
gcd.slc <- function(long1, lat1, long2, lat2) {
  R <- 6371 # Earth mean radius [km]
  d <- acos(sin(lat1)*sin(lat2) + cos(lat1)*cos(lat2) * cos(long2-long1)) * R
  return(d) # Distance in km
}

ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
lats <- ped_loc$Latitude
lons <- ped_loc$Longitude
ids <- ped_loc$`Sensor Description`

## name fix dictionary
{
ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
}
loc <- data.frame(lons, lats, ids)
rownames(loc) <- ids
```

```{r alg_lm}
for (i in large_miss_sensors){
  sensor_dists <- numeric()
  for (j in small_miss_sensors){
      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
      sensor_dists[j] <- dists
  }
  names(sensor_dists) <- small_miss_sensors
  closest_sensor <- sensor_dists %>% sort %>%
                    head(2) %>% names
  close_df_scale <- dfa[, closest_sensor[1]] %>% unlist %>% scale
  close_df_scale2 <- dfa[, closest_sensor[2]] %>% unlist %>% scale

  sensor_df <- dfa[, i] %>% unlist
  model <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2, data = dfa, family = quasipoisson())
  models[[i]] <- stripGlmLR(model)
}

```

Using the two geographically closest sensors with complete cases, we use another GLM specified as:
$$E \left[ \text{Hourly_Counts}_{\text{sensor}} \vert \text{Time}, \text{Hourly_Counts}_{\text{closest sensors}} \right] = \mu_{\text{Time}, \text{Hourly_Counts}_{\text{closest sensors}}}$$
$$\text{where } \quad  \mu_{\text{Time}, \text{Hourly_Counts}_{\text{closest sensors}}} \sim \text{Time} \times \text{Hourly_Counts}_{\text{closest sensor 1}} + \text{Time} \times \text{Hourly_Counts}_{\text{closest sensor 2}}$$

We would expect that the amount of people passing through the neighbouring sensors will be a strong predictor of the pedestrian counts as they are likely to also pass through. Using geographical neighbours, we can capture rare events effectively. Any large shocks to the counts at the neighbouring sensors will have an effect on the counts to be imputed.

In order to use a mixture of the patterns from two different sensors, we need to use the scaled counts at the neighbouring sensors to avoid the magnitude of the counts having an effect on the predicted counts. We scale the counts by standardising to $\mu = 0, \sigma^2 = 1$. Because the covariance with neighbouring sensors may differ over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.

### 2017-09-26 code/longtowide.R

@@ -5,15 +5,14 @@ library(foreach)
 library(doSNOW)
 library(rwalkr)
 
-##############################################
+
 ped_data <- NULL
 if(file.exists("data/ped_df.csv")){
   ped_data <- read_csv("data/ped_df.csv")
   ped_data$X1 <- NULL
   }
 
-###############################################
-## update data from latest to 
+## update data
 last_df_date <- ifelse(file.exists("data/ped_df.csv"),
                         ped_data$Date %>% sort() %>%
                         tail(1),
@@ -22,10 +21,9 @@ last_df_date <- ifelse(file.exists("data/ped_df.csv"),
 if(last_df_date != today(tzone = "Australia/Melbourne")){
   new_data <- rwalkr::walk_melb(from = last_df_date, to = today()-1)
   ped_data <- rbind(ped_data, new_data)
-  write_csv(new_data, path = "data/ped_df.csv", append = TRUE)
+  write_csv(ped_data, "data/ped_df.csv")
   }
 
-###############################################
 
 ped_data$Sensor_ID <- ped_data$Sensor
 ped_data$Hourly_Counts <- ped_data$Count
@@ -34,12 +32,9 @@ ped_data$Day <- lubridate::wday(ped_data$Date_Time, label = TRUE, abbr = F)
 ped_data$Month <- lubridate::month(ped_data$Date_Time, label = TRUE, abbr = F)
 ped_data$Time <- as.factor(ped_data$Time)
 
-###############################################
 ## long to wide
 ped_data %>% select(Sensor_ID, Count, Date_Time, Time, Month, Day) %>% 
   spread(Sensor_ID, Count) -> dfa
-
-###############################################
 
 pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
 pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
[…]

### 2017-09-27 thesis/draft.Rmd
fixed replacement of NA to imputed values

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
+                      message = FALSE, warning = FALSE)
 library(tidyverse)
@@
 library(pander)
+library(RColorBrewer)
 ```
@@
 dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
+original_df <- dfa
+dfa <- filter(dfa, Date_Time < "2017-01-01")
 ```
@@
 
-Replacement with "good" historical values, or using historical average values was found to have resulted in very poor fit with high MARE (mean absolute relative error).
-Thet did find that using a moving average worked the best out of all the heuristic methods used.
+Replacement with "good" historical values, or using historical average values was found to have resulted in very poor fit with high MARE/MARD (mean absolute relative error/difference).
+They did find that using a moving average worked the best out of all the heuristic methods used.
 The largest problem with heuristic methods, however, is the inherent inability to reflect sudden fluctuations/shocks during abnormal periods.
@@
 For the purposes of that paper, it was found to be inappropriate with the data being analysed.
-However, unlike _Roh et al._ where patterns were compared to traffic volumes in different juristictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller.
+However, unlike _Roh et al._ where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller.
 Another difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data.
@@
 used_sensors <- list()
+dfa2 <- dfa
 p <- progress_estimated(length(unique(ped_data$Sensor)))
-cl <- makeCluster(2)
-registerDoSNOW(cl)
-
-models <- foreach(i = small_miss_sensors) %dopar%
-{
+for(i in small_miss_sensors){
   p$tick()$print()
@@
   model <- glm(sensor_df ~  Month + DayType*Time, family = 'quasipoisson', data = dfa)
+  models[[i]] <- (stripGlmLR(model))
+}
+```
 
+```{r alg_glm_impute}
   options(na.action = na.pass)
-  fitted_data[[i]] <- predict(model, newdata = dfa, type = "response")
+for(i in small_miss_sensors){
+  ts_dat <- data.frame(Month = dfa$Month, DayType = dfa$DayType, Time = dfa$Time)
+  fitted_data[[i]] <- predict.glm(models[[i]],newdata = ts_dat, type = "response")
+}
   options(na.action = na.omit)
-  dfa[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
-  return(stripGlmLR(model))
+  for(i in small_miss_sensors){
+    dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
     }
+
 ```
@@
 for (i in large_miss_sensors){
+  # p$tick()$print()
   sensor_dists <- numeric()
@@
                     head(2) %>% names
-  close_df_scale <- dfa[, closest_sensor[1]] %>% unlist %>% scale
-  close_df_scale2 <- dfa[, closest_sensor[2]] %>% unlist %>% scale
+  close_df_scale <- dfa2[, closest_sensor[1]] %>% unlist %>% scale
+  close_df_scale2 <- dfa2[, closest_sensor[2]] %>% unlist %>% scale
 
-  sensor_df <- dfa[, i] %>% unlist
-  model <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2, data = dfa, family = quasipoisson())
+  sensor_df <- dfa2[, i] %>% unlist
+
+  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa$Time)
+  options(na.action = na.omit)
+  model <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
+               data = tr_dat, family = quasipoisson())
   models[[i]] <- stripGlmLR(model)
+
+  options(na.action = na.pass)
+  fitted_data[[i]] <- predict(model, newdata = tr_dat, type = "response")
+  options(na.action = na.omit)
+
+  dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
+
 }
@@
 Because the covariance with neighbouring sensors may differ over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
+
+## Imputation Results
+
+For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station.
+Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).
+
+```{r}
+ggplot(data = NULL, aes(x = original_df %>% 
+             filter(Date_Time < "2017-01-01") %>%
+             .[, "Southern Cross Station"],
+           y = fitted_data$`Southern Cross Station`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = brewer.pal(3, 'Dark2')[1],
+                       alpha = 0.7, size = 1.5,) +
+           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
+                       colour = brewer.pal(3, 'Dark2')[2]) +
+           theme_minimal() + theme(aspect.ratio = 1) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Actual against Predicted Counts at Southern Cross Station")
+
+```
+
+\begin{center}
+\textbf{Figure }:  A plot of the Actual vs Predicted pedestrian counts at Southern Cross Station.
+The green line is $x=y$, representing a reference for perfect fit, while orange line is the linear fit.
+\end{center}
+
+Visually, we can see that the fitted values are good estimates as the relationship between the actuals vs fitted is very close to 1:1.
+This is seen when we perform a least squares estimate of $\widehat{Fitted}_t = \hat{\beta}_0 + \hat{\beta}_1 Actual_t$, and the estimated $\hat{\beta}_0$ is close to 0 and $\hat{\beta}_1$ is close to 1.
+
+Focusing on the date 04/07/2014, which has 3 hours of missing data:
+```{r}
+fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
+                           fit_ = dfa2$`Southern Cross Station`) %>%
+                  filter(Date_Time < "2014-07-06 00:00:00",
+                         Date_Time > "2014-07-04 12:00:00") %>%
+                        .$fit_
+
+ggplot(original_df %>%  filter(Date_Time < "2014-07-06 00:00:00",
+                               Date_Time > "2014-07-04 12:00:00")) +
+  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = 'red',
+            size = 2) +
+  geom_line(aes(x = Date_Time, y = `Southern Cross Station`),
+            size  = 2) +
+  theme_minimal() + xlab("Time") +
+  ylab("Pedestrian Counts at Southern Cross Station") +
+  ggtitle("Pedestrian Counts at Southern Cross Station on 04/07/2014 - 05/07/2014")
+```
+\begin{center}
+\textbf{Figure }:  A plot of the pedestrian counts at Southern Cross Station. The black line is the observed counts, where the red line is the imputed counts.  
+\end{center}
+
+Here we see that the GLM model with only time and date based variables work well for fitting small periods of missing data.

### 2017-09-28 thesis/draft.Rmd
corrected NA fix code, where true NA was incorrectly made 0 values

@@
 dfa <- filter(dfa, Date_Time < "2017-01-01")
+ori_2 <- filter(original_df, Date_Time < "2017-01-01")
 ```
@@
 registerDoSNOW(cl)
-models <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
-  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
+fitted_base <- list()
+models_base <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
+    tr_dat <- data.frame(dfa[, i],  HDay = dfa$HDay, Time = dfa$Time)
     colnames(tr_dat)[1] <- "Hourly_Counts"
     model <- glm(Hourly_Counts ~ HDay*Time, data = tr_dat, family = quasipoisson())
-  model <- stripGlmLR(model)
-  return(model)
+    return(stripGlmLR(model))
   }
-
+names(models_base) <- unique(ped_data$Sensor_ID)
 # ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels
+for (i in unique(ped_data$Sensor)){  
+  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
+  options(na.action = na.pass)
+  fitted_base[[i]] <- predict.glm(models_base[[i]], newdata = tr_dat, type = "response")
+  options(na.action = na.omit)
+  }
 
@@
 
-run_lengths <- foreach(i = unique(ped_data$Sensor)) %do%
-  {
+run_lengths <- list()
+
+for(i in unique(ped_data$Sensor)){
     dfa[is.na(dfa[, i]), i] <- 0
@@
     dfa[maybeNA, i] <- NA
-    return(rl)
+    defsNA <- is.na(ori_2[, i])
+    dfa[defsNA, i] <- NA
+    run_lengths[[i]] <- rl
+  }
 
-    }
 run_lengths %>% lapply(., `[[`, 1) %>%
@@
 fitted_data <- list()
-used_sensors <- list()
 dfa2 <- dfa
@@
 
+Firstly, we evaluate the initial model used to impute:
+$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
+
+Note, this is also before the adjustments made to treat suspiciously long sequences of 0 values as `NA`.
+
 For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station.
-Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).
+At this sensor, the proportion of missing values (adjusted) is `r round(na_prop["Southern Cross Station"]*100,2)`Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).
 
@@
              .[, "Southern Cross Station"],
-           y = fitted_data$`Southern Cross Station`)) +
+           y = fitted_base$`Southern Cross Station`)) +
            geom_point(alpha = 0.05) +
@@
 This is seen when we perform a least squares estimate of $\widehat{Fitted}_t = \hat{\beta}_0 + \hat{\beta}_1 Actual_t$, and the estimated $\hat{\beta}_0$ is close to 0 and $\hat{\beta}_1$ is close to 1.
+
+Here we see that the GLM model with only time and date based variables work well for fitting small periods of missing data.
+
+However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor. We use the sensor at Australia on Collins to 
+```{r}
+options(na.action = na.omit)
+ggplot(data = NULL, aes(x = original_df[-(is.na(original_df$`Australia on Collins`)), ] %>% 
+             filter(Date_Time < "2017-01-01") %>%
+             .[, "Australia on Collins"],
+           y = fitted_base$`Australia on Collins`[-(is.na(original_df$`Australia on Collins`))])) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = brewer.pal(3, 'Dark2')[1],
+                       alpha = 0.7, size = 1.5,) +
+           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
+                       colour = brewer.pal(3, 'Dark2')[2]) +
+           theme_minimal() + theme(aspect.ratio = 1) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Actual against Predicted Counts at Australia on Collins")
+
+```
+We can see that the model is heavily biased by the false 0 values, causing the predicted values to be underestimate (slope of the fitted line is $<1$).
 
@@
 \end{center}
-
-Here we see that the GLM model with only time and date based variables work well for fitting small periods of missing data.

### 2017-09-28 thesis/draft.Rmd
fixed for LaTeX

@@
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
   word_document: default
@@
 The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
-$$E \left[\text{Hourly_Counts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
+$$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
 $$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
@@
 For sensors with a small proportion of missing values, we use a GLM quasi-poisson model again. To improve on the previous model used, we use the model:
-$$E \left[\text{Hourly_Counts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
+$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
 
@@
 Using the two geographically closest sensors with complete cases, we use another GLM specified as:
-$$E \left[ \text{Hourly_Counts}_{\text{sensor}} \vert \text{Time}, \text{Hourly_Counts}_{\text{closest sensors}} \right] = \mu_{\text{Time}, \text{Hourly_Counts}_{\text{closest sensors}}}$$
-$$\text{where } \quad  \mu_{\text{Time}, \text{Hourly_Counts}_{\text{closest sensors}}} \sim \text{Time} \times \text{Hourly_Counts}_{\text{closest sensor 1}} + \text{Time} \times \text{Hourly_Counts}_{\text{closest sensor 2}}$$
+$$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{closest sensors}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{closest sensors}}}$$
+$$\text{where } \quad  \mu_{\text{Time}, \text{HourlyCounts}_{\text{closest sensors}}} \sim \text{Time} \times \text{HourlyCounts}_{\text{closest sensor 1}} + \text{Time} \times \text{HourlyCounts}_{\text{closest sensor 2}}$$
 
@@
            theme_minimal() + theme(aspect.ratio = 1) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
            xlab("Actual Pedestrian Count") +
@@
 
-However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor. We use the sensor at Australia on Collins to 
+However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor.
+We use the sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%.
 ```{r}
@@
 We can see that the model is heavily biased by the false 0 values, causing the predicted values to be underestimate (slope of the fitted line is $<1$).
+This demonstrates the importance of ensuring outliers are removed from the data.
 

### 2017-09-28 thesis/draft.Rmd
fixed for LaTeX

@@
 Focusing on the date 04/07/2014, which has 3 hours of missing data:
+
 ```{r}

### 2017-09-28 thesis/draft.Rmd

@@
                        colour = brewer.pal(3, 'Dark2')[1],
-                       alpha = 0.7, size = 1.5,) +
+                       alpha = 0.7, size = 1.5) +
            geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
@@
 This demonstrates the importance of ensuring outliers are removed from the data.
+
+The criterion we will use to measure goodness of fit is MARE. It is calculated by:
+
+$$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \vert \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \vert}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$
+
+We calculate a seperate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data).
+This will help to identify overfitting of the data.
+We use MARE as it will be a comparable measure between sensor locations as well as comparable between models.
+
+```{r}
+MARE_base_glm_tr <- numeric()
+for (i in unique(ped_data$Sensor)){
+  MARE_base_glm_tr[i] <- mean(abs(fitted_base[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
+}
+pander(data.frame(Tr_MARE = MARE_base_glm_tr))
+```
 

### 2017-09-29 presentation/final.Rmd
presentation - still WIP

---
title: "Modelling the City of Melbourne Pedestrian Data"
subtitle: "Thesis Presentation"
author: "Gavin Chin, supervised by Di Cook"
date: "2 October 2017"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "myremark.css"]
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE,
        scipen = 100, cache = TRUE)
library(tidyverse)
library(ggmap)
library(knitr)
library(lubridate)
library(gridExtra)
library(viridis)
library(pander)
set.seed(13131313)
## quicker data source compared to rwalkr::walk_melb() 
## ped_data <- read_csv("http://209.148.91.227/files/ped_df_28sep17.csv")

  ped_data <- read_csv("../data/ped_df.csv")

  ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")

  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
                               mean(range(ped_loc$Latitude))),
                                zoom=14)
```

```{r, include = FALSE, echo = FALSE, warning=FALSE, message=FALSE}
ped_data$Sensor_ID <- ped_data$Sensor
ped_data$Hourly_Counts <- ped_data$Count
ped_data$Date_Time <- ymd_h(paste(ped_data$Date, ped_data$Time, sep = " "))
ped_data$Day <- lubridate::wday(ped_data$Date_Time, label = TRUE, abbr = F)
ped_data$Month <- lubridate::month(ped_data$Date_Time, label = TRUE, abbr = F)
ped_data$Time <- as.factor(ped_data$Time)
ped_data$WkYr <- paste(week(ped_data$Date_Time), year(ped_data$Date_Time), sep = "-")

## long to wide
ped_data %>% select(Sensor_ID, Count, Date_Time, Time, Month, Day) %>% 
  spread(Sensor_ID, Count) -> dfa

pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
pub_hday16 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)

pub_hdays$Date <- ymd(pub_hdays$Date)
pub_hdays$Month <- lubridate::month(pub_hdays$Date, label = TRUE, abbr = F)
pub_hdays$VIC <- 0
pub_hdays$VIC[grep(glob2rx("*VIC*"), pub_hdays$`Applicable.To`)] <- 1
pub_hdays$VIC[grep("NAT", pub_hdays$`Applicable.To`)] <- 1

dfa <- dfa %>% mutate(IsHDay = (date(Date_Time) %in% pub_hdays$Date[pub_hdays$VIC == 1]))

dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
                     ifelse(dfa$Day == "Saturday", "Saturday",
                    ifelse(dfa$Day == "Monday", "Monday",
                     ifelse(dfa$Day == "Friday", "Friday",
                                  "Midweek"))))
dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
dfa <- dfa %>% mutate(WkYr = paste(week(Date_Time), year(Date_Time)), sep = "-")
original_df <- dfa
dfa <- filter(dfa, Date_Time < "2017-01-01")
ori_2 <- filter(original_df, Date_Time < "2017-01-01")
```

class: center

# Introduction

The City of Melbourne has an Open Data Platform which makes council data available to the public

This project will focus on the pedestrian sensor data

```{r, echo = FALSE, fig.height=5, fig.retina = 3}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
                         data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

---
class: center

## How is the data collected?

The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
- These sensors are positioned under an awning or street pole, and count the number of people passing each hour

![figure of pedestrian sensors](../img/ped_sensors.png)

_source: http://www.pedestrian.melbourne.vic.gov.au/ _

---
class: middle
## Accessing the data

- The data is available as a `.csv` at https://data.melbourne.vic.gov.au/, and is updated monthly
- Alternatively, the data is also available as daily data from http://www.pedestrian.melbourne.vic.gov.au/
  * The `rwalkr` R package by Earo Wang provides an API to access the data from both sources easily in R in a tidy format

- Both sources of data provide the same counts data, but have different forms

We use the daily data sourced from `pedestrian.melbourne.vic.gov.au` rather than the data available at the Open Data Platform as missing values are explicit

---

## Format of the data
#### Variables available:
- Time and Date
- Sensor ID/Location
- Hourly Pedestrian Count

```{r, echo=FALSE}
ped_data %>% as.tibble
```

---

# Why do we want to model pedestrian traffic?

### Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector

#### Examples ways it can be used:

- Government example uses:  
  * Infrastructure planning
  * Security planning
- Private example uses:
  * Marketing campaign planning
  * Resource management
  * Investment planning
---
class: center
## The Problem: Prediction of pedestrian traffic
```{r, echo=FALSE, fig.retina = 3, fig.height=6, fig.width= 10}
p1 <- ggplot(dfa %>% filter(Date_Time > "2016-07-12 00:00:00",
                            Date_Time < "2016-07-25 00:00:00")) +
        geom_line(aes(x = Date_Time, y = `Flagstaff Station`)) +
        scale_y_continuous(limits = c(0, 7000))
p2 <- ggplot(dfa %>% filter(Date_Time > "2016-07-12 00:00:00",
                            Date_Time < "2016-07-25 00:00:00")) +
        geom_line(aes(x = Date_Time, y = `Southbank`)) +
        scale_y_continuous(limits = c(0, 7000))
p3 <- ggplot(dfa %>% filter(Date_Time > "2016-07-12 00:00:00",
                            Date_Time < "2016-07-25 00:00:00")) +
        geom_line(aes(x = Date_Time, y = `State Library`)) +
        scale_y_continuous(limits = c(0, 7000))
grid.arrange(p2, p1, p3, nrow = 3, ncol = 1)
```

--

#### Different locations have different counts and patterns

---
class: middle, center
## Missing Data
```{r, echo = FALSE, fig.retina=2, fig.width= 12, message=FALSE, warning=FALSE}
miss_plot_dat <- ped_data %>% filter(Date_Time > "2014-06-01",
                                     Date_Time < "2015-06-01") %>%
  group_by(WkYr, Sensor_ID) %>%
  summarise(Weekly_Count = sum(Hourly_Counts))

miss_plot_dat2 <- ped_data %>% group_by(Date) %>%
  filter(Date > "2014-06-01", Date < "2015-06-01") %>% left_join(miss_plot_dat)

random_sensors <- sample(unique(ped_data$Sensor_ID), 10, replace = FALSE)

ggplot(miss_plot_dat2 %>% filter(Sensor_ID %in% random_sensors)) +
  geom_point(aes(y = Sensor_ID, x = Date, colour = Weekly_Count), shape = 15, size = 8, alpha = 0.15) +
  ylab("Location") + theme_minimal() + theme(legend.position = "bottom") +
  scale_color_viridis(option = "magma", guide = guide_colorbar(barwidth = 10)) +
  ggtitle("Weekly Counts for at various locations")

```

---

### 2017-09-29 presentation/final.Rmd
added model to presentation file, still need to add imputation steps, predictive model

@@
 library(pander)
+library(png)
+library(grid)
 set.seed(13131313)
@@
 
-```{r, include = FALSE, echo = FALSE, warning=FALSE, message=FALSE}
+class: center
+
+# Introduction
+
+The City of Melbourne has an Open Data Platform which makes council data available to the public
+
+This project will focus on the pedestrian sensor data
+
+```{r, echo = FALSE, fig.height=5, fig.retina = 3}
+ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
+                         data = ped_loc, colour = "red", alpha = 0.7) +
+              labs(title = "Melbourne Pedestrian Sensor Locations") +
+              xlab(NULL) + ylab(NULL) +
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.ticks = element_blank())
+```
+
+---
+class: middle
+
+## How is the data collected?
+
+The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
+- These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+
+```{r, echo = FALSE, fig.align= 'center', fig.retina = 3, out.height= "369px", out.width = "852px"}
+include_graphics("../img/ped_sensors.png")
+```
+
+_source: http://www.pedestrian.melbourne.vic.gov.au/_
+
+---
+class: middle
+## Accessing the data
+
+- The data is available as a `.csv` at https://data.melbourne.vic.gov.au/, and is updated monthly
+- Alternatively, the data is also available as daily data from http://www.pedestrian.melbourne.vic.gov.au/
+  * The `rwalkr` R package by Earo Wang provides an API to access the data from both sources easily in R in a tidy format
+
+- Both sources of data provide the same counts data, but have different forms
+
+We use the daily data sourced from `pedestrian.melbourne.vic.gov.au` rather than the data available at the Open Data Platform as missing values are explicit
+
+---
+
+## Format of the data
+#### Variables available:
+- Time and Date
+- Sensor ID/Location
+- Hourly Pedestrian Count
+
+```{r, echo=FALSE}
+ped_data %>% as.tibble
+```
+
+---
+```{r, include = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache = TRUE}
 ped_data$Sensor_ID <- ped_data$Sensor
@@
 ```
-
-class: center
-
-# Introduction
-
-The City of Melbourne has an Open Data Platform which makes council data available to the public
-
-This project will focus on the pedestrian sensor data
-
-```{r, echo = FALSE, fig.height=5, fig.retina = 3}
-ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
-                         data = ped_loc, colour = "red", alpha = 0.7) +
-              labs(title = "Melbourne Pedestrian Sensor Locations") +
-              xlab(NULL) + ylab(NULL) +
-              theme(plot.title = element_text(hjust = 0.5),
-                    plot.caption = element_text(hjust = 0.5),
-                    axis.text.x = element_blank(),
-                    axis.text.y = element_blank(),
-                    axis.ticks = element_blank())
-```
-
----
-class: center
-
-## How is the data collected?
-
-The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
-- These sensors are positioned under an awning or street pole, and count the number of people passing each hour
-
-![figure of pedestrian sensors](../img/ped_sensors.png)
-
-_source: http://www.pedestrian.melbourne.vic.gov.au/ _
-
----
-class: middle
-## Accessing the data
-
-- The data is available as a `.csv` at https://data.melbourne.vic.gov.au/, and is updated monthly
-- Alternatively, the data is also available as daily data from http://www.pedestrian.melbourne.vic.gov.au/
-  * The `rwalkr` R package by Earo Wang provides an API to access the data from both sources easily in R in a tidy format
-
-- Both sources of data provide the same counts data, but have different forms
-
-We use the daily data sourced from `pedestrian.melbourne.vic.gov.au` rather than the data available at the Open Data Platform as missing values are explicit
-
----
-
-## Format of the data
-#### Variables available:
-- Time and Date
-- Sensor ID/Location
-- Hourly Pedestrian Count
-
-```{r, echo=FALSE}
-ped_data %>% as.tibble
-```
-
----
-
 # Why do we want to model pedestrian traffic?
@@
 ## Missing Data
-```{r, echo = FALSE, fig.retina=2, fig.width= 12, message=FALSE, warning=FALSE}
+```{r, echo = FALSE, fig.retina=2, fig.width= 12, message=FALSE, warning=FALSE, cache = TRUE}
 miss_plot_dat <- ped_data %>% filter(Date_Time > "2014-06-01",
@@
 
-random_sensors <- sample(unique(ped_data$Sensor_ID), 10, replace = FALSE)
+random_sensors <- c("Collins Place (North)","Lygon St (West)","Spring St-Lonsdale St (South)","Flinders St-Spring St (West)","Convention/Exhibition Centre","Sandridge Bridge")
 
-ggplot(miss_plot_dat2 %>% filter(Sensor_ID %in% random_sensors)) +
-  geom_point(aes(y = Sensor_ID, x = Date, colour = Weekly_Count), shape = 15, size = 8, alpha = 0.15) +
+mdf <- miss_plot_dat2 %>% filter(Sensor_ID %in% random_sensors)
+
+ggplot(mdf) +
+  geom_point(aes(y = Sensor_ID, x = Date, colour = Weekly_Count), shape = 15, size = 15, alpha = 0.15) +
   ylab("Location") + theme_minimal() + theme(legend.position = "bottom") +
@@
 
+- Some small and large periods of missing data
+- We need to impute these values before building a predictive model
 ---
+
+## Imputation of missing values
+
+- A simple approach of using a single model specification at all locations did not work well
+  * Some sensors had large proportions of missing data, particularly pre-2015
+- Need to use different model for sensors with large proportion of missing values and those with a small proportion of missing values
+
+```{r model_alg, echo = FALSE, warning=FALSE, message=FALSE, cache = TRUE}
+n_obs <- nrow(dfa)
+threshold <- 0.1
+threshold_miss_hrs <- 50
+
+run_lengths <- list()
+
+### code from:
+## http://www.win-vector.com/blog/2014/05/trimming-the-fat-from-glm-models-in-r/
+stripGlmLR = function(cm) {
+  cm$y = c()
+  cm$model = c()
+
+  cm$residuals = c()
+  cm$fitted.values = c()
+  cm$effects = c()
+  cm$qr$qr = c()  
+  cm$linear.predictors = c()
+  cm$weights = c()
+  cm$prior.weights = c()
+  cm$data = c()
+
+  cm$family$variance = c()
+  cm$family$dev.resids = c()
+  cm$family$aic = c()
+  cm$family$validmu = c()
+  cm$family$simulate = c()
+  attr(cm$terms,".Environment") = c()
+  attr(cm$formula,".Environment") = c()
+
+  cm
+}
+
+for(i in unique(ped_data$Sensor)){
+    dfa[is.na(dfa[, i]), i] <- 0
+    na_length_check <- rle(as.numeric(unlist(dfa[,i])))
+    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs),
+                   times = na_length_check$lengths)
+    rl <- dfa[, i] %>% unlist %>% rle()
+    dfa[maybeNA, i] <- NA
+    defsNA <- is.na(ori_2[, i])
+    dfa[defsNA, i] <- NA
+    run_lengths[[i]] <- rl
+}
+
+na_prop <- numeric()
+
+for (i in unique(ped_data$Sensor)){
+  na_prop[i] <- sum(is.na(dfa[, i])) / n_obs
+}
+
+large_miss_sensors <- names(na_prop[na_prop > threshold])
+small_miss_sensors <- setdiff(names(na_prop), large_miss_sensors)
+
+models <- list()
+fitted_data <- list()
+dfa2 <- dfa
+for(i in small_miss_sensors){
+  options(na.action = na.omit)
+  sensor_df <- unlist(dfa[, i])
+  model <- glm(sensor_df ~  Month + DayType*Time,
+               family = 'quasipoisson', data = dfa)
+  models[[i]] <- (stripGlmLR(model))
+}
+
+  options(na.action = na.pass)
+for(i in small_miss_sensors){
+  ts_dat <- data.frame(Month = dfa$Month, DayType = dfa$DayType, Time = dfa$Time)
+  fitted_data[[i]] <- predict.glm(models[[i]],newdata = ts_dat, type = "response")
+}
+  options(na.action = na.omit)
+  for(i in small_miss_sensors){
+    dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
+  }
+
+  lats <- ped_loc$Latitude
+lons <- ped_loc$Longitude
+ids <- ped_loc$`Sensor Description`
+
+## name fix dictionary
+{
+ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
+ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
+ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
+ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
+ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
+ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
+ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
+ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
+ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
+}
+loc <- data.frame(lons, lats, ids)
+rownames(loc) <- ids
+
+# Calculates the geodesic distance between two points specified by radian latitude/longitude using the
+# Spherical Law of Cosines (slc)
+# source: https://www.r-bloggers.com/great-circle-distance-calculations-in-r/
+gcd.slc <- function(long1, lat1, long2, lat2) {
+  R <- 6371 # Earth mean radius [km]
+  d <- acos(sin(lat1)*sin(lat2) + cos(lat1)*cos(lat2) * cos(long2-long1)) * R
+  return(d) # Distance in km
+}
+
+for (i in large_miss_sensors){
+  # p$tick()$print()
+  sensor_dists <- numeric()
+  for (j in small_miss_sensors){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  close_df_scale <- dfa2[, closest_sensor[1]] %>% unlist %>% scale
+  close_df_scale2 <- dfa2[, closest_sensor[2]] %>% unlist %>% scale
+
+  sensor_df <- dfa2[, i] %>% unlist
+
+  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa$Time)
+  options(na.action = na.omit)
+  model <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
+               data = tr_dat, family = quasipoisson())
+  models[[i]] <- stripGlmLR(model)
+
+  options(na.action = na.pass)
+  fitted_data[[i]] <- predict(model, newdata = tr_dat, type = "response")
+  options(na.action = na.omit)
+
+  dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
+
+}
+
+```
+
+```{r, echo = FALSE, fig.width = 7, fig.height = 4.5, fig.retina = 3, fig.align = 'center'}
+fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
+                           fit_ = dfa2$`Southern Cross Station`) %>%
+                  filter(Date_Time < "2014-07-06 00:00:00",
+                         Date_Time > "2014-07-04 12:00:00") %>%
+                        .$fit_
+
+ggplot(original_df %>%  filter(Date_Time < "2014-07-06 00:00:00",
+                               Date_Time > "2014-07-04 12:00:00")) +
+  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = 'red',
+            size = 2) +
+  geom_line(aes(x = Date_Time, y = `Southern Cross Station`),
+            size  = 2) +
+  theme_minimal() + xlab("Time") +
+  ylab("Pedestrian Counts") +
+  ggtitle("Pedestrian Counts at Southern Cross Station on 04/07/2014 - 05/07/2014")
+```
+
+---
+class: middle
+## Imputation Algorithm
+
+#### Needed to develop an algorithm to:
+
+- Find which sensors have a "large" proportion of missing values  
+- At each sensor location, run appropriate GLM model with quasipoisson error distribution for imputation  
+  * if "small", use GLM with `Hourly_Counts` as response variable and time and date based variables as predictors
+  * if "large", use GLM with `Hourly_Counts` as response variable and neighbouring sensor counts as predictors
+- Replace missing values with imputed values
+
+---

### 2017-09-29 presentation/final.Rmd
added first step to algorithm for imputation

@@
   * The `rwalkr` R package by Earo Wang provides an API to access the data from both sources easily in R in a tidy format
-
 - Both sources of data provide the same counts data, but have different forms
@@
 class: middle
-## Imputation Algorithm
+# Imputation Algorithm
 
@@
 
-- Find which sensors have a "large" proportion of missing values  
-- At each sensor location, run appropriate GLM model with quasipoisson error distribution for imputation  
+- **Step 1:** Find which sensors have a "large" proportion of missing values  
+- **Step 2:** At each sensor location, run appropriate GLM model with quasipoisson error distribution for imputation  
   * if "small", use GLM with `Hourly_Counts` as response variable and time and date based variables as predictors
   * if "large", use GLM with `Hourly_Counts` as response variable and neighbouring sensor counts as predictors
-- Replace missing values with imputed values
+- **Step 3:** Replace missing values with imputed values  
+---
+## Step 1: Calculating proportion of missing values
+- Need to first fix potential false zero-count values
+  * Some observations of `Count = 0` may actually be missing values
+  * Include a check for long periods with zero count, and replace with `NA` if too long
+- Classify sensors as "small" or "large" proportion of missing values with a threshold proportion
+  * We used 10% as the threshold proportion
+```{r, fig.retina = 3, echo = FALSE, message = FALSE, fig.height = 4, fig.align = 'center'}
+loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
+                                  "Small", "Large"))
+melbbw <- get_map(location=c(mean(range(ped_loc$Longitude)),
+                            mean(range(ped_loc$Latitude))),
+                            zoom=14, source = "stamen", maptype = "toner")
+
+ggmap(melbbw) + geom_point(aes(x = lons, y = lats, colour = cl),
+                         data = loc, alpha = .6, size = 4, shape = 15) +
+              labs(title = "Melbourne Pedestrian Sensor Locations") +
+              xlab(NULL) + ylab(NULL) +
+              scale_color_viridis(name = "Proportion of missing values",
+                                  discrete = TRUE) +
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.ticks = element_blank(),
+                    legend.position = "bottom")
+```
 
 ---
+
+## Step 2: Fit models with "small" proportion of missing values
+
+At each sensor location with small proportion of missing values, we estimate a generalised linear model with:
+$$\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType}$$
+and a quaispoisson error distribution.
+- Quasipoisson regression allows for $Var[\text{HourlyCount}] = \theta \mu$
+- These models only use time and date based variables to predict pedestrian counts
+  * `DayType` is a factor level/categorical variable based on the day of the week and public holidays
+  * `Time` is treated as a factor level, as it has a non-linear relationship with counts

### 2017-09-29 presentation/final.Rmd
update

@@
 ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
-                         data = ped_loc, colour = "red", alpha = 0.7) +
+                         data = ped_loc, colour = "red",
+                         size = 4, alpha = 0.7, shape = 17) +
               labs(title = "Melbourne Pedestrian Sensor Locations") +
@@
                                Date_Time > "2014-07-04 12:00:00")) +
-  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = 'red',
+  geom_line(aes(x = Date_Time, y = fit_dat_plot),
+            colour = plasma(1, begin = .75, end = 0),
             size = 2) +
@@
   * We used 10% as the threshold proportion
-```{r, fig.retina = 3, echo = FALSE, message = FALSE, fig.height = 4, fig.align = 'center'}
+```{r, fig.retina = 4, echo = FALSE, message = FALSE, fig.height = 4, fig.align = 'center', warning=FALSE}
 loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
                                   "Small", "Large"))
-melbbw <- get_map(location=c(mean(range(ped_loc$Longitude)),
+melb2 <- get_map(location = c(mean(range(ped_loc$Longitude)),
                             mean(range(ped_loc$Latitude))),
-                            zoom=14, source = "stamen", maptype = "toner")
+                            zoom=14, maptype = "satellite")
+perf_loc <- c(144.937, -37.826, 144.977, -37.794)
 
-ggmap(melbbw) + geom_point(aes(x = lons, y = lats, colour = cl),
+ggmap(melb2) + geom_point(aes(x = lons, y = lats, colour = cl),
                          data = loc, alpha = .6, size = 4, shape = 15) +
@@
               scale_color_viridis(name = "Proportion of missing values",
-                                  discrete = TRUE) +
+                                  discrete = TRUE, option = "C") +
+              scale_x_continuous(limits = c(144.937, 144.977)) +
+              scale_y_continuous(limits = c(-37.83, -37.79)) +
               theme(plot.title = element_text(hjust = 0.5),
@@
                     axis.ticks = element_blank(),
-                    legend.position = "bottom")
+                    legend.position = "right")
 ```
@@
   * `Time` is treated as a factor level, as it has a non-linear relationship with counts
+
+---
+
+## Step 2: Replacing missing values with imputed values
+
+- Need to fill the missing data at the sensors with small proportions of missing values first
+  * The imputation models at sensors with large proportions of missing values are based on neighbouring sensors
+  * Requires complete data at neighbours to work properly
+
+```{r, echo = FALSE, fig.retina = 3, fig.width = 10, fig.height = 4}
+fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
+                           fit_ = dfa2$`Flagstaff Station`) %>%
+                  filter(Date_Time < "2014-07-05 15:00:00",
+                         Date_Time > "2014-07-03 23:00:00") %>%
+                        .$fit_
+
+p1 <-  ggplot(original_df %>%  filter(Date_Time < "2014-07-05 15:00:00",
+                                 Date_Time > "2014-07-03 23:00:00")) +
+        geom_line(aes(x = Date_Time, y = `Flagstaff Station`),
+                  size  = 2) +
+        theme_minimal() + xlab("Time") +
+        ylab("Pedestrian Counts") +
+        ggtitle("Flagstaff Station")
+p2 <-  ggplot(original_df %>%  filter(Date_Time < "2014-07-05 15:00:00",
+                                 Date_Time > "2014-07-03 23:00:00")) +
+        geom_line(aes(x = Date_Time, y = fit_dat_plot),
+                  colour = plasma(1, begin = .335, end = 0),
+                  size = 2) +
+        geom_line(aes(x = Date_Time, y = `Flagstaff Station`),
+                  size  = 2) +
+        theme_minimal() + xlab("Time") +
+        ylab("Pedestrian Counts") +
+        ggtitle("Flagstaff Station with imputed values")
+grid.arrange(p1, p2, nrow = 1)
+```

### 2017-09-29 presentation/final.Rmd
added visualisation for closest neighbours

@@
 class: middle, center
-## Missing Data
+# Missing Data
 ```{r, echo = FALSE, fig.retina=2, fig.width= 12, message=FALSE, warning=FALSE, cache = TRUE}
@@
 
-## Step 2: Fit models with "small" proportion of missing values
+## Step 2a: Fit models at sensors with "small" proportion of missing values
 
@@
 
-## Step 2: Replacing missing values with imputed values
+## Step 2b: Replacing missing values with imputed values
 
@@
 
-```{r, echo = FALSE, fig.retina = 3, fig.width = 10, fig.height = 4}
+```{r, echo=FALSE, fig.height=4, fig.retina=3, fig.width=10, message=FALSE, warning=FALSE}
 fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
@@
         geom_line(aes(x = Date_Time, y = `Flagstaff Station`),
+                  size  = 1.5) +
+        geom_point(aes(x = Date_Time, y = `Flagstaff Station`),
                   size  = 2) +
@@
         geom_line(aes(x = Date_Time, y = fit_dat_plot),
-                  colour = plasma(1, begin = .335, end = 0),
-                  size = 2) +
+                  colour = plasma(1, begin = .375, end = 0),
+                  size = 1.5) +
         geom_line(aes(x = Date_Time, y = `Flagstaff Station`),
+                  size  = 1.5) +
+        geom_point(aes(x = Date_Time, y = `Flagstaff Station`),
                   size  = 2) +
@@
 ```
+
+---
+class: middle
+## Step 2c: Fit models at sensors with "large" proportion of missing values
+
+```{r, echo = FALSE, fig.retina = 3, message=FALSE, warning=F}
+rownames(loc) <- loc$ids
+melb3 <- get_map(location = c(loc["Southbank",]$lons,
+                              loc["Southbank",]$lats),
+                 zoom = 17, maptype = "roadmap")
+
+sens_lines_lon <- as.numeric(c(loc["Flinders St Station Underpass",]$lons,
+                loc["Southbank",]$lons,
+                loc["Sandridge Bridge",]$lons))
+
+sens_lines_lat <- as.numeric(c(loc["Flinders St Station Underpass",]$lats,
+                loc["Southbank",]$lats,
+                loc["Sandridge Bridge",]$lats))
+
+sens_lines <- data.frame(lon = sens_lines_lon,
+                         lat = sens_lines_lat)
+```
+
+- Need to use complete data from neighbouring sensors
+  * Use data from imputation in previous step
+- Neighbour defined as geographical neighbours for simplicity
+  * Find two closest sensors by great-circle (haversine) distance
+---
+class: center, middle
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width=10}
+ggmap(melb3) + geom_line(aes(x = lon, y = lat),
+                        data = sens_lines, size = 2.5,
+                        colour = viridis(3)[2], alpha = 0.5) +
+               geom_point(aes(x = lons, y = lats, colour = cl),
+                          size = 12,
+                          data = loc) +
+               scale_colour_viridis(discrete = TRUE,
+                                    name = "Sensor NA Prop.",
+                                    begin = 0.5, end = 0) +
+               geom_point(aes(x = loc["Southbank",]$lons,
+                              y = loc["Southbank",]$lats),
+                          size = 26, colour = viridis(3)[2]) +
+               geom_text(aes(label = ids, x = lons, y = lats),
+                         data = loc, size = 8,
+                         nudge_y = -0.0002) +
+              xlab(NULL) + ylab(NULL) +
+              scale_y_continuous(limits = c(-37.8215, -37.818)) +
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.ticks = element_blank(),
+                    legend.position = "right")
+```

### 2017-09-29 presentation/final.Rmd
update - map of sensors by missing values proportion on separate slide

@@
 - **Step 3:** Replace missing values with imputed values
+
 ---
@@
   * We used 10% as the threshold proportion
-```{r, fig.retina = 4, echo = FALSE, message = FALSE, fig.height = 4, fig.align = 'center', warning=FALSE}
+
+```{r, echo = FALSE, fig.retina = 3, fig.height = 3, fig.align='center'}
+ggplot() + geom_histogram(aes(x = na_prop*100), bins = 12) +
+          xlab("Percentage of Missing Values (%)") +
+          ylab("Number of Sensors") + theme_minimal() +
+          geom_vline(xintercept = 10, size = 2,
+                     colour = 'red') +
+          geom_text(aes(label = "10% Threshold", x = 10, y = 1),
+                    colour = "white") +
+          ggtitle("Histogram of Proportions of Missing Values")
+```
+
+---
+class: middle, center
+## Step 1: Classification of sensors by "small" and "large" proportion of  missing values
+
+```{r, fig.retina = 4, echo = FALSE, message = FALSE, fig.width = 11, fig.align = 'center', warning=FALSE}
 loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
@@
 ggmap(melb2) + geom_point(aes(x = lons, y = lats, colour = cl),
-                         data = loc, alpha = .6, size = 4, shape = 15) +
+                         data = loc, alpha = .6, size = 8, shape = 15) +
               labs(title = "Melbourne Pedestrian Sensor Locations") +

### 2017-09-30 presentation/final.Rmd
sat morning update - still finishing off imputation slides

@@
                     colour = "white") +
-          ggtitle("Histogram of Proportions of Missing Values")
+          ggtitle("Histogram of Proportions of Missing Values during 2014 - 2016")
 ```
@@
 
-```{r, echo = FALSE, fig.retina = 3, message=FALSE, warning=F}
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.height = 4, fig.align='center'}
 rownames(loc) <- loc$ids
@@
                          lat = sens_lines_lat)
-```
-
-- Need to use complete data from neighbouring sensors
-  * Use data from imputation in previous step
-- Neighbour defined as geographical neighbours for simplicity
-  * Find two closest sensors by great-circle (haversine) distance
----
-class: center, middle
-```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width=10}
 ggmap(melb3) + geom_line(aes(x = lon, y = lat),
@@
                geom_point(aes(x = lons, y = lats, colour = cl),
-                          size = 12,
+                          size = 8,
                           data = loc) +
@@
                               y = loc["Southbank",]$lats),
-                          size = 26, colour = viridis(3)[2]) +
+                          size = 16, colour = viridis(3)[2]) +
                geom_text(aes(label = ids, x = lons, y = lats),
-                         data = loc, size = 8,
+                         data = loc, size = 4,
                          nudge_y = -0.0002) +
               xlab(NULL) + ylab(NULL) +
-              scale_y_continuous(limits = c(-37.8215, -37.818)) +
               theme(plot.title = element_text(hjust = 0.5),
@@
                     axis.ticks = element_blank(),
-                    legend.position = "right")
+                    legend.position = c(.8,0.2))
+
 ```
+- Need to use complete data from neighbouring sensors
+  * Use data from imputation in previous step
+- Neighbour defined as geographical neighbours for simplicity
+  * Find two closest sensors by great-circle (haversine) distance
+
+---
+class: center, middle
+## _Step 2c: Fit models at sensors with "large" proportion of missing values_
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, include=FALSE}
+i <- "Southbank"
+  sensor_dists <- numeric()
+  for (j in small_miss_sensors){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  dfa3 <- original_df %>% filter(Date_Time > "2017-01-01")
+  close_df <- dfa3[, closest_sensor[1]] %>% unlist
+  close_df2 <- dfa3[, closest_sensor[2]] %>% unlist
+
+  close_dfsc <- scale(close_df)
+  close_dfsc2 <- scale(close_df2)
+
+  sensor_df <- dfa3[, i] %>% unlist
+  sensor_dfsc <- scale(sensor_df)
+
+  tr_dat <- data.frame(sensor_df, close_df, close_df2,
+                       Time = as.integer(dfa3$Time),
+                       Date_Time = dfa3$Date_Time,
+                       HDay = dfa3$HDay) %>%
+            gather(key = Sensor, value = Count, sensor_df:close_df2)
+
+  tr_dat_sc <- data.frame(sensor_dfsc, close_dfsc, close_dfsc2,
+                       Time = as.integer(dfa3$Time),
+                       Date_Time = dfa3$Date_Time,
+                       HDay = dfa3$HDay) %>%
+            gather(key = Sensor, value = Count, sensor_dfsc:close_dfsc2)
+
+p1 <- ggplot(tr_dat %>% filter(Date_Time < "2017-03-02",
+                         Date_Time > "2017-02-26")) +
+      geom_line(aes(x = Date_Time, y = Count,
+                    colour = Sensor), size = 2) +
+      scale_colour_viridis(discrete = TRUE,
+                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
+                                      "Neighbour 2 (Sandridge Bridge)",
+                                      "Target (Southbank)")) +
+      xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
+      theme(legend.position = "bottom")
+
+p2 <- ggplot(tr_dat_sc %>% filter(Date_Time < "2017-03-02",
+                         Date_Time > "2017-02-26")) +
+      geom_line(aes(x = Date_Time, y = Count,
+                    colour = Sensor), size = 2) +
+      scale_colour_viridis(discrete = TRUE,
+                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
+                                      "Neighbour 2 (Sandridge Bridge)",
+                                      "Target (Southbank)")) +
+      xlab("Time") + ggtitle("Scaled/Stadardised Neighbours") +
+      ylab("Standard Deviation") + theme_minimal() +
+      theme(legend.position = "none")
+```
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width=10, fig.height=6.5}
+grid.arrange(p1, p2, ncol = 1)
+```

### 2017-09-30 presentation/final.Rmd
update - fixed slide titles

@@
 
-## Format of the data
+# Format of the data
 #### Variables available:
-- Time and Date
-- Sensor ID/Location
-- Hourly Pedestrian Count
+- Time and date, sensor ID/location and hourly pedestrian count (volume)
 
 ```{r, echo=FALSE}
-ped_data %>% as.tibble
+print(ped_data %>% as.tibble, n = 4)
 ```
-
----
 ```{r, include = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache = TRUE}
@@
 ```
-# Why do we want to model pedestrian traffic?
+
+```{r, echo=FALSE, fig.retina = 3, fig.height=3, fig.width= 10, warning=F, message = F}
+pd2 <- ggplot(dfa %>% 
+              filter(Date_Time > "2016-07-12 00:00:00",
+                     Date_Time < "2016-07-25 00:00:00")) +
+        scale_y_continuous(limits = c(0, 7000)) +
+        xlab("Time and Date") +
+        theme_minimal()
+
+p1 <-  pd2 + geom_line(aes(x = Date_Time, y = `Flagstaff Station`))
+p2 <-  pd2 + geom_line(aes(x = Date_Time, y = `Southbank`))
+p3 <-  pd2 + geom_line(aes(x = Date_Time, y = `State Library`))
+
+grid.arrange(p1, p2, p3, ncol = 1)
+```
+
+---
+class: center
+# The Problem: Prediction of pedestrian traffic
+```{r, echo=FALSE, fig.retina = 3, fig.height=5, fig.width= 10, warning=F, message = F}
+pd <- ggplot(dfa %>% filter(Date_Time < "2017-01-01 00:00:00",
+                            Date_Time > "2013-12-31 00:00:00")) +
+        scale_y_continuous(limits = c(0, 7000)) +
+        xlab(NULL) +
+        theme_minimal()
+
+p1 <-  pd + geom_line(aes(x = Date_Time, y = `Flagstaff Station`))
+p2 <-  pd + geom_line(aes(x = Date_Time, y = `Southbank`))
+p3 <-  pd + geom_line(aes(x = Date_Time, y = `State Library`)) +
+            xlab("Time and Date")
+
+grid.arrange(p2, p1, p3, nrow = 3, ncol = 1)
+```
+
+- Different locations have different counts and patterns
+- We want to predict using 2014-2016 data to train our model
+
+---
+## Why do we want to model pedestrian traffic?
 
@@
 ---
-class: center
-## The Problem: Prediction of pedestrian traffic
-```{r, echo=FALSE, fig.retina = 3, fig.height=6, fig.width= 10}
-p1 <- ggplot(dfa %>% filter(Date_Time > "2016-07-12 00:00:00",
-                            Date_Time < "2016-07-25 00:00:00")) +
-        geom_line(aes(x = Date_Time, y = `Flagstaff Station`)) +
-        scale_y_continuous(limits = c(0, 7000))
-p2 <- ggplot(dfa %>% filter(Date_Time > "2016-07-12 00:00:00",
-                            Date_Time < "2016-07-25 00:00:00")) +
-        geom_line(aes(x = Date_Time, y = `Southbank`)) +
-        scale_y_continuous(limits = c(0, 7000))
-p3 <- ggplot(dfa %>% filter(Date_Time > "2016-07-12 00:00:00",
-                            Date_Time < "2016-07-25 00:00:00")) +
-        geom_line(aes(x = Date_Time, y = `State Library`)) +
-        scale_y_continuous(limits = c(0, 7000))
-grid.arrange(p2, p1, p3, nrow = 3, ncol = 1)
-```
-
---
-
-#### Different locations have different counts and patterns
-
----
 class: middle, center
@@
 
-random_sensors <- c("Collins Place (North)","Lygon St (West)","Spring St-Lonsdale St (South)","Flinders St-Spring St (West)","Convention/Exhibition Centre","Sandridge Bridge")
+random_sensors <- c("Collins Place (North)","Southbank","Spring St-Lonsdale St (South)",
+                    "Flinders St-Spring St (West)","Convention/Exhibition Centre","Sandridge Bridge")
 
-mdf <- miss_plot_dat2 %>% filter(Sensor_ID %in% random_sensors)
+mdf <- miss_plot_dat2 %>% filter(Sensor_ID %in% random_sensors) %>%
+        mutate(Sensor_ID = as.factor(Sensor_ID))
 
@@
 class: middle
-## Step 2c: Fit models at sensors with "large" proportion of missing values
+## Step 2c: Find neighbours for sensors with "large" proportion of missing values
 
@@
 class: center, middle
-## _Step 2c: Fit models at sensors with "large" proportion of missing values_
+## _Step 2c: Fit find neighbours sensors with "large" proportion of missing values_
 
@@
       xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
-      theme(legend.position = "bottom")
+      theme(legend.position = "none")
 
@@
       ylab("Standard Deviation") + theme_minimal() +
-      theme(legend.position = "none")
+      theme(legend.position = "bottom")
 ```
 
-```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width=10, fig.height=6.5}
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width=10, fig.height=6}
 grid.arrange(p1, p2, ncol = 1)
 ```
+
+- Need to standardise counts at neighbours to match patterns
+
+---
+## _Step 2c: Fit models using neighbours_
+At each sensor location with "large" proportion of missing values, we estimate a generalised linear model with:
+
+$$\mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}^{SC}} \sim \text{Time} \times \text{HourlyCounts}_{\text{neighbour 1}}^{SC} + \text{Time} \times \text{HourlyCounts}_{\text{neighbour 2}}^{SC}$$
+
+and a quaispoisson error distribution.
+
+- use interaction with `Time`, the hour of the day, and the scaled counts at neighbours
+  * Again, `Time` is treated as a factor variable
+  * Counts scaled by standardising each sensor to $\mu = 0$ and $\sigma^2 = 1$

### 2017-09-30 presentation/final.Rmd
update

@@
 We use the daily data sourced from `pedestrian.melbourne.vic.gov.au` rather than the data available at the Open Data Platform as missing values are explicit
+- accessible using `rwalkr::walk_melb()`
 
@@
 threshold <- 0.1
-threshold_miss_hrs <- 50
+threshold_miss_hrs <- 17
 
@@
 and a quaispoisson error distribution.
-- Quasipoisson regression allows for $Var[\text{HourlyCount}] = \theta \mu$
+- Quasipoisson regression allows for $Var[\text{HourlyCount}] = \theta \mu$ (ie. overdispersion)
 - These models only use time and date based variables to predict pedestrian counts
@@
 ```
-- Need to use complete data from neighbouring sensors
-  * Use data from imputation in previous step
+- Need to use data from neighbouring sensors
 - Neighbour defined as geographical neighbours for simplicity
@@
 ---
+class: middle
 ## _Step 2c: Fit models using neighbours_
@@
   * Counts scaled by standardising each sensor to $\mu = 0$ and $\sigma^2 = 1$
+
+---
+## Step 3: Replacing missing values with imputed values
+
+Using the models estimated, we can impute the missing values
+- For sensors with large proportion of missing values, we needed to have complete data at the neighbours
+  * Can't use neighbours with large proportion of missing values
+  * Use imputed data from sensors with small proportion of missing values
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width=10, fig.height = 4}
+p1 <-   ggplot(original_df %>% filter(Date_Time < "2017-01-01")) +
+        geom_line(aes(x = Date_Time, y = `Southbank`)) +
+        scale_y_continuous(limits = c(0, 7500)) +
+        theme_minimal() + ylab("Pedestrian Count") +
+        ggtitle("Southbank Pedestrian Counts before Imputation")
+
+p2 <-   ggplot(original_df %>% filter(Date_Time < "2017-01-01")) +
+        geom_line(aes(x = Date_Time, y = dfa2$Southbank),
+                  colour = viridis(3)[1]) +
+        geom_line(aes(x = Date_Time, y = `Southbank`)) +
+        scale_y_continuous(limits = c(0, 7500)) +
+        theme_minimal() + ylab("Pedestrian Count") +
+        ggtitle("Southbank Pedestrian Counts after Imputation")
+grid.arrange(p1, p2, ncol = 1)
+```
+
+---
+class: middle
+# Implementation of the imputation algorithm
+
+- general code which will run the algorithm
+- inputs required: 
+  * data in wide format
+  * threshold value for missing values proportion (_default:_ `threshold = 0.1`)
+  * max length of repeated values (_default:_ `maxlength = 17`)
+- use R and RStudio with packages:
+  * `dplyr`, `lubridate` and `tidyr` for data manipulation
+  * `foreach` and `doSNOW` for parallel computing
+  * `rwalkr` for accessing latest data
+---
+## Imputation Model Evaluation
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 9}
+p1 <-   ggplot(data = NULL, aes(x = original_df %>% 
+             filter(Date_Time < "2017-01-01") %>%
+             .[, "Southern Cross Station"],
+           y = fitted_data$`Southern Cross Station`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.5, size = 1.5,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + theme(aspect.ratio = 1) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Actual against Predicted at Southern Cross Station")
+
+p2 <-   ggplot(data = NULL, aes(x = dfa %>% 
+             filter(Date_Time < "2017-01-01") %>%
+             .[, "Southbank"],
+           y = fitted_data$`Southbank`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.5, size = 1.5,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + theme(aspect.ratio = 1) +
+           scale_x_continuous(limits = c(0, 6000)) +
+           scale_y_continuous(limits = c(0, 6000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Actual against Predicted at Southbank")
+grid.arrange(p1, p2, nrow = 1)
+```

### 2017-09-30 presentation/final.Rmd
started slides of imputation evaluation: not sure if should be incldued

@@
 
+neighbours <- list()
+
 for (i in large_miss_sensors){
@@
   close_df_scale2 <- dfa2[, closest_sensor[2]] %>% unlist %>% scale
+
+  neighbours[[i]] <- closest_sensor
 
@@
 ## Imputation Model Evaluation
+### In-sample Fit
 
@@
 ```
+
+```{r}
+MARE_glm_tr <- numeric()
+for (i in unique(ped_data$Sensor)){
+  MARE_glm_tr[i] <- mean(abs(fitted_data[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
+}
+```
+
+---
+class: center, middle
+## Imputation Model Evaluation
+### Out-of-sample cross validation
+```{r, echo = FALSE}
+test_fit <- list()
+test_df <- original_df %>% filter(Date_Time > "2016-12-31")
+  for(i in small_miss_sensors){
+    test_fit[[i]] <- predict.glm(models[[i]],
+                                 newdata = test_df,
+                                 type = "response")
+    }
+
+  for(i in large_miss_sensors){
+    ts_data <- data.frame(close_dfsc = (test_df[, neighbours[[i]][1]] %>% scale),
+                          close_dfsc2 = (test_df[, neighbours[[i]][2]] %>% scale),
+                          Time = test_df$Time)
+    colnames(ts_data) <- c("close_df_scale", "close_df_scale2", "Time")
+    test_fit[[i]] <- predict.glm(models[[i]],
+                                 newdata = ts_data,
+                                 type = "response")
+  }
+MARE_glm_ts <- numeric()
+  for (i in unique(ped_data$Sensor)){
+    MARE_glm_ts[i] <- mean(abs(test_fit[[i]] - unlist(test_df[, i])),
+                           na.rm = T) / mean(unlist(test_df[, i]), na.rm = T)
+    }
+
+```
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 3.5, fig.height = 3.5}
+ggplot(test_df, aes(x = `Bourke Street Mall (North)`,
+                    y = test_fit$`Bourke Street Mall (North)`)) +
+                  geom_point(alpha = 0.5) +
+                  scale_x_continuous(limits = c(0, 5000)) +
+                  scale_y_continuous(limits = c(0, 5000)) +
+                  xlab("Actual") + ylab("Fitted") + theme(aspect.ratio = 1) +
+                  geom_abline(aes(slope = 1, intercept = 0),
+                              colour = viridis(3)[1],
+                              alpha = 0.7, size = 1.5,
+                              linetype = 2) +
+                  geom_smooth(method = "lm", size = 1.5, 
+                       colour = viridis(3)[3]) +
+                  ggtitle("Actual vs Fitted \n at Bourke Street Mall (North) in 2017")
+
+```
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 10, fig.height = 3}
+fit_t_plot <- data.frame(fit = test_fit$`Bourke Street Mall (North)`,
+                         Date_Time = test_df$Date_Time)  
+ggplot(test_df %>% filter(Date_Time > "2017-08-31")) +
+  geom_line(aes(x = Date_Time,
+                y = fit), data = (fit_t_plot %>%
+                                  filter(Date_Time > "2017-08-31")),
+            colour = magma(3)[2], alpha = 0.6) +
+  geom_line(aes(x = Date_Time,
+                y = `Bourke Street Mall (North)`),
+            alpha = 0.6) +
+  scale_y_continuous(limits = c(0, 5000)) +
+  theme_minimal() + xlab("Time") + ylab("Count") +
+  ggtitle("Bourke Street Mall (North)")
+
+```
+
+---

### 2017-09-30 presentation/final.Rmd
added slide about prediction - maybe need to say future work, or present a basic initial prediction model

@@
 ---
+
+# Predictive model using imputed data
+- can perform analysis at all locations now with complete data
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 10, fig.height = 3}
+melb_cent_mon9 <- dfa2 %>%
+  filter(Day == "Monday",
+         Time == 9) %>%
+         .$`Melbourne Central`
+
+monday_hdays <- dfa2 %>%
+  filter(Day == "Monday",
+         Time == 9) %>%
+  .$IsHDay
+
+p1 <- ggplot() + geom_line(aes(x = 1:length(melb_cent_mon9),
+                             y = melb_cent_mon9)) +
+      theme_minimal() + xlab("Time (nth Week since 2014)") +
+      geom_point(aes(x = 1:length(melb_cent_mon9),
+                     y = 0, alpha = monday_hdays),
+                 size = 3) + 
+      scale_alpha_discrete(name = "Is public holiday") + 
+      ylab("Counts at Monday 9am")
+p2 <- ggplot() + geom_density(aes(x = melb_cent_mon9)) +
+                 theme_minimal() + xlab("Counts at Monday 9am")
+
+grid.arrange(p1, p2, nrow = 1, top = "Monday 9am Counts at Melbourne Central")
+
+```
+- can treat each specific hour of the week as a time series
+  * find covariates outside the counts data to help predict counts
+  * eg. weather data, public events

### 2017-10-01 presentation/final.Rmd
added final/future steps for the prject - may need to either add more points or elaborate on points already there
update - more detailed future steps

@@
 ---
+class: center
 ## Imputation Model Evaluation
-### In-sample Fit
+### In-sample fit
 
@@
   * eg. weather data, public events
+
+### Objective: On the fly predictions of pedestrian traffic
+- requires efficient data storage
+
+---
+# Future steps for final paper
+- develop Shiny app wih RStudio to provide on the fly predictions
+  * make the predictive models more accessible
+  * provide visualisations to present the predictions
+
+```{r, cache = TRUE, fig.retina = 4, echo = FALSE, message = FALSE, fig.align= 'center', fig.width=5, fig.height=4, warning=F}
+mar13_16_8am <- ped_data[ped_data$Date == "2016-03-13", ] %>%
+                filter(Time == 8)
+
+loc$Sensor <- loc$ids
+
+mar13_16_8am %>% left_join(loc) -> mar13_16_8am
+
+bourkest_loc <- loc[loc$ids == "Flinders St Station Underpass",]
+
+  melb_dark <- get_map(location=c(bourkest_loc$lons, bourkest_loc$lats),
+                                zoom=16, maptype = "toner-lite")
+
+ggmap(melb_dark) + geom_point(aes(x = lons, y = lats, alpha = Count),
+                              colour = magma(8)[6],
+                      data = mar13_16_8am, shape = 15, size = 16) +
+              scale_alpha_continuous(range = c(.1, .6)) +
+              theme_minimal() +
+                theme(plot.title = element_text(hjust = 0.5),
+                    plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.ticks = element_blank(),
+                    legend.position = "right") +
+                xlab(NULL) + ylab(NULL) + ggtitle("Pedestrian Traffic around Flinders St \n at 13 Mar 2016, 8AM")
+
+```
+
+- explore predictors outside of counts data

### 2017-10-01 presentation/final.Rmd
fixed plot on slide 21

@@
 
-```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 3.5, fig.height = 3.5}
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 4.5, fig.height = 3.5}
 ggplot(test_df, aes(x = `Bourke Street Mall (North)`,

### 2017-10-01 presentation/final.Rmd
update

@@
   ylab("Pedestrian Counts") +
-  ggtitle("Pedestrian Counts at Southern Cross Station on 04/07/2014 - 05/07/2014")
+  ggtitle("Pedestrian Counts at Southern Cross Station \n on 04/07/2014 - 05/07/2014") +
+  theme(plot.title = element_text(hjust = 0.5))
 ```
@@
                          data = loc, size = 4,
-                         nudge_y = -0.0002) +
+                         nudge_y = -0.0003) +
               xlab(NULL) + ylab(NULL) +
@@
       xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
-      theme(legend.position = "none")
+      theme(legend.position = "none",
+            plot.title = element_text(hjust = 0.5))
 
@@
       ylab("Standard Deviation") + theme_minimal() +
-      theme(legend.position = "bottom")
+      theme(legend.position = "bottom",
+            plot.title = element_text(hjust = 0.5))
 ```
@@
   * max length of repeated values (_default:_ `maxlength = 17`)
+  * `threshold` and `maxlength` are tuning parameters
 - use R and RStudio with packages:
@@
                     y = test_fit$`Bourke Street Mall (North)`)) +
-                  geom_point(alpha = 0.5) +
+                  geom_point(alpha = 0.2) +
                   scale_x_continuous(limits = c(0, 5000)) +
@@
                   geom_smooth(method = "lm", size = 1.5, 
-                       colour = viridis(3)[3]) +
+                       colour = viridis(3)[3], alpha = 0.7) +
                   ggtitle("Actual vs Fitted \n at Bourke Street Mall (North) in 2017")
@@
 
-# Predictive model using imputed data
+## Predictive model using imputed data
 - can perform analysis at all locations now with complete data
-```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 10, fig.height = 3}
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.width = 10, fig.height = 2.5}
 melb_cent_mon9 <- dfa2 %>%
@@
                  size = 3) + 
-      scale_alpha_discrete(name = "Is public holiday") + 
+      scale_alpha_discrete(name = "Is public holiday",
+                           range = c(0.01,1)) + 
       ylab("Counts at Monday 9am")
@@
   * eg. weather data, public events
-
-### Objective: On the fly predictions of pedestrian traffic
+- can use clustering methods to investigate similarities between locations
+  * potentially use different models based on cluster membership
+  * eg. train stations, parks and recreation, shopping
+#### Objective: On the fly predictions of pedestrian traffic
 - requires efficient data storage
@@
 # Future steps for final paper
+- perform simulations to test imputation algorithm
 - develop Shiny app wih RStudio to provide on the fly predictions
@@
 
-```{r, cache = TRUE, fig.retina = 4, echo = FALSE, message = FALSE, fig.align= 'center', fig.width=5, fig.height=4, warning=F}
+```{r, cache = TRUE, fig.retina = 4, echo = FALSE, message = FALSE, fig.align= 'center', fig.width=5, fig.height=3.5, warning=F}
 mar13_16_8am <- ped_data[ped_data$Date == "2016-03-13", ] %>%
@@
 - explore predictors outside of counts data
+
+---
+# Summary
+- investigated issues with the Melbourne CBD pedestrian data
+- proposed algorithm for imputation of missing values in pedestrian data using split approach based on missing values proportion
+-
+
+---
+# Acknowledgements
+
+References:
+- City of Melbourne Open Data Portal - https://data.melbourne.vic.gov.au/
+- An Introduction to Statistical Learning - _G. James, D. Witten, T. Hastie and R. Tibshirani_
+-
+
+R packages used:
+  * `dplyr`, `lubridate` and `tidyr` for data manipulation
+  * `foreach` and `doSNOW` for parallel computing
+  * `rwalkr` for accessing latest data (Thanks Earo!)
+  * `ggplot2` and `ggmap` for visualisations
+  * `xaringan` for compliling this presentation
+
+GitHub repository for this project: https://github.com/gavinchin/honours2017

### 2017-10-01 presentation/final.Rmd
updated with di's formatting corrections

--- 2017-10-01 presentation/final.Rmd
@@
 author: "Gavin Chin, supervised by Di Cook"
-date: "2 October 2017"
+date: "3 October 2017"
 output:
@@
 ```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
-options(htmltools.dir.version = FALSE,
-        scipen = 100, cache = TRUE)
+knitr::opts_chunk$set(htmltools.dir.version = FALSE,
+        scipen = 100, 
+        cache = TRUE,
+        echo = FALSE,
+        warning = FALSE,
+        message = FALSE,
+        error = FALSE,
+        fig.align = 'center')
+```
+
+```{r}
 library(tidyverse)
@@
 ---
-class: middle
-
 ## How is the data collected?
 
-The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
-- These sensors are positioned under an awning or street pole, and count the number of people passing each hour
+The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD. These sensors are positioned under an awning or street pole, and count the number of people passing each hour.
 
@@
 - Alternatively, the data is also available as daily data from http://www.pedestrian.melbourne.vic.gov.au/
-  * The `rwalkr` R package by Earo Wang provides an API to access the data from both sources easily in R in a tidy format
+- The `rwalkr` R package by Earo Wang provides an API to access the data from both sources easily in R in a tidy format
 - Both sources of data provide the same counts data, but have different forms
 
-We use the daily data sourced from `pedestrian.melbourne.vic.gov.au` rather than the data available at the Open Data Platform as missing values are explicit
-- accessible using `rwalkr::walk_melb()`
+We use the daily data sourced from `pedestrian.melbourne.vic.gov.au` because missing values are explicit, accessible using `rwalkr::walk_melb()`
 
@@
 
-# Format of the data
-#### Variables available:
-- Time and date, sensor ID/location and hourly pedestrian count (volume)
+## Format of the data
+
+Time and date, sensor ID/location and hourly pedestrian count (volume)
 
@@
 ---
-class: center
-# The Problem: Prediction of pedestrian traffic
-```{r, echo=FALSE, fig.retina = 3, fig.height=5, fig.width= 10, warning=F, message = F}
+## The Problem: Prediction of pedestrian traffic
+```{r, echo=FALSE, fig.retina = 3, fig.height=4.5, fig.width= 10, warning=F, message = F}
 pd <- ggplot(dfa %>% filter(Date_Time < "2017-01-01 00:00:00",
@@
 - Different locations have different counts and patterns
-- We want to predict using 2014-2016 data to train our model
+- We want to build the prediction model using 2014-2016 data to train it
 
@@
 ---
-class: middle, center
-# Missing Data
-```{r, echo = FALSE, fig.retina=2, fig.width= 12, message=FALSE, warning=FALSE, cache = TRUE}
+## Missing Data
+
+```{r, echo = FALSE, fig.retina=2, fig.width= 12, fig.height=5, message=FALSE, warning=FALSE, cache = TRUE}
 miss_plot_dat <- ped_data %>% filter(Date_Time > "2014-06-01",
@@
 
-- Some small and large periods of missing data
+- Some small and large periods of missing data (grey indicates missing)
 - We need to impute these values before building a predictive model
@@
 ---
-class: middle
-# Imputation Algorithm
+## Imputation Algorithm
 
-#### Needed to develop an algorithm to:
-
-- **Step 1:** Find which sensors have a "large" proportion of missing values  
+- **Step 1:** Check proportion of missing values  
 - **Step 2:** At each sensor location, run appropriate GLM model with quasipoisson error distribution for imputation  
@@
 $$\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType}$$
-and a quaispoisson error distribution.
+and a quasipoisson error distribution.
 - Quasipoisson regression allows for $Var[\text{HourlyCount}] = \theta \mu$ (ie. overdispersion)
@@
 - Need to fill the missing data at the sensors with small proportions of missing values first
-  * The imputation models at sensors with large proportions of missing values are based on neighbouring sensors
-  * Requires complete data at neighbours to work properly
 
@@
 
-```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.height = 4, fig.align='center'}
+  * The imputation models at sensors with large proportions of missing values are based on neighbouring sensors
+  * Requires complete data at neighbours to work properly
+
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.height = 3.25, fig.align='center'}
 rownames(loc) <- loc$ids
@@
                          data = loc, size = 4,
-                         nudge_y = -0.0002) +
+                         nudge_y = -0.000325) +
               xlab(NULL) + ylab(NULL) +
@@
 ```
-- Need to use data from neighbouring sensors
-- Neighbour defined as geographical neighbours for simplicity
-  * Find two closest sensors by great-circle (haversine) distance
+- Need to use data from neighbouring sensors, defined as geographical neighbours for simplicity
+- Find two closest sensors by great-circle (haversine) distance
 
 ---
-class: center, middle
-## _Step 2c: Fit find neighbours sensors with "large" proportion of missing values_
+class: middle
+## Step 2c: Fit find neighbours sensors with "large" proportion of missing values
 
@@
 class: middle
-## _Step 2c: Fit models using neighbours_
+## Step 2c: Fit models using neighbours
 At each sensor location with "large" proportion of missing values, we estimate a generalised linear model with:
@@
 
-and a quaispoisson error distribution.
+and a quasipoisson error distribution.
 
@@
   * `foreach` and `doSNOW` for parallel computing
-  * `rwalkr` for accessing latest data
+
 ---
@@
 
-```{r}
+```{r, echo = FALSE}
 MARE_glm_tr <- numeric()
@@
 
-### Objective: On the fly predictions of pedestrian traffic
-- requires efficient data storage
+---
+### Objective: On-the-fly predictions of pedestrian traffic
+
+- On-the-fly predictions based on glm computed on complete (imputed) training data (2014-2016), for each location
+- Model objects are hundreds of megabytes
+- Requires efficient data storage
+- Modify the model object, to minimise the structure containing only what is necessary for prediction.
 
 ---
-# Future steps for final paper
-- develop Shiny app wih RStudio to provide on the fly predictions
+# Future steps for finishing
+
+- package into a Shiny app for easy on-the-fly predictions
   * make the predictive models more accessible
-  * provide visualisations to present the predictions
+  * provide visualisations of the predictions
 
@@
 
-- explore predictors outside of counts data
+- perform simulations to test imputation algorithm for robustness
+- explore predictions outside of counts data
+
+---
+## Summary of my contributions
+- Explored issues in the Melbourne CBD pedestrian data available from the City of Melbourne Open Data Portal
+- Algorithm for imputing missings
+- Efficient training model structures
+
+---
+## Acknowledgements
+- R and RStudio
+- Packages used for this work were: 
+  * `dplyr`, `lubridate` and `tidyr` for data manipulation
+  * `foreach` and `doSNOW` for parallel computing
+  * `rwalkr` for accessing latest data (Thanks Earo!)
+  * `ggplot2` and `ggmap` for visualisations
+  * `xaringan` for compliling this presentation
+- Code, presentation and thesis at [github.com:gavinchin/honours2017](github.com:gavinchin/honours2017)

### 2017-10-01 presentation/final.Rmd
added slide for questions

@@
   * `xaringan` for compliling this presentation
-- Code, presentation and thesis at [github.com:gavinchin/honours2017](github.com:gavinchin/honours2017)
+- Code, presentation and thesis at [github.com/gavinchin/honours2017](https://github.com/gavinchin/honours2017)
+
+I'd like to thank Di Cook and Earo Wang for helping me with the project, as well as my fellow honours classmates.
+
+---
+class: inverse, center
+# Questions or comments?
+
+--
+
+## Thanks for listening!

### 2017-10-02 presentation/final.Rmd
added mean abs. relative error histogram at end. rough transcript, WIP

@@
 threshold <- 0.1
-threshold_miss_hrs <- 17
+threshold_miss_hrs <- 6
 
@@
 
-```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.height = 3.25, fig.align='center'}
+```{r, echo = FALSE, fig.retina = 6, message=FALSE, warning=F, fig.height = 3.5, fig.height=3.5, fig.align='center'}
 rownames(loc) <- loc$ids
@@
                           size = 16, colour = viridis(3)[2]) +
-               geom_text(aes(label = ids, x = lons, y = lats),
-                         data = loc, size = 4,
+               geom_label(aes(label = ids, x = lons, y = lats,
+                              fill = cl),
+                         fontface = 'bold', colour = 'white',
+                         data = loc, size = 3.5,
                          nudge_y = -0.000325) +
+              scale_fill_viridis(discrete = TRUE,
+                                  name = "Sensor NA Prop.",
+                                  begin = 0.5, end = 0) +
               xlab(NULL) + ylab(NULL) +
@@
   * threshold value for missing values proportion (_default:_ `threshold = 0.1`)
-  * max length of repeated values (_default:_ `maxlength = 17`)
+  * max length of repeated values (_default:_ `maxlength = 6`)
 - use R and RStudio with packages:
@@
   * `rwalkr` for accessing latest data (Thanks Earo!)
-  * `ggplot2` and `ggmap` for visualisations
+  * `ggplot2`, `ggmap` and `viridis` for visualisations
   * `xaringan` for compliling this presentation
@@
 ## Thanks for listening!
+
+---
+## MARE of imputation
+```{r, fig.height=4, fig.retina = 3, fig.align='center'}
+MAE_tr <- numeric()
+for (i in unique(ped_data$Sensor)){
+    MAE_tr[i] <- mean(abs(fitted_data[[i]] - unlist(dfa[, i])),
+                           na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
+    }
+ggplot() + geom_histogram(aes(x = MAE_tr), binwidth = 0.05) +
+  geom_density(aes(x = MAE_tr),
+               adjust = 1.2, size = 2) +
+  labs(x = "Mean Absolute Relative Error",
+       y = "Number of Sensors",
+       title = "Histogram of MARE of algorithm") +
+  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
+```
+
+$$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \vert \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \vert}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$

### 2017-10-02 presentation/transcript_notes.Rmd
update pseudo transcript

---
title: "Untitled"
output: github_document
---

# Presentation Transcript

I'm Gavin and I've been working on modelling Melbourne CBD pedestrian data.

The City of Melbourne provide an open data portal, which makes council data publicly available.  
This project will focus on the pedestrian sensor data.

Currently, the City of Melbourne has 43 sensors installed around the Melbourne CBD. These sensors are positioned under an awning or street pole, and count the number of people pass through the counting zone each hour. This data is then sent to a server.

The public can access the pedestrian data a number of ways. A raw csv file can be downloaded at the official City of Melbourne open data portal website. The data available there is updated on a monthly basis.  
Alternatively, the City of Melbourne also hosts an interative visualisation of the pedestrian data, which provides csv files containing data for a single day.  
Using the rwalkr package by Earo, we can easily access the Melbourne pedestrian data to use for analysis. This package imports the data in tidy and easy to use format. 
For this project, we use the single day data provided by the visualisation site through the walk_melb function in R. We use this dataset instead of the one available at the Open Data Portal because missing values are explicit, as well as because it is updated daily.

The data we are working with provides three main variables. The time and date, sensor id or location, and the hourly pedestrian count. The data is provided in a long format. For most of the analysis, a wide data format has been used instead, providing a single observation for each hour, and the counts for the hour at each location being a column in the data.  
This reduces the number of rows in the data, and makes data manipulations, such as generating time and date based variables, a lot more efficient because we only need to do a single calculation for each hour of data rather 43 calculations for each sensor.

The objective of this project is to be able to predict how many people will pass through each sensor for a given time and date. We want to built the prediction model using 2014-2016 data to train it, leaving the current 2017 data as a test set. We have chosen not to use data prior to 2014 because of a very high proportion of missing data as many sensors were not installed before 2013.

Well, why would we want to predict pedestrian traffic? Being able to predict how many people pass through a certain location can be used by both the government sector as wel as the private sector. Examples of some uses may be infrastructure planning, or security planning (which has become quite important in recent times). Private uses of such information include marketing, for finding good locations to advertise, resource management for businesses, and investment planning.

The major issue we faced when trying to model the pedestrian data is the missing data. Even when we take a smaller period of the data, by only using data from 2014 onwards, we see that most sensors having missing data. In particular, we see that the periods of missing data can vary between sensors, and the lengths of missing data as well. Some of these missing values are caused by sensor malfunction or technical problems, while others are missing as the sensor has not been installed yet. Before we can built our predictive model, we need to impute these values.

Initially, when a simple approach of estimating a model at each sensor location with the same model specification or parameterisation, we found that it did not work very well. In particular, sensors which had a "large" proportion of missing values did not have enough data to build a good imputation model.  
So a different model for sensors with a "large" proportion of missing values was required.

This is the proposed algorithm for imputating the pedestrian data.
The first step is to check the proportion of missing values at each sensor for decide which type of model is used to impute missing values.  
Then, for each sensor location, we run the appropriate generalised linear model with quasipoisson errors to fit the data.  
If the proportion of missing values in the data is small, then we estimate a GLM model with hourly counts as a response variable and use deterministic time and date based variables as predictors.  
If the proportion of missing values is large, then we will estimate a GLM model with hourly counts as a response variable while using neighbouring sensor counts as predictors.  
Then using these models, we can replace missing values with the imputed values.

For the first step, we need to fix potential false zero-count values before we can calculate the proportion of missing values. This is where we have some issues in the data where some of the pedestrian counts of zero may actually be missing values. A way to check if a zero count could be a missing value is by checking the lengths of repeated values of zero. If a sequence of zero counts is longer than 6 hours, all the values in that sequence of repeated values are flagged as missing values.
Once the corrections to potential false zero-count values has been made, we can calculate the proportion of missing values at each sensor. Using this proportion, we classify a sensor to have either a "small" or "large" proportion of missing values based on a threshold value. We used 10% as a threshold proportion, where any sensor missing more than 10% of its data was considered to have a large proportion of missing values.

When classified, we can see the geographical location of sensors with small and large proportion of missing values. We see that many sensors with large proportions of missing values are in close proximity to sensors with small proportions of missing values.

So moving on to the next step in the algorithm, we fit models at all the sensors with a small proportion of missing values. We estimate a generalised linear model with quasipoisson errors where we include Month as a factor level with an additive effect, and two way interaction term between the time of the day and the day type. Time must be considered as a factor, not a numerical value, because hour of the day does not have a linear relationship with counts. Instead of using the day of the week, day type is used, where the weekdays Tuesday, Wednesday and Thursday are presumed to have the same daily pattern, which we call "midweek". The other levels the variable can take are Monday, Friday, Saturday, Sunday and Public Holiday. By collapsing the midweek days into a single factor level, we decrease the number of parameters which need to be estimated by 46.
We use a quasipoisson distribution because we have counts data. We use this instead of a standard poisson distribution because we do not want to restrict the variance of the hourly counts to be equal to the mean of the counts. A quasipoisson regression estimates an additional parameter, theta, to allow for overdispersion of the counts data.

Once we have estimated the model, we fill in the missing data with imputed values from the model. We need to perform this before proceeding to the next step in the algorithm.

We need to do this because the models at the sensors with large proportions of missing values use the data from neighbouring sensors. In order for this to work, we require complete data at the neighbours, otherwise we will not be able to impute values if the neighbour's count is not observed. Therefore, the set of possible neighbours only contains sensors with a small proportion of missing values.
For simplicity, we are using geographical neighbours as the neighbouring sensors. We define neighbours as the closest sensors by great circle distance. This is a measure of geographical distance.

Looking at the plot of counts over time between sensors, we can see that the counts at each hour move similarly. To perform pattern matching of pedestrian counts, we need to standardise counts at the neighbours.

When we do, we model the counts with a GLM quasipoisson model with an interaction between time and the scaled counts at the two neighbouring sensors. These scaled counts are effectively standard deviations at the neighbouring sensors, as we standardise the counts at each sensor to have mean = 0 and variance = 1.

After estimating the models at each of the sensors with large proportion of missing values, we can replace their missing values with values imputed by the models. To ensure it is possible to impute these values, we needed complete data at the neighbours. As a consequence, we need to make sure we had imputed the missing values at the neighbours before being able to estimate the models and impute values.

Implementing this imputation algorithm, we want to make the code as general as possible. This will ensure that the algorithm will work as the City of Melbourne adds more sensors around the CBD. Because we want to minimise the subjectivity if the models used for imputation, we want minimal user inputs to the algorithm. The resulting code only requires the data itself in the correct format, a threshold value for missing values proportion and the max length of repeated values. These two inputs can be viewed as tuning parameters to the algorithm, as it will affect which model is run at some sensors. We chose 10% missing value proportion as this seemed both reasonable, as well as giving sufficient number of sensors being considered to have a "small" proportion of missing values. A maximum length of 6 for repeated values is also used as we would not expect more than 6 hours of zero pedestrian activity.  
We implement this algorithm using R and RStudio with packages: dplyr, lubridate and tidyr for data manipulation, and foreach and doSNOW for parallel computing. Parallel computing here significantly improves the performance of the algorithm in R, utilising more of the computer's capabilities and increasing efficiency.

Here we have some preliminary results for evaluation of the imputation method. Here we have actual values plotted against fitted values at Southern Cross Station, a sensor location with a small proportion of missing values, and Southbank, a sensor location with a large proportion of missing values. The yellow line is a linear fit line, while the dashed purple line indicates a perfect fit, where predicted values are equal to actual values.

### 2017-10-02 presentation/transcript_notes.Rmd
updated

@@
 
-This is the proposed algorithm for imputating the pedestrian data.
+This is the proposed algorithm for imputing the pedestrian data.
 The first step is to check the proportion of missing values at each sensor for decide which type of model is used to impute missing values.  
@@
 The yellow line is a linear fit line, while the dashed purple line indicates a perfect fit, where predicted values are equal to actual values.
+As these are in-sample fits - the data on which the models are trained on, we expected quite good accuracy.
+
+This is an example of the out-of-sample fit, where we plot the 2017 actual counts against the fitted counts.
+We observe that the the imputation is underestimating the counts, where the slope of the fitted linear curve is less than 1.
+When we look a small window of the time series, we can also see the underestimation.
+But overall, we see the pattern tends to match the actual pattern.
+
+After we have a complete dataset at all the sensor locations, we can perform more meaningful analysis on the data.
+Foe example, we can see what happens when we treat a specific hour of the week as a time series.
+When we build the predictive model, we are looking to include predictors which will explain these deviations from the mean.
+Looking at the distribution of the counts at Melbourne Central on Mondays at 9 am from a kernel desity estimation, we see that it is reasonably close to being normally distributed.
+We look to explore covariates outside the counts data to help predict the counts, such as looking at weather data, or events data.
+
+The objective of building the predictive model is to be able to provide on-the-fly predictions based on a glm computed on complete training data for each location.
+Because of the size of the data, with 3 years of hourly data at 43 sensor locations, the estimated model objects given by the base R GLM estimation function are very large.
+A complete model which contains a

### 2017-10-02 presentation/transcript_notes.Rmd
fixed typo in slide 24, updated pseudo transcript

@@
 Because of the size of the data, with 3 years of hourly data at 43 sensor locations, the estimated model objects given by the base R GLM estimation function are very large.
-A complete model which contains a
+A complete model which contains all the GLMs at each sensor adds up to a 1.7gb object in R.
+This is unnecessarily large, as the bare minimum we need to store is the coefficients and the model formula and some other small details to retain the predict function in R.
+This brings down the object size to under 20mb.
+
+Future steps which will be taken in the next couple weeks in order to finish this project is to package the predictive model into a Shiny app.
+This will provide an accessible interface which can be used by end users to make predictions about the pedestrian activity around the city at any particular time.
+This will also allow for inputs for covariates, such as weather.
+Within the app, we want to provide visualisations which will help to show which locations are expected to be busy or quiet.
+Other steps which need to be taken include diagnostics on the imputation algorithm. This involves testing the robustness of the method by simulating different levels of missingness.

### 2017-10-02 presentation/transcript_notes.Rmd
transcript done.

@@
 Other steps which need to be taken include diagnostics on the imputation algorithm. This involves testing the robustness of the method by simulating different levels of missingness.
+
+So a brief summary of the work I've done for my honours research project:
+I've explored issues in the Melbourne CBD pedestrian data provided by the City of Melbourne, proposed an algorithm for imputing missings in the data and investigated efficient training model structures within R in this application.

### 2017-10-03 presentation/final.Rmd
fixed typo in slide 24, updated pseudo transcript
added slides for question prep

@@
 - perform simulations to test imputation algorithm for robustness
-- explore predictions outside of counts data
+- explore predictors outside of counts data
 
@@
 $$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \vert \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \vert}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$
+
+---
+## Proportion of missing values prior to correction:
+
+```{r, fig.retina = 3, fig.align= 'center'}
+na_prop_bad <- numeric()
+
+for (i in unique(ped_data$Sensor)){
+  na_prop_bad[i] <- sum(is.na(original_df[, i])) / n_obs
+}
+
+ggplot() + geom_histogram(aes(x = na_prop_bad*100), bins = 12) +
+          xlab("Percentage of Missing Values (%)") +
+          ylab("Number of Sensors") + theme_minimal() +
+          geom_vline(xintercept = 10, size = 2,
+                     colour = 'red') +
+          geom_text(aes(label = "10% Threshold", x = 10, y = 1),
+                    colour = "white") +
+          ggtitle("Histogram of Proportions of Missing Values \n during 2014 - 2016 prior to correction") +
+          theme(plot.title = element_text(hjust = 0.5))
+```

### 2017-10-03 presentation/final.Rmd
update

@@
 ```
+
+---
+## Specification of imputation models:
+Model at sensors with small proportion of missing values
+$$E[HourlyCounts_{sensor, t}|.] = \exp \{ \beta_{0, sensor} + \sum_{j = 1}^{11}(\beta_{1,j,sensor}Month^j_{t}) + \sum_{k = 1}^{23}(\beta_{2,k,sensor}Time_{t})$$
+
+$$+\sum_{l = 1}^{5}(\beta_{3,l,sensor}DayType_t) + \sum_{k = 1}^{23} \sum_{l = 1}^{5} (\beta_{4,k,l,sensor}Time_t DayType_t) \} \qquad \forall \mathtt{NA}_{prop}(sensor) < \mathtt{threshold}$$

### 2017-10-12 thesis/draft.Rmd
pdf for draft
working through imputation evaluation

@@
 library(RColorBrewer)
+library(gridExtra)
+library(viridis)
+library(ggmap)
 ```
@@
   }
+
+ ## ped_data <- read_csv("http://209.148.91.227/files/ped_df_28sep17.csv")
 
@@
 
+ori_2 <- dfa
+
 run_lengths %>% lapply(., `[[`, 1) %>%
@@
 pander(summary(sequences[sequences > 1]))
-
-# ggplot() + geom_boxplot(aes(y = sequences, x = "A"), alpha = 0.2) +
-#     xlab(NULL) + ylab("Length of same counts value") +
-#     theme(axis.text.x = element_blank(),
-#           axis.ticks.x = element_blank(),
-#           axis.title.x = element_blank()) +
-#     theme_minimal()
 ```
+Table: Summary statistics of the sequence lengths of repeated values
 
@@
 For example, we may want to classify any sequence of `Hourly_Counts = 0` running for longer than 24 hours as being considered `NA`.
-From the summary of lengths of repeated values (which are repeated at least once), we see that the distribution of the lengths are very positively skewed.
-With a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data.
-Classifying any sequence of repeated values, typically 0, which has a length greater than 24 as `NA` should allow true zero counts to avoid being misclassified as `NA`.
+From the summary of lengths of repeated values (which are repeated at least once), we see that the distribution of the lengths are very positively skewed with a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data.
+Classifying any sequence of repeated values, typically 0, which has a length greater than 6 as `NA` should allow true zero counts to avoid being misclassified as `NA`.
+We select the length of 6 hours as it would be unreasonable to assume any sensor location to have the exact same number of pedestrian counts over such a long period.
 
-After we have replaced questionable zero values with `NA`, we can calculate the proportion of missing values for each sensor location.
-From these calculated proportion, we can classify sensors as either having a "large" or "small" proportion of missing values, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.05`.
+After we have replaced questionable zero values with the value `NA`, we can calculate the proportion of missing values for each sensor location.
+From these calculated proportion, we can classify sensors as either having a "large" or "small" proportion of missing values, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.1`.
 
-```{r}
+```{r na_prop_compute}
 na_prop <- numeric()
@@
 
+na_prop_pre <- numeric()
+
+for (i in unique(ped_data$Sensor)){
+  na_prop_pre[i] <- sum(is.na(ori_2[, i])) / n_obs
+}
+
 large_miss_sensors <- names(na_prop[na_prop > threshold])
 small_miss_sensors <- setdiff(names(na_prop), large_miss_sensors)
+```
+
+```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Comparison of distribution of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data"}
+theme_g <- theme_minimal() +
+           theme(plot.title = element_text(hjust = 0.5,
+                                           size = 10))
+p1 <- ggplot() + geom_histogram(aes(x = na_prop*100), bins = 12) +
+          xlab("Percentage of Missing Values (%)") +
+          ylab("Number of Sensors") + theme_g +
+          geom_vline(xintercept = 10, size = 2,
+                     colour = 'red') +
+          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
+                    colour = "white", fill = 'black', alpha = 0.4) +
+          scale_y_continuous(limits = c(0,35)) +
+          ggtitle("Histogram of Proportions of Missing Values \n during 2014 - 2016")
+p2 <- ggplot() + geom_histogram(aes(x = na_prop_pre*100), bins = 12) +
+          xlab("Percentage of Missing Values (%)") +
+          ylab("Number of Sensors") + theme_g +
+          geom_vline(xintercept = 10, size = 2,
+                     colour = 'red') +
+          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
+                    colour = "white", fill = 'black', alpha = 0.4) +
+          scale_y_continuous(limits = c(0,35)) +
+          ggtitle("Histogram of Proportions of Missing Values \n during 2014 - 2016 prior to false zero counts fix")
+grid.arrange(p2, p1, nrow = 1)
 ```
@@
 
-For sensors with a small proportion of missing values, we use a GLM quasi-poisson model again. To improve on the previous model used, we use the model:
+For sensors with a small proportion of missing values, we use a GLM quasipoisson model. To improve on the previous model used, we use the model:
 $$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
@@
 
+```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
+ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
+lats <- ped_loc$Latitude
+lons <- ped_loc$Longitude
+ids <- ped_loc$`Sensor Description`
+
+{
+ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
+ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
+ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
+ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
+ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
+ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
+ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
+ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
+ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
+}
+loc <- data.frame(lons, lats, ids)
+rownames(loc) <- ids
+loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
+                                  "Small", "Large"))
+rownames(loc) <- loc$ids
+
+melb3 <- get_map(location = c(loc["Southbank",]$lons,
+                              loc["Southbank",]$lats),
+                 zoom = 17, maptype = "roadmap")
+
+sens_lines_lon <- as.numeric(c(loc["Flinders St Station Underpass",]$lons,
+                loc["Southbank",]$lons,
+                loc["Sandridge Bridge",]$lons))
+
+sens_lines_lat <- as.numeric(c(loc["Flinders St Station Underpass",]$lats,
+                loc["Southbank",]$lats,
+                loc["Sandridge Bridge",]$lats))
+
+sens_lines <- data.frame(lon = sens_lines_lon,
+                         lat = sens_lines_lat)
+options(na.action = 'na.omit')
+ggmap(melb3) + geom_line(aes(x = lon, y = lat),
+                        data = sens_lines, size = 2.5,
+                        colour = viridis(3)[2], alpha = 0.5) +
+               geom_point(aes(x = lons, y = lats, colour = cl),
+                          size = 8,
+                          data = loc) +
+               scale_colour_viridis(discrete = TRUE,
+                                    name = "Sensor NA Prop.",
+                                    begin = 0.5, end = 0) +
+               geom_point(aes(x = loc["Southbank",]$lons,
+                              y = loc["Southbank",]$lats),
+                          size = 16, colour = viridis(3)[2]) +
+               geom_label(aes(label = ids, x = lons, y = lats,
+                              fill = cl),
+                         fontface = 'bold', colour = 'white',
+                         data = loc, size = 3.5,
+                         nudge_y = -0.000325) +
+              scale_fill_viridis(discrete = TRUE,
+                                  name = "Sensor NA Prop.",
+                                  begin = 0.5, end = 0) +
+              xlab(NULL) + ylab(NULL) +
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.ticks = element_blank(),
+                    legend.position = c(.8,0.2))
+```
+
 For each of the sensor locations with high proportions of missings, we need to decide which sensor to use for prediction.
@@
 }
-
-ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
-lats <- ped_loc$Latitude
-lons <- ped_loc$Longitude
-ids <- ped_loc$`Sensor Description`
 
@@
 Using the two geographically closest sensors with complete cases, we use another GLM specified as:
-$$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{closest sensors}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{closest sensors}}}$$
-$$\text{where } \quad  \mu_{\text{Time}, \text{HourlyCounts}_{\text{closest sensors}}} \sim \text{Time} \times \text{HourlyCounts}_{\text{closest sensor 1}} + \text{Time} \times \text{HourlyCounts}_{\text{closest sensor 2}}$$
+$$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{neighbours}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}}$$
+$$\text{where } \quad  \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}} \sim \text{Time} \times \text{HourlyCounts}_{\text{neighbour 1}}^{SC} + \text{Time} \times \text{HourlyCounts}_{\text{neighbour 2}}^{SC}$$
 
@@
 
+```{r, echo = FALSE, message=FALSE, warning=F, include=FALSE}
+i <- "Southbank"
+  sensor_dists <- numeric()
+  for (j in small_miss_sensors){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  dfa3 <- original_df %>% filter(Date_Time > "2017-01-01")
+  close_df <- dfa3[, closest_sensor[1]] %>% unlist
+  close_df2 <- dfa3[, closest_sensor[2]] %>% unlist
+
+  close_dfsc <- scale(close_df)
+  close_dfsc2 <- scale(close_df2)
+
+  sensor_df <- dfa3[, i] %>% unlist
+  sensor_dfsc <- scale(sensor_df)
+
+  tr_dat <- data.frame(sensor_df, close_df, close_df2,
+                       Time = as.integer(dfa3$Time),
+                       Date_Time = dfa3$Date_Time,
+                       HDay = dfa3$HDay) %>%
+            gather(key = Sensor, value = Count, sensor_df:close_df2)
+
+  tr_dat_sc <- data.frame(sensor_dfsc, close_dfsc, close_dfsc2,
+                       Time = as.integer(dfa3$Time),
+                       Date_Time = dfa3$Date_Time,
+                       HDay = dfa3$HDay) %>%
+            gather(key = Sensor, value = Count, sensor_dfsc:close_dfsc2)
+
+p1 <- ggplot(tr_dat %>% filter(Date_Time < "2017-03-02",
+                         Date_Time > "2017-02-26")) +
+      geom_line(aes(x = Date_Time, y = Count,
+                    colour = Sensor), size = 2) +
+      scale_colour_viridis(discrete = TRUE,
+                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
+                                      "Neighbour 2 (Sandridge Bridge)",
+                                      "Target (Southbank)")) +
+      xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
+      theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
+
+p2 <- ggplot(tr_dat_sc %>% filter(Date_Time < "2017-03-02",
+                         Date_Time > "2017-02-26")) +
+      geom_line(aes(x = Date_Time, y = Count,
+                    colour = Sensor), size = 2) +
+      scale_colour_viridis(discrete = TRUE,
+                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
+                                      "Neighbour 2 (Sandridge Bridge)",
+                                      "Target (Southbank)")) +
+      xlab("Time") + ggtitle("Scaled/Stadardised Neighbours") +
+      ylab("Standard Deviations") + theme_minimal() +
+      theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
+```
+
+```{r, echo = FALSE, message=FALSE, warning=F, fig.width=10, fig.height=5}
+grid.arrange(p1, p2, ncol = 1)
+```
+
 In order to use a mixture of the patterns from two different sensors, we need to use the scaled counts at the neighbouring sensors to avoid the magnitude of the counts having an effect on the predicted counts.
 We scale the counts by standardising to $\mu = 0, \sigma^2 = 1$.
-Because the covariance with neighbouring sensors may differ over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
+Because the covariance with neighbouring sensors may vary over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
 
@@
 
-Note, this is also before the adjustments made to treat suspiciously long sequences of 0 values as `NA`.
+Note, this is also after the adjustments made to treat suspiciously long sequences of 0 values as `NA` in the actual counts.
+The simple GLM model is trained before this correction, while the improved model is trained on the adjusted data.
 
@@
 
-```{r}
-ggplot(data = NULL, aes(x = original_df %>% 
+```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
+p1 <- ggplot(data = NULL, aes(x = original_df %>% 
              filter(Date_Time < "2017-01-01") %>%
@@
            geom_abline(aes(slope = 1, intercept = 0),
-                       colour = brewer.pal(3, 'Dark2')[1],
-                       alpha = 0.7, size = 1.5,) +
-           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
-                       colour = brewer.pal(3, 'Dark2')[2]) +
-           theme_minimal() + theme(aspect.ratio = 1) +
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + 
+           theme(aspect.ratio = 1,
+                 plot.title = element_text(hjust = 0.5)) +
            scale_x_continuous(limits = c(0, 4000)) +
@@
            ylab("Predicted Pedestrian Count") +
-           ggtitle("Actual against Predicted Counts at Southern Cross Station")
-
+           ggtitle("Simple GLM Model: Actual v. Predicted \n at Southern Cross Station")
+p2 <- ggplot(data = NULL, aes(x = original_df %>% 
+             filter(Date_Time < "2017-01-01") %>%
+             .[, "Southern Cross Station"],
+           y = fitted_data$`Southern Cross Station`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + 
+           theme(aspect.ratio = 1,
+                 plot.title = element_text(hjust = 0.5)) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Improved GLM Model: Actual v. Predicted \n at Southern Cross Station")
+grid.arrange(p1, p2, nrow = 1)
 ```
-
-\begin{center}
-\textbf{Figure }:  A plot of the Actual vs Predicted pedestrian counts at Southern Cross Station.
-The green line is $x=y$, representing a reference for perfect fit, while orange line is the linear fit.
-\end{center}
 
@@
 options(na.action = na.omit)
-ggplot(data = NULL, aes(x = original_df[-(is.na(original_df$`Australia on Collins`)), ] %>% 
+p1 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
              filter(Date_Time < "2017-01-01") %>%
-             .[, "Australia on Collins"],
-           y = fitted_base$`Australia on Collins`[-(is.na(original_df$`Australia on Collins`))])) +
-           geom_point(alpha = 0.05) +
+             .[, "Australia on Collins"])) +
+           geom_point(aes(y = fitted_base$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
+                      alpha = 0.05) +
            geom_abline(aes(slope = 1, intercept = 0),
-                       colour = brewer.pal(3, 'Dark2')[1],
-                       alpha = 0.7, size = 1.5) +
-           geom_smooth(method = "lm", size = 1.5, alpha = 0.7, 
-                       colour = brewer.pal(3, 'Dark2')[2]) +
-           theme_minimal() + theme(aspect.ratio = 1) +
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(aes(y = fitted_base$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
+                       method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + theme(aspect.ratio = 1,
+                                   plot.title = element_text(hjust = 0.5)) +
            scale_x_continuous(limits = c(0, 4000)) +
@@
            ylab("Predicted Pedestrian Count") +
-           ggtitle("Actual against Predicted Counts at Australia on Collins")
+           ggtitle("Simple GLM Model: Actual v. Predicted \n at Australia on Collins")
+p2 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
+             filter(Date_Time < "2017-01-01") %>%
+             .[, "Australia on Collins"])) +
+           geom_point(aes(y = fitted_data$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
+                      alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(aes(y = fitted_data$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
+                       method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + theme(aspect.ratio = 1,
+                                   plot.title = element_text(hjust = 0.5)) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Improved GLM Model: Actual v. Predicted \n at Australia on Collins")
+```
+
+```{r, fig.height=4, fig.width=9, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}
+
+grid.arrange(p1, p2, nrow = 1)
 
 ```
-We can see that the model is heavily biased by the false 0 values, causing the predicted values to be underestimate (slope of the fitted line is $<1$).
-This demonstrates the importance of ensuring outliers are removed from the data.
+
+We can see that the simple model has bias caused by the false zero counts, resulting in the predicted values to be underestimated.
+This is shown by the slope of the fitted line on the actual counts against predicted counts being $<1$, emphasising the importance of the first step taken in the algorithm of the improved model to class the false zero counts as missing values.
 
@@
 }
-pander(data.frame(Tr_MARE = MARE_base_glm_tr))
+
+MARE_tr <- numeric()
+for (i in unique(ped_data$Sensor)){
+  MARE_tr[i] <- mean(abs(fitted_data[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
+}
+
+mare_df <- data.frame(simple = MARE_base_glm_tr, improved = MARE_tr, ids) %>%
+           gather(model, mare, simple:improved)
+p1 <- ggplot(mare_df) + geom_boxplot(aes(x = model, y = mare, fill = model),
+                                     width = .5, alpha = 0.5) +
+      labs(x = "Model", y = "In-Sample MARE") +
+      scale_fill_viridis(discrete = TRUE, name = "Model")
+
+p2 <- ggplot(mare_df) + geom_density(aes(x = mare, fill = model,
+                                         colour = model), adjust = 1.3,
+                                     alpha = 0.5) +
+      scale_fill_viridis(discrete = TRUE) +
+      scale_colour_viridis(discrete = TRUE) +
+      theme_minimal() + theme(plot.title = element_text(hjust = 0.5),
+                              legend.position = "none") +
+      labs(x = "In-sample MARE")
+p3 <- ggplot() + geom_point(aes(x = MARE_base_glm_tr, y = MARE_tr,
+                                colour = (MARE_tr<MARE_base_glm_tr))) +
+      geom_abline(aes(slope = 1, intercept = 0)) +
+      scale_colour_viridis(discrete = TRUE, name = "Improved Model is better") +
+      theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
+      theme(aspect.ratio = 1, legend.position = "bottom",
+            legend.direction = "vertical")
+```
+
+```{r, fig.height=3.5, fig.width=9, fig.cap="Distribution of in-sample (2014-2016 data) MARE by imputation model"}
+
+grid.arrange(p1, p2, p3, nrow = 1)
+
 ```
@@
 
-```{r}
+```{r, fig.cap="A plot of the pedestrian counts at Southern Cross Station. The black line is the observed counts, where the red line is the imputed counts."}
 fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
@@
 ```
-\begin{center}
-\textbf{Figure }:  A plot of the pedestrian counts at Southern Cross Station. The black line is the observed counts, where the red line is the imputed counts.  
-\end{center}

### 2017-10-12 thesis/draft.Rmd
update.

@@
 ori_2 <- filter(original_df, Date_Time < "2017-01-01")
+ori_ts <- filter(original_df, Date_Time > "2016-12-31")
+ori_ts2 <- ori_ts
 ```
@@
 
-```{r, echo = FALSE, message=FALSE, warning=F, fig.width=10, fig.height=5}
+```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 8, fig.height = 4}
 grid.arrange(p1, p2, ncol = 1)
@@
       geom_abline(aes(slope = 1, intercept = 0)) +
-      scale_colour_viridis(discrete = TRUE, name = "Improved Model is better") +
+      scale_colour_viridis(discrete = TRUE, option = "C",
+                           begin = 0.15, end = 0.85,
+                           direction = -1,
+                           name = "Improved Model is better") +
       theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
@@
 
-```{r, fig.height=3.5, fig.width=9, fig.cap="Distribution of in-sample (2014-2016 data) MARE by imputation model"}
+```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}
 
@@
 
-Focusing on the date 04/07/2014, which has 3 hours of missing data:
+While it is not very clear when comparing the distribution of in-sample MARE between imputation models, when we compare the in-sample MARE at each location we can see that the majority of locations had better in-sample fit with the improved model using the algorithm.
+We do see that some locations did have worse in-sample fit, but overall we see improvement with `r sum(MARE_base_glm_tr > MARE_tr)` having better fit with the improved model over the simple model.
 
-```{r, fig.cap="A plot of the pedestrian counts at Southern Cross Station. The black line is the observed counts, where the red line is the imputed counts."}
+Focusing on the date 04/07/2014, which has 3 hours of missing data at Southern Cross Station:
+
+```{r, fig.height = 2, fig.width = 5, fig.align='center', fig.cap="A plot of the pedestrian counts at Southern Cross Station.
+The black line is the observed counts, where the yellow line is the imputed counts."}
 fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
@@
                                Date_Time > "2014-07-04 12:00:00")) +
-  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = 'red',
+  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = viridis(3)[3],
             size = 2) +
@@
   theme_minimal() + xlab("Time") +
-  ylab("Pedestrian Counts at Southern Cross Station") +
-  ggtitle("Pedestrian Counts at Southern Cross Station on 04/07/2014 - 05/07/2014")
+  ylab("Pedestrian Counts") +
+  ggtitle("Pedestrian Counts at Southern Cross Station \n on 04/07/2014 - 05/07/2014") +
+  theme(plot.title = element_text(hjust = 0.5)) 
 ```
+
+More importantly, we can evaluate the performance of the model on the out of sample observations (2017 data).
+Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, we need to again apply the correction by long sequences of repeated values to be able to properly evaluate the goodness of fit of out-of-sample observations.
+
+```{r ts_fix0}
+for(i in unique(ped_data$Sensor)){
+    ori_ts[is.na(ori_ts[, i]), i] <- 0
+    na_length_check <- rle(as.numeric(unlist(ori_ts[,i])))
+    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
+    rl <- ori_ts[, i] %>% unlist %>% rle()
+    ori_ts[maybeNA, i] <- NA
+    defsNA <- is.na(ori_ts2[, i])
+    ori_ts[defsNA, i] <- NA
+  }
+```
+
+```{r test_fit}
+test_base <- list()
+test_fit <- list()
+ts <- original_df %>% filter(Date_Time > "2016-12-31")
+
+ for(i in unique(ped_data$Sensor)){
+    test_base[[i]] <- predict.glm(models_base[[i]], newdata = ts,
+                                  type = "response")
+ }
+
+ for(i in small_miss_sensors){
+   test_fit[[i]]  <- predict.glm(models[[i]], newdata = ts,
+                                 type = "response")
+ }
+ for(i in large_miss_sensors){
+   sensor_dists <- numeric()
+  for (j in small_miss_sensors){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  close_df_scale <- ts[, closest_sensor[1]] %>% unlist %>% scale
+  close_df_scale2 <- ts[, closest_sensor[2]] %>% unlist %>% scale
+
+  ts_sc <- data.frame(close_df_scale, close_df_scale2, Time = ts$Time)
+
+  test_fit[[i]] <- predict.glm(models[[i]], newdata = ts_sc,
+                               type = "response")
+
+ }
+
+MARE_base_ts <- numeric()
+for (i in unique(ped_data$Sensor)){
+  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts[, i])), na.rm = T) / mean(unlist(ori_ts[, i]), na.rm = T)
+}
+
+MARE_ts <- numeric()
+for (i in unique(ped_data$Sensor)){
+  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts[, i])), na.rm = T) / mean(unlist(ori_ts[, i]), na.rm = T)
+}
+```
+
+Again, we start by comparing the fit at a sensor location with a small proportion of missing values, Vic Arts Centre:
+```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
+p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
+           y = test_base$`Melbourne Central`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + 
+           theme(aspect.ratio = 1,
+                 plot.title = element_text(hjust = 0.5)) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Simple GLM Model: Actual v. Predicted \n at Melbourne Central")
+p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
+           y = test_fit$`Melbourne Central`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + 
+           theme(aspect.ratio = 1,
+                 plot.title = element_text(hjust = 0.5)) +
+           scale_x_continuous(limits = c(0, 4000)) +
+           scale_y_continuous(limits = c(0, 4000)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
+grid.arrange(p1, p2, nrow = 1)
+```
+
+We find that the out-of-sample predictions of at Melbourne Central by the simple model has a tendency to underestimate the pedestrian counts.
+By comparison, the predictions made by the improved imputation model does not underestimate the counts as much.
+Comparing the out-of-sample MARE between models at Melbourne Central, MARE decreases from `r round(MARE_base_ts["Melbourne Central"]*100, 2)`% to `r round(MARE_ts["Melbourne Central"]*100, 2)`% with the use of the improved model.
+
+```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
+p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Collins Place (South)"],
+           y = test_base$`Collins Place (South)`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + 
+           theme(aspect.ratio = 1,
+                 plot.title = element_text(hjust = 0.5)) +
+           scale_x_continuous(limits = c(0, 2500)) +
+           scale_y_continuous(limits = c(0, 2500)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Simple GLM Model: Actual v. Predicted \n at Collins Place (South)")
+p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Collins Place (South)"],
+           y = test_fit$`Collins Place (South)`)) +
+           geom_point(alpha = 0.05) +
+           geom_abline(aes(slope = 1, intercept = 0),
+                       colour = viridis(3)[1],
+                       alpha = 0.7, size = 1.25,
+                       linetype = 2) +
+           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
+                       colour = viridis(3)[3]) +
+           theme_minimal() + 
+           theme(aspect.ratio = 1,
+                 plot.title = element_text(hjust = 0.5)) +
+           scale_x_continuous(limits = c(0, 2500)) +
+           scale_y_continuous(limits = c(0, 2500)) +
+           xlab("Actual Pedestrian Count") +
+           ylab("Predicted Pedestrian Count") +
+           ggtitle("Improved GLM Model: Actual v. Predicted \n at Collins Place (South)")
+grid.arrange(p1, p2, nrow = 1)
+```
+
+At a sensor location with a large proportion of missing values, Collins Place (South), we again see a tendency to underestimate counts with the simple imputation model.
+The improved imputation method reduces the underestimation of the predictions and provides better predictions.
+The MARE is greatly improved, where it decreases from `r round(MARE_base_ts["Collins Place (South)"]*100, 2)`% to `r round(MARE_ts["Collins Place (South)"]*100, 2)`% with the use of the improved model.

### 2017-10-13 thesis/draft.Rmd
imputation robustness testing

@@
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
-                      message = FALSE, warning = FALSE)
+                      message = FALSE, warning = FALSE,
+                      fig.pos = 'h')
 library(tidyverse)
@@
 library(ggmap)
+set.seed(12345)
 ```
@@
 
-More importantly, we can evaluate the performance of the model on the out of sample observations (2017 data).
+More importantly, we can evaluate the performance of the model on the out of sample observations (2017 data) as cross validation.
 Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, we need to again apply the correction by long sequences of repeated values to be able to properly evaluate the goodness of fit of out-of-sample observations.
+The method used for this is identical to that used on the training set in the algorithm, replacing sequences of length 6 hours or longer with `NA`.
 
@@
 The MARE is greatly improved, where it decreases from `r round(MARE_base_ts["Collins Place (South)"]*100, 2)`% to `r round(MARE_ts["Collins Place (South)"]*100, 2)`% with the use of the improved model.
+
+```{r}
+MARE_base_ts2 <- MARE_base_ts[setdiff(unique(ped_data$Sensor),
+                                      c("Spencer St-Collins St (South)",
+                                        "Spencer St-Collins St (North)"))]
+MARE_ts2 <- MARE_ts[setdiff(unique(ped_data$Sensor),
+                                      c("Spencer St-Collins St (South)",
+                                        "Spencer St-Collins St (North)"))]
+
+mare_df <- data.frame(simple = MARE_base_ts2, improved = MARE_ts2) %>%
+           gather(model, mare, simple:improved)
+
+p1 <- ggplot(mare_df) + geom_boxplot(aes(x = model, y = mare, fill = model),
+                                     width = .5, alpha = 0.5) +
+      labs(x = "Model", y = "Out-of-sample MARE") +
+      scale_fill_viridis(discrete = TRUE, name = "Model")
+
+p2 <- ggplot(mare_df) + geom_density(aes(x = mare, fill = model,
+                                         colour = model), adjust = 1.3,
+                                     alpha = 0.5) +
+      scale_fill_viridis(discrete = TRUE) +
+      scale_colour_viridis(discrete = TRUE) +
+      theme_minimal() + theme(plot.title = element_text(hjust = 0.5),
+                              legend.position = "none") +
+      labs(x = "Out-of-sample MARE")
+p3 <- ggplot() + geom_point(aes(x = MARE_base_ts2, y = MARE_ts2,
+                                colour = (MARE_ts2<MARE_base_ts2))) +
+      geom_abline(aes(slope = 1, intercept = 0)) +
+      scale_colour_viridis(discrete = TRUE, option = "C",
+                           begin = 0.15, end = 0.85,
+                           direction = -1,
+                           name = "Improved Model is better") +
+      scale_x_continuous(limits = c(0,1)) +
+      scale_y_continuous(limits = c(0,1)) +
+      theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
+      theme(aspect.ratio = 1, legend.position = "bottom",
+            legend.direction = "vertical")
+```
+
+```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) sensors")}
+
+grid.arrange(p1, p2, p3, nrow = 1)
+
+```
+
+Spencer St-Collins St (North)/(South) was omitted from the the plots above as the out-of-sample MARE is extremely high at `r round(MARE_ts["Spencer St-Collins St (North)"]*100, 2)`% and `r round(MARE_ts["Spencer St-Collins St (South)"]*100, 2)`% respectively.
+
+As these sensor locations were both classified as locations with a large proportion of missing values, it would appear that the relationship between these locations and their neighbouring sensors has changed.
+This is because the in-smaple MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`% and `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% respectively.
+
+```{r, fig.width = 8, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
+ggplot() + geom_smooth(aes(x = as.integer(Time),
+                           y = `Spencer St-Collins St (South)`),
+                       colour = viridis(3)[1],
+                       data = dfa2) +
+           geom_smooth(aes(x = as.integer(Time),
+                           y = `Spencer St-Collins St (South)`),
+                       colour = viridis(3)[3],
+                       data = ori_ts2) +
+           theme_minimal() + labs(x = "Time of Day") +
+           facet_wrap(~ DayType)
+```
+
+The plot of the average counts by time of day and type of day and data set shows that this location saw an increase in counts on weekdays which did not occur at the neghbouring sensors.
+We also see that the pattern is slightly different during weekdays, with a higher afternoon/evening peak.
+These types of change in pedestrian traffic behaviour is hard to capture in our imputation model, especially when its neighbours did not have a similar change.
+While we can add a year variable to allow for year to year growth, the short period of training data doesn't make this appropriate at this stage.
+
+Another way we can perform cross-validation is to remove data at a sensor, and evaulate the predictions made.
+This will allow us to evaluate the imputation model's performance within sample, which is the primary concern with the imputation process.
+We select a sensor which has a small proportion of missing values, Bourke Street Mall (South), which has `r round((1 - na_prop["Bourke Street Mall (South)"])*100, 3)`% of data available.
+As observed in the data, missing data is present in different ways: random, short periods of missing data (no more that 6 hours) or long periods of missing data (multiple months).
+We will simulate both types of missingness at this sensor location.
+
+First, we remove randomly remove data such that 20% of the data is missing, then estimate the imputation model. Using this model, we calculate the MARE within the training data.
+```{r}
+miss_idx <- sample(1:nrow(dfa2), round(nrow(dfa2)*0.2))
+
+mall_missing_1 <- dfa2$`Bourke Street Mall (South)`
+mall_missing_1[miss_idx] <- NA
+
+i <- "Bourke Street Mall (South)"
+sensor_dists <- numeric()
+  for (j in small_miss_sensors[-6]){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors[-6]
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  close_df <- dfa2[, closest_sensor[1]] %>% unlist
+  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
+
+  close_df_scale <- scale(close_df)
+  close_df_scale2 <- scale(close_df2)
+
+  sensor_df <- mall_missing_1 %>% unlist
+  sensor_dfsc <- scale(sensor_df)
+
+  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
+
+  options(na.action = na.omit)
+  mall_model_1 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
+               data = tr_dat, family = quasipoisson())
+  options(na.action = na.pass)
+  mall_fit_1 <- predict.glm(mall_model_1, newdata = tr_dat,
+                            type = "response")
+  options(na.action = na.omit)
+
+  mall_mare_1 <- mean(abs(mall_fit_1 - unlist(dfa2$`Bourke Street Mall (South)`)),
+                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
+                                            na.rm = T)
+
+```
+
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South)"}
+ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1)) +
+           labs(y = "Pedestrian Count", x = "Time and Date",
+                title = "Simulated Missingness 1") +
+          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
+```
+
+With random hours of missing data, we have very good predictions by using neighbouring sensors with a low MARE of `r round(mall_mare_1*100, 2)`%.
+Unfortunately, this type of missingness is not consistent with what we observe at most sensors with large proportions of missing data.
+
+Instead, we need to remove 20% of the data in a single period to simulate what we typically observe.
+We take a random date and remove 2629 hours of data before and after this date.
+Then we proceed to estimate the imputation model again and calculate the MARE.
+
+```{r}
+miss_seed <- max(sample(1:nrow(dfa2),1) - 2629, 0)
+miss_idx <- seq(miss_seed, miss_seed + 5258)
+
+mall_missing_2 <- dfa2$`Bourke Street Mall (South)`
+mall_missing_2[miss_idx] <- NA
+
+i <- "Bourke Street Mall (South)"
+sensor_dists <- numeric()
+  for (j in small_miss_sensors[-6]){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors[-6]
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  close_df <- dfa2[, closest_sensor[1]] %>% unlist
+  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
+
+  close_df_scale <- scale(close_df)
+  close_df_scale2 <- scale(close_df2)
+
+  sensor_df <- mall_missing_2 %>% unlist
+  sensor_dfsc <- scale(sensor_df)
+
+  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
+
+  options(na.action = na.omit)
+  mall_model_2 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
+               data = tr_dat, family = quasipoisson())
+  options(na.action = na.pass)
+  mall_fit_2 <- predict.glm(mall_model_2, newdata = tr_dat,
+                            type = "response")
+  options(na.action = na.omit)
+
+  mall_mare_2 <- mean(abs(mall_fit_2 - unlist(dfa2$`Bourke Street Mall (South)`)),
+                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
+                                            na.rm = T)
+
+```
+
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South)"}
+ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2)) +
+           labs(y = "Pedestrian Count", x = "Time and Date",
+                title = "Simulated Missingness 2") +
+          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
+```

### 2017-10-15 thesis/draft.Rmd
fixed test data MARE calculation to usse false-zero count corrected data

@@
   word_document: default
+fontsize: 12pt
+header-includes:
+- \usepackage{setspace}
+- \doublespacing
 ---
@@
 
-```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
+```{r, fig.width=3, fig.height=3, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
 ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
@@
 ggmap(melb3) + geom_line(aes(x = lon, y = lat),
-                        data = sens_lines, size = 2.5,
+                        data = sens_lines, size = 2,
                         colour = viridis(3)[2], alpha = 0.5) +
                geom_point(aes(x = lons, y = lats, colour = cl),
-                          size = 8,
+                          size = 6,
                           data = loc) +
@@
                               y = loc["Southbank",]$lats),
-                          size = 16, colour = viridis(3)[2]) +
+                          size = 12, colour = viridis(3)[2]) +
                geom_label(aes(label = ids, x = lons, y = lats,
@@
                          fontface = 'bold', colour = 'white',
-                         data = loc, size = 3.5,
+                         data = loc, size = 3,
                          nudge_y = -0.000325) +
@@
 for (i in unique(ped_data$Sensor)){
-  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts[, i])), na.rm = T) / mean(unlist(ori_ts[, i]), na.rm = T)
+  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
 }
@@
 for (i in unique(ped_data$Sensor)){
-  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts[, i])), na.rm = T) / mean(unlist(ori_ts[, i]), na.rm = T)
-}
-```
-
-Again, we start by comparing the fit at a sensor location with a small proportion of missing values, Vic Arts Centre:
+  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
+}
+```
+
+Again, we start by comparing the fit at a sensor location with a small proportion of missing values, Melbourne Central:
+
 ```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
@@
                                       c("Spencer St-Collins St (South)",
-                                        "Spencer St-Collins St (North)"))]
+                                        "Spencer St-Collins St (North)",
+                                        "Flinders St-Swanston St (West)"))]
 MARE_ts2 <- MARE_ts[setdiff(unique(ped_data$Sensor),
                                       c("Spencer St-Collins St (South)",
-                                        "Spencer St-Collins St (North)"))]
+                                        "Spencer St-Collins St (North)",
+                                        "Flinders St-Swanston St (West)"))]
 
@@
 
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) sensors")}
+```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}
 
@@
 
-Spencer St-Collins St (North)/(South) was omitted from the the plots above as the out-of-sample MARE is extremely high at `r round(MARE_ts["Spencer St-Collins St (North)"]*100, 2)`% and `r round(MARE_ts["Spencer St-Collins St (South)"]*100, 2)`% respectively.
+Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) was omitted from the the plots above as the out-of-sample MARE is extremely high at `r round(MARE_ts["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_ts["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_ts["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
 
 As these sensor locations were both classified as locations with a large proportion of missing values, it would appear that the relationship between these locations and their neighbouring sensors has changed.
-This is because the in-smaple MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`% and `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% respectively.
+This is because the in-saple MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_tr["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
 
@@
 Another way we can perform cross-validation is to remove data at a sensor, and evaulate the predictions made.
-This will allow us to evaluate the imputation model's performance within sample, which is the primary concern with the imputation process.
+This will allow us to evaluate the imputation model's performance within sample, which is the primary concern with them imputation process.
 We select a sensor which has a small proportion of missing values, Bourke Street Mall (South), which has `r round((1 - na_prop["Bourke Street Mall (South)"])*100, 3)`% of data available.
@@
 ```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South)"}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1)) +
+ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1),
+                     size = 0.1, alpha = 1) +
            labs(y = "Pedestrian Count", x = "Time and Date",
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South)"}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2)) +
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% data missing"}
+ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2),
+                size = 0.1, alpha = 1) +
            labs(y = "Pedestrian Count", x = "Time and Date",
@@
 ```
+
+The MARE of the estimated model is `r round(mall_mare_2*100, 2)`%, which is similar to what was obtained when the missing values were random.
+This demonstrates that the model quite well when we remove 20% of the observations from the training data.
+
+```{r}
+miss_seed <- min(max(sample(1:nrow(dfa2),1) - 6573, 0), 13146)
+miss_idx <- seq(miss_seed, miss_seed + 13146)
+
+mall_missing_3 <- dfa2$`Bourke Street Mall (South)`
+mall_missing_3[miss_idx] <- NA
+
+i <- "Bourke Street Mall (South)"
+sensor_dists <- numeric()
+  for (j in small_miss_sensors[-6]){
+      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
+      sensor_dists[j] <- dists
+  }
+  names(sensor_dists) <- small_miss_sensors[-6]
+  closest_sensor <- sensor_dists %>% sort %>%
+                    head(2) %>% names
+  close_df <- dfa2[, closest_sensor[1]] %>% unlist
+  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
+
+  close_df_scale <- scale(close_df)
+  close_df_scale2 <- scale(close_df2)
+
+  sensor_df <- mall_missing_3 %>% unlist
+
+  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
+
+  options(na.action = na.omit)
+  mall_model_3 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
+               data = tr_dat, family = quasipoisson())
+  options(na.action = na.pass)
+  mall_fit_3 <- predict.glm(mall_model_3, newdata = tr_dat,
+                            type = "response")
+  options(na.action = na.omit)
+
+  mall_mare_3 <- mean(abs(mall_fit_3 - unlist(dfa2$`Bourke Street Mall (South)`)),
+                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
+                                            na.rm = T)
+
+```
+
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% data missing"}
+ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_3),
+                size = 0.1, alpha = 1) +
+           labs(y = "Pedestrian Count", x = "Time and Date",
+                title = "Simulated Missingness 3") +
+          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
+```
+
+Removing 50% of the data using the same method, we find that the in-sample MARE is not affected very much by the significantly large proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
+
+# Predictive Model
+
+For the purpose of prediction, we want to use the most simplistic model possible to make quick predictions of the expected pedestrian counts at each location.
+The model should not be expected to be able to predict large deviations from time and date based estimates.
+
+Similar to the simple imputation model used, we will also use a generalised linear model with quasi-poisson errors using time and date based variables as predictors. The model specification used is:
+
+$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
+
+$$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$
+
+When we estimate the GLM at each sensor location with R, we only need to save the elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function.
+Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets.
+This is due to the model object providing redundancies.
+In particular, elements which are unnecessarily saved for our purposes: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
+
+The result of removing all the unnecessary redundant data elements in the GLM object is a storage size reduction from `r mall_model_2 %>% object.size() %>% format(units = "auto")` to `r mall_model_2 %>% stripGlmLR() %>% object.size() %>% format(units = "auto")`.
+This is only `r round((mall_model_2 %>% stripGlmLR() %>% object.size() %>% as.numeric/ mall_model_2 %>% object.size() %>% as.numeric)*100, 2)`% of the size of the original GLM object.
+With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.

### 2017-10-16 code/pred_model.R
draft predict function, glm model exported to file

@@ -1,0 +1,45 @@
+## predictive model estimation
+## requires imputation to have been already completed
+
+library(dplyr)
+library(pryr)
+
+### code from:
+## http://www.win-vector.com/blog/2014/05/trimming-the-fat-from-glm-models-in-r/
+stripGlmLR = function(cm) {
+  cm$y = c()
+  cm$model = c()
+  
+  cm$residuals = c()
+  cm$fitted.values = c()
+  cm$effects = c()
+  cm$qr$qr = c()  
+  cm$linear.predictors = c()
+  cm$weights = c()
+  cm$prior.weights = c()
+  cm$data = c()
+  
+  
+  cm$family$variance = c()
+  cm$family$dev.resids = c()
+  cm$family$aic = c()
+  cm$family$validmu = c()
+  cm$family$simulate = c()
+  attr(cm$terms,".Environment") = c()
+  attr(cm$formula,".Environment") = c()
+  
+  cm
+}
+#######################################
+
+tr_df <- dfa2
+locs <- loc$ids %>% as.character()
+
+pred_model <- list()
+  for(i in locs) {
+  model <- glm(eval(as.name(i)) ~ Month + DayType*Time, data = tr_df,
+               family = quasipoisson())
+  pred_model[[i]] <- stripGlmLR(model)
+}
+
+save(pred_model, file = "data/model_est.RData")

### 2017-10-16 code/predictions.R

@@ -1,0 +1,38 @@
+library(purrr)
+library(dplyr)
+library(tidyr)
+library(lubridate)
+load("data/model_est.RData")
+
+daytype <- function(date){
+  if(!is.Date(date)){stop("Object is not a date.")}
+  x_day <- wday(date, label = TRUE, abbr = FALSE)
+  x_daytype <- ifelse(x_day == "Sunday", "Sunday",
+                ifelse(x_day == "Saturday", "Saturday",
+                ifelse(x_day == "Monday", "Monday",
+                ifelse(x_day == "Friday", "Friday",
+                          "Midweek"))))
+  return(x_daytype)
+}
+
+ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
+                        t_hour = 12,
+                        is_pub_hol = FALSE) {
+  if(!(t_hour %in% seq(0, 23))){
+    stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
+  if(!is.Date(pred_date)){stop("Date given to predict is not a date. Use lubridate::ymd()")}
+ 
+  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
+               mutate(DayType = daytype(pred_date),
+                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
+                      Date = pred_date)
+  
+  preds <- numeric()
+  for(i in names(pred_model)){
+    preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
+  }
+  preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
+    spread(Sensor_ID, fit) %>% cbind(date_pred)
+  
+  return(preds_df)
+}

### 2017-10-16 code/predictions.R
need to improve efficiency - 3.3 seconds for full day predictions, 130ms for single hour prediction. need to get roxygen working

@@ -2,6 +2,7 @@ library(purrr)
 library(dplyr)
 library(tidyr)
 library(lubridate)
+library(roxygen2)
 load("data/model_est.RData")
 
 daytype <- function(date){
@@ -14,13 +15,25 @@ daytype <- function(date){
                           "Midweek"))))
   return(x_daytype)
 }
-
+#' Predicted pedestrian counts at various sensor locations in Melbourne
+#'
+#' @param pred_date input date
+#' @param t_hour input integer
+#' @param is_pub_hol input logical
+#' @return data frame of the fitted values at each location in wide format, with time and date
+#' @export
+#' @examples
+#' ped_predict()
+#' ped_predict("2017-12-25", 18, TRUE)
 ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
                         t_hour = 12,
                         is_pub_hol = FALSE) {
   if(!(t_hour %in% seq(0, 23))){
     stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
-  if(!is.Date(pred_date)){stop("Date given to predict is not a date. Use lubridate::ymd()")}
+  pred_date <- ymd(pred_date)
+  if(!is.Date(pred_date)){
+    stop(paste("Date given to predict cannot be parsed as a date. Use the date format 'yyyy-mm-dd'. Example: 25 January 2018 is '2018-01-25'."))
+    }
  
   date_pred <- data.frame(Time = as.factor(t_hour)) %>%
                mutate(DayType = daytype(pred_date),
@@ -32,7 +45,17 @@ ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
     preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
   }
   preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
-    spread(Sensor_ID, fit) %>% cbind(date_pred)
+    spread(Sensor_ID, fit) %>% cbind(date_pred) %>% as_tibble
   
   return(preds_df)
+}
+
+ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),
+                            is_pub_hol = FALSE){
+  day_df <- data.frame()
+      for(t in 0:23){
+        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
+        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
+      }
+  return(as.tibble(day_df))
 }

### 2017-10-18 thesis/draft.Rmd
prediction function input, is_pub_hol, now working

@@
 fontsize: 12pt
-header-includes:
-- \usepackage{setspace}
-- \doublespacing
 ---
@@
 With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.
+
+The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves.
+Using the estimated model parameters, we can write a new function to facilitate the predictions in an easy to use format.
+
+```{r, echo = FALSE}
+library(purrr)
+library(dplyr)
+library(tidyr)
+library(lubridate)
+library(roxygen2)
+load("../data/model_est.RData")
+
+daytype <- function(date){
+  if(!is.Date(date)){stop("Object is not a date.")}
+  x_day <- wday(date, label = TRUE, abbr = FALSE)
+  x_daytype <- ifelse(x_day == "Sunday", "Sunday",
+                ifelse(x_day == "Saturday", "Saturday",
+                ifelse(x_day == "Monday", "Monday",
+                ifelse(x_day == "Friday", "Friday",
+                          "Midweek"))))
+  return(x_daytype)
+}
+#' Predicted pedestrian counts at various sensor locations in Melbourne
+#'
+#' @param pred_date input date
+#' @param t_hour input integer
+#' @param is_pub_hol input logical
+#' @return data frame of the fitted values at each location in wide format, with time and date
+#' @export
+#' @examples
+#' ped_predict()
+#' ped_predict("2017-12-25", 18, TRUE)
+#' 
+ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
+                        t_hour = 12,
+                        is_pub_hol = FALSE,
+                        long = FALSE) {
+  if(!(t_hour %in% seq(0, 23))){
+    stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
+  pred_date <- ymd(pred_date)
+  if(!is.Date(pred_date)){
+    stop(paste("Date given to predict cannot be parsed as a date. Use the date format 'yyyy-mm-dd'. Example: 25 January 2018 is '2018-01-25'."))
+    }
+
+  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
+               mutate(DayType = ifelse(is_pub_hol,
+                                       "Holiday",
+                                       daytype(pred_date)),
+                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
+                      Date = pred_date)
+
+  preds <- numeric()
+  for(i in names(pred_model)){
+    preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
+  }
+  if(long){
+    return(preds)
+  }
+    else
+  preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
+    spread(Sensor_ID, fit) %>% cbind(date_pred) %>% as_tibble
+
+  return(preds_df)
+}
+
+ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),
+                            is_pub_hol = FALSE){
+  day_df <- data.frame()
+      system.time(for(t in 0:23){
+        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
+        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
+      })
+  return(as.tibble(day_df))
+}
+
+```
+
+```{r, echo = TRUE, eval = FALSE}
+ped_predict(pred_date = "2017-10-24", t_hour = 13)
+```
+This is function will return tibble (trimmed down version of `data.frame()` in R) containing the predicted counts at each sensor location for the 13th hour (13:00/1:00 PM) on 24 October 2017.
+This tibble can then be easily used for visualisations or analysis.
+
+Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predic.t

### 2017-10-18 code/predictions.R

@@ -25,9 +25,11 @@ daytype <- function(date){
 #' @examples
 #' ped_predict()
 #' ped_predict("2017-12-25", 18, TRUE)
+#' 
 ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
                         t_hour = 12,
-                        is_pub_hol = FALSE) {
+                        is_pub_hol = FALSE,
+                        long = FALSE) {
   if(!(t_hour %in% seq(0, 23))){
     stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
   pred_date <- ymd(pred_date)
@@ -36,7 +38,9 @@ ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
     }
  
   date_pred <- data.frame(Time = as.factor(t_hour)) %>%
-               mutate(DayType = daytype(pred_date),
+               mutate(DayType = ifelse(is_pub_hol,
+                                       "Holiday",
+                                       daytype(pred_date)),
                       Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
                       Date = pred_date)
   
@@ -44,6 +48,10 @@ ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
   for(i in names(pred_model)){
     preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
   }
+  if(long){
+    return(preds)
+  }
+    else
   preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
     spread(Sensor_ID, fit) %>% cbind(date_pred) %>% as_tibble
   
@@ -53,9 +61,9 @@ ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
 ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),
                             is_pub_hol = FALSE){
   day_df <- data.frame()
-      for(t in 0:23){
+      system.time(for(t in 0:23){
         t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
         suppressWarnings(day_df <- bind_rows(day_df, t_preds))
-      }
+      })
   return(as.tibble(day_df))
 }

### 2017-10-18 thesis/draft.Rmd
typo fix

@@
 ```{r, echo = TRUE, eval = FALSE}
-ped_predict(pred_date = "2017-10-24", t_hour = 13)
-```
-This is function will return tibble (trimmed down version of `data.frame()` in R) containing the predicted counts at each sensor location for the 13th hour (13:00/1:00 PM) on 24 October 2017.
+ped_predict(pred_date = "2017-12-26", t_hour = 13, is_pub_hol = TRUE)
+```
+This function will return tibble (trimmed down version of `data.frame()` in R) containing the predicted counts at each sensor location for the 13th hour (13:00/1:00 PM) on Boxing Day (26 December) 2017.
 This tibble can then be easily used for visualisations or analysis.
-
-Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predic.t
+Due to the lack of data on future public holiday dates, the function assumes the date given is not a public holiday unless `is_pub_hol = TRUE` is provided.
+Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predict.

### 2017-10-20 thesis/draft.Rmd
need to add: abstract, updated introduction, motivation, lit review,  completed predictive model section with glm justification, conclusion, references

@@
 ---
-title: "Thesis Draft"
+title: "Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods"
 author: "Gavin Chin"
@@
 library(ggmap)
+library(broom)
 set.seed(12345)
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South)"}
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South).
+20% of the observations are removed from the traning data for the imputation model at random."}
 ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1),
                      size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date",
-                title = "Simulated Missingness 1") +
+           labs(y = "Pedestrian Count", x = "Time and Date") +
           theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
@@
 $$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$
-
-When we estimate the GLM at each sensor location with R, we only need to save the elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function.
-Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets.
-This is due to the model object providing redundancies.
-In particular, elements which are unnecessarily saved for our purposes: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
-
-The result of removing all the unnecessary redundant data elements in the GLM object is a storage size reduction from `r mall_model_2 %>% object.size() %>% format(units = "auto")` to `r mall_model_2 %>% stripGlmLR() %>% object.size() %>% format(units = "auto")`.
-This is only `r round((mall_model_2 %>% stripGlmLR() %>% object.size() %>% as.numeric/ mall_model_2 %>% object.size() %>% as.numeric)*100, 2)`% of the size of the original GLM object.
-With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.
-
-The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves.
-Using the estimated model parameters, we can write a new function to facilitate the predictions in an easy to use format.
 
@@
 
+Using a GLM predicts well compared to estimating with other models such as AR(p) or other time series models when considering the tradeoff between model complexity and prediction accuracy.
+
+As 
+```{r, fig.width=6, fig.height=4}
+ggplot(dfa2 %>%
+         mutate(Date = date(Date_Time))) +
+  geom_path(aes(x = as.integer(Time), y = `Melbourne Central`,
+                group = Date, colour = Month),
+            alpha = 0.1) +
+  scale_colour_viridis(discrete = TRUE, option = "C") +
+  facet_wrap(~ DayType) + theme_minimal() +
+  theme(legend.position = "none") +
+  labs(x = "Time of Day", y = "Pedestrian Counts \n at Melbourne Central")
+```
+
+```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM at each sensor for each variable. Variables (x-axis) is ordered by the scagnostic measure, Outlying.")}
+dfcoeff <- data.frame(terms = factor(), coeff = numeric(), sensor = factor())
+for(i in names(pred_model)){
+  terms <- names(coef(pred_model[[i]]))
+  coeff <- coef(pred_model[[i]])
+  sensor <- rep(i, length(terms))
+  df_b <- data.frame(terms, coeff, sensor)
+  dfcoeff <- bind_rows(dfcoeff, df_b)
+}
+dfcoeff <- dfcoeff %>% spread(terms, coeff)
+
+library(GGally)
+library(scagnostics)
+ggparcoord(dfcoeff[, -1],
+           order = "Outlying", alphaLines = 0.3)
+```
+
+When we estimate the GLM at each sensor location with R, we only need to save the elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function.
+Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets.
+This is due to the model object providing redundancies.
+In particular, elements which are unnecessarily saved for our purposes: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
+
+The result of removing all the unnecessary redundant data elements in the GLM object is a storage size reduction from `r mall_model_2 %>% object.size() %>% format(units = "auto")` to `r mall_model_2 %>% stripGlmLR() %>% object.size() %>% format(units = "auto")`.
+This is only `r round((mall_model_2 %>% stripGlmLR() %>% object.size() %>% as.numeric/ mall_model_2 %>% object.size() %>% as.numeric)*100, 2)`% of the size of the original GLM object.
+With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.
+
+The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves.
+Using the estimated model parameters, we can write a new function to facilitate the predictions in an easy to use format.
+
 ```{r, echo = TRUE, eval = FALSE}

### 2017-10-20 thesis/draft.Rmd
update..

@@
 
-For the purpose of prediction, we want to use the most simplistic model possible to make quick predictions of the expected pedestrian counts at each location.
-The model should not be expected to be able to predict large deviations from time and date based estimates.
-
-Similar to the simple imputation model used, we will also use a generalised linear model with quasi-poisson errors using time and date based variables as predictors. The model specification used is:
+For the purpose of prediction, the use of the most simplistic model possible to make quick predictions of the expected pedestrian counts at each location, accessible to the public.
+The model is not expected to be able to predict large deviations from time and date based estimates.
+
+Similar to the simple imputation model used,  a generalised linear model with quasi-poisson errors is estimated using time and date based variables as predictors. The model specification used is:
 
@@
 
-Using a GLM predicts well compared to estimating with other models such as AR(p) or other time series models when considering the tradeoff between model complexity and prediction accuracy.
-
-As 
-```{r, fig.width=6, fig.height=4}
+Using a GLM predicts well compared to estimating with other models such as AR($p$) or other time series models when considering the tradeoff between model complexity and prediction accuracy.
+Due to multiple seasonalities, AR($p$) models would need to be considered on a hour of the week basis at each sensor location.
+Prediction with time series models would also require of historical data.
+For the objective of this predictive model, the additional imputs required by end users to predict pedestrian counts when making out of sample predictions is not desirable.
+
+```{r, fig.width=6, fig.height=4, fig.cap = paste("Plot of daily pedestrian counts series at Melbourne Central by DayType and Month.
+Month of the year is mapped to colour to help identify annual seasonal patterns, where month has an additive effect on the pedestrian counts.
+At Melbourne Central, different months can be seen to have different mean counts for a given type of day and hour of the day.")}
 ggplot(dfa2 %>%
@@
 
+A plot of the daily pedestrian counts series at Melbourne Central by type of day and month show that time and date variables as deterministic predictors adequately explains most variation in the pedestrian counts.
+It can be seen that month has an additive effect as the pattern of the hourly counts for a given date are not affected by month.
+The six different factor levels that `DayType`, the type of day, is seen to be a sufficiently explain six different patterns.
+The type of day, Midweek, shows evidence that no additional information is gained from identifying the day to be Tuesday, Wednesday or Thursday.
+
 ```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM at each sensor for each variable. Variables (x-axis) is ordered by the scagnostic measure, Outlying.")}

### 2017-10-21 thesis/draft.Rmd
12pm update

@@
 
-```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Comparison of distribution of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data"}
+```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
+The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data.
+This emphasises the necessity to correct for false zero counts as it will affect "}
 theme_g <- theme_minimal() +
@@
                     colour = "white", fill = 'black', alpha = 0.4) +
-          scale_y_continuous(limits = c(0,35)) +
-          ggtitle("Histogram of Proportions of Missing Values \n during 2014 - 2016")
+          scale_y_continuous(limits = c(0,35))
 p2 <- ggplot() + geom_histogram(aes(x = na_prop_pre*100), bins = 12) +
@@
                     colour = "white", fill = 'black', alpha = 0.4) +
-          scale_y_continuous(limits = c(0,35)) +
-          ggtitle("Histogram of Proportions of Missing Values \n during 2014 - 2016 prior to false zero counts fix")
+          scale_y_continuous(limits = c(0,35))
 grid.arrange(p2, p1, nrow = 1)
@@
 
-```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM at each sensor for each variable. Variables (x-axis) is ordered by the scagnostic measure, Outlying.")}
+```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM for the interaction term $\text{DayType} \times \text{Time}$ for each hour at each sensor.
+Variables (x-axis) is ordered by the scagnostic measure, Outlying.
+"), include=FALSE, eval = FALSE}
 dfcoeff <- data.frame(terms = factor(), coeff = numeric(), sensor = factor())
@@
 library(scagnostics)
-ggparcoord(dfcoeff[, -1],
-           order = "Outlying", alphaLines = 0.3)
-```
-
-When we estimate the GLM at each sensor location with R, we only need to save the elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function.
+ggparcoord(dfcoeff[, 51:74], scale = "std", 
+           order = "Clumpy", alphaLines = 0.3) + theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, size = 6)) +
+  labs(x = "Parameter")
+```
+
+When estimating the GLM at each sensor location with R, only the necessary elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function are saved to file.
 Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets.
 This is due to the model object providing redundancies.
-In particular, elements which are unnecessarily saved for our purposes: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
+In particular, elements which are unnecessarily saved for the purposes of the predictive model: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
 
@@
 The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves.
-Using the estimated model parameters, we can write a new function to facilitate the predictions in an easy to use format.
+Using the estimated model parameters, a new function is written to facilitate the predictions in an easy to use format.
 

### 2017-10-21 thesis/draft.Rmd
WIP: removing plot titles, improving figure captions

@@
 
-With random hours of missing data, we have very good predictions by using neighbouring sensors with a low MARE of `r round(mall_mare_1*100, 2)`%.
-Unfortunately, this type of missingness is not consistent with what we observe at most sensors with large proportions of missing data.
-
-Instead, we need to remove 20% of the data in a single period to simulate what we typically observe.
-We take a random date and remove 2629 hours of data before and after this date.
-Then we proceed to estimate the imputation model again and calculate the MARE.
+With random hours of missing data, the predictions made by using neighbouring sensors  are estimated with good accuracy with a low MARE of `r round(mall_mare_1*100, 2)`%.
+Unfortunately, this type of missingness is not consistent with what is generally observed at most sensors with large proportions of missing data.
+
+Instead, a more meaningful method is to remove 20% of the data in a single period to simulate what is typically observed.
+A random date is taken and then 2629 hours of data is removed before and after this date (10% before, 10% after).
+The imputation model is estimated again in order to calculate a MARE.
 
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% data missing"}
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence.
+This method of simulating missing data is a better representation of the missing data observed at sensors with a large proportion of missing data compared to randomly removing hourly observations from the training data."}
 ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2),
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% data missing"}
+```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence.
+A large proportion of the training data is removed to test the robustness of the imputation method used at sensors with extremely large proportions of missing data during the training period defined."}
 ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_3),
@@
 
-Removing 50% of the data using the same method, we find that the in-sample MARE is not affected very much by the significantly large proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
+Removing 50% of the data using the same method, it is found that the in-sample MARE is not affected significantly by the larger proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
 

### 2017-10-21 thesis/draft.Rmd
WIP: removing plot titles, improving figure captions

@@
                 size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date",
-                title = "Simulated Missingness 2") +
+           labs(y = "Pedestrian Count", x = "Time and Date") +
           theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
@@
                 size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date",
-                title = "Simulated Missingness 3") +
+           labs(y = "Pedestrian Count", x = "Time and Date") +
           theme_minimal() + theme(plot.title = element_text(hjust = 0.5))

### 2017-10-21 thesis/draft.Rmd
conclusion wip

@@
 
-First, we remove randomly remove data such that 20% of the data is missing, then estimate the imputation model. Using this model, we calculate the MARE within the training data.
+First, data is randomly removed such that 20% of the training data is missing, then the imputation model is estimated.
+Using the estimated model, the MARE within the training data is calculated (in-sample MARE).
 ```{r}
@@
                       Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
-                      Date = pred_date)
+                      Date = pred_date,
+                      Date_Time = ymd_h(paste(Date, t_hour, sep = " ")))
 
@@
 
-ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),
-                            is_pub_hol = FALSE){
+ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),is_pub_hol = FALSE){
   day_df <- data.frame()
-      system.time(for(t in 0:23){
+      for(t in 0:23){
         t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
         suppressWarnings(day_df <- bind_rows(day_df, t_preds))
-      })
+      }
   return(as.tibble(day_df))
@@
 Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predict.
+
+Using these function as a basis, visualisations of pedestrian counts can be easily generated in R:
+```{r, echo = TRUE, fig.cap = "Prediction for 21/05/2018 in R using `ped_predict_day()`, then plotting the time series using `ggplot2`.
+This shows the user friendly nature of the function, where the function returns predictions in a tidy, wide format.
+No data manipulation is required in order to plot the predicted counts.", fig.height=2}
+ped_predict_day("2018-05-21") %>% 
+  ggplot() + geom_line(aes(x = Date_Time, y = Southbank))
+```
+
+# Conclusion  
+Analysis of the Melbourne CBD pedestrian data made publicly available by the City of Melbourne has revealed issues with the dataset with relation to missing values in particular.
+A major issue which was found when using data sourced from the City of Melbourne's official pedestrian data visualisation (also accessible using `rwalkr::walk_melb()`) is potential for false zero-counts where missing values were given the value of zero.
+As outliers have major implications on analysis, a correction for the false counts in the data was proposed, by checking for lengths of repeated value sequences.
+An imputation method for the pedestrian data was developed with a theory of sensor locations with a large proportion of missing values require modelling with non-deterministic predictors due to the lack of data.
+This method involved using neighbouring sensor counts data for prediction.
+Sensors with a small proportion of missing values used imputation models with only time and date variables as predictors.
+The proposed algorithm was evaluated in comparison to a single model specification of a generalised linear model with only time and date based variables at each sensor location.

### 2017-10-21 thesis/draft.Rmd
implementation write up started

@@
 threshold <- 0.1
-threshold_miss_hrs <- 50
+threshold_miss_hrs <- 6
 
@@
 Because the covariance with neighbouring sensors may vary over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
+
+## Implementation of improved imputation algorithm
+### False zero-counts correction
+The false zero-counts correction is performed in R using the run length encoding function, `rle()`.
+This function computes the length and values of runs of equal values in a vector.
+Unfortunately, this function does not recognise `NA` in the its computation, so all NA values are initially recoded to take the value 0 at each sensor.
+This results in all missing values and true zero counts to take the value zero when running `rle()`.
+With the returned run lengths, all sequences of repeated values longer than the defined length of 6 hours are replaced with `NA`.
+Any values which were also originally missing, but were not missing for longer than 6 hours (and hence are still coded as 0), are recoded to be `NA`.
+It must be noted that sequences of repeated values not equal to zero which run for longer that 6 hours would be incorrectly replaced with `NA`.
+This is highly unlikely, however, as 6 hours of the exact same pedestrian counts would be extremely rare.
+
+### Caculation of proportion of missing values and classification of sensors
+In the working data, a wide format is used.
+This makes it easier to calculate missing value proportion at each sensor location, as it is the proportion of missing values in the column for each sensor.
+Using `is.na()` returns a logical vector indicating whether the sensor count is missing/`NA` or not.
+The sum of the instances where `is.na = TRUE` divided by the hours in the period of 2014-2016 is computed to get the proportion of missing values.
+
+These proportions are saved for each sensor and then classified as either "large" or "small" based on a threshold value for the proportion.
+A list of sensors with large proportions of missing data and a list of sensors with small proportions of missing data is generated.
+
+### Estimation of GLM at sensor locations with a small proportion of missing
+A for loop
 
@@
 Sensors with a small proportion of missing values used imputation models with only time and date variables as predictors.
-The proposed algorithm was evaluated in comparison to a single model specification of a generalised linear model with only time and date based variables at each sensor location.
+The proposed algorithm was evaluated in comparison to a single model specification of a generalised linear model with only time and date based variables at each sensor location and uncorrected training data (false zero counts not corrected).
+This method was found to have improved prediction accuracy, as well as being robust at high levels of missingness.

### 2017-10-21 thesis/draft.Rmd
update...

@@
 
-### Small-Large Split Approach
+### Improved imputation algorithm using small-large split approach
 
@@
 ## Implementation of improved imputation algorithm
+
+The exact code used for this paper can be found on http://github.com/gavinchin/honours2017
+
 ### False zero-counts correction
@@
 ### Estimation of GLM at sensor locations with a small proportion of missing
-A for loop
+A for loop is used to estimate the GLM specified:
+```{r, echo = TRUE, eval = FALSE}
+glm(dfa[, i] ~ Month + DayType*Time, data = dfa,
+    family = quasipoisson())
+```
+where `dfa` is the object name given to the data frame containing the training data after the false zero-count correction, and `i` is the $i^{th{}}$ sensor in the list of sensors with a small proportion of missing values.
+These models are stored in a list, forming a list of models.
+An alternative approach to using a for loop in R is splitting the data into separate data frames for each sensor location (forming a list of data frames) and applying the `glm()` function on each data frame in the list using the function `purrr::map()`.
+However, when this was attempted, performance issues were encountered due to the large size of the estimated model objects in R.
+This was not an issue when a for loop was used.
+Additional code for removing unnecessary elements of the GLM model object before storing to the list of models, reduucing the memory usage.
+
+### Imputation at sensor locations with a small proportion of missing values
+
+The predicted values are obtained using `predict.glm()`, using the option `type = "response"` to have predicted values returned predicted counts.
+The default option returns the predicted values of $\hat{\lambda}$ in the quasipoisson process, not the counts, $\exp (\hat{\lambda})$.
 

### 2017-10-21 thesis/draft.Rmd
update

@@
 ### False zero-counts correction
+
 The false zero-counts correction is performed in R using the run length encoding function, `rle()`.
@@
 ### Caculation of proportion of missing values and classification of sensors
+
 In the working data, a wide format is used.
@@
 
-### Estimation of GLM at sensor locations with a small proportion of missing
+### Estimation of GLM at sensor locations with a small proportion of missing values
+
 A for loop is used to estimate the GLM specified:
@@
 The default option returns the predicted values of $\hat{\lambda}$ in the quasipoisson process, not the counts, $\exp (\hat{\lambda})$.
+Missing values are then replaced using the imputed values, generating a data frame of complete data as sensor locations for with a small proportion of missing values to be used for prediction at sensors with a large proportion of missing values.
+
+### Estimation of GLM at sensor locations with a large proportion of missing values
+The following steps are performed on each sensor with a large proportion of missing values inside a for loop.
+
+##### Finding neighbouring sensors
+Using location data of the sensors provided by the City of Melbourne, a data frame containing the latitude and longitude coordinates of all the sensors which were classified as having a small proportion of missing values (and now have imputed, complete data), which are the candidates to be neighbours used for prediction.
+Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()`, an approximation of distance between each of the candidate sensors is calculated in and stored as a vector.
+Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.
 

### 2017-10-21 thesis/draft.Rmd
update

@@
 
-### Estimation of GLM at sensor locations with a large proportion of missing values
+### Estimation procedure at sensor locations with a large proportion of missing values
 The following steps are performed on each sensor with a large proportion of missing values inside a for loop.
@@
 Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.
+
+##### Extracting neighbouring sensors counts and standardising
+
+Using the sensor location names, the neighbouring sensors counts are extracted from the imputed training data and each vector is standardised to $\mu = 0, \sigma^2 = 1$ using the base R function `scale()` then stored.
+
+##### GLM estimation
+
+The GLM for sensor locations with a large proportion of missing values also uses `glm()` as follows:
+```{r, echo = TRUE, eval = FALSE}
+glm(dfa[, i] ~ Time*close_dfsc + Time*close_dfsc2,
+    data = dfa, family = quasipoisson())
+```
+where `dfa[, i]` is the vector of pedestrian counts for the sensor being estimated, `Time` is the factor level for the hour of the day and `close_dfsc` and `close_dfsc2` are the standardised pedestrian counts at two geographically closest sensors.
 

### 2017-10-21 thesis/draft.Rmd
bibliography updated

@@
 
-[^1]: from _'About Melbourne Data'_ page at https://data.melbourne.vic.gov.au/about
-
-[^2]: the specific pedestrian data is available at   https://data.melbourne.vic.gov.au/Transport-Movement/Pedestrian-volume-updated-monthly-/b2ak-trbp
-
-[^3]: https://github.com/earowang/rwalkr/
-
-### Imputation Literature Review Notes
+### Literature Review
 <!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->
@@
 where `dfa[, i]` is the vector of pedestrian counts for the sensor being estimated, `Time` is the factor level for the hour of the day and `close_dfsc` and `close_dfsc2` are the standardised pedestrian counts at two geographically closest sensors.
+Like the GLM for sensor locations with a small proportion of missing values, the model object is slim down to remove unnecessary elements to reduce memory usage.
+It is also noteworthy that the computational burden of the procedure for imputation using neighbouring sensors is significantly less due to less parameters being estimated.
+
+##### Replacing missings with imputed values  
+Unlike the GLM used for imputation for sensors with a small proportion of missing values, the model for neighbouring sensors requires transformation of the counts at the neighbours for imputation.
+As a consequence, it is more efficient to use `predict.glm()` inside the for loop as the neighbouring sensor counts have already been standardised when used for training the GLM.
+Again, the option `type = "response"` is specified to return predicted counts.
 

### 2017-10-21 thesis_draft/00-abstract.Rmd
bookdown update - need to fix formatting, improve introduction, write abstract, finish off lit review and conclusion

--- 2017-10-21 thesis/draft.Rmd
@@
 ---
-title: "Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods"
-author: "Gavin Chin"
-date: "21 September 2017"
-output:
-  pdf_document: default
-  html_document: default
-  word_document: default
-fontsize: 12pt
+knit: "bookdown::render_book"
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
-                      message = FALSE, warning = FALSE,
-                      fig.pos = 'h')
-library(tidyverse)
-library(lubridate)
-library(foreach)
-library(doSNOW)
-library(rwalkr)
-library(pander)
-library(RColorBrewer)
-library(gridExtra)
-library(viridis)
-library(ggmap)
-library(broom)
-set.seed(12345)
-```
-
-```{r data_loading, warning=F, message=F}
-ped_data <- NULL
-if(file.exists("../data/ped_df.csv")){
-  ped_data <- read_csv("../data/ped_df.csv")
-  ped_data$X1 <- NULL
-  }
-
- ## ped_data <- read_csv("http://209.148.91.227/files/ped_df_28sep17.csv")
-
-## update data
-last_df_date <- ifelse(file.exists("../data/ped_df.csv"),
-                        ped_data$Date %>% sort() %>%
-                        tail(1),
-                        as.Date("2013-12-31")) + 1 %>% as.Date(origin = "1970-01-01")
-
-if(last_df_date != today(tzone = "Australia/Melbourne")){
-  new_data <- rwalkr::walk_melb(from = last_df_date, to = today()-1)
-  ped_data <- rbind(ped_data, new_data)
-  write_csv(ped_data, "../data/ped_df.csv")
-  }
-
-ped_data$Sensor_ID <- ped_data$Sensor
-ped_data$Hourly_Counts <- ped_data$Count
-ped_data$Date_Time <- ymd_h(paste(ped_data$Date, ped_data$Time, sep = " "))
-ped_data$Day <- lubridate::wday(ped_data$Date_Time, label = TRUE, abbr = F)
-ped_data$Month <- lubridate::month(ped_data$Date_Time, label = TRUE, abbr = F)
-ped_data$Time <- as.factor(ped_data$Time)
-
-## long to wide
-ped_data %>% select(Sensor_ID, Count, Date_Time, Time, Month, Day) %>% 
-  spread(Sensor_ID, Count) -> dfa
-
-pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
-pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
-pub_hday16 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
-pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)
-
-pub_hdays$Date <- ymd(pub_hdays$Date)
-pub_hdays$Month <- lubridate::month(pub_hdays$Date, label = TRUE, abbr = F)
-pub_hdays$VIC <- 0
-pub_hdays$VIC[grep(glob2rx("*VIC*"), pub_hdays$`Applicable.To`)] <- 1
-pub_hdays$VIC[grep("NAT", pub_hdays$`Applicable.To`)] <- 1
-
-dfa <- dfa %>% mutate(IsHDay = (date(Date_Time) %in% pub_hdays$Date[pub_hdays$VIC == 1]))
-
-dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
-                     ifelse(dfa$Day == "Saturday", "Saturday",
-                    ifelse(dfa$Day == "Monday", "Monday",
-                     ifelse(dfa$Day == "Friday", "Friday",
-                                  "Midweek"))))
-dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
-dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
-original_df <- dfa
-dfa <- filter(dfa, Date_Time < "2017-01-01")
-ori_2 <- filter(original_df, Date_Time < "2017-01-01")
-ori_ts <- filter(original_df, Date_Time > "2016-12-31")
-ori_ts2 <- ori_ts
-```
-## Background
-
-#### The Data: City of Melbourne Pedestrian Data
+# Abstract {-}
 
@@
 
-The dataset being investigated can be obtained from the official City of Melbourne's open data[^2].
-The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations.
-In total, there is `r nrow(ped_data)` observations in the dataset, representing `r nrow(dfa)` hours of data.
-It is available in `.csv` format, allowing for easy data import into R.
-We can also access the pedestrian data using the the `rwalkr` package by Earo Wang[^3].
-This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.
-
-```{r}
-ped_data %>% select(Sensor, Hourly_Counts) %>% split(.$Sensor)  %>%
-            map(is.na) %>% map(sum) %>% unlist -> sensor_missvals
-```
-
-Not even an single sensor location has complete data for the period between between 1/1/2014 and `r paste(day(today()), month(today()), year(today()), sep = "/")`.
-
-[^1]: from _'About Melbourne Data'_ page at https://data.melbourne.vic.gov.au/about
-
-[^2]: the specific pedestrian data is available at   https://data.melbourne.vic.gov.au/Transport-Movement/Pedestrian-volume-updated-monthly-/b2ak-trbp
-
-[^3]: https://github.com/earowang/rwalkr/
-
-### Imputation Literature Review Notes
-<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->
-
-<!-- http://ir.lib.uwo.ca/cgi/viewcontent.cgi?article=1250&context=csce2016   -->
-<!-- - traffic counters with 40% to 60% missing values -->
-<!-- - heuristic methods -->
-<!--   - using "good historical values" or historical average values resulted -->
-<!--     in MARE (mean absolute relative error) reaching up to 80% -->
-<!--   - using moving average values worked better than other heuristic methods -->
-<!--   - all heuristic methods inhenrently cannot reflect sudden fluctuations during abnormal periods -->
-<!-- - pattern matching methods -->
-<!--   - find "candidate" volume pattern to compare to "study curve" and use the candidate which best matches the study curve -->
-<!-- - ARIMA -->
-<!--   - did not work well with multiple seasonalities -->
-<!-- - AI methods -->
-<!--   - genetic algorithms (GAs) and artificial neural networks (ANNs) -->
-
-<!-- Method used in the study was kNN with an interaction using a weather model, which they found to work best with large variations in traffic during winter -->
-<!-- (severe weather conditions affecting traffic volume)   -->
-
-Reviewing current research on imputation of missing values in traffic data, a paper on imputation of missing classified traffic data during winter season (Imputation of Missing Classified Traffic Data During Winter Season - _Hyuk-Jae Roh, Satish Sharma, Prasanta K.
-Sahu_) stated some methods which were found to be poor for imputation.
-The data they were attempting imputation for were traffic counters located on the highway network in Alberta, Canada with between 40% and 60% missing data.
-
-Replacement with "good" historical values, or using historical average values was found to have resulted in very poor fit with high MARE/MARD (mean absolute relative error/difference).
-They did find that using a moving average worked the best out of all the heuristic methods used.
-The largest problem with heuristic methods, however, is the inherent inability to reflect sudden fluctuations/shocks during abnormal periods.
-With such a large proportion of missing values, pattern matching methods were investigated.
-This involves comparing the _study curve_, the pattern at the counter which is to be imputed, to candidate patterns (patterns at other locations).
-
-For the purposes of that paper, it was found to be inappropriate with the data being analysed.
-However, unlike _Roh et al._ where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller.
-Another difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data.
-They proposed using non-parametric estimation methods with k-Nearest Neighbours for their imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.
-
-<!-- #### Flexible and Robust Method for Missing Loop Detector Data Imputation - _K. C. Henrickson, Y. Zou, and Y. Wang_   -->
-
-<!-- http://docs.trb.org/prp/15-5804.pdf   -->
-<!-- - also use kNN -->
-
-<!-- #### Key Factors of K-nearest Neighbors Nonparametric Regression in Short-time Traffic Flow Forecasting -->
-<!-- http://www.springer.com/cda/content/document/cda_downloaddocument/9789462391017-c2.pdf?SGWID=0-0-45-1492476-p177147492 -->
-
-<!-- #### K-NEAREST NEIGHBOR ALGORITHM FOR UNIVARIATE TIME SERIES PREDICTION -->
-
-<!-- http://webbut.unitbv.ro/BU2012/Series%20III/BULETIN%20III/Sasu%20A.pdf -->
-
-## Methodology: Imputation of Missing Values
-
-### Basic GLM Approach
-
-The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
-$$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
-$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
-`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
-Both these variables are categorical/factors. We cannot treat time of the day as an integer or numeric because it does not have a linear relationship with counts.s
-
-We use the Quasi-Poisson error distribution as opposed to normal/Gaussian errors due to the response variables being count data.
-The Poisson distribution allows for only non-negative integer values, while estimating a Quasi-Poisson model estimates an additional parameter for overdispersion.
-This allows more flexibility in the model by allowing the assumption that $E[Y]=Var[Y]=\mu$ to be relaxed.
-Instead, the Quasi-Poisson distribution allows for $Var[Y] = \theta \mu$.
-
-```{r}
-## Base GLM models
-
-### code from:
-## http://www.win-vector.com/blog/2014/05/trimming-the-fat-from-glm-models-in-r/
-stripGlmLR = function(cm) {
-  cm$y = c()
-  cm$model = c()
-
-  cm$residuals = c()
-  cm$fitted.values = c()
-  cm$effects = c()
-  cm$qr$qr = c()  
-  cm$linear.predictors = c()
-  cm$weights = c()
-  cm$prior.weights = c()
-  cm$data = c()
-
-  cm$family$variance = c()
-  cm$family$dev.resids = c()
-  cm$family$aic = c()
-  cm$family$validmu = c()
-  cm$family$simulate = c()
-  attr(cm$terms,".Environment") = c()
-  attr(cm$formula,".Environment") = c()
-
-  cm
-}
-#######################################
-```
-
-```{r}
-cl <- makeCluster(2)
-registerDoSNOW(cl)
-fitted_base <- list()
-models_base <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
-    tr_dat <- data.frame(dfa[, i],  HDay = dfa$HDay, Time = dfa$Time)
-    colnames(tr_dat)[1] <- "Hourly_Counts"
-    model <- glm(Hourly_Counts ~ HDay*Time, data = tr_dat, family = quasipoisson())
-    return(stripGlmLR(model))
-  }
-names(models_base) <- unique(ped_data$Sensor_ID)
-# ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels
-for (i in unique(ped_data$Sensor)){  
-  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
-  options(na.action = na.pass)
-  fitted_base[[i]] <- predict.glm(models_base[[i]], newdata = tr_dat, type = "response")
-  options(na.action = na.omit)
-  }
-
-```
-
-While this works well with a small proportion of missing values, it also requires a large number of observations.
-Specifically, with this paramaterisation, we have $23 + 7 + (23 \times 7) + 1=192$ coefficients to estimate.
-Another disadvantage of using this model is the lack of robustness to outliers due to the sensitivity of some parameters.
-This becomes particularly problematic at sensor locations which have been recently installed, where there is a lack of historical data.
-
-### Improved imputation algorithm using small-large split approach
-
-The proposed alternative approach to imputing the missing values in the data focuses on improving the models used at locations with a large proportion of missing values.
-A potential threshold value which could be used to class a sensor as having a "large" proportion of missing values vs a "small" proportion of missing values is 10%.
-At locations with a "small" proportion of missings, we can continue to use the GLM quasi-possion regression specified in $(1)$.
-At the locations with a large proportion of missing values, we need to use information from neighbouring sensors.
-
-The first issue which needs to be addressed is the potential of values which are actually missing due to sensor failure or malfunction having a zero count.
-This issue will cause bias in the estimates, as well as affect the classification of a sensor.
-Of course, we cannot simply classify all zero counts as `NA`, as true zero values are possible.
-
-```{r}
-n_obs <- nrow(dfa)
-threshold <- 0.1
-threshold_miss_hrs <- 6
-
-run_lengths <- list()
-
-for(i in unique(ped_data$Sensor)){
-    dfa[is.na(dfa[, i]), i] <- 0
-    na_length_check <- rle(as.numeric(unlist(dfa[,i])))
-    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
-    rl <- dfa[, i] %>% unlist %>% rle()
-    dfa[maybeNA, i] <- NA
-    defsNA <- is.na(ori_2[, i])
-    dfa[defsNA, i] <- NA
-    run_lengths[[i]] <- rl
-  }
-
-ori_2 <- dfa
-
-run_lengths %>% lapply(., `[[`, 1) %>%
-    unlist -> sequences
-
-pander(summary(sequences[sequences > 1]))
-```
-Table: Summary statistics of the sequence lengths of repeated values
-
-Instead, a simple check we can implement is to look for long sequences of zero counts.
-For example, we may want to classify any sequence of `Hourly_Counts = 0` running for longer than 24 hours as being considered `NA`.
-From the summary of lengths of repeated values (which are repeated at least once), we see that the distribution of the lengths are very positively skewed with a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data.
-Classifying any sequence of repeated values, typically 0, which has a length greater than 6 as `NA` should allow true zero counts to avoid being misclassified as `NA`.
-We select the length of 6 hours as it would be unreasonable to assume any sensor location to have the exact same number of pedestrian counts over such a long period.
-
-After we have replaced questionable zero values with the value `NA`, we can calculate the proportion of missing values for each sensor location.
-From these calculated proportion, we can classify sensors as either having a "large" or "small" proportion of missing values, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.1`.
-
-```{r na_prop_compute}
-na_prop <- numeric()
-
-for (i in unique(ped_data$Sensor)){
-  na_prop[i] <- sum(is.na(dfa[, i])) / n_obs
-}
-
-na_prop_pre <- numeric()
-
-for (i in unique(ped_data$Sensor)){
-  na_prop_pre[i] <- sum(is.na(ori_2[, i])) / n_obs
-}
-
-large_miss_sensors <- names(na_prop[na_prop > threshold])
-small_miss_sensors <- setdiff(names(na_prop), large_miss_sensors)
-```
-
-```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
-The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data.
-This emphasises the necessity to correct for false zero counts as it will affect "}
-theme_g <- theme_minimal() +
-           theme(plot.title = element_text(hjust = 0.5,
-                                           size = 10))
-p1 <- ggplot() + geom_histogram(aes(x = na_prop*100), bins = 12) +
-          xlab("Percentage of Missing Values (%)") +
-          ylab("Number of Sensors") + theme_g +
-          geom_vline(xintercept = 10, size = 2,
-                     colour = 'red') +
-          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
-                    colour = "white", fill = 'black', alpha = 0.4) +
-          scale_y_continuous(limits = c(0,35))
-p2 <- ggplot() + geom_histogram(aes(x = na_prop_pre*100), bins = 12) +
-          xlab("Percentage of Missing Values (%)") +
-          ylab("Number of Sensors") + theme_g +
-          geom_vline(xintercept = 10, size = 2,
-                     colour = 'red') +
-          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
-                    colour = "white", fill = 'black', alpha = 0.4) +
-          scale_y_continuous(limits = c(0,35))
-grid.arrange(p2, p1, nrow = 1)
-```
-
-Because the models for the sensors with a large proportion of missing values rely on having neighbouring sensors having complete cases (no missing values), we want to impute these values first.
-This will then allow us to have as much information to be available for the models which use neighbouring sensors to to train on.
-
-For sensors with a small proportion of missing values, we use a GLM quasipoisson model. To improve on the previous model used, we use the model:
-$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
-
-$$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$
-
-```{r alg_glm}
-models <- list()
-fitted_data <- list()
-dfa2 <- dfa
-p <- progress_estimated(length(unique(ped_data$Sensor)))
-for(i in small_miss_sensors){
-  p$tick()$print()
-  options(na.action = na.omit)
-  # if(max(na_length_check$lengths) > max_miss_hrs) 
-  # {
-  #   model <- NULL
-  # }
-
-  sensor_df <- unlist(dfa[, i])
-
-  model <- glm(sensor_df ~  Month + DayType*Time, family = 'quasipoisson', data = dfa)
-  models[[i]] <- (stripGlmLR(model))
-}
-```
-
-```{r alg_glm_impute}
-  options(na.action = na.pass)
-for(i in small_miss_sensors){
-  ts_dat <- data.frame(Month = dfa$Month, DayType = dfa$DayType, Time = dfa$Time)
-  fitted_data[[i]] <- predict.glm(models[[i]],newdata = ts_dat, type = "response")
-}
-  options(na.action = na.omit)
-  for(i in small_miss_sensors){
-    dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
-    }
-
-```
-
-Instead of `HDay`, we use `DayType` which classifies Tuesday, Wednesday and Thursday as a single factor level, Midweek.
-We do this as the daily patterns on these weekdays are similar.
-This reduces the number of parameters to estimate by 46.
-A result of this is we also have more robust estimates for the midweek days, being less sensitive to outliers.
-Because this model will be estimated for sensor locations with few missing values,  we have enough data to add in month as an additive effect.
-This will help capture the annual seasonality.
-
-Using these estimated models, we impute the missing values to produce complete data at all these sensor locations.
-For the locations with high proportions of missing values can be predicted using the now complete data from all the locations which had a small proportion of missing values.
-Using a threshold of 10%, this gives us `r length(small_miss_sensors)` sensor locations to use.
-
-```{r, fig.width=3, fig.height=3, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
-ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
-lats <- ped_loc$Latitude
-lons <- ped_loc$Longitude
-ids <- ped_loc$`Sensor Description`
-
-{
-ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
-ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
-ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
-ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
-ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
-ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
-ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
-ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
-ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
-}
-loc <- data.frame(lons, lats, ids)
-rownames(loc) <- ids
-loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
-                                  "Small", "Large"))
-rownames(loc) <- loc$ids
-
-melb3 <- get_map(location = c(loc["Southbank",]$lons,
-                              loc["Southbank",]$lats),
-                 zoom = 17, maptype = "roadmap")
-
-sens_lines_lon <- as.numeric(c(loc["Flinders St Station Underpass",]$lons,
-                loc["Southbank",]$lons,
-                loc["Sandridge Bridge",]$lons))
-
-sens_lines_lat <- as.numeric(c(loc["Flinders St Station Underpass",]$lats,
-                loc["Southbank",]$lats,
-                loc["Sandridge Bridge",]$lats))
-
-sens_lines <- data.frame(lon = sens_lines_lon,
-                         lat = sens_lines_lat)
-options(na.action = 'na.omit')
-ggmap(melb3) + geom_line(aes(x = lon, y = lat),
-                        data = sens_lines, size = 2,
-                        colour = viridis(3)[2], alpha = 0.5) +
-               geom_point(aes(x = lons, y = lats, colour = cl),
-                          size = 6,
-                          data = loc) +
-               scale_colour_viridis(discrete = TRUE,
-                                    name = "Sensor NA Prop.",
-                                    begin = 0.5, end = 0) +
-               geom_point(aes(x = loc["Southbank",]$lons,
-                              y = loc["Southbank",]$lats),
-                          size = 12, colour = viridis(3)[2]) +
-               geom_label(aes(label = ids, x = lons, y = lats,
-                              fill = cl),
-                         fontface = 'bold', colour = 'white',
-                         data = loc, size = 3,
-                         nudge_y = -0.000325) +
-              scale_fill_viridis(discrete = TRUE,
-                                  name = "Sensor NA Prop.",
-                                  begin = 0.5, end = 0) +
-              xlab(NULL) + ylab(NULL) +
-              theme(plot.title = element_text(hjust = 0.5),
-                    plot.caption = element_text(hjust = 0.5),
-                    axis.text.x = element_blank(),
-                    axis.text.y = element_blank(),
-                    axis.ticks = element_blank(),
-                    legend.position = c(.8,0.2))
-```
-
-For each of the sensor locations with high proportions of missings, we need to decide which sensor to use for prediction.
-A simple method of finding a geographical closest neighbour is to take the haversine/great-circle distance between the sensor to be imputed and all the possible candidates.
-```{r, message=F}
-# Calculates the geodesic distance between two points specified by radian latitude/longitude using the
-# Spherical Law of Cosines (slc)
-# source: https://www.r-bloggers.com/great-circle-distance-calculations-in-r/
-gcd.slc <- function(long1, lat1, long2, lat2) {
-  R <- 6371 # Earth mean radius [km]
-  d <- acos(sin(lat1)*sin(lat2) + cos(lat1)*cos(lat2) * cos(long2-long1)) * R
-  return(d) # Distance in km
-}
-
-## name fix dictionary
-{
-ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
-ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
-ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
-ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
-ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
-ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
-ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
-ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
-ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
-}
-loc <- data.frame(lons, lats, ids)
-rownames(loc) <- ids
-```
-
-```{r alg_lm}
-for (i in large_miss_sensors){
-  # p$tick()$print()
-  sensor_dists <- numeric()
-  for (j in small_miss_sensors){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df_scale <- dfa2[, closest_sensor[1]] %>% unlist %>% scale
-  close_df_scale2 <- dfa2[, closest_sensor[2]] %>% unlist %>% scale
-
-  sensor_df <- dfa2[, i] %>% unlist
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa$Time)
-  options(na.action = na.omit)
-  model <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  models[[i]] <- stripGlmLR(model)
-
-  options(na.action = na.pass)
-  fitted_data[[i]] <- predict(model, newdata = tr_dat, type = "response")
-  options(na.action = na.omit)
-
-  dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
-
-}
-
-```
-
-Using the two geographically closest sensors with complete cases, we use another GLM specified as:
-$$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{neighbours}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}}$$
-$$\text{where } \quad  \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}} \sim \text{Time} \times \text{HourlyCounts}_{\text{neighbour 1}}^{SC} + \text{Time} \times \text{HourlyCounts}_{\text{neighbour 2}}^{SC}$$
-
-We would expect that the amount of people passing through the neighbouring sensors will be a strong predictor of the pedestrian counts as they are likely to also pass through.
-Using geographical neighbours, we can capture rare events effectively.
-Any large shocks to the counts at the neighbouring sensors will have an effect on the counts to be imputed.
-
-```{r, echo = FALSE, message=FALSE, warning=F, include=FALSE}
-i <- "Southbank"
-  sensor_dists <- numeric()
-  for (j in small_miss_sensors){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  dfa3 <- original_df %>% filter(Date_Time > "2017-01-01")
-  close_df <- dfa3[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa3[, closest_sensor[2]] %>% unlist
-
-  close_dfsc <- scale(close_df)
-  close_dfsc2 <- scale(close_df2)
-
-  sensor_df <- dfa3[, i] %>% unlist
-  sensor_dfsc <- scale(sensor_df)
-
-  tr_dat <- data.frame(sensor_df, close_df, close_df2,
-                       Time = as.integer(dfa3$Time),
-                       Date_Time = dfa3$Date_Time,
-                       HDay = dfa3$HDay) %>%
-            gather(key = Sensor, value = Count, sensor_df:close_df2)
-
-  tr_dat_sc <- data.frame(sensor_dfsc, close_dfsc, close_dfsc2,
-                       Time = as.integer(dfa3$Time),
-                       Date_Time = dfa3$Date_Time,
-                       HDay = dfa3$HDay) %>%
-            gather(key = Sensor, value = Count, sensor_dfsc:close_dfsc2)
-
-p1 <- ggplot(tr_dat %>% filter(Date_Time < "2017-03-02",
-                         Date_Time > "2017-02-26")) +
-      geom_line(aes(x = Date_Time, y = Count,
-                    colour = Sensor), size = 2) +
-      scale_colour_viridis(discrete = TRUE,
-                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
-                                      "Neighbour 2 (Sandridge Bridge)",
-                                      "Target (Southbank)")) +
-      xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
-      theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
-
-p2 <- ggplot(tr_dat_sc %>% filter(Date_Time < "2017-03-02",
-                         Date_Time > "2017-02-26")) +
-      geom_line(aes(x = Date_Time, y = Count,
-                    colour = Sensor), size = 2) +
-      scale_colour_viridis(discrete = TRUE,
-                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
-                                      "Neighbour 2 (Sandridge Bridge)",
-                                      "Target (Southbank)")) +
-      xlab("Time") + ggtitle("Scaled/Stadardised Neighbours") +
-      ylab("Standard Deviations") + theme_minimal() +
-      theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
-```
-
-```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 8, fig.height = 4}
-grid.arrange(p1, p2, ncol = 1)
-```
-
-In order to use a mixture of the patterns from two different sensors, we need to use the scaled counts at the neighbouring sensors to avoid the magnitude of the counts having an effect on the predicted counts.
-We scale the counts by standardising to $\mu = 0, \sigma^2 = 1$.
-Because the covariance with neighbouring sensors may vary over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
-
-## Implementation of improved imputation algorithm
-
-The exact code used for this paper can be found on http://github.com/gavinchin/honours2017
-
-### False zero-counts correction
-
-The false zero-counts correction is performed in R using the run length encoding function, `rle()`.
-This function computes the length and values of runs of equal values in a vector.
-Unfortunately, this function does not recognise `NA` in the its computation, so all NA values are initially recoded to take the value 0 at each sensor.
-This results in all missing values and true zero counts to take the value zero when running `rle()`.
-With the returned run lengths, all sequences of repeated values longer than the defined length of 6 hours are replaced with `NA`.
-Any values which were also originally missing, but were not missing for longer than 6 hours (and hence are still coded as 0), are recoded to be `NA`.
-It must be noted that sequences of repeated values not equal to zero which run for longer that 6 hours would be incorrectly replaced with `NA`.
-This is highly unlikely, however, as 6 hours of the exact same pedestrian counts would be extremely rare.
-
-### Caculation of proportion of missing values and classification of sensors
-
-In the working data, a wide format is used.
-This makes it easier to calculate missing value proportion at each sensor location, as it is the proportion of missing values in the column for each sensor.
-Using `is.na()` returns a logical vector indicating whether the sensor count is missing/`NA` or not.
-The sum of the instances where `is.na = TRUE` divided by the hours in the period of 2014-2016 is computed to get the proportion of missing values.
-
-These proportions are saved for each sensor and then classified as either "large" or "small" based on a threshold value for the proportion.
-A list of sensors with large proportions of missing data and a list of sensors with small proportions of missing data is generated.
-
-### Estimation of GLM at sensor locations with a small proportion of missing values
-
-A for loop is used to estimate the GLM specified:
-```{r, echo = TRUE, eval = FALSE}
-glm(dfa[, i] ~ Month + DayType*Time, data = dfa,
-    family = quasipoisson())
-```
-where `dfa` is the object name given to the data frame containing the training data after the false zero-count correction, and `i` is the $i^{th{}}$ sensor in the list of sensors with a small proportion of missing values.
-These models are stored in a list, forming a list of models.
-An alternative approach to using a for loop in R is splitting the data into separate data frames for each sensor location (forming a list of data frames) and applying the `glm()` function on each data frame in the list using the function `purrr::map()`.
-However, when this was attempted, performance issues were encountered due to the large size of the estimated model objects in R.
-This was not an issue when a for loop was used.
-Additional code for removing unnecessary elements of the GLM model object before storing to the list of models, reduucing the memory usage.
-
-### Imputation at sensor locations with a small proportion of missing values
-
-The predicted values are obtained using `predict.glm()`, using the option `type = "response"` to have predicted values returned predicted counts.
-The default option returns the predicted values of $\hat{\lambda}$ in the quasipoisson process, not the counts, $\exp (\hat{\lambda})$.
-Missing values are then replaced using the imputed values, generating a data frame of complete data as sensor locations for with a small proportion of missing values to be used for prediction at sensors with a large proportion of missing values.
-
-### Estimation of GLM at sensor locations with a large proportion of missing values
-The following steps are performed on each sensor with a large proportion of missing values inside a for loop.
-
-##### Finding neighbouring sensors
-Using location data of the sensors provided by the City of Melbourne, a data frame containing the latitude and longitude coordinates of all the sensors which were classified as having a small proportion of missing values (and now have imputed, complete data), which are the candidates to be neighbours used for prediction.
-Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()`, an approximation of distance between each of the candidate sensors is calculated in and stored as a vector.
-Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.
-
-## Imputation Results
-
-Firstly, we evaluate the initial model used to impute:
-$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
-
-Note, this is also after the adjustments made to treat suspiciously long sequences of 0 values as `NA` in the actual counts.
-The simple GLM model is trained before this correction, while the improved model is trained on the adjusted data.
-
-For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station.
-At this sensor, the proportion of missing values (adjusted) is `r round(na_prop["Southern Cross Station"]*100,2)`Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).
-
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
-p1 <- ggplot(data = NULL, aes(x = original_df %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Southern Cross Station"],
-           y = fitted_base$`Southern Cross Station`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Southern Cross Station")
-p2 <- ggplot(data = NULL, aes(x = original_df %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Southern Cross Station"],
-           y = fitted_data$`Southern Cross Station`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Southern Cross Station")
-grid.arrange(p1, p2, nrow = 1)
-```
-
-Visually, we can see that the fitted values are good estimates as the relationship between the actuals vs fitted is very close to 1:1.
-This is seen when we perform a least squares estimate of $\widehat{Fitted}_t = \hat{\beta}_0 + \hat{\beta}_1 Actual_t$, and the estimated $\hat{\beta}_0$ is close to 0 and $\hat{\beta}_1$ is close to 1.
-
-Here we see that the GLM model with only time and date based variables work well for fitting small periods of missing data.
-
-However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor.
-We use the sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%.
-```{r}
-options(na.action = na.omit)
-p1 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Australia on Collins"])) +
-           geom_point(aes(y = fitted_base$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                      alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(aes(y = fitted_base$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                       method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + theme(aspect.ratio = 1,
-                                   plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Australia on Collins")
-p2 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Australia on Collins"])) +
-           geom_point(aes(y = fitted_data$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                      alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(aes(y = fitted_data$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                       method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + theme(aspect.ratio = 1,
-                                   plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Australia on Collins")
-```
-
-```{r, fig.height=4, fig.width=9, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}
-
-grid.arrange(p1, p2, nrow = 1)
-
-```
-
-We can see that the simple model has bias caused by the false zero counts, resulting in the predicted values to be underestimated.
-This is shown by the slope of the fitted line on the actual counts against predicted counts being $<1$, emphasising the importance of the first step taken in the algorithm of the improved model to class the false zero counts as missing values.
-
-The criterion we will use to measure goodness of fit is MARE. It is calculated by:
-
-$$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \vert \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \vert}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$
-
-We calculate a seperate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data).
-This will help to identify overfitting of the data.
-We use MARE as it will be a comparable measure between sensor locations as well as comparable between models.
-
-```{r}
-MARE_base_glm_tr <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_base_glm_tr[i] <- mean(abs(fitted_base[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
-}
-
-MARE_tr <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_tr[i] <- mean(abs(fitted_data[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
-}
-
-mare_df <- data.frame(simple = MARE_base_glm_tr, improved = MARE_tr, ids) %>%
-           gather(model, mare, simple:improved)
-p1 <- ggplot(mare_df) + geom_boxplot(aes(x = model, y = mare, fill = model),
-                                     width = .5, alpha = 0.5) +
-      labs(x = "Model", y = "In-Sample MARE") +
-      scale_fill_viridis(discrete = TRUE, name = "Model")
-
-p2 <- ggplot(mare_df) + geom_density(aes(x = mare, fill = model,
-                                         colour = model), adjust = 1.3,
-                                     alpha = 0.5) +
-      scale_fill_viridis(discrete = TRUE) +
-      scale_colour_viridis(discrete = TRUE) +
-      theme_minimal() + theme(plot.title = element_text(hjust = 0.5),
-                              legend.position = "none") +
-      labs(x = "In-sample MARE")
-p3 <- ggplot() + geom_point(aes(x = MARE_base_glm_tr, y = MARE_tr,
-                                colour = (MARE_tr<MARE_base_glm_tr))) +
-      geom_abline(aes(slope = 1, intercept = 0)) +
-      scale_colour_viridis(discrete = TRUE, option = "C",
-                           begin = 0.15, end = 0.85,
-                           direction = -1,
-                           name = "Improved Model is better") +
-      theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
-      theme(aspect.ratio = 1, legend.position = "bottom",
-            legend.direction = "vertical")
-```
-
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}
-
-grid.arrange(p1, p2, p3, nrow = 1)
-
-```
-
-While it is not very clear when comparing the distribution of in-sample MARE between imputation models, when we compare the in-sample MARE at each location we can see that the majority of locations had better in-sample fit with the improved model using the algorithm.
-We do see that some locations did have worse in-sample fit, but overall we see improvement with `r sum(MARE_base_glm_tr > MARE_tr)` having better fit with the improved model over the simple model.
-
-Focusing on the date 04/07/2014, which has 3 hours of missing data at Southern Cross Station:
-
-```{r, fig.height = 2, fig.width = 5, fig.align='center', fig.cap="A plot of the pedestrian counts at Southern Cross Station.
-The black line is the observed counts, where the yellow line is the imputed counts."}
-fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
-                           fit_ = dfa2$`Southern Cross Station`) %>%
-                  filter(Date_Time < "2014-07-06 00:00:00",
-                         Date_Time > "2014-07-04 12:00:00") %>%
-                        .$fit_
-
-ggplot(original_df %>%  filter(Date_Time < "2014-07-06 00:00:00",
-                               Date_Time > "2014-07-04 12:00:00")) +
-  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = viridis(3)[3],
-            size = 2) +
-  geom_line(aes(x = Date_Time, y = `Southern Cross Station`),
-            size  = 2) +
-  theme_minimal() + xlab("Time") +
-  ylab("Pedestrian Counts") +
-  ggtitle("Pedestrian Counts at Southern Cross Station \n on 04/07/2014 - 05/07/2014") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-More importantly, we can evaluate the performance of the model on the out of sample observations (2017 data) as cross validation.
-Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, we need to again apply the correction by long sequences of repeated values to be able to properly evaluate the goodness of fit of out-of-sample observations.
-The method used for this is identical to that used on the training set in the algorithm, replacing sequences of length 6 hours or longer with `NA`.
-
-```{r ts_fix0}
-for(i in unique(ped_data$Sensor)){
-    ori_ts[is.na(ori_ts[, i]), i] <- 0
-    na_length_check <- rle(as.numeric(unlist(ori_ts[,i])))
-    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
-    rl <- ori_ts[, i] %>% unlist %>% rle()
-    ori_ts[maybeNA, i] <- NA
-    defsNA <- is.na(ori_ts2[, i])
-    ori_ts[defsNA, i] <- NA
-  }
-```
-
-```{r test_fit}
-test_base <- list()
-test_fit <- list()
-ts <- original_df %>% filter(Date_Time > "2016-12-31")
-
- for(i in unique(ped_data$Sensor)){
-    test_base[[i]] <- predict.glm(models_base[[i]], newdata = ts,
-                                  type = "response")
- }
-
- for(i in small_miss_sensors){
-   test_fit[[i]]  <- predict.glm(models[[i]], newdata = ts,
-                                 type = "response")
- }
- for(i in large_miss_sensors){
-   sensor_dists <- numeric()
-  for (j in small_miss_sensors){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df_scale <- ts[, closest_sensor[1]] %>% unlist %>% scale
-  close_df_scale2 <- ts[, closest_sensor[2]] %>% unlist %>% scale
-
-  ts_sc <- data.frame(close_df_scale, close_df_scale2, Time = ts$Time)
-
-  test_fit[[i]] <- predict.glm(models[[i]], newdata = ts_sc,
-                               type = "response")
-
- }
-
-MARE_base_ts <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
-}
-
-MARE_ts <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
-}
-```
-
-Again, we start by comparing the fit at a sensor location with a small proportion of missing values, Melbourne Central:
-
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
-p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
-           y = test_base$`Melbourne Central`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Melbourne Central")
-p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
-           y = test_fit$`Melbourne Central`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
-grid.arrange(p1, p2, nrow = 1)
-```
-
-We find that the out-of-sample predictions of at Melbourne Central by the simple model has a tendency to underestimate the pedestrian counts.
-By comparison, the predictions made by the improved imputation model does not underestimate the counts as much.
-Comparing the out-of-sample MARE between models at Melbourne Central, MARE decreases from `r round(MARE_base_ts["Melbourne Central"]*100, 2)`% to `r round(MARE_ts["Melbourne Central"]*100, 2)`% with the use of the improved model.
-
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
-p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Collins Place (South)"],
-           y = test_base$`Collins Place (South)`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 2500)) +
-           scale_y_continuous(limits = c(0, 2500)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Collins Place (South)")
-p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Collins Place (South)"],
-           y = test_fit$`Collins Place (South)`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 2500)) +
-           scale_y_continuous(limits = c(0, 2500)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Collins Place (South)")
-grid.arrange(p1, p2, nrow = 1)
-```
-
-At a sensor location with a large proportion of missing values, Collins Place (South), we again see a tendency to underestimate counts with the simple imputation model.
-The improved imputation method reduces the underestimation of the predictions and provides better predictions.
-The MARE is greatly improved, where it decreases from `r round(MARE_base_ts["Collins Place (South)"]*100, 2)`% to `r round(MARE_ts["Collins Place (South)"]*100, 2)`% with the use of the improved model.
-
-```{r}
-MARE_base_ts2 <- MARE_base_ts[setdiff(unique(ped_data$Sensor),
-                                      c("Spencer St-Collins St (South)",
-                                        "Spencer St-Collins St (North)",
-                                        "Flinders St-Swanston St (West)"))]
-MARE_ts2 <- MARE_ts[setdiff(unique(ped_data$Sensor),
-                                      c("Spencer St-Collins St (South)",
-                                        "Spencer St-Collins St (North)",
-                                        "Flinders St-Swanston St (West)"))]
-
-mare_df <- data.frame(simple = MARE_base_ts2, improved = MARE_ts2) %>%
-           gather(model, mare, simple:improved)
-
-p1 <- ggplot(mare_df) + geom_boxplot(aes(x = model, y = mare, fill = model),
-                                     width = .5, alpha = 0.5) +
-      labs(x = "Model", y = "Out-of-sample MARE") +
-      scale_fill_viridis(discrete = TRUE, name = "Model")
-
-p2 <- ggplot(mare_df) + geom_density(aes(x = mare, fill = model,
-                                         colour = model), adjust = 1.3,
-                                     alpha = 0.5) +
-      scale_fill_viridis(discrete = TRUE) +
-      scale_colour_viridis(discrete = TRUE) +
-      theme_minimal() + theme(plot.title = element_text(hjust = 0.5),
-                              legend.position = "none") +
-      labs(x = "Out-of-sample MARE")
-p3 <- ggplot() + geom_point(aes(x = MARE_base_ts2, y = MARE_ts2,
-                                colour = (MARE_ts2<MARE_base_ts2))) +
-      geom_abline(aes(slope = 1, intercept = 0)) +
-      scale_colour_viridis(discrete = TRUE, option = "C",
-                           begin = 0.15, end = 0.85,
-                           direction = -1,
-                           name = "Improved Model is better") +
-      scale_x_continuous(limits = c(0,1)) +
-      scale_y_continuous(limits = c(0,1)) +
-      theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
-      theme(aspect.ratio = 1, legend.position = "bottom",
-            legend.direction = "vertical")
-```
-
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}
-
-grid.arrange(p1, p2, p3, nrow = 1)
-
-```
-
-Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) was omitted from the the plots above as the out-of-sample MARE is extremely high at `r round(MARE_ts["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_ts["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_ts["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
-
-As these sensor locations were both classified as locations with a large proportion of missing values, it would appear that the relationship between these locations and their neighbouring sensors has changed.
-This is because the in-saple MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_tr["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
-
-```{r, fig.width = 8, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
-ggplot() + geom_smooth(aes(x = as.integer(Time),
-                           y = `Spencer St-Collins St (South)`),
-                       colour = viridis(3)[1],
-                       data = dfa2) +
-           geom_smooth(aes(x = as.integer(Time),
-                           y = `Spencer St-Collins St (South)`),
-                       colour = viridis(3)[3],
-                       data = ori_ts2) +
-           theme_minimal() + labs(x = "Time of Day") +
-           facet_wrap(~ DayType)
-```
-
-The plot of the average counts by time of day and type of day and data set shows that this location saw an increase in counts on weekdays which did not occur at the neghbouring sensors.
-We also see that the pattern is slightly different during weekdays, with a higher afternoon/evening peak.
-These types of change in pedestrian traffic behaviour is hard to capture in our imputation model, especially when its neighbours did not have a similar change.
-While we can add a year variable to allow for year to year growth, the short period of training data doesn't make this appropriate at this stage.
-
-Another way we can perform cross-validation is to remove data at a sensor, and evaulate the predictions made.
-This will allow us to evaluate the imputation model's performance within sample, which is the primary concern with them imputation process.
-We select a sensor which has a small proportion of missing values, Bourke Street Mall (South), which has `r round((1 - na_prop["Bourke Street Mall (South)"])*100, 3)`% of data available.
-As observed in the data, missing data is present in different ways: random, short periods of missing data (no more that 6 hours) or long periods of missing data (multiple months).
-We will simulate both types of missingness at this sensor location.
-
-First, data is randomly removed such that 20% of the training data is missing, then the imputation model is estimated.
-Using the estimated model, the MARE within the training data is calculated (in-sample MARE).
-```{r}
-miss_idx <- sample(1:nrow(dfa2), round(nrow(dfa2)*0.2))
-
-mall_missing_1 <- dfa2$`Bourke Street Mall (South)`
-mall_missing_1[miss_idx] <- NA
-
-i <- "Bourke Street Mall (South)"
-sensor_dists <- numeric()
-  for (j in small_miss_sensors[-6]){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors[-6]
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df <- dfa2[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
-
-  close_df_scale <- scale(close_df)
-  close_df_scale2 <- scale(close_df2)
-
-  sensor_df <- mall_missing_1 %>% unlist
-  sensor_dfsc <- scale(sensor_df)
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
-
-  options(na.action = na.omit)
-  mall_model_1 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  options(na.action = na.pass)
-  mall_fit_1 <- predict.glm(mall_model_1, newdata = tr_dat,
-                            type = "response")
-  options(na.action = na.omit)
-
-  mall_mare_1 <- mean(abs(mall_fit_1 - unlist(dfa2$`Bourke Street Mall (South)`)),
-                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
-                                            na.rm = T)
-
-```
-
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South).
-20% of the observations are removed from the traning data for the imputation model at random."}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1),
-                     size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date") +
-          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
-```
-
-With random hours of missing data, the predictions made by using neighbouring sensors  are estimated with good accuracy with a low MARE of `r round(mall_mare_1*100, 2)`%.
-Unfortunately, this type of missingness is not consistent with what is generally observed at most sensors with large proportions of missing data.
-
-Instead, a more meaningful method is to remove 20% of the data in a single period to simulate what is typically observed.
-A random date is taken and then 2629 hours of data is removed before and after this date (10% before, 10% after).
-The imputation model is estimated again in order to calculate a MARE.
-
-```{r}
-miss_seed <- max(sample(1:nrow(dfa2),1) - 2629, 0)
-miss_idx <- seq(miss_seed, miss_seed + 5258)
-
-mall_missing_2 <- dfa2$`Bourke Street Mall (South)`
-mall_missing_2[miss_idx] <- NA
-
-i <- "Bourke Street Mall (South)"
-sensor_dists <- numeric()
-  for (j in small_miss_sensors[-6]){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors[-6]
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df <- dfa2[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
-
-  close_df_scale <- scale(close_df)
-  close_df_scale2 <- scale(close_df2)
-
-  sensor_df <- mall_missing_2 %>% unlist
-  sensor_dfsc <- scale(sensor_df)
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
-
-  options(na.action = na.omit)
-  mall_model_2 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  options(na.action = na.pass)
-  mall_fit_2 <- predict.glm(mall_model_2, newdata = tr_dat,
-                            type = "response")
-  options(na.action = na.omit)
-
-  mall_mare_2 <- mean(abs(mall_fit_2 - unlist(dfa2$`Bourke Street Mall (South)`)),
-                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
-                                            na.rm = T)
-
-```
-
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence.
-This method of simulating missing data is a better representation of the missing data observed at sensors with a large proportion of missing data compared to randomly removing hourly observations from the training data."}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2),
-                size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date") +
-          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
-```
-
-The MARE of the estimated model is `r round(mall_mare_2*100, 2)`%, which is similar to what was obtained when the missing values were random.
-This demonstrates that the model quite well when we remove 20% of the observations from the training data.
-
-```{r}
-miss_seed <- min(max(sample(1:nrow(dfa2),1) - 6573, 0), 13146)
-miss_idx <- seq(miss_seed, miss_seed + 13146)
-
-mall_missing_3 <- dfa2$`Bourke Street Mall (South)`
-mall_missing_3[miss_idx] <- NA
-
-i <- "Bourke Street Mall (South)"
-sensor_dists <- numeric()
-  for (j in small_miss_sensors[-6]){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors[-6]
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df <- dfa2[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
-
-  close_df_scale <- scale(close_df)
-  close_df_scale2 <- scale(close_df2)
-
-  sensor_df <- mall_missing_3 %>% unlist
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
-
-  options(na.action = na.omit)
-  mall_model_3 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  options(na.action = na.pass)
-  mall_fit_3 <- predict.glm(mall_model_3, newdata = tr_dat,
-                            type = "response")
-  options(na.action = na.omit)
-
-  mall_mare_3 <- mean(abs(mall_fit_3 - unlist(dfa2$`Bourke Street Mall (South)`)),
-                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
-                                            na.rm = T)
-
-```
-
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence.
-A large proportion of the training data is removed to test the robustness of the imputation method used at sensors with extremely large proportions of missing data during the training period defined."}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_3),
-                size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date") +
-          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
-```
-
-Removing 50% of the data using the same method, it is found that the in-sample MARE is not affected significantly by the larger proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
-
-# Predictive Model
-
-For the purpose of prediction, the use of the most simplistic model possible to make quick predictions of the expected pedestrian counts at each location, accessible to the public.
-The model is not expected to be able to predict large deviations from time and date based estimates.
-
-Similar to the simple imputation model used,  a generalised linear model with quasi-poisson errors is estimated using time and date based variables as predictors. The model specification used is:
-
-$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
-
-$$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$
-
-```{r, echo = FALSE}
-library(purrr)
-library(dplyr)
-library(tidyr)
-library(lubridate)
-library(roxygen2)
-load("../data/model_est.RData")
-
-daytype <- function(date){
-  if(!is.Date(date)){stop("Object is not a date.")}
-  x_day <- wday(date, label = TRUE, abbr = FALSE)
-  x_daytype <- ifelse(x_day == "Sunday", "Sunday",
-                ifelse(x_day == "Saturday", "Saturday",
-                ifelse(x_day == "Monday", "Monday",
-                ifelse(x_day == "Friday", "Friday",
-                          "Midweek"))))
-  return(x_daytype)
-}
-#' Predicted pedestrian counts at various sensor locations in Melbourne
-#'
-#' @param pred_date input date
-#' @param t_hour input integer
-#' @param is_pub_hol input logical
-#' @return data frame of the fitted values at each location in wide format, with time and date
-#' @export
-#' @examples
-#' ped_predict()
-#' ped_predict("2017-12-25", 18, TRUE)
-#' 
-ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
-                        t_hour = 12,
-                        is_pub_hol = FALSE,
-                        long = FALSE) {
-  if(!(t_hour %in% seq(0, 23))){
-    stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
-  pred_date <- ymd(pred_date)
-  if(!is.Date(pred_date)){
-    stop(paste("Date given to predict cannot be parsed as a date. Use the date format 'yyyy-mm-dd'. Example: 25 January 2018 is '2018-01-25'."))
-    }
-
-  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
-               mutate(DayType = ifelse(is_pub_hol,
-                                       "Holiday",
-                                       daytype(pred_date)),
-                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
-                      Date = pred_date,
-                      Date_Time = ymd_h(paste(Date, t_hour, sep = " ")))
-
-  preds <- numeric()
-  for(i in names(pred_model)){
-    preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
-  }
-  if(long){
-    return(preds)
-  }
-    else
-  preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
-    spread(Sensor_ID, fit) %>% cbind(date_pred) %>% as_tibble
-
-  return(preds_df)
-}
-
-ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),is_pub_hol = FALSE){
-  day_df <- data.frame()
-      for(t in 0:23){
-        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
-        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
-      }
-  return(as.tibble(day_df))
-}
-
-```
-
-Using a GLM predicts well compared to estimating with other models such as AR($p$) or other time series models when considering the tradeoff between model complexity and prediction accuracy.
-Due to multiple seasonalities, AR($p$) models would need to be considered on a hour of the week basis at each sensor location.
-Prediction with time series models would also require of historical data.
-For the objective of this predictive model, the additional imputs required by end users to predict pedestrian counts when making out of sample predictions is not desirable.
-
-```{r, fig.width=6, fig.height=4, fig.cap = paste("Plot of daily pedestrian counts series at Melbourne Central by DayType and Month.
-Month of the year is mapped to colour to help identify annual seasonal patterns, where month has an additive effect on the pedestrian counts.
-At Melbourne Central, different months can be seen to have different mean counts for a given type of day and hour of the day.")}
-ggplot(dfa2 %>%
-         mutate(Date = date(Date_Time))) +
-  geom_path(aes(x = as.integer(Time), y = `Melbourne Central`,
-                group = Date, colour = Month),
-            alpha = 0.1) +
-  scale_colour_viridis(discrete = TRUE, option = "C") +
-  facet_wrap(~ DayType) + theme_minimal() +
-  theme(legend.position = "none") +
-  labs(x = "Time of Day", y = "Pedestrian Counts \n at Melbourne Central")
-```
-
-A plot of the daily pedestrian counts series at Melbourne Central by type of day and month show that time and date variables as deterministic predictors adequately explains most variation in the pedestrian counts.
-It can be seen that month has an additive effect as the pattern of the hourly counts for a given date are not affected by month.
-The six different factor levels that `DayType`, the type of day, is seen to be a sufficiently explain six different patterns.
-The type of day, Midweek, shows evidence that no additional information is gained from identifying the day to be Tuesday, Wednesday or Thursday.
-
-```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM for the interaction term $\text{DayType} \times \text{Time}$ for each hour at each sensor.
-Variables (x-axis) is ordered by the scagnostic measure, Outlying.
-"), include=FALSE, eval = FALSE}
-dfcoeff <- data.frame(terms = factor(), coeff = numeric(), sensor = factor())
-for(i in names(pred_model)){
-  terms <- names(coef(pred_model[[i]]))
-  coeff <- coef(pred_model[[i]])
-  sensor <- rep(i, length(terms))
-  df_b <- data.frame(terms, coeff, sensor)
-  dfcoeff <- bind_rows(dfcoeff, df_b)
-}
-dfcoeff <- dfcoeff %>% spread(terms, coeff)
-
-library(GGally)
-library(scagnostics)
-ggparcoord(dfcoeff[, 51:74], scale = "std", 
-           order = "Clumpy", alphaLines = 0.3) + theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, size = 6)) +
-  labs(x = "Parameter")
-```
-
-When estimating the GLM at each sensor location with R, only the necessary elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function are saved to file.
-Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets.
-This is due to the model object providing redundancies.
-In particular, elements which are unnecessarily saved for the purposes of the predictive model: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
-
-The result of removing all the unnecessary redundant data elements in the GLM object is a storage size reduction from `r mall_model_2 %>% object.size() %>% format(units = "auto")` to `r mall_model_2 %>% stripGlmLR() %>% object.size() %>% format(units = "auto")`.
-This is only `r round((mall_model_2 %>% stripGlmLR() %>% object.size() %>% as.numeric/ mall_model_2 %>% object.size() %>% as.numeric)*100, 2)`% of the size of the original GLM object.
-With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.
-
-The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves.
-Using the estimated model parameters, a new function is written to facilitate the predictions in an easy to use format.
-
-```{r, echo = TRUE, eval = FALSE}
-ped_predict(pred_date = "2017-12-26", t_hour = 13, is_pub_hol = TRUE)
-```
-This function will return tibble (trimmed down version of `data.frame()` in R) containing the predicted counts at each sensor location for the 13th hour (13:00/1:00 PM) on Boxing Day (26 December) 2017.
-This tibble can then be easily used for visualisations or analysis.
-Due to the lack of data on future public holiday dates, the function assumes the date given is not a public holiday unless `is_pub_hol = TRUE` is provided.
-Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predict.
-
-Using these function as a basis, visualisations of pedestrian counts can be easily generated in R:
-```{r, echo = TRUE, fig.cap = "Prediction for 21/05/2018 in R using `ped_predict_day()`, then plotting the time series using `ggplot2`.
-This shows the user friendly nature of the function, where the function returns predictions in a tidy, wide format.
-No data manipulation is required in order to plot the predicted counts.", fig.height=2}
-ped_predict_day("2018-05-21") %>% 
-  ggplot() + geom_line(aes(x = Date_Time, y = Southbank))
-```
-
-# Conclusion  
-Analysis of the Melbourne CBD pedestrian data made publicly available by the City of Melbourne has revealed issues with the dataset with relation to missing values in particular.
-A major issue which was found when using data sourced from the City of Melbourne's official pedestrian data visualisation (also accessible using `rwalkr::walk_melb()`) is potential for false zero-counts where missing values were given the value of zero.
-As outliers have major implications on analysis, a correction for the false counts in the data was proposed, by checking for lengths of repeated value sequences.
-An imputation method for the pedestrian data was developed with a theory of sensor locations with a large proportion of missing values require modelling with non-deterministic predictors due to the lack of data.
-This method involved using neighbouring sensor counts data for prediction.
-Sensors with a small proportion of missing values used imputation models with only time and date variables as predictors.
-The proposed algorithm was evaluated in comparison to a single model specification of a generalised linear model with only time and date based variables at each sensor location and uncorrected training data (false zero counts not corrected).
-This method was found to have improved prediction accuracy, as well as being robust at high levels of missingness.
+<!-- 
+The following line is required to re-set page numbering after preliminary material. Do not remove
+-->
+\clearpage\pagenumbering{arabic}\setcounter{page}{0}

### 2017-10-21 thesis_draft/01-chap1.Rmd
bookdown update - need to fix formatting, improve introduction, write abstract, finish off lit review and conclusion

--- 2017-10-21 thesis/draft.Rmd
@@
+---
+chapter: 1
+knit: "bookdown::render_book"
+---
 ---
@@
 ```
-## Background
-
-#### The Data: City of Melbourne Pedestrian Data
+# Background
+
+### The Data: City of Melbourne Pedestrian Data
 
@@
 
-### Literature Review
+# Literature Review
 <!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->
@@
 
-## Methodology: Imputation of Missing Values
+# Methodology: Imputation of Missing Values
 
@@
 
-```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
+```{r naprop_hist, fig.width=6, fig.height=3, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
 The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data.
@@
 
-```{r, fig.width=3, fig.height=3, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
+```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
 ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
@@
 
-```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 8, fig.height = 4}
+```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 6, fig.height = 3}
 grid.arrange(p1, p2, ncol = 1)
@@
 
-## Implementation of improved imputation algorithm
+# Implementation of improved imputation algorithm
 
@@
 Using location data of the sensors provided by the City of Melbourne, a data frame containing the latitude and longitude coordinates of all the sensors which were classified as having a small proportion of missing values (and now have imputed, complete data), which are the candidates to be neighbours used for prediction.
-Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()`, an approximation of distance between each of the candidate sensors is calculated in and stored as a vector.
+Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()` from , an approximation of distance between each of the candidate sensors is calculated in and stored as a vector.
 Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.
@@
 
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
+```{r, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
 The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Southern Cross Station")
+           ylab("Predicted Pedestrian Count")
 p2 <- ggplot(data = NULL, aes(x = original_df %>% 
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Southern Cross Station")
+           ylab("Predicted Pedestrian Count")
 grid.arrange(p1, p2, nrow = 1)
@@
 
-```{r, fig.height=4, fig.width=9, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
+```{r, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
 The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}
@@
 
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}
+```{r, fig.height=1, fig.width=6, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}
 
@@
 
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
+```{r, fig.height=3, fig.width=6, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
 The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
@@
 
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
+```{r, fig.height=3, fig.width=6, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
 The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
@@
 
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}
+```{r, fig.height=1, fig.width=6, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}
 
@@
 
-```{r, fig.width = 8, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
+```{r, fig.width = 6, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
 ggplot() + geom_smooth(aes(x = as.integer(Time),
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South).
+```{r, fig.width = 5, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South).
 20% of the observations are removed from the traning data for the imputation model at random."}
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence.
+```{r, fig.width = 5, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence.
 This method of simulating missing data is a better representation of the missing data observed at sensors with a large proportion of missing data compared to randomly removing hourly observations from the training data."}
@@
 
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence.
+```{r, fig.width = 5, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence.
 A large proportion of the training data is removed to test the robustness of the imputation method used at sensors with extremely large proportions of missing data during the training period defined."}
@@
 Removing 50% of the data using the same method, it is found that the in-sample MARE is not affected significantly by the larger proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
-
-# Predictive Model
-
-For the purpose of prediction, the use of the most simplistic model possible to make quick predictions of the expected pedestrian counts at each location, accessible to the public.
-The model is not expected to be able to predict large deviations from time and date based estimates.
-
-Similar to the simple imputation model used,  a generalised linear model with quasi-poisson errors is estimated using time and date based variables as predictors. The model specification used is:
-
-$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
-
-$$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$
-
-```{r, echo = FALSE}
-library(purrr)
-library(dplyr)
-library(tidyr)
-library(lubridate)
-library(roxygen2)
-load("../data/model_est.RData")
-
-daytype <- function(date){
-  if(!is.Date(date)){stop("Object is not a date.")}
-  x_day <- wday(date, label = TRUE, abbr = FALSE)
-  x_daytype <- ifelse(x_day == "Sunday", "Sunday",
-                ifelse(x_day == "Saturday", "Saturday",
-                ifelse(x_day == "Monday", "Monday",
-                ifelse(x_day == "Friday", "Friday",
-                          "Midweek"))))
-  return(x_daytype)
-}
-#' Predicted pedestrian counts at various sensor locations in Melbourne
-#'
-#' @param pred_date input date
-#' @param t_hour input integer
-#' @param is_pub_hol input logical
-#' @return data frame of the fitted values at each location in wide format, with time and date
-#' @export
-#' @examples
-#' ped_predict()
-#' ped_predict("2017-12-25", 18, TRUE)
-#' 
-ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
-                        t_hour = 12,
-                        is_pub_hol = FALSE,
-                        long = FALSE) {
-  if(!(t_hour %in% seq(0, 23))){
-    stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
-  pred_date <- ymd(pred_date)
-  if(!is.Date(pred_date)){
-    stop(paste("Date given to predict cannot be parsed as a date. Use the date format 'yyyy-mm-dd'. Example: 25 January 2018 is '2018-01-25'."))
-    }
-
-  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
-               mutate(DayType = ifelse(is_pub_hol,
-                                       "Holiday",
-                                       daytype(pred_date)),
-                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
-                      Date = pred_date,
-                      Date_Time = ymd_h(paste(Date, t_hour, sep = " ")))
-
-  preds <- numeric()
-  for(i in names(pred_model)){
-    preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
-  }
-  if(long){
-    return(preds)
-  }
-    else
-  preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
-    spread(Sensor_ID, fit) %>% cbind(date_pred) %>% as_tibble
-
-  return(preds_df)
-}
-
-ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),is_pub_hol = FALSE){
-  day_df <- data.frame()
-      for(t in 0:23){
-        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
-        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
-      }
-  return(as.tibble(day_df))
-}
-
-```
-
-Using a GLM predicts well compared to estimating with other models such as AR($p$) or other time series models when considering the tradeoff between model complexity and prediction accuracy.
-Due to multiple seasonalities, AR($p$) models would need to be considered on a hour of the week basis at each sensor location.
-Prediction with time series models would also require of historical data.
-For the objective of this predictive model, the additional imputs required by end users to predict pedestrian counts when making out of sample predictions is not desirable.
-
-```{r, fig.width=6, fig.height=4, fig.cap = paste("Plot of daily pedestrian counts series at Melbourne Central by DayType and Month.
-Month of the year is mapped to colour to help identify annual seasonal patterns, where month has an additive effect on the pedestrian counts.
-At Melbourne Central, different months can be seen to have different mean counts for a given type of day and hour of the day.")}
-ggplot(dfa2 %>%
-         mutate(Date = date(Date_Time))) +
-  geom_path(aes(x = as.integer(Time), y = `Melbourne Central`,
-                group = Date, colour = Month),
-            alpha = 0.1) +
-  scale_colour_viridis(discrete = TRUE, option = "C") +
-  facet_wrap(~ DayType) + theme_minimal() +
-  theme(legend.position = "none") +
-  labs(x = "Time of Day", y = "Pedestrian Counts \n at Melbourne Central")
-```
-
-A plot of the daily pedestrian counts series at Melbourne Central by type of day and month show that time and date variables as deterministic predictors adequately explains most variation in the pedestrian counts.
-It can be seen that month has an additive effect as the pattern of the hourly counts for a given date are not affected by month.
-The six different factor levels that `DayType`, the type of day, is seen to be a sufficiently explain six different patterns.
-The type of day, Midweek, shows evidence that no additional information is gained from identifying the day to be Tuesday, Wednesday or Thursday.
-
-```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM for the interaction term $\text{DayType} \times \text{Time}$ for each hour at each sensor.
-Variables (x-axis) is ordered by the scagnostic measure, Outlying.
-"), include=FALSE, eval = FALSE}
-dfcoeff <- data.frame(terms = factor(), coeff = numeric(), sensor = factor())
-for(i in names(pred_model)){
-  terms <- names(coef(pred_model[[i]]))
-  coeff <- coef(pred_model[[i]])
-  sensor <- rep(i, length(terms))
-  df_b <- data.frame(terms, coeff, sensor)
-  dfcoeff <- bind_rows(dfcoeff, df_b)
-}
-dfcoeff <- dfcoeff %>% spread(terms, coeff)
-
-library(GGally)
-library(scagnostics)
-ggparcoord(dfcoeff[, 51:74], scale = "std", 
-           order = "Clumpy", alphaLines = 0.3) + theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, size = 6)) +
-  labs(x = "Parameter")
-```
-
-When estimating the GLM at each sensor location with R, only the necessary elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function are saved to file.
-Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets.
-This is due to the model object providing redundancies.
-In particular, elements which are unnecessarily saved for the purposes of the predictive model: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.
-
-The result of removing all the unnecessary redundant data elements in the GLM object is a storage size reduction from `r mall_model_2 %>% object.size() %>% format(units = "auto")` to `r mall_model_2 %>% stripGlmLR() %>% object.size() %>% format(units = "auto")`.
-This is only `r round((mall_model_2 %>% stripGlmLR() %>% object.size() %>% as.numeric/ mall_model_2 %>% object.size() %>% as.numeric)*100, 2)`% of the size of the original GLM object.
-With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.
-
-The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves.
-Using the estimated model parameters, a new function is written to facilitate the predictions in an easy to use format.
-
-```{r, echo = TRUE, eval = FALSE}
-ped_predict(pred_date = "2017-12-26", t_hour = 13, is_pub_hol = TRUE)
-```
-This function will return tibble (trimmed down version of `data.frame()` in R) containing the predicted counts at each sensor location for the 13th hour (13:00/1:00 PM) on Boxing Day (26 December) 2017.
-This tibble can then be easily used for visualisations or analysis.
-Due to the lack of data on future public holiday dates, the function assumes the date given is not a public holiday unless `is_pub_hol = TRUE` is provided.
-Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predict.
-
-Using these function as a basis, visualisations of pedestrian counts can be easily generated in R:
-```{r, echo = TRUE, fig.cap = "Prediction for 21/05/2018 in R using `ped_predict_day()`, then plotting the time series using `ggplot2`.
-This shows the user friendly nature of the function, where the function returns predictions in a tidy, wide format.
-No data manipulation is required in order to plot the predicted counts.", fig.height=2}
-ped_predict_day("2018-05-21") %>% 
-  ggplot() + geom_line(aes(x = Date_Time, y = Southbank))
-```
-
-# Conclusion  
-Analysis of the Melbourne CBD pedestrian data made publicly available by the City of Melbourne has revealed issues with the dataset with relation to missing values in particular.
-A major issue which was found when using data sourced from the City of Melbourne's official pedestrian data visualisation (also accessible using `rwalkr::walk_melb()`) is potential for false zero-counts where missing values were given the value of zero.
-As outliers have major implications on analysis, a correction for the false counts in the data was proposed, by checking for lengths of repeated value sequences.
-An imputation method for the pedestrian data was developed with a theory of sensor locations with a large proportion of missing values require modelling with non-deterministic predictors due to the lack of data.
-This method involved using neighbouring sensor counts data for prediction.
-Sensors with a small proportion of missing values used imputation models with only time and date variables as predictors.
-The proposed algorithm was evaluated in comparison to a single model specification of a generalised linear model with only time and date based variables at each sensor location and uncorrected training data (false zero counts not corrected).
-This method was found to have improved prediction accuracy, as well as being robust at high levels of missingness.

### 2017-10-21 thesis_draft/02-chap2.Rmd
bookdown update - need to fix formatting, improve introduction, write abstract, finish off lit review and conclusion

--- 2017-10-21 thesis/draft.Rmd
@@
 ---
-title: "Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods"
-author: "Gavin Chin"
-date: "21 September 2017"
-output:
-  pdf_document: default
-  html_document: default
-  word_document: default
-fontsize: 12pt
+chapter: 2
+knit: "bookdown::render_book"
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
-                      message = FALSE, warning = FALSE,
-                      fig.pos = 'h')
-library(tidyverse)
-library(lubridate)
-library(foreach)
-library(doSNOW)
-library(rwalkr)
-library(pander)
-library(RColorBrewer)
-library(gridExtra)
-library(viridis)
-library(ggmap)
-library(broom)
-set.seed(12345)
-```
-
-```{r data_loading, warning=F, message=F}
-ped_data <- NULL
-if(file.exists("../data/ped_df.csv")){
-  ped_data <- read_csv("../data/ped_df.csv")
-  ped_data$X1 <- NULL
-  }
-
- ## ped_data <- read_csv("http://209.148.91.227/files/ped_df_28sep17.csv")
-
-## update data
-last_df_date <- ifelse(file.exists("../data/ped_df.csv"),
-                        ped_data$Date %>% sort() %>%
-                        tail(1),
-                        as.Date("2013-12-31")) + 1 %>% as.Date(origin = "1970-01-01")
-
-if(last_df_date != today(tzone = "Australia/Melbourne")){
-  new_data <- rwalkr::walk_melb(from = last_df_date, to = today()-1)
-  ped_data <- rbind(ped_data, new_data)
-  write_csv(ped_data, "../data/ped_df.csv")
-  }
-
-ped_data$Sensor_ID <- ped_data$Sensor
-ped_data$Hourly_Counts <- ped_data$Count
-ped_data$Date_Time <- ymd_h(paste(ped_data$Date, ped_data$Time, sep = " "))
-ped_data$Day <- lubridate::wday(ped_data$Date_Time, label = TRUE, abbr = F)
-ped_data$Month <- lubridate::month(ped_data$Date_Time, label = TRUE, abbr = F)
-ped_data$Time <- as.factor(ped_data$Time)
-
-## long to wide
-ped_data %>% select(Sensor_ID, Count, Date_Time, Time, Month, Day) %>% 
-  spread(Sensor_ID, Count) -> dfa
-
-pub_hday14 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/56a5ee91-8e94-416e-81f7-3fe626958f7e/download/australianpublicholidays-201415.csv---australianpublicholidays.csv.csv")
-pub_hday15 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/13ca6df3-f6c9-42a1-bb20-6e2c12fe9d94/download/australianpublicholidays-201516.csv")
-pub_hday16 <- read.csv("http://data.gov.au/dataset/b1bc6077-dadd-4f61-9f8c-002ab2cdff10/resource/a24ecaf2-044a-4e66-989c-eacc81ded62f/download/australianpublicholidays-201617.csv")
-pub_hdays <- rbind(pub_hday14, pub_hday15, pub_hday16)
-
-pub_hdays$Date <- ymd(pub_hdays$Date)
-pub_hdays$Month <- lubridate::month(pub_hdays$Date, label = TRUE, abbr = F)
-pub_hdays$VIC <- 0
-pub_hdays$VIC[grep(glob2rx("*VIC*"), pub_hdays$`Applicable.To`)] <- 1
-pub_hdays$VIC[grep("NAT", pub_hdays$`Applicable.To`)] <- 1
-
-dfa <- dfa %>% mutate(IsHDay = (date(Date_Time) %in% pub_hdays$Date[pub_hdays$VIC == 1]))
-
-dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
-                     ifelse(dfa$Day == "Saturday", "Saturday",
-                    ifelse(dfa$Day == "Monday", "Monday",
-                     ifelse(dfa$Day == "Friday", "Friday",
-                                  "Midweek"))))
-dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
-dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
-original_df <- dfa
-dfa <- filter(dfa, Date_Time < "2017-01-01")
-ori_2 <- filter(original_df, Date_Time < "2017-01-01")
-ori_ts <- filter(original_df, Date_Time > "2016-12-31")
-ori_ts2 <- ori_ts
-```
-## Background
-
-#### The Data: City of Melbourne Pedestrian Data
-
-The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives"[^1].
-This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.
-In particular, we want to be able to model the pedestrian traffic at different locations in the CBD, and forecast the traffic.
-In addition, we aim to have an interactive, dynamic data visualisation of the model to help us to understand how Melbourne operates.
-
-The dataset being investigated can be obtained from the official City of Melbourne's open data[^2].
-The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations.
-In total, there is `r nrow(ped_data)` observations in the dataset, representing `r nrow(dfa)` hours of data.
-It is available in `.csv` format, allowing for easy data import into R.
-We can also access the pedestrian data using the the `rwalkr` package by Earo Wang[^3].
-This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.
-
-```{r}
-ped_data %>% select(Sensor, Hourly_Counts) %>% split(.$Sensor)  %>%
-            map(is.na) %>% map(sum) %>% unlist -> sensor_missvals
-```
-
-Not even an single sensor location has complete data for the period between between 1/1/2014 and `r paste(day(today()), month(today()), year(today()), sep = "/")`.
-
-[^1]: from _'About Melbourne Data'_ page at https://data.melbourne.vic.gov.au/about
-
-[^2]: the specific pedestrian data is available at   https://data.melbourne.vic.gov.au/Transport-Movement/Pedestrian-volume-updated-monthly-/b2ak-trbp
-
-[^3]: https://github.com/earowang/rwalkr/
-
-### Imputation Literature Review Notes
-<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->
-
-<!-- http://ir.lib.uwo.ca/cgi/viewcontent.cgi?article=1250&context=csce2016   -->
-<!-- - traffic counters with 40% to 60% missing values -->
-<!-- - heuristic methods -->
-<!--   - using "good historical values" or historical average values resulted -->
-<!--     in MARE (mean absolute relative error) reaching up to 80% -->
-<!--   - using moving average values worked better than other heuristic methods -->
-<!--   - all heuristic methods inhenrently cannot reflect sudden fluctuations during abnormal periods -->
-<!-- - pattern matching methods -->
-<!--   - find "candidate" volume pattern to compare to "study curve" and use the candidate which best matches the study curve -->
-<!-- - ARIMA -->
-<!--   - did not work well with multiple seasonalities -->
-<!-- - AI methods -->
-<!--   - genetic algorithms (GAs) and artificial neural networks (ANNs) -->
-
-<!-- Method used in the study was kNN with an interaction using a weather model, which they found to work best with large variations in traffic during winter -->
-<!-- (severe weather conditions affecting traffic volume)   -->
-
-Reviewing current research on imputation of missing values in traffic data, a paper on imputation of missing classified traffic data during winter season (Imputation of Missing Classified Traffic Data During Winter Season - _Hyuk-Jae Roh, Satish Sharma, Prasanta K.
-Sahu_) stated some methods which were found to be poor for imputation.
-The data they were attempting imputation for were traffic counters located on the highway network in Alberta, Canada with between 40% and 60% missing data.
-
-Replacement with "good" historical values, or using historical average values was found to have resulted in very poor fit with high MARE/MARD (mean absolute relative error/difference).
-They did find that using a moving average worked the best out of all the heuristic methods used.
-The largest problem with heuristic methods, however, is the inherent inability to reflect sudden fluctuations/shocks during abnormal periods.
-With such a large proportion of missing values, pattern matching methods were investigated.
-This involves comparing the _study curve_, the pattern at the counter which is to be imputed, to candidate patterns (patterns at other locations).
-
-For the purposes of that paper, it was found to be inappropriate with the data being analysed.
-However, unlike _Roh et al._ where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller.
-Another difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data.
-They proposed using non-parametric estimation methods with k-Nearest Neighbours for their imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.
-
-<!-- #### Flexible and Robust Method for Missing Loop Detector Data Imputation - _K. C. Henrickson, Y. Zou, and Y. Wang_   -->
-
-<!-- http://docs.trb.org/prp/15-5804.pdf   -->
-<!-- - also use kNN -->
-
-<!-- #### Key Factors of K-nearest Neighbors Nonparametric Regression in Short-time Traffic Flow Forecasting -->
-<!-- http://www.springer.com/cda/content/document/cda_downloaddocument/9789462391017-c2.pdf?SGWID=0-0-45-1492476-p177147492 -->
-
-<!-- #### K-NEAREST NEIGHBOR ALGORITHM FOR UNIVARIATE TIME SERIES PREDICTION -->
-
-<!-- http://webbut.unitbv.ro/BU2012/Series%20III/BULETIN%20III/Sasu%20A.pdf -->
-
-## Methodology: Imputation of Missing Values
-
-### Basic GLM Approach
-
-The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
-$$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
-$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
-`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
-Both these variables are categorical/factors. We cannot treat time of the day as an integer or numeric because it does not have a linear relationship with counts.s
-
-We use the Quasi-Poisson error distribution as opposed to normal/Gaussian errors due to the response variables being count data.
-The Poisson distribution allows for only non-negative integer values, while estimating a Quasi-Poisson model estimates an additional parameter for overdispersion.
-This allows more flexibility in the model by allowing the assumption that $E[Y]=Var[Y]=\mu$ to be relaxed.
-Instead, the Quasi-Poisson distribution allows for $Var[Y] = \theta \mu$.
-
-```{r}
-## Base GLM models
-
-### code from:
-## http://www.win-vector.com/blog/2014/05/trimming-the-fat-from-glm-models-in-r/
-stripGlmLR = function(cm) {
-  cm$y = c()
-  cm$model = c()
-
-  cm$residuals = c()
-  cm$fitted.values = c()
-  cm$effects = c()
-  cm$qr$qr = c()  
-  cm$linear.predictors = c()
-  cm$weights = c()
-  cm$prior.weights = c()
-  cm$data = c()
-
-  cm$family$variance = c()
-  cm$family$dev.resids = c()
-  cm$family$aic = c()
-  cm$family$validmu = c()
-  cm$family$simulate = c()
-  attr(cm$terms,".Environment") = c()
-  attr(cm$formula,".Environment") = c()
-
-  cm
-}
-#######################################
-```
-
-```{r}
-cl <- makeCluster(2)
-registerDoSNOW(cl)
-fitted_base <- list()
-models_base <- foreach(i = unique(ped_data$Sensor_ID)) %dopar% {
-    tr_dat <- data.frame(dfa[, i],  HDay = dfa$HDay, Time = dfa$Time)
-    colnames(tr_dat)[1] <- "Hourly_Counts"
-    model <- glm(Hourly_Counts ~ HDay*Time, data = tr_dat, family = quasipoisson())
-    return(stripGlmLR(model))
-  }
-names(models_base) <- unique(ped_data$Sensor_ID)
-# ped_models[[31]]$xlevels <- ped_models[[1]]$xlevels
-for (i in unique(ped_data$Sensor)){  
-  tr_dat <- data.frame(dfa [, i],  HDay = dfa$HDay, Time = dfa$Time)
-  options(na.action = na.pass)
-  fitted_base[[i]] <- predict.glm(models_base[[i]], newdata = tr_dat, type = "response")
-  options(na.action = na.omit)
-  }
-
-```
-
-While this works well with a small proportion of missing values, it also requires a large number of observations.
-Specifically, with this paramaterisation, we have $23 + 7 + (23 \times 7) + 1=192$ coefficients to estimate.
-Another disadvantage of using this model is the lack of robustness to outliers due to the sensitivity of some parameters.
-This becomes particularly problematic at sensor locations which have been recently installed, where there is a lack of historical data.
-
-### Improved imputation algorithm using small-large split approach
-
-The proposed alternative approach to imputing the missing values in the data focuses on improving the models used at locations with a large proportion of missing values.
-A potential threshold value which could be used to class a sensor as having a "large" proportion of missing values vs a "small" proportion of missing values is 10%.
-At locations with a "small" proportion of missings, we can continue to use the GLM quasi-possion regression specified in $(1)$.
-At the locations with a large proportion of missing values, we need to use information from neighbouring sensors.
-
-The first issue which needs to be addressed is the potential of values which are actually missing due to sensor failure or malfunction having a zero count.
-This issue will cause bias in the estimates, as well as affect the classification of a sensor.
-Of course, we cannot simply classify all zero counts as `NA`, as true zero values are possible.
-
-```{r}
-n_obs <- nrow(dfa)
-threshold <- 0.1
-threshold_miss_hrs <- 6
-
-run_lengths <- list()
-
-for(i in unique(ped_data$Sensor)){
-    dfa[is.na(dfa[, i]), i] <- 0
-    na_length_check <- rle(as.numeric(unlist(dfa[,i])))
-    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
-    rl <- dfa[, i] %>% unlist %>% rle()
-    dfa[maybeNA, i] <- NA
-    defsNA <- is.na(ori_2[, i])
-    dfa[defsNA, i] <- NA
-    run_lengths[[i]] <- rl
-  }
-
-ori_2 <- dfa
-
-run_lengths %>% lapply(., `[[`, 1) %>%
-    unlist -> sequences
-
-pander(summary(sequences[sequences > 1]))
-```
-Table: Summary statistics of the sequence lengths of repeated values
-
-Instead, a simple check we can implement is to look for long sequences of zero counts.
-For example, we may want to classify any sequence of `Hourly_Counts = 0` running for longer than 24 hours as being considered `NA`.
-From the summary of lengths of repeated values (which are repeated at least once), we see that the distribution of the lengths are very positively skewed with a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data.
-Classifying any sequence of repeated values, typically 0, which has a length greater than 6 as `NA` should allow true zero counts to avoid being misclassified as `NA`.
-We select the length of 6 hours as it would be unreasonable to assume any sensor location to have the exact same number of pedestrian counts over such a long period.
-
-After we have replaced questionable zero values with the value `NA`, we can calculate the proportion of missing values for each sensor location.
-From these calculated proportion, we can classify sensors as either having a "large" or "small" proportion of missing values, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.1`.
-
-```{r na_prop_compute}
-na_prop <- numeric()
-
-for (i in unique(ped_data$Sensor)){
-  na_prop[i] <- sum(is.na(dfa[, i])) / n_obs
-}
-
-na_prop_pre <- numeric()
-
-for (i in unique(ped_data$Sensor)){
-  na_prop_pre[i] <- sum(is.na(ori_2[, i])) / n_obs
-}
-
-large_miss_sensors <- names(na_prop[na_prop > threshold])
-small_miss_sensors <- setdiff(names(na_prop), large_miss_sensors)
-```
-
-```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
-The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data.
-This emphasises the necessity to correct for false zero counts as it will affect "}
-theme_g <- theme_minimal() +
-           theme(plot.title = element_text(hjust = 0.5,
-                                           size = 10))
-p1 <- ggplot() + geom_histogram(aes(x = na_prop*100), bins = 12) +
-          xlab("Percentage of Missing Values (%)") +
-          ylab("Number of Sensors") + theme_g +
-          geom_vline(xintercept = 10, size = 2,
-                     colour = 'red') +
-          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
-                    colour = "white", fill = 'black', alpha = 0.4) +
-          scale_y_continuous(limits = c(0,35))
-p2 <- ggplot() + geom_histogram(aes(x = na_prop_pre*100), bins = 12) +
-          xlab("Percentage of Missing Values (%)") +
-          ylab("Number of Sensors") + theme_g +
-          geom_vline(xintercept = 10, size = 2,
-                     colour = 'red') +
-          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
-                    colour = "white", fill = 'black', alpha = 0.4) +
-          scale_y_continuous(limits = c(0,35))
-grid.arrange(p2, p1, nrow = 1)
-```
-
-Because the models for the sensors with a large proportion of missing values rely on having neighbouring sensors having complete cases (no missing values), we want to impute these values first.
-This will then allow us to have as much information to be available for the models which use neighbouring sensors to to train on.
-
-For sensors with a small proportion of missing values, we use a GLM quasipoisson model. To improve on the previous model used, we use the model:
-$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
-
-$$\text{where } \quad \mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}} \sim \text{Month} + \text{Time}\times \text{DayType} \tag{2}$$
-
-```{r alg_glm}
-models <- list()
-fitted_data <- list()
-dfa2 <- dfa
-p <- progress_estimated(length(unique(ped_data$Sensor)))
-for(i in small_miss_sensors){
-  p$tick()$print()
-  options(na.action = na.omit)
-  # if(max(na_length_check$lengths) > max_miss_hrs) 
-  # {
-  #   model <- NULL
-  # }
-
-  sensor_df <- unlist(dfa[, i])
-
-  model <- glm(sensor_df ~  Month + DayType*Time, family = 'quasipoisson', data = dfa)
-  models[[i]] <- (stripGlmLR(model))
-}
-```
-
-```{r alg_glm_impute}
-  options(na.action = na.pass)
-for(i in small_miss_sensors){
-  ts_dat <- data.frame(Month = dfa$Month, DayType = dfa$DayType, Time = dfa$Time)
-  fitted_data[[i]] <- predict.glm(models[[i]],newdata = ts_dat, type = "response")
-}
-  options(na.action = na.omit)
-  for(i in small_miss_sensors){
-    dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
-    }
-
-```
-
-Instead of `HDay`, we use `DayType` which classifies Tuesday, Wednesday and Thursday as a single factor level, Midweek.
-We do this as the daily patterns on these weekdays are similar.
-This reduces the number of parameters to estimate by 46.
-A result of this is we also have more robust estimates for the midweek days, being less sensitive to outliers.
-Because this model will be estimated for sensor locations with few missing values,  we have enough data to add in month as an additive effect.
-This will help capture the annual seasonality.
-
-Using these estimated models, we impute the missing values to produce complete data at all these sensor locations.
-For the locations with high proportions of missing values can be predicted using the now complete data from all the locations which had a small proportion of missing values.
-Using a threshold of 10%, this gives us `r length(small_miss_sensors)` sensor locations to use.
-
-```{r, fig.width=3, fig.height=3, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
-ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
-lats <- ped_loc$Latitude
-lons <- ped_loc$Longitude
-ids <- ped_loc$`Sensor Description`
-
-{
-ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
-ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
-ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
-ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
-ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
-ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
-ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
-ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
-ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
-}
-loc <- data.frame(lons, lats, ids)
-rownames(loc) <- ids
-loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
-                                  "Small", "Large"))
-rownames(loc) <- loc$ids
-
-melb3 <- get_map(location = c(loc["Southbank",]$lons,
-                              loc["Southbank",]$lats),
-                 zoom = 17, maptype = "roadmap")
-
-sens_lines_lon <- as.numeric(c(loc["Flinders St Station Underpass",]$lons,
-                loc["Southbank",]$lons,
-                loc["Sandridge Bridge",]$lons))
-
-sens_lines_lat <- as.numeric(c(loc["Flinders St Station Underpass",]$lats,
-                loc["Southbank",]$lats,
-                loc["Sandridge Bridge",]$lats))
-
-sens_lines <- data.frame(lon = sens_lines_lon,
-                         lat = sens_lines_lat)
-options(na.action = 'na.omit')
-ggmap(melb3) + geom_line(aes(x = lon, y = lat),
-                        data = sens_lines, size = 2,
-                        colour = viridis(3)[2], alpha = 0.5) +
-               geom_point(aes(x = lons, y = lats, colour = cl),
-                          size = 6,
-                          data = loc) +
-               scale_colour_viridis(discrete = TRUE,
-                                    name = "Sensor NA Prop.",
-                                    begin = 0.5, end = 0) +
-               geom_point(aes(x = loc["Southbank",]$lons,
-                              y = loc["Southbank",]$lats),
-                          size = 12, colour = viridis(3)[2]) +
-               geom_label(aes(label = ids, x = lons, y = lats,
-                              fill = cl),
-                         fontface = 'bold', colour = 'white',
-                         data = loc, size = 3,
-                         nudge_y = -0.000325) +
-              scale_fill_viridis(discrete = TRUE,
-                                  name = "Sensor NA Prop.",
-                                  begin = 0.5, end = 0) +
-              xlab(NULL) + ylab(NULL) +
-              theme(plot.title = element_text(hjust = 0.5),
-                    plot.caption = element_text(hjust = 0.5),
-                    axis.text.x = element_blank(),
-                    axis.text.y = element_blank(),
-                    axis.ticks = element_blank(),
-                    legend.position = c(.8,0.2))
-```
-
-For each of the sensor locations with high proportions of missings, we need to decide which sensor to use for prediction.
-A simple method of finding a geographical closest neighbour is to take the haversine/great-circle distance between the sensor to be imputed and all the possible candidates.
-```{r, message=F}
-# Calculates the geodesic distance between two points specified by radian latitude/longitude using the
-# Spherical Law of Cosines (slc)
-# source: https://www.r-bloggers.com/great-circle-distance-calculations-in-r/
-gcd.slc <- function(long1, lat1, long2, lat2) {
-  R <- 6371 # Earth mean radius [km]
-  d <- acos(sin(lat1)*sin(lat2) + cos(lat1)*cos(lat2) * cos(long2-long1)) * R
-  return(d) # Distance in km
-}
-
-## name fix dictionary
-{
-ids[ids == "Flinders Street Station Underpass"] <- "Flinders St Station Underpass"
-ids[ids == "Flinders St-Spark La"] <- "Flinders St-Spark Lane"
-ids[ids == "Queen St (West)"] <-  "Queen Street (West)"
-ids[ids == "The Arts Centre"] <- "Vic Arts Centre"
-ids[ids == "Lonsdale St-Spring St (West)"] <- "Spring St-Lonsdale St (South)"
-ids[ids == "Lygon St (East)"] <- "Lygon Street (East)"
-ids[ids == "QV Market-Elizabeth St (West)"] <- "QV Market-Elizabeth (West)"
-ids[ids == "Melbourne Convention Exhibition Centre"] <- "Convention/Exhibition Centre"
-ids[ids == "St Kilda Rd-Alexandra Gardens"] <- "St. Kilda-Alexandra Gardens"
-}
-loc <- data.frame(lons, lats, ids)
-rownames(loc) <- ids
-```
-
-```{r alg_lm}
-for (i in large_miss_sensors){
-  # p$tick()$print()
-  sensor_dists <- numeric()
-  for (j in small_miss_sensors){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df_scale <- dfa2[, closest_sensor[1]] %>% unlist %>% scale
-  close_df_scale2 <- dfa2[, closest_sensor[2]] %>% unlist %>% scale
-
-  sensor_df <- dfa2[, i] %>% unlist
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa$Time)
-  options(na.action = na.omit)
-  model <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  models[[i]] <- stripGlmLR(model)
-
-  options(na.action = na.pass)
-  fitted_data[[i]] <- predict(model, newdata = tr_dat, type = "response")
-  options(na.action = na.omit)
-
-  dfa2[is.na(dfa[, i]), i] <- fitted_data[[i]][is.na(dfa[, i])]
-
-}
-
-```
-
-Using the two geographically closest sensors with complete cases, we use another GLM specified as:
-$$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{neighbours}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}}$$
-$$\text{where } \quad  \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}} \sim \text{Time} \times \text{HourlyCounts}_{\text{neighbour 1}}^{SC} + \text{Time} \times \text{HourlyCounts}_{\text{neighbour 2}}^{SC}$$
-
-We would expect that the amount of people passing through the neighbouring sensors will be a strong predictor of the pedestrian counts as they are likely to also pass through.
-Using geographical neighbours, we can capture rare events effectively.
-Any large shocks to the counts at the neighbouring sensors will have an effect on the counts to be imputed.
-
-```{r, echo = FALSE, message=FALSE, warning=F, include=FALSE}
-i <- "Southbank"
-  sensor_dists <- numeric()
-  for (j in small_miss_sensors){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  dfa3 <- original_df %>% filter(Date_Time > "2017-01-01")
-  close_df <- dfa3[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa3[, closest_sensor[2]] %>% unlist
-
-  close_dfsc <- scale(close_df)
-  close_dfsc2 <- scale(close_df2)
-
-  sensor_df <- dfa3[, i] %>% unlist
-  sensor_dfsc <- scale(sensor_df)
-
-  tr_dat <- data.frame(sensor_df, close_df, close_df2,
-                       Time = as.integer(dfa3$Time),
-                       Date_Time = dfa3$Date_Time,
-                       HDay = dfa3$HDay) %>%
-            gather(key = Sensor, value = Count, sensor_df:close_df2)
-
-  tr_dat_sc <- data.frame(sensor_dfsc, close_dfsc, close_dfsc2,
-                       Time = as.integer(dfa3$Time),
-                       Date_Time = dfa3$Date_Time,
-                       HDay = dfa3$HDay) %>%
-            gather(key = Sensor, value = Count, sensor_dfsc:close_dfsc2)
-
-p1 <- ggplot(tr_dat %>% filter(Date_Time < "2017-03-02",
-                         Date_Time > "2017-02-26")) +
-      geom_line(aes(x = Date_Time, y = Count,
-                    colour = Sensor), size = 2) +
-      scale_colour_viridis(discrete = TRUE,
-                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
-                                      "Neighbour 2 (Sandridge Bridge)",
-                                      "Target (Southbank)")) +
-      xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
-      theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
-
-p2 <- ggplot(tr_dat_sc %>% filter(Date_Time < "2017-03-02",
-                         Date_Time > "2017-02-26")) +
-      geom_line(aes(x = Date_Time, y = Count,
-                    colour = Sensor), size = 2) +
-      scale_colour_viridis(discrete = TRUE,
-                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
-                                      "Neighbour 2 (Sandridge Bridge)",
-                                      "Target (Southbank)")) +
-      xlab("Time") + ggtitle("Scaled/Stadardised Neighbours") +
-      ylab("Standard Deviations") + theme_minimal() +
-      theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
-```
-
-```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 8, fig.height = 4}
-grid.arrange(p1, p2, ncol = 1)
-```
-
-In order to use a mixture of the patterns from two different sensors, we need to use the scaled counts at the neighbouring sensors to avoid the magnitude of the counts having an effect on the predicted counts.
-We scale the counts by standardising to $\mu = 0, \sigma^2 = 1$.
-Because the covariance with neighbouring sensors may vary over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
-
-## Implementation of improved imputation algorithm
-
-The exact code used for this paper can be found on http://github.com/gavinchin/honours2017
-
-### False zero-counts correction
-
-The false zero-counts correction is performed in R using the run length encoding function, `rle()`.
-This function computes the length and values of runs of equal values in a vector.
-Unfortunately, this function does not recognise `NA` in the its computation, so all NA values are initially recoded to take the value 0 at each sensor.
-This results in all missing values and true zero counts to take the value zero when running `rle()`.
-With the returned run lengths, all sequences of repeated values longer than the defined length of 6 hours are replaced with `NA`.
-Any values which were also originally missing, but were not missing for longer than 6 hours (and hence are still coded as 0), are recoded to be `NA`.
-It must be noted that sequences of repeated values not equal to zero which run for longer that 6 hours would be incorrectly replaced with `NA`.
-This is highly unlikely, however, as 6 hours of the exact same pedestrian counts would be extremely rare.
-
-### Caculation of proportion of missing values and classification of sensors
-
-In the working data, a wide format is used.
-This makes it easier to calculate missing value proportion at each sensor location, as it is the proportion of missing values in the column for each sensor.
-Using `is.na()` returns a logical vector indicating whether the sensor count is missing/`NA` or not.
-The sum of the instances where `is.na = TRUE` divided by the hours in the period of 2014-2016 is computed to get the proportion of missing values.
-
-These proportions are saved for each sensor and then classified as either "large" or "small" based on a threshold value for the proportion.
-A list of sensors with large proportions of missing data and a list of sensors with small proportions of missing data is generated.
-
-### Estimation of GLM at sensor locations with a small proportion of missing values
-
-A for loop is used to estimate the GLM specified:
-```{r, echo = TRUE, eval = FALSE}
-glm(dfa[, i] ~ Month + DayType*Time, data = dfa,
-    family = quasipoisson())
-```
-where `dfa` is the object name given to the data frame containing the training data after the false zero-count correction, and `i` is the $i^{th{}}$ sensor in the list of sensors with a small proportion of missing values.
-These models are stored in a list, forming a list of models.
-An alternative approach to using a for loop in R is splitting the data into separate data frames for each sensor location (forming a list of data frames) and applying the `glm()` function on each data frame in the list using the function `purrr::map()`.
-However, when this was attempted, performance issues were encountered due to the large size of the estimated model objects in R.
-This was not an issue when a for loop was used.
-Additional code for removing unnecessary elements of the GLM model object before storing to the list of models, reduucing the memory usage.
-
-### Imputation at sensor locations with a small proportion of missing values
-
-The predicted values are obtained using `predict.glm()`, using the option `type = "response"` to have predicted values returned predicted counts.
-The default option returns the predicted values of $\hat{\lambda}$ in the quasipoisson process, not the counts, $\exp (\hat{\lambda})$.
-Missing values are then replaced using the imputed values, generating a data frame of complete data as sensor locations for with a small proportion of missing values to be used for prediction at sensors with a large proportion of missing values.
-
-### Estimation of GLM at sensor locations with a large proportion of missing values
-The following steps are performed on each sensor with a large proportion of missing values inside a for loop.
-
-##### Finding neighbouring sensors
-Using location data of the sensors provided by the City of Melbourne, a data frame containing the latitude and longitude coordinates of all the sensors which were classified as having a small proportion of missing values (and now have imputed, complete data), which are the candidates to be neighbours used for prediction.
-Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()`, an approximation of distance between each of the candidate sensors is calculated in and stored as a vector.
-Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.
-
-## Imputation Results
-
-Firstly, we evaluate the initial model used to impute:
-$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
-
-Note, this is also after the adjustments made to treat suspiciously long sequences of 0 values as `NA` in the actual counts.
-The simple GLM model is trained before this correction, while the improved model is trained on the adjusted data.
-
-For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station.
-At this sensor, the proportion of missing values (adjusted) is `r round(na_prop["Southern Cross Station"]*100,2)`Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).
-
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
-p1 <- ggplot(data = NULL, aes(x = original_df %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Southern Cross Station"],
-           y = fitted_base$`Southern Cross Station`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Southern Cross Station")
-p2 <- ggplot(data = NULL, aes(x = original_df %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Southern Cross Station"],
-           y = fitted_data$`Southern Cross Station`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Southern Cross Station")
-grid.arrange(p1, p2, nrow = 1)
-```
-
-Visually, we can see that the fitted values are good estimates as the relationship between the actuals vs fitted is very close to 1:1.
-This is seen when we perform a least squares estimate of $\widehat{Fitted}_t = \hat{\beta}_0 + \hat{\beta}_1 Actual_t$, and the estimated $\hat{\beta}_0$ is close to 0 and $\hat{\beta}_1$ is close to 1.
-
-Here we see that the GLM model with only time and date based variables work well for fitting small periods of missing data.
-
-However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor.
-We use the sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%.
-```{r}
-options(na.action = na.omit)
-p1 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Australia on Collins"])) +
-           geom_point(aes(y = fitted_base$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                      alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(aes(y = fitted_base$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                       method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + theme(aspect.ratio = 1,
-                                   plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Australia on Collins")
-p2 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
-             filter(Date_Time < "2017-01-01") %>%
-             .[, "Australia on Collins"])) +
-           geom_point(aes(y = fitted_data$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                      alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(aes(y = fitted_data$`Australia on Collins`[-(is.na(ori_2$`Australia on Collins`))]),
-                       method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + theme(aspect.ratio = 1,
-                                   plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Australia on Collins")
-```
-
-```{r, fig.height=4, fig.width=9, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}
-
-grid.arrange(p1, p2, nrow = 1)
-
-```
-
-We can see that the simple model has bias caused by the false zero counts, resulting in the predicted values to be underestimated.
-This is shown by the slope of the fitted line on the actual counts against predicted counts being $<1$, emphasising the importance of the first step taken in the algorithm of the improved model to class the false zero counts as missing values.
-
-The criterion we will use to measure goodness of fit is MARE. It is calculated by:
-
-$$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \vert \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \vert}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$
-
-We calculate a seperate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data).
-This will help to identify overfitting of the data.
-We use MARE as it will be a comparable measure between sensor locations as well as comparable between models.
-
-```{r}
-MARE_base_glm_tr <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_base_glm_tr[i] <- mean(abs(fitted_base[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
-}
-
-MARE_tr <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_tr[i] <- mean(abs(fitted_data[[i]] - unlist(dfa[, i])), na.rm = T) / mean(unlist(dfa[, i]), na.rm = T)
-}
-
-mare_df <- data.frame(simple = MARE_base_glm_tr, improved = MARE_tr, ids) %>%
-           gather(model, mare, simple:improved)
-p1 <- ggplot(mare_df) + geom_boxplot(aes(x = model, y = mare, fill = model),
-                                     width = .5, alpha = 0.5) +
-      labs(x = "Model", y = "In-Sample MARE") +
-      scale_fill_viridis(discrete = TRUE, name = "Model")
-
-p2 <- ggplot(mare_df) + geom_density(aes(x = mare, fill = model,
-                                         colour = model), adjust = 1.3,
-                                     alpha = 0.5) +
-      scale_fill_viridis(discrete = TRUE) +
-      scale_colour_viridis(discrete = TRUE) +
-      theme_minimal() + theme(plot.title = element_text(hjust = 0.5),
-                              legend.position = "none") +
-      labs(x = "In-sample MARE")
-p3 <- ggplot() + geom_point(aes(x = MARE_base_glm_tr, y = MARE_tr,
-                                colour = (MARE_tr<MARE_base_glm_tr))) +
-      geom_abline(aes(slope = 1, intercept = 0)) +
-      scale_colour_viridis(discrete = TRUE, option = "C",
-                           begin = 0.15, end = 0.85,
-                           direction = -1,
-                           name = "Improved Model is better") +
-      theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
-      theme(aspect.ratio = 1, legend.position = "bottom",
-            legend.direction = "vertical")
-```
-
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}
-
-grid.arrange(p1, p2, p3, nrow = 1)
-
-```
-
-While it is not very clear when comparing the distribution of in-sample MARE between imputation models, when we compare the in-sample MARE at each location we can see that the majority of locations had better in-sample fit with the improved model using the algorithm.
-We do see that some locations did have worse in-sample fit, but overall we see improvement with `r sum(MARE_base_glm_tr > MARE_tr)` having better fit with the improved model over the simple model.
-
-Focusing on the date 04/07/2014, which has 3 hours of missing data at Southern Cross Station:
-
-```{r, fig.height = 2, fig.width = 5, fig.align='center', fig.cap="A plot of the pedestrian counts at Southern Cross Station.
-The black line is the observed counts, where the yellow line is the imputed counts."}
-fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
-                           fit_ = dfa2$`Southern Cross Station`) %>%
-                  filter(Date_Time < "2014-07-06 00:00:00",
-                         Date_Time > "2014-07-04 12:00:00") %>%
-                        .$fit_
-
-ggplot(original_df %>%  filter(Date_Time < "2014-07-06 00:00:00",
-                               Date_Time > "2014-07-04 12:00:00")) +
-  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = viridis(3)[3],
-            size = 2) +
-  geom_line(aes(x = Date_Time, y = `Southern Cross Station`),
-            size  = 2) +
-  theme_minimal() + xlab("Time") +
-  ylab("Pedestrian Counts") +
-  ggtitle("Pedestrian Counts at Southern Cross Station \n on 04/07/2014 - 05/07/2014") +
-  theme(plot.title = element_text(hjust = 0.5)) 
-```
-
-More importantly, we can evaluate the performance of the model on the out of sample observations (2017 data) as cross validation.
-Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, we need to again apply the correction by long sequences of repeated values to be able to properly evaluate the goodness of fit of out-of-sample observations.
-The method used for this is identical to that used on the training set in the algorithm, replacing sequences of length 6 hours or longer with `NA`.
-
-```{r ts_fix0}
-for(i in unique(ped_data$Sensor)){
-    ori_ts[is.na(ori_ts[, i]), i] <- 0
-    na_length_check <- rle(as.numeric(unlist(ori_ts[,i])))
-    maybeNA <- rep((na_length_check$lengths > threshold_miss_hrs), times = na_length_check$lengths)
-    rl <- ori_ts[, i] %>% unlist %>% rle()
-    ori_ts[maybeNA, i] <- NA
-    defsNA <- is.na(ori_ts2[, i])
-    ori_ts[defsNA, i] <- NA
-  }
-```
-
-```{r test_fit}
-test_base <- list()
-test_fit <- list()
-ts <- original_df %>% filter(Date_Time > "2016-12-31")
-
- for(i in unique(ped_data$Sensor)){
-    test_base[[i]] <- predict.glm(models_base[[i]], newdata = ts,
-                                  type = "response")
- }
-
- for(i in small_miss_sensors){
-   test_fit[[i]]  <- predict.glm(models[[i]], newdata = ts,
-                                 type = "response")
- }
- for(i in large_miss_sensors){
-   sensor_dists <- numeric()
-  for (j in small_miss_sensors){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df_scale <- ts[, closest_sensor[1]] %>% unlist %>% scale
-  close_df_scale2 <- ts[, closest_sensor[2]] %>% unlist %>% scale
-
-  ts_sc <- data.frame(close_df_scale, close_df_scale2, Time = ts$Time)
-
-  test_fit[[i]] <- predict.glm(models[[i]], newdata = ts_sc,
-                               type = "response")
-
- }
-
-MARE_base_ts <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
-}
-
-MARE_ts <- numeric()
-for (i in unique(ped_data$Sensor)){
-  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
-}
-```
-
-Again, we start by comparing the fit at a sensor location with a small proportion of missing values, Melbourne Central:
-
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
-p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
-           y = test_base$`Melbourne Central`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Melbourne Central")
-p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
-           y = test_fit$`Melbourne Central`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 4000)) +
-           scale_y_continuous(limits = c(0, 4000)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
-grid.arrange(p1, p2, nrow = 1)
-```
-
-We find that the out-of-sample predictions of at Melbourne Central by the simple model has a tendency to underestimate the pedestrian counts.
-By comparison, the predictions made by the improved imputation model does not underestimate the counts as much.
-Comparing the out-of-sample MARE between models at Melbourne Central, MARE decreases from `r round(MARE_base_ts["Melbourne Central"]*100, 2)`% to `r round(MARE_ts["Melbourne Central"]*100, 2)`% with the use of the improved model.
-
-```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
-p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Collins Place (South)"],
-           y = test_base$`Collins Place (South)`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 2500)) +
-           scale_y_continuous(limits = c(0, 2500)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Collins Place (South)")
-p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Collins Place (South)"],
-           y = test_fit$`Collins Place (South)`)) +
-           geom_point(alpha = 0.05) +
-           geom_abline(aes(slope = 1, intercept = 0),
-                       colour = viridis(3)[1],
-                       alpha = 0.7, size = 1.25,
-                       linetype = 2) +
-           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
-                       colour = viridis(3)[3]) +
-           theme_minimal() + 
-           theme(aspect.ratio = 1,
-                 plot.title = element_text(hjust = 0.5)) +
-           scale_x_continuous(limits = c(0, 2500)) +
-           scale_y_continuous(limits = c(0, 2500)) +
-           xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Collins Place (South)")
-grid.arrange(p1, p2, nrow = 1)
-```
-
-At a sensor location with a large proportion of missing values, Collins Place (South), we again see a tendency to underestimate counts with the simple imputation model.
-The improved imputation method reduces the underestimation of the predictions and provides better predictions.
-The MARE is greatly improved, where it decreases from `r round(MARE_base_ts["Collins Place (South)"]*100, 2)`% to `r round(MARE_ts["Collins Place (South)"]*100, 2)`% with the use of the improved model.
-
-```{r}
-MARE_base_ts2 <- MARE_base_ts[setdiff(unique(ped_data$Sensor),
-                                      c("Spencer St-Collins St (South)",
-                                        "Spencer St-Collins St (North)",
-                                        "Flinders St-Swanston St (West)"))]
-MARE_ts2 <- MARE_ts[setdiff(unique(ped_data$Sensor),
-                                      c("Spencer St-Collins St (South)",
-                                        "Spencer St-Collins St (North)",
-                                        "Flinders St-Swanston St (West)"))]
-
-mare_df <- data.frame(simple = MARE_base_ts2, improved = MARE_ts2) %>%
-           gather(model, mare, simple:improved)
-
-p1 <- ggplot(mare_df) + geom_boxplot(aes(x = model, y = mare, fill = model),
-                                     width = .5, alpha = 0.5) +
-      labs(x = "Model", y = "Out-of-sample MARE") +
-      scale_fill_viridis(discrete = TRUE, name = "Model")
-
-p2 <- ggplot(mare_df) + geom_density(aes(x = mare, fill = model,
-                                         colour = model), adjust = 1.3,
-                                     alpha = 0.5) +
-      scale_fill_viridis(discrete = TRUE) +
-      scale_colour_viridis(discrete = TRUE) +
-      theme_minimal() + theme(plot.title = element_text(hjust = 0.5),
-                              legend.position = "none") +
-      labs(x = "Out-of-sample MARE")
-p3 <- ggplot() + geom_point(aes(x = MARE_base_ts2, y = MARE_ts2,
-                                colour = (MARE_ts2<MARE_base_ts2))) +
-      geom_abline(aes(slope = 1, intercept = 0)) +
-      scale_colour_viridis(discrete = TRUE, option = "C",
-                           begin = 0.15, end = 0.85,
-                           direction = -1,
-                           name = "Improved Model is better") +
-      scale_x_continuous(limits = c(0,1)) +
-      scale_y_continuous(limits = c(0,1)) +
-      theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
-      theme(aspect.ratio = 1, legend.position = "bottom",
-            legend.direction = "vertical")
-```
-
-```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}
-
-grid.arrange(p1, p2, p3, nrow = 1)
-
-```
-
-Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) was omitted from the the plots above as the out-of-sample MARE is extremely high at `r round(MARE_ts["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_ts["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_ts["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
-
-As these sensor locations were both classified as locations with a large proportion of missing values, it would appear that the relationship between these locations and their neighbouring sensors has changed.
-This is because the in-saple MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_tr["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
-
-```{r, fig.width = 8, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
-ggplot() + geom_smooth(aes(x = as.integer(Time),
-                           y = `Spencer St-Collins St (South)`),
-                       colour = viridis(3)[1],
-                       data = dfa2) +
-           geom_smooth(aes(x = as.integer(Time),
-                           y = `Spencer St-Collins St (South)`),
-                       colour = viridis(3)[3],
-                       data = ori_ts2) +
-           theme_minimal() + labs(x = "Time of Day") +
-           facet_wrap(~ DayType)
-```
-
-The plot of the average counts by time of day and type of day and data set shows that this location saw an increase in counts on weekdays which did not occur at the neghbouring sensors.
-We also see that the pattern is slightly different during weekdays, with a higher afternoon/evening peak.
-These types of change in pedestrian traffic behaviour is hard to capture in our imputation model, especially when its neighbours did not have a similar change.
-While we can add a year variable to allow for year to year growth, the short period of training data doesn't make this appropriate at this stage.
-
-Another way we can perform cross-validation is to remove data at a sensor, and evaulate the predictions made.
-This will allow us to evaluate the imputation model's performance within sample, which is the primary concern with them imputation process.
-We select a sensor which has a small proportion of missing values, Bourke Street Mall (South), which has `r round((1 - na_prop["Bourke Street Mall (South)"])*100, 3)`% of data available.
-As observed in the data, missing data is present in different ways: random, short periods of missing data (no more that 6 hours) or long periods of missing data (multiple months).
-We will simulate both types of missingness at this sensor location.
-
-First, data is randomly removed such that 20% of the training data is missing, then the imputation model is estimated.
-Using the estimated model, the MARE within the training data is calculated (in-sample MARE).
-```{r}
-miss_idx <- sample(1:nrow(dfa2), round(nrow(dfa2)*0.2))
-
-mall_missing_1 <- dfa2$`Bourke Street Mall (South)`
-mall_missing_1[miss_idx] <- NA
-
-i <- "Bourke Street Mall (South)"
-sensor_dists <- numeric()
-  for (j in small_miss_sensors[-6]){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors[-6]
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df <- dfa2[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
-
-  close_df_scale <- scale(close_df)
-  close_df_scale2 <- scale(close_df2)
-
-  sensor_df <- mall_missing_1 %>% unlist
-  sensor_dfsc <- scale(sensor_df)
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
-
-  options(na.action = na.omit)
-  mall_model_1 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  options(na.action = na.pass)
-  mall_fit_1 <- predict.glm(mall_model_1, newdata = tr_dat,
-                            type = "response")
-  options(na.action = na.omit)
-
-  mall_mare_1 <- mean(abs(mall_fit_1 - unlist(dfa2$`Bourke Street Mall (South)`)),
-                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
-                                            na.rm = T)
-
-```
-
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South).
-20% of the observations are removed from the traning data for the imputation model at random."}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1),
-                     size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date") +
-          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
-```
-
-With random hours of missing data, the predictions made by using neighbouring sensors  are estimated with good accuracy with a low MARE of `r round(mall_mare_1*100, 2)`%.
-Unfortunately, this type of missingness is not consistent with what is generally observed at most sensors with large proportions of missing data.
-
-Instead, a more meaningful method is to remove 20% of the data in a single period to simulate what is typically observed.
-A random date is taken and then 2629 hours of data is removed before and after this date (10% before, 10% after).
-The imputation model is estimated again in order to calculate a MARE.
-
-```{r}
-miss_seed <- max(sample(1:nrow(dfa2),1) - 2629, 0)
-miss_idx <- seq(miss_seed, miss_seed + 5258)
-
-mall_missing_2 <- dfa2$`Bourke Street Mall (South)`
-mall_missing_2[miss_idx] <- NA
-
-i <- "Bourke Street Mall (South)"
-sensor_dists <- numeric()
-  for (j in small_miss_sensors[-6]){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors[-6]
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df <- dfa2[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
-
-  close_df_scale <- scale(close_df)
-  close_df_scale2 <- scale(close_df2)
-
-  sensor_df <- mall_missing_2 %>% unlist
-  sensor_dfsc <- scale(sensor_df)
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
-
-  options(na.action = na.omit)
-  mall_model_2 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  options(na.action = na.pass)
-  mall_fit_2 <- predict.glm(mall_model_2, newdata = tr_dat,
-                            type = "response")
-  options(na.action = na.omit)
-
-  mall_mare_2 <- mean(abs(mall_fit_2 - unlist(dfa2$`Bourke Street Mall (South)`)),
-                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
-                                            na.rm = T)
-
-```
-
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence.
-This method of simulating missing data is a better representation of the missing data observed at sensors with a large proportion of missing data compared to randomly removing hourly observations from the training data."}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2),
-                size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date") +
-          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
-```
-
-The MARE of the estimated model is `r round(mall_mare_2*100, 2)`%, which is similar to what was obtained when the missing values were random.
-This demonstrates that the model quite well when we remove 20% of the observations from the training data.
-
-```{r}
-miss_seed <- min(max(sample(1:nrow(dfa2),1) - 6573, 0), 13146)
-miss_idx <- seq(miss_seed, miss_seed + 13146)
-
-mall_missing_3 <- dfa2$`Bourke Street Mall (South)`
-mall_missing_3[miss_idx] <- NA
-
-i <- "Bourke Street Mall (South)"
-sensor_dists <- numeric()
-  for (j in small_miss_sensors[-6]){
-      dists <- gcd.slc(loc[i,]$lons, loc[i,]$lats, loc[j,]$lons, loc[j,]$lats)
-      sensor_dists[j] <- dists
-  }
-  names(sensor_dists) <- small_miss_sensors[-6]
-  closest_sensor <- sensor_dists %>% sort %>%
-                    head(2) %>% names
-  close_df <- dfa2[, closest_sensor[1]] %>% unlist
-  close_df2 <- dfa2[, closest_sensor[2]] %>% unlist
-
-  close_df_scale <- scale(close_df)
-  close_df_scale2 <- scale(close_df2)
-
-  sensor_df <- mall_missing_3 %>% unlist
-
-  tr_dat <- data.frame(sensor_df, close_df_scale, close_df_scale2, Time = dfa2$Time)
-
-  options(na.action = na.omit)
-  mall_model_3 <- glm(sensor_df ~ Time*close_df_scale + Time*close_df_scale2,
-               data = tr_dat, family = quasipoisson())
-  options(na.action = na.pass)
-  mall_fit_3 <- predict.glm(mall_model_3, newdata = tr_dat,
-                            type = "response")
-  options(na.action = na.omit)
-
-  mall_mare_3 <- mean(abs(mall_fit_3 - unlist(dfa2$`Bourke Street Mall (South)`)),
-                          na.rm = T) / mean(unlist(dfa2$`Bourke Street Mall (South)`),
-                                            na.rm = T)
-
-```
-
-```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence.
-A large proportion of the training data is removed to test the robustness of the imputation method used at sensors with extremely large proportions of missing data during the training period defined."}
-ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_3),
-                size = 0.1, alpha = 1) +
-           labs(y = "Pedestrian Count", x = "Time and Date") +
-          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
-```
-
-Removing 50% of the data using the same method, it is found that the in-sample MARE is not affected significantly by the larger proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
 
@@
 This method was found to have improved prediction accuracy, as well as being robust at high levels of missingness.
+
+\nocite{*}

### 2017-10-21 thesis_draft/index.Rmd
bookdown update - need to fix formatting, improve introduction, write abstract, finish off lit review and conclusion

---
title: 'Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods'
degreetype: 'Bachelor of Economics Advanced with Honours'
author: 'Gavin Chin'
output: bookdown::pdf_book
site: bookdown::bookdown_site
bibliography: ["thesisrefs.bib"]
link-citations: yes
knit: "bookdown::render_book"
---

<!-- 
Edit these lines as appropriate.
The actual thesis content is in several Rmd files.

You'll need to edit the _bookdown.yml file to set the order in which you'd like them to appear.

If you have specific LaTeX packages to add, put them in monashthesis.tex.

You will need to ensure you have installed the knitr and bookdown packages for R.

You will also need LaTeX installed on your computer.
-->

### 2017-10-21 thesis_draft/01-chap1.Rmd
fixing figure captions

@@
 
-```{r naprop_hist, fig.width=6, fig.height=3, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
+```{r naprop_hist, fig.width=6, fig.height=3, fig.cap=paste("Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
 The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data.
-This emphasises the necessity to correct for false zero counts as it will affect "}
+This emphasises the necessity to correct for false zero counts as it will cause biased estimates.")}
 theme_g <- theme_minimal() +
@@
 
-```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
+```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = paste("Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours by geographical distance to be used to predict the pedestrian counts at Southbank.
+The algorithm will not choose sensors with large proportions of missing values to be used as neighbours for prediction.")}
 ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
@@
       scale_colour_viridis(discrete = TRUE,
-                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
-                                      "Neighbour 2 (Sandridge Bridge)",
+                           labels = c("Flinders St Station Underpass",
+                                      "Sandridge Bridge",
                                       "Target (Southbank)")) +
-      xlab("Time") + theme_minimal() + ggtitle("Unscaled Neighbours") +
+      xlab("Time") + theme_minimal() +
       theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
@@
       scale_colour_viridis(discrete = TRUE,
-                           labels = c("Neighbour 1 (Flinders St Station Underpass)",
-                                      "Neighbour 2 (Sandridge Bridge)",
+                           labels = c("Flinders St Station Underpass",
+                                      "Sandridge Bridge",
                                       "Target (Southbank)")) +
-      xlab("Time") + ggtitle("Scaled/Stadardised Neighbours") +
+      xlab("Time") + 
       ylab("Standard Deviations") + theme_minimal() +
-      theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
-```
-
-```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 6, fig.height = 3}
+      theme(legend.position = "bottom", legend.direction = "vertical")
+```
+
+```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 6, fig.height = 5, fig.cap=paste("Plot of the time series at the sensor with missing values, Southbank, and it's neighbouring sensors, as pedestrian counts and standardised.
+Because the counts are standardised to `mean = 0` and `var = 1`, the standardised counts are also standard deviations.
+The curves for the standardised counts at neighbouring sensors of Southbank closely match that of the standardised counts at Southbank.
+This demonstrates the importance of scaling counts at the neighbours when using them as predictors.")}
 grid.arrange(p1, p2, ncol = 1)
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count")
+           ylab("Predicted Pedestrian Count /n by simple imputation method")
 p2 <- ggplot(data = NULL, aes(x = original_df %>% 
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count")
+           ylab("Predicted Pedestrian Count /n by improved imputation method")
 grid.arrange(p1, p2, nrow = 1)
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Australia on Collins")
+           ylab("Predicted Pedestrian Count \n by simple imputation method")
 p2 <- ggplot(data = NULL, aes(x = ori_2[-(is.na(ori_2$`Australia on Collins`)), ] %>% 
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Australia on Collins")
+           ylab("Predicted Pedestrian Count /n by improved imputation method")
 ```

### 2017-10-21 thesis_draft/01-chap1.Rmd
update

@@
 ```{r, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.
+The improvement is very minor at Southern Cross Station as the data at this sensor is close to complete.
+Inclusion of month as a predictor is a main source of improvement in the model in this case."}
 p1 <- ggplot(data = NULL, aes(x = original_df %>% 
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count /n by improved imputation method")
+           ylab("Predicted Pedestrian Count \n by improved imputation method")
 grid.arrange(p1, p2, nrow = 1)
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count /n by improved imputation method")
+           ylab("Predicted Pedestrian Count \n by improved imputation method")
 ```
@@
 ```{r, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.
+The improvement in the imputation model is significant, where the predicted values lie closer to the dashed line.
+The simple model shows bias caused by the false zero-counts, causing underestimation.
+")}
 
@@
       labs(x = "Model", y = "In-Sample MARE") +
-      scale_fill_viridis(discrete = TRUE, name = "Model")
+      scale_fill_viridis(discrete = TRUE, name = "Model") +
+      theme_minimal() + theme(legend.position = "bottom",
+                              legend.direction = "vertical",
+                              legend.text = element_text(size = 5),
+                              legend.title = element_blank())
 
@@
                            direction = -1,
-                           name = "Improved Model is better") +
+                           name = "MARE Improved") +
       theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
       theme(aspect.ratio = 1, legend.position = "bottom",
-            legend.direction = "vertical")
-```
-
-```{r, fig.height=1, fig.width=6, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}
+            legend.direction = "vertical", legend.text = element_text(size = 5))
+```
+
+```{r, fig.height=3, fig.width=6, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model.
+It is seen that the distribution of MARE in the improved model is generally lower, with 34 locations having better fit with the improved model")}
 
@@
 
-```{r, fig.height = 2, fig.width = 5, fig.align='center', fig.cap="A plot of the pedestrian counts at Southern Cross Station.
-The black line is the observed counts, where the yellow line is the imputed counts."}
+```{r, fig.height = 2, fig.width = 5, fig.align='center', fig.cap="A plot of the pedestrian counts at Southern Cross Station for 04/07/2014 - 05/07/2014.
+The black line is the observed counts, where the yellow line is the imputed counts.
+This illustrates the replacement of missing values with imputed values."}
 fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
@@
   ylab("Pedestrian Counts") +
-  ggtitle("Pedestrian Counts at Southern Cross Station \n on 04/07/2014 - 05/07/2014") +
   theme(plot.title = element_text(hjust = 0.5)) 
@@
 ```{r, fig.height=3, fig.width=6, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.
+The improved method at Melbourne Central is seen to correct underestimation in the simple model by adding month of the year as a predictor in the model."}
 p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Melbourne Central")
+           ylab("Predicted Pedestrian Count \n by simple imputation method")
 p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
+           ylab("Predicted Pedestrian Count \n by improved imputation method") +
            ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
@@
 ```{r, fig.height=3, fig.width=6, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model.
-The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
+The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.
+Using neighbouring sensor counts as predictors, as well as correcting for the false zero-counts in the training data, are seen to reduce the degree of underestimation in the predicted values.
+"}
 p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Collins Place (South)"],
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Simple GLM Model: Actual v. Predicted \n at Collins Place (South)")
+           ylab("Predicted Pedestrian Count \n by simple imputation method")
 p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Collins Place (South)"],
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Collins Place (South)")
+           ylab("Predicted Pedestrian Count \n by improved imputation method")
 grid.arrange(p1, p2, nrow = 1)
@@
                                      width = .5, alpha = 0.5) +
-      labs(x = "Model", y = "Out-of-sample MARE") +
-      scale_fill_viridis(discrete = TRUE, name = "Model")
+      labs(x = "Model", y = "In-Sample MARE") +
+      scale_fill_viridis(discrete = TRUE, name = "Model") +
+      theme_minimal() + theme(legend.position = "bottom",
+                              legend.direction = "vertical",
+                              legend.text = element_text(size = 5),
+                              legend.title = element_blank())
 
@@
                               legend.position = "none") +
-      labs(x = "Out-of-sample MARE")
+      labs(x = "In-sample MARE")
 p3 <- ggplot() + geom_point(aes(x = MARE_base_ts2, y = MARE_ts2,
-                                colour = (MARE_ts2<MARE_base_ts2))) +
+                                colour = (MARE_tr<MARE_base_glm_tr))) +
       geom_abline(aes(slope = 1, intercept = 0)) +
@@
                            direction = -1,
-                           name = "Improved Model is better") +
-      scale_x_continuous(limits = c(0,1)) +
-      scale_y_continuous(limits = c(0,1)) +
+                           name = "MARE Improved") +
       theme_minimal() + labs(x = "Simple MARE", y = "Improved MARE") +
       theme(aspect.ratio = 1, legend.position = "bottom",
-            legend.direction = "vertical")
-```
-
-```{r, fig.height=1, fig.width=6, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}
+            legend.direction = "vertical", legend.text = element_text(size = 5))
+```
+
+```{r, fig.height=3, fig.width=6, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors.
+Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) was omitted from the the plots above as the out-of-sample MARE is extremely high.
+Improvement in the MARE is seen at most sensors again, but some sensors did not have improved prediction accuracy due to their patterns varying from thier neighbours over time.")}
 
@@
 
-```{r, fig.width = 6, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
+```{r, fig.width = 6, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line).
+The pedestrian counts characteristics are seen to be different between the in-sample observations (2014-2016) and the out-of-sample observations (2017).
+This explains the much larger MARE in the out-of-sample predictions, as the model cannot account for a change in pattern over time."}
 ggplot() + geom_smooth(aes(x = as.integer(Time),

### 2017-10-21 _book/thesis.tex
update

% This is a LaTeX thesis template for Monash University.
% to be used with Rmarkdown
% This template was produced by Rob Hyndman
% Version: 6 September 2016

\documentclass{monashthesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Add any LaTeX packages and other preamble here if required
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Joe Bloggs}
\title{Thesis title}
\degrees{B.Sc. (Hons), University of Tangambalanga}
\def\degreetitle{Doctor of Philosophy}
% Add subject and keywords below
\hypersetup{
     %pdfsubject={The Subject},
     %pdfkeywords={Some Keywords},
     pdfauthor={Joe Bloggs},
     pdftitle={Thesis title},
     pdfproducer={Bookdown with LaTeX}
}

\bibliography{thesisrefs}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newtheorem{proposition}{Proposition}[chapter]
\theoremstyle{definition}
\newtheorem{example}{Example}[chapter]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[chapter]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\pagenumbering{roman}

\titlepage

{\sf\tighttoc}

\chapter*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my pet goldfish for \dots

\chapter*{Declaration}\label{declaration}
\addcontentsline{toc}{chapter}{Declaration}

I hereby declare that this thesis contains no material which has been
accepted for the award of any other degree or diploma in any university
or equivalent institution, and that, to the best of my knowledge and
belief, this thesis contains no material previously published or written
by another person, except where due reference is made in the text of the
thesis.

\vspace*{2cm}

\par

\authorname

\chapter*{Preface}\label{preface}
\addcontentsline{toc}{chapter}{Preface}

The material in Chapter \ref{ch:intro} has been submitted to the journal
\emph{Journal of Impossible Results} for possible publication.

The contribution in Chapter \ref{ch:litreview} of this thesis was
presented in the International Symposium on Nonsense held in Dublin,
Ireland, in July 2015.

\chapter*{Abstract}\label{abstract}
\addcontentsline{toc}{chapter}{Abstract}

This thesis is about \ldots{}

\clearpage\pagenumbering{arabic}\setcounter{page}{0}

\chapter{Introduction}\label{ch:intro}

This is where you introduce the main ideas of your thesis, and an
overview of the context and background.

In a PhD, Chapter 2 would normally contain a literature review.
Typically, Chapters 3--5 would contain your own contributions. Think of
each of these as potential papers to be submitted to journals. Finally,
Chapter 6 provides some concluding remarks, discussion, ideas for future
research, and so on. Appendixes can contain additional material that
don't fit into any chapters, but that you want to put on record. For
example, additional tables, output, etc.

\section{Rmarkdown}\label{rmarkdown}

In this template, the rest of the chapter shows how to use Rmarkdown.
The big advantage of using Rmarkdown is that it allows you to include
your R code directly into your thesis, to ensure there are no errors in
copying and pasting, and that everything is reproducible. It also helps
you stay better organized.

For details on using \emph{R Markdown} see
\url{http://rmarkdown.rstudio.com}.

\section{Data}\label{data}

Included in this template is a file called \texttt{sales.csv}. This
contains quarterly data on Sales and Advertising budget for a small
company over the period 1981--2005. It also contains the GDP (gross
domestic product) over the same period. All series have been adjusted
for inflation. We can load in this data set using the following command:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sales <-}\StringTok{ }\KeywordTok{ts}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\StringTok{"data/sales.csv"}\NormalTok{)[,-}\DecValTok{1}\NormalTok{], }\DataTypeTok{start=}\DecValTok{1981}\NormalTok{, }\DataTypeTok{frequency=}\DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Any data you use in your thesis can go into the data directory. The data
should be in exactly the format you obtained it. Do no editing or
manipulation of the data outside of R. Any data munging should be
scripted in R and form part of your thesis files (possibly hidden in the
output).

\section{Figures}\label{figures}

Figure \ref{fig:deaths} shows time plots of the data we just loaded.
Notice how figure captions and references work. Chunk names can be used
as figure labels with \texttt{fig:} prefixed. Never manually type figure
numbers, as they can change when you add or delete figures. This way,
the figure numbering is always correct.

\begin{figure}[htbp]
\centering
\includegraphics{thesis_files/figure-latex/deaths-1.pdf}
\caption{\label{fig:deaths}Quarterly sales, advertising and GDP data.}
\end{figure}

\section{Results from analyses}\label{results-from-analyses}

We can fit a dynamic regression model to the sales data.

If \(y_t\) denotes the sales in quarter \(t\), \(x_t\) denotes the
corresponding advertising budget and \(z_t\) denotes the GDP, then the
resulting model is:

\begin{equation}
  y_t - y_{t-4} = \beta (x_t-x_{t-4}) + \gamma (z_t-z_{t-4}) + \theta_1 \varepsilon_{t-1} + \Theta_1 \varepsilon_{t-4} + \varepsilon_t
\end{equation}

where \(\beta = 2.29\), \(\gamma = 0.99\), \(\theta_1 = 0.20\), and
\(\Theta_1 = -0.90\).

\section{Tables}\label{tables}

Let's assume future advertising spend and GDP are at the current levels.
Then forecasts for the next year are given in Table
\ref{tab:salesforecasts}.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lllll}
\toprule
Point Forecast & Lo 80 & Hi 80 & Lo 95 & Hi 95 \\
\midrule
1000.6 &  948.0 & 1053.3 & 920.2 & 1081.1 \\
1013.3 &  959.6 & 1066.9 & 931.2 & 1095.3 \\
1076.8 & 1023.1 & 1130.4 & 994.7 & 1158.8 \\
1003.3 &  949.6 & 1057.0 & 921.2 & 1085.4 \\
\bottomrule
\end{tabular}
\caption{Forecasts for the next year assuming Advertising budget and GDP are unchanged.}
\label{tab:salesforecasts}
\end{center}
\end{table}

Again, notice the use of labels and references to automatically generate
Table numbers. In this case, we need to generate the label ourselves.

The \texttt{knitLatex} package is useful for generating tables from R
output. Other packages can do similar things including the
\texttt{kable} function in \texttt{knitr} which is somewhat simpler but
you have less control over the result. If you use \texttt{knitLatex} to
generate tables, don't forget to include \texttt{results="asis"} in the
chunk settings.

\chapter{Literature Review}\label{ch:litreview}

This chapter contains a summary of the context in which your research is
set.

Imagine you are writing for your fellow PhD students. Topics that are
well-known to them do not have to be included here. But things that they
may not know about should be included.

Resist the temptation to discuss everything you've read in the last few
years. And you are not writing a textbook either. This chapter is meant
to provide the background necessary to understand the material in
subsequent chapters. Stick to that.

You will need to organize the literature review around themes, and
within each theme provide a story explaining the development of ideas to
date. In each theme, you should get to the point where your ideas will
fit in. But leave your ideas to later chapters. This way it is clear
what has been done beforehand, and what new contributions you are making
to the research field.

All citations should be done using markdown notation as shown below.
This way, your bibliography will be compiled automatically and
correctly.

\section{Exponential Smoothing}\label{sec:expsmooth}

Exponential smoothing was originally developed in the late 1950s
\autocites{Brown59}{Brown63}{Holt57}{Winters60}. Because of their
computational simplicity and interpretability, they became widely used
in practice.

Empirical studies by \textcite{MH79} and \textcite{Metal82} found little
difference in forecast accuracy between exponential smoothing and ARIMA
models. This made the family of exponential smoothing procedures an
attractive proposition \autocite[see][]{CKOS01}.

The methods were less popular in academic circles until \textcite{OKS97}
introduced a state space formulation of some of the methods, which was
extended in \textcite{HKSG02} to cover the full range of exponential
smoothing methods.

\appendix

\chapter{Additional stuff}\label{additional-stuff}

You might put some computer output here, or maybe additional tables.

Note that line 5 must appear before your first appendix. But other
appendices can just start like any other chapter.

\printbibliography[heading=bibintoc]

\end{document}

### 2017-10-21 thesis_draft/00-abstract.Rmd
update

@@
 
-The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives"[^1].
+The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives".
 This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.

### 2017-10-21 thesis_draft/01-chap1.Rmd
update

@@
 ```
-# Background
-
-### The Data: City of Melbourne Pedestrian Data
-
-The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives"[^1].
+# Introduction
+
+The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives".
 This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.
@@
 
-The dataset being investigated can be obtained from the official City of Melbourne's open data[^2].
+# Motivation
+
+#Background
+
+### The Data: City of Melbourne Pedestrian Data
+
+The dataset being investigated can be obtained from the official City of Melbourne's open data[@com_data].
 The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations.
@@
 It is available in `.csv` format, allowing for easy data import into R.
-We can also access the pedestrian data using the the `rwalkr` package by Earo Wang[^3].
+We can also access the pedestrian data using the the `rwalkr` package [@_wang_rwalkr].
 This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.
@@
 The improved method at Melbourne Central is seen to correct underestimation in the simple model by adding month of the year as a predictor in the model."}
-p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
+p1 <- ggplot(data = NULL, aes(x = ori_ts2[, "Melbourne Central"],
            y = test_base$`Melbourne Central`)) +
@@
            ylab("Predicted Pedestrian Count \n by simple imputation method")
-p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
+p2 <- ggplot(data = NULL, aes(x =ori_ts2[, "Melbourne Central"],
            y = test_fit$`Melbourne Central`)) +
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count \n by improved imputation method") +
-           ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
+           ylab("Predicted Pedestrian Count \n by improved imputation method")
 grid.arrange(p1, p2, nrow = 1)
@@
 p3 <- ggplot() + geom_point(aes(x = MARE_base_ts2, y = MARE_ts2,
-                                colour = (MARE_tr<MARE_base_glm_tr))) +
+                                colour = (MARE_ts2<MARE_base_ts2))) +
       geom_abline(aes(slope = 1, intercept = 0)) +

### 2017-10-21 thesis_draft/index.Rmd
update

@@
 bibliography: ["thesisrefs.bib"]
+biblio-style: "apalike"
 link-citations: yes

### 2017-10-21 thesis_draft/01-chap1.Rmd
update

@@
 This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.
-In particular, we want to be able to model the pedestrian traffic at different locations in the CBD, and forecast the traffic.
-In addition, we aim to have an interactive, dynamic data visualisation of the model to help us to understand how Melbourne operates.
+In particular, the aim is to model the pedestrian behaviour at different locations in the CBD, and predict pedestrian counts.
+
+The main issue which is encountered with traffic data in general is the presence of missing values. For the Melbourne CBD pedestrian data, data is missing as a result of the
+
+Imputation methods are explored due to the lack of complete data in the pedestrian dataset.
 
 # Motivation
+Modelling pedestrian activity in the Melbourne CBD will provide insights to the social and consumer behaviour of the people within the city.
+In particular, prediction of pedestrian traffic will provide information about the way the city operates.
+Examples of some uses of such information include infrastructure planning, or security planning (which has become quite important in recent times) within the government sector.
+Private uses of such information include marketing, such as finding optimal locations to launch advertising campaigns, resource management for businesses to improve staffing based on pedestrian traffic, and investment planning to analyse business plan feasibility.
+
+Throughout the data exploration and analysis, issues with respect to missing values in the pedestrian data are explored and solutions to these issues are proposed.
+These issues have been seen to cause biased estimates when building models for prediction.
+Efficient methods for solving these issues is also a priority due to the size of the data.
 
@@
 
-The dataset being investigated can be obtained from the official City of Melbourne's open data[@com_data].
+The dataset being analysed can be obtained from the official City of Melbourne's open data[@com_data].
 The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations.
@@
 
-Not even an single sensor location has complete data for the period between between 1/1/2014 and `r paste(day(today()), month(today()), year(today()), sep = "/")`.
+No sensor location has complete data for the period between between 1/1/2014 and `r paste(day(today()), month(today()), year(today()), sep = "/")`.
 
@@
 
-Reviewing current research on imputation of missing values in traffic data, a paper on imputation of missing classified traffic data during winter season (Imputation of Missing Classified Traffic Data During Winter Season - _Hyuk-Jae Roh, Satish Sharma, Prasanta K.
-Sahu_) stated some methods which were found to be poor for imputation.
-The data they were attempting imputation for were traffic counters located on the highway network in Alberta, Canada with between 40% and 60% missing data.
+Reviewing current research on imputation of missing values in traffic data, a paper on imputation of missing classified traffic data during winter season (Roh et al., 2016) stated some methods which were found to be poor for imputation.
+The data used in this research were from traffic counters located on the highway network in Alberta, Canada with between 40% and 60% missing data.
 
 Replacement with "good" historical values, or using historical average values was found to have resulted in very poor fit with high MARE/MARD (mean absolute relative error/difference).
-They did find that using a moving average worked the best out of all the heuristic methods used.
-The largest problem with heuristic methods, however, is the inherent inability to reflect sudden fluctuations/shocks during abnormal periods.
+Roh et al.
+found that using a moving average worked the best out of all the heuristic methods used.
+The largest problem with heuristic methods, however, was the inherent inability to reflect sudden fluctuations/shocks during abnormal periods.
 With such a large proportion of missing values, pattern matching methods were investigated.
@@
 For the purposes of that paper, it was found to be inappropriate with the data being analysed.
-However, unlike _Roh et al._ where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller.
-Another difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data.
-They proposed using non-parametric estimation methods with k-Nearest Neighbours for their imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.
-
-<!-- #### Flexible and Robust Method for Missing Loop Detector Data Imputation - _K. C. Henrickson, Y. Zou, and Y. Wang_   -->
-
-<!-- http://docs.trb.org/prp/15-5804.pdf   -->
-<!-- - also use kNN -->
-
-<!-- #### Key Factors of K-nearest Neighbors Nonparametric Regression in Short-time Traffic Flow Forecasting -->
-<!-- http://www.springer.com/cda/content/document/cda_downloaddocument/9789462391017-c2.pdf?SGWID=0-0-45-1492476-p177147492 -->
-
-<!-- #### K-NEAREST NEIGHBOR ALGORITHM FOR UNIVARIATE TIME SERIES PREDICTION -->
-
-<!-- http://webbut.unitbv.ro/BU2012/Series%20III/BULETIN%20III/Sasu%20A.pdf -->
+However, unlike the highway traffic data, where patterns were compared to traffic volumes in different jurisdictions (large geographical distance), the geographic distances between the sensor locations in the Melbourne CBD pedestrian data are much smaller.
+Another major difference is the randomness of the missing values, where the periods of missing data run longer in the Melbourne CBD pedestrian data compared to that in the Alberta traffic data.
+Roh et al.
+proposed using non-parametric estimation methods with k-Nearest Neighbours (kNN) for imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.
 

### 2017-10-21 thesis_draft/01-chap1.Rmd
update

@@
 `Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
-Both these variables are categorical/factors. We cannot treat time of the day as an integer or numeric because it does not have a linear relationship with counts.s
-
-We use the Quasi-Poisson error distribution as opposed to normal/Gaussian errors due to the response variables being count data.
-The Poisson distribution allows for only non-negative integer values, while estimating a Quasi-Poisson model estimates an additional parameter for overdispersion.
+Both these variables are categorical/factors. Time of the day  cannot be treated as an integer or numeric because it does not have a linear relationship with the pedestrian counts.
+
+The quasipoisson error distribution is used as opposed to normal/Gaussian errors due to the response variables being count data.
+The poisson distribution allows for only non-negative integer values, while estimating a quasipoisson model estimates an additional parameter for overdispersion.
 This allows more flexibility in the model by allowing the assumption that $E[Y]=Var[Y]=\mu$ to be relaxed.
-Instead, the Quasi-Poisson distribution allows for $Var[Y] = \theta \mu$.
+Instead, the quasipoisson distribution allows for $Var[Y] = \theta \mu$.
 
@@
 
-While this works well with a small proportion of missing values, it also requires a large number of observations.
-Specifically, with this paramaterisation, we have $23 + 7 + (23 \times 7) + 1=192$ coefficients to estimate.
+While this simple model works well with a small proportion of missing values, a large number of observations is required for estimation.
+Specifically, with this paramaterisation, $23 + 7 + (23 \times 7) + 1=192$ parameters need to be estimated.
 Another disadvantage of using this model is the lack of robustness to outliers due to the sensitivity of some parameters.
@@
 A potential threshold value which could be used to class a sensor as having a "large" proportion of missing values vs a "small" proportion of missing values is 10%.
-At locations with a "small" proportion of missings, we can continue to use the GLM quasi-possion regression specified in $(1)$.
-At the locations with a large proportion of missing values, we need to use information from neighbouring sensors.
+At locations with a "small" proportion of missings, we can continue to use a GLM quasipossion regression with only time and date based variables as predictors.
+At the locations with a large proportion of missing values, neighbouring sensor counts are used for prediction.
 

### 2017-10-21 thesis_draft/01-chap1.Rmd
fixing first person to third person

@@
 A potential threshold value which could be used to class a sensor as having a "large" proportion of missing values vs a "small" proportion of missing values is 10%.
-At locations with a "small" proportion of missings, we can continue to use a GLM quasipossion regression with only time and date based variables as predictors.
-At the locations with a large proportion of missing values, neighbouring sensor counts are used for prediction.
+At locations with a "small" proportion of missings, a GLM quasipossion regression with only time and date based variables as predictors is still an appropriate model to use.
+At the locations with a large proportion of missing values, however, neighbouring sensor counts are used for prediction.
 
@@
 This issue will cause bias in the estimates, as well as affect the classification of a sensor.
-Of course, we cannot simply classify all zero counts as `NA`, as true zero values are possible.
+Of course, it is not valid to simply classify all zero counts as `NA`, as true zero values are possible.
 
@@
 
-Instead, a simple check we can implement is to look for long sequences of zero counts.
-For example, we may want to classify any sequence of `Hourly_Counts = 0` running for longer than 24 hours as being considered `NA`.
-From the summary of lengths of repeated values (which are repeated at least once), we see that the distribution of the lengths are very positively skewed with a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data.
-Classifying any sequence of repeated values, typically 0, which has a length greater than 6 as `NA` should allow true zero counts to avoid being misclassified as `NA`.
-We select the length of 6 hours as it would be unreasonable to assume any sensor location to have the exact same number of pedestrian counts over such a long period.
-
-After we have replaced questionable zero values with the value `NA`, we can calculate the proportion of missing values for each sensor location.
-From these calculated proportion, we can classify sensors as either having a "large" or "small" proportion of missing values, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.1`.
+Instead, a simple check which is implemented is to look for long sequences of zero counts.
+In this case, any sequence of `Hourly_Counts = 0` running for longer than 6 hours can be considered `NA`.
+From the summary of lengths of repeated values (which are repeated at least once), it is seen that the distribution of the lengths are very positively skewed with a mean length of `r round(mean(sequences[sequences > 1]), 2)` hours of missing data.
+Classifying any sequence of repeated values, typically 0, which has a length greater than 6 as `NA` will allow true zero counts to avoid being misclassified as `NA`.
+The length of 6 hours is selected as the limit as it would be unreasonable to assume any sensor location to have the exact same number of pedestrian counts over such a long period.
+
+After replacement of questionable zero values with the value `NA`, only then can the calculation of the proportion of missing values for each sensor location be performed.
+From these calculated proportion, classification of sensors as either having a "large" or "small" proportion of missing values is performed, where large is defined as $\mathtt{NAprop}_{sensor} > \mathtt{threshold}$ and `threshold = 0.1`.
 
@@
 
-Because the models for the sensors with a large proportion of missing values rely on having neighbouring sensors having complete cases (no missing values), we want to impute these values first.
-This will then allow us to have as much information to be available for the models which use neighbouring sensors to to train on.
-
-For sensors with a small proportion of missing values, we use a GLM quasipoisson model. To improve on the previous model used, we use the model:
+Because the models for the sensors with a large proportion of missing values rely on having neighbouring sensors having complete cases (no missing values), it is necessary  to impute these values (counts at sensors with small proportion of missing values) first.
+This will ensure as much information is available for the models which use neighbouring sensors to to train on.
+
+For sensors with a small proportion of missing values, a GLM quasipoisson model is estimated. To improve on the previous model used in the simple method, the model used is:
 $$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$
@@
 
-Instead of `HDay`, we use `DayType` which classifies Tuesday, Wednesday and Thursday as a single factor level, Midweek.
-We do this as the daily patterns on these weekdays are similar.
-This reduces the number of parameters to estimate by 46.
-A result of this is we also have more robust estimates for the midweek days, being less sensitive to outliers.
-Because this model will be estimated for sensor locations with few missing values,  we have enough data to add in month as an additive effect.
-This will help capture the annual seasonality.
-
-Using these estimated models, we impute the missing values to produce complete data at all these sensor locations.
+Instead of `HDay`, `DayType` is used which classifies Tuesday, Wednesday and Thursday as a single factor level, Midweek.
+This can be interpreted as the type of day.
+This can be done as the daily patterns on the Mideweek weekdays are similar to each other.
+The intuition for the use of this variable is for the reduction in the number of parameters to estimate ($23 \times 2 = 46$ less parameters).
+Another consequence of this is more robust estimates for the midweek days, being less sensitive to outliers on a particular day.
+Because this model will be estimated for sensor locations with few missing values,  enough data is available to add in month as an additive effect tol help capture the annual seasonality.
+This is not possible at locations with a large proportion of missing values, particularly at newer installations where less than 12 months of data is available in the training period specified.
+
+Using these estimated models, imputation of the missing values to produce complete data is performed at all these sensor locations.
 For the locations with high proportions of missing values can be predicted using the now complete data from all the locations which had a small proportion of missing values.
-Using a threshold of 10%, this gives us `r length(small_miss_sensors)` sensor locations to use.
+Using a threshold of 10%, there are `r length(small_miss_sensors)` sensor locations to use as potential candidates to be used as neighbouring sensors for prediction.
 
@@
 
-For each of the sensor locations with high proportions of missings, we need to decide which sensor to use for prediction.
-A simple method of finding a geographical closest neighbour is to take the haversine/great-circle distance between the sensor to be imputed and all the possible candidates.
+For each of the sensor locations with high proportions of missings, the algorithm needs to decide which sensor to use for prediction.
+A simple method which is applied for finding the geographically closest neighbours is to take the haversine/great-circle distances between the sensor to be imputed and all the possible candidates.
+
 ```{r, message=F}
@@
 
-Using the two geographically closest sensors with complete cases, we use another GLM specified as:
+Using the two geographically closest sensors (with complete cases) as defined by the haversine distance, the GLM specified for estimation is:
 $$E \left[ \text{HourlyCounts}_{\text{sensor}} \vert \text{Time}, \text{HourlyCounts}_{\text{neighbours}} \right] = \mu_{\text{Time}, \text{HourlyCounts}_{\text{neighbours}}}$$

### 2017-10-21 thesis_draft/01-chap1.Rmd
update

@@
 
-We would expect that the amount of people passing through the neighbouring sensors will be a strong predictor of the pedestrian counts as they are likely to also pass through.
-Using geographical neighbours, we can capture rare events effectively.
-Any large shocks to the counts at the neighbouring sensors will have an effect on the counts to be imputed.
+The model uses and interaction term between the standardised/scaled counts at the neighbouring sensors and the time of the day (hour factor).
+
+The amount of people passing through the neighbouring sensors will be expected to be a strong predictor of the pedestrian counts as they are likely to also pass through.
+Using geographical neighbours, it is possible capture rare events effectively as the predictors are also random variables.
+Any large shocks to the counts at the neighbouring sensors will have an effect on the predicted counts.
 
@@
 
-In order to use a mixture of the patterns from two different sensors, we need to use the scaled counts at the neighbouring sensors to avoid the magnitude of the counts having an effect on the predicted counts.
-We scale the counts by standardising to $\mu = 0, \sigma^2 = 1$.
-Because the covariance with neighbouring sensors may vary over time, we also add an interaction term between the neighbouring sensor counts and the time of the day.
+In order to use a mixture of the patterns from two different sensors, the use of scaled counts at the neighbouring sensors is required to avoid the magnitude of the counts having an effect on the predicted counts.
+The counts are scaled by standardising to $\mu = 0, \sigma^2 = 1$ at each sensor (standardised within sensors).
+Because the covariance with neighbouring sensors may vary over time, an interaction term between the neighbouring sensor counts and the time of the day is included.
 
@@
 
-Firstly, we evaluate the initial model used to impute:
+Firstly, the initial, simple method of imputation used is evaluated:
 $$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
@@
 
-For the purposes of evaluating the imputation method at a location with a small proportion of missing values, we will first look at the counts from Southern Cross Station.
-At this sensor, the proportion of missing values (adjusted) is `r round(na_prop["Southern Cross Station"]*100,2)`Firstly, we evaluate the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016).
+For the purposes of evaluating the imputation method at a location with a small proportion of missing values, the counts from Southern Cross Station will be used to evaluated prediction accuracy.
+At this sensor, the proportion of missing values (adjusted) is `r round(na_prop["Southern Cross Station"]*100,2)`.
+Firstly, the goodness of fit of the predictions made within the training period (01/01/2014 to 31/12/2016) are evaluated.
 
@@
            xlab("Actual Pedestrian Count") +
-           ylab("Predicted Pedestrian Count /n by simple imputation method")
+           ylab("Predicted Pedestrian Count \n by simple imputation method")
 p2 <- ggplot(data = NULL, aes(x = original_df %>% 
@@
 
-Visually, we can see that the fitted values are good estimates as the relationship between the actuals vs fitted is very close to 1:1.
-This is seen when we perform a least squares estimate of $\widehat{Fitted}_t = \hat{\beta}_0 + \hat{\beta}_1 Actual_t$, and the estimated $\hat{\beta}_0$ is close to 0 and $\hat{\beta}_1$ is close to 1.
-
-Here we see that the GLM model with only time and date based variables work well for fitting small periods of missing data.
+Visually, it can be seen that the fitted values are good estimates of the pedestrian counts as the relationship between the actual values and the fitted values is very close to 1:1.
+This is best observed when performing a least squares estimate (linear fit) of $\widehat{Fitted}_t = \hat{\beta}_0 + \hat{\beta}_1 Actual_t$, where the estimated $\hat{\beta}_0$ is close to 0 and $\hat{\beta}_1$ is close to 1.
+
+This provides evidence that the GLM model with only time and date based variables work well when there is only small periods of missing data.
 

### 2017-10-22 thesis_draft/01-chap1.Rmd
fixed error in plots using uncorrected  out of sample data

@@
 
-However, if we try to use this same model specification at sensor locations with large proportions of missing values, the fit is poor.
+However, using this same model specification at sensor locations with large proportions of missing values, the fit is poor.
 We use the sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%.
@@
 
-$$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \vert \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \vert}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$
+$$MARE_{sensor} =  \frac{1}{T^2} \frac{\sum_{t=1}^T \left| \widehat{\text{HourlyCounts}}_{sensor,t} - \text{HourlyCounts}_{sensor,t} \right|}{\sum_{t=1}^T \text{HourlyCounts}_{sensor,t}}$$
 
@@
 for (i in unique(ped_data$Sensor)){
-  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
+  MARE_base_ts[i] <- mean(abs(test_base[[i]] - unlist(ori_ts[, i])), na.rm = T) / mean(unlist(ori_ts[, i]), na.rm = T)
 }
@@
 for (i in unique(ped_data$Sensor)){
-  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts2[, i])), na.rm = T) / mean(unlist(ori_ts2[, i]), na.rm = T)
+  MARE_ts[i] <- mean(abs(test_fit[[i]] - unlist(ori_ts[, i])), na.rm = T) / mean(unlist(ori_ts[, i]), na.rm = T)
 }
@@
 The improved method at Melbourne Central is seen to correct underestimation in the simple model by adding month of the year as a predictor in the model."}
-p1 <- ggplot(data = NULL, aes(x = ori_ts2[, "Melbourne Central"],
+p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
            y = test_base$`Melbourne Central`)) +
@@
            ylab("Predicted Pedestrian Count \n by simple imputation method")
-p2 <- ggplot(data = NULL, aes(x =ori_ts2[, "Melbourne Central"],
+p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
            y = test_fit$`Melbourne Central`)) +

### 2017-10-22 thesis_draft/01-chap1.Rmd
update

@@
 # Implementation of improved imputation algorithm
+
+All data manipulation, exploration, visualisation and model estimation has been done in R using the RStudio IDE. A full list of R packages used is available in the Acknowledgements.
 
@@
 
-However, using this same model specification at sensor locations with large proportions of missing values, the fit is poor.
-We use the sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%.
+However, using this same model specification at sensor locations with large proportions of missing values, the prediction accuracy is poor.
+The sensor at Australia on Collins, where the proportion of missing values is `r round(na_prop["Australia on Collins"]*100, 2)`%, is used for the evaluation of goodness of fits.
 ```{r}
@@
 
-We can see that the simple model has bias caused by the false zero counts, resulting in the predicted values to be underestimated.
+It is observed that the simple model has biased estimates caused by the false zero counts, resulting in the predicted values to be underestimated.
 This is shown by the slope of the fitted line on the actual counts against predicted counts being $<1$, emphasising the importance of the first step taken in the algorithm of the improved model to class the false zero counts as missing values.
 
-The criterion we will use to measure goodness of fit is MARE. It is calculated by:
+The criterion used to measure goodness of fit is mean absolute relative error (MARE). It is calculated by:
 
@@
 
-We calculate a seperate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data).
-This will help to identify overfitting of the data.
-We use MARE as it will be a comparable measure between sensor locations as well as comparable between models.
+Absolute errors are used instead of squared errors as penalising larger errors relative to smaller errors was not desirable. A relative measure is used so that comparisons can be made between sensors.
+
+A seperate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data) is calculated. This is to help identify overfitting of the training data.
 
@@
 
-While it is not very clear when comparing the distribution of in-sample MARE between imputation models, when we compare the in-sample MARE at each location we can see that the majority of locations had better in-sample fit with the improved model using the algorithm.
-We do see that some locations did have worse in-sample fit, but overall we see improvement with `r sum(MARE_base_glm_tr > MARE_tr)` having better fit with the improved model over the simple model.
-
-Focusing on the date 04/07/2014, which has 3 hours of missing data at Southern Cross Station:
+While it is not very clear when comparing the distribution of in-sample MARE between imputation models, comparisons between the in-sample MARE at each location shows that the majority of locations had better in-sample fit with the improved model using the algorithm.
+It was observed that some locations did have worse in-sample fit, but overall, an improvement in model accuracy is seen with `r sum(MARE_base_glm_tr > MARE_tr)` locations having better fit with the improved model over the simple model.
+
+Focusing on the date 04/07/2014, which has 3 hours of missing data at Southern Cross Station, the replacement of missing values with imputed values is illustrated.
+It can be seen that the general pattern of pedestrian counts is retained, and thus should have minimal impact when using the imputed data for training predictive models.
 
@@
 
-More importantly, we can evaluate the performance of the model on the out of sample observations (2017 data) as cross validation.
-Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, we need to again apply the correction by long sequences of repeated values to be able to properly evaluate the goodness of fit of out-of-sample observations.
-The method used for this is identical to that used on the training set in the algorithm, replacing sequences of length 6 hours or longer with `NA`.
+More importantly, the performance of the model's predictions of the out of sample observations (2017 data) is evavaluated as cross validation.
+Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, correction of false zero-counts are required to be able to properly evaluate the goodness of fit of out-of-sample observations.
+The method used for this is identical to that used on the training set in the algorithm, with replacement of repeated value sequences of length 6 hours or longer with the value `NA`.
 
@@
 
-Again, we start by comparing the fit at a sensor location with a small proportion of missing values, Melbourne Central:
+Again, comparing the fit at a sensor location with a small proportion of missing values, Melbourne Central shows the bias in the estimates of the simple model caused by false zero-counts.
+There is good improvement in the prediction accuracy using the improved method.
 
@@
 
-We find that the out-of-sample predictions of at Melbourne Central by the simple model has a tendency to underestimate the pedestrian counts.
-By comparison, the predictions made by the improved imputation model does not underestimate the counts as much.
-Comparing the out-of-sample MARE between models at Melbourne Central, MARE decreases from `r round(MARE_base_ts["Melbourne Central"]*100, 2)`% to `r round(MARE_ts["Melbourne Central"]*100, 2)`% with the use of the improved model.
+When comparing the out-of-sample MARE between models at Melbourne Central, the improvement in prediction accuracy is seen by the MARE decreasing from `r round(MARE_base_ts["Melbourne Central"]*100, 2)`% to `r round(MARE_ts["Melbourne Central"]*100, 2)`% with the use of the improved model.
 
@@
 
-At a sensor location with a large proportion of missing values, Collins Place (South), we again see a tendency to underestimate counts with the simple imputation model.
-The improved imputation method reduces the underestimation of the predictions and provides better predictions.
+At a sensor location with a large proportion of missing values, Collins Place (South), the tendency to underestimate counts with the simple imputation model is also observed.
+The improved imputation method reduces the underestimation of the predictions and provides better predictions as a result of the false-zero correction and using neighbouring sensor counts as predictors.
 The MARE is greatly improved, where it decreases from `r round(MARE_base_ts["Collins Place (South)"]*100, 2)`% to `r round(MARE_ts["Collins Place (South)"]*100, 2)`% with the use of the improved model.
@@
 The plot of the average counts by time of day and type of day and data set shows that this location saw an increase in counts on weekdays which did not occur at the neghbouring sensors.
-We also see that the pattern is slightly different during weekdays, with a higher afternoon/evening peak.
+It was also seen that the pattern is slightly different during weekdays, with a higher afternoon/evening peak.
 These types of change in pedestrian traffic behaviour is hard to capture in our imputation model, especially when its neighbours did not have a similar change.
-While we can add a year variable to allow for year to year growth, the short period of training data doesn't make this appropriate at this stage.
-
-Another way we can perform cross-validation is to remove data at a sensor, and evaulate the predictions made.
-This will allow us to evaluate the imputation model's performance within sample, which is the primary concern with them imputation process.
-We select a sensor which has a small proportion of missing values, Bourke Street Mall (South), which has `r round((1 - na_prop["Bourke Street Mall (South)"])*100, 3)`% of data available.
+While it is possible to add a variable for year to allow for year to year growth, the short period of training data doesn't make this appropriate at this stage.
+
+Another method to perform cross-validation of the robustness of the model is to remove data at a sensor, and evaulate the predictions made.
+This will allow us to evaluate the imputation model's performance of out of sample but within training period observations, which is the primary concern with them imputation process.
+The sensor selected with a small proportion of missing values, Bourke Street Mall (South), has `r round((1 - na_prop["Bourke Street Mall (South)"])*100, 3)`% of data available.
 As observed in the data, missing data is present in different ways: random, short periods of missing data (no more that 6 hours) or long periods of missing data (multiple months).
-We will simulate both types of missingness at this sensor location.
+Simulation of both types of missingness at this sensor location is performed.
 
@@
 The MARE of the estimated model is `r round(mall_mare_2*100, 2)`%, which is similar to what was obtained when the missing values were random.
-This demonstrates that the model quite well when we remove 20% of the observations from the training data.
+This demonstrates that the model performs quite well when 20% of the observations from the training data is missing/removed.
 
@@
 
-Removing 50% of the data using the same method, it is found that the in-sample MARE is not affected significantly by the larger proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
+Removing 50% of the data using the same method, it is also found that the in-sample MARE is not affected significantly by the larger proportion of missing training data where it is only `r round(mall_mare_3*100, 2)`%.
+As such, it can be concluded that, at least for this sensor location, the imputation using neighbouring sensor counts as predictors is robust to high proportions of missing data.

### 2017-10-22 thesis_draft/01-chap1.Rmd
need to fix references, add more lit review, longer intro, longer conclusion, longer motivation, add acknowledgements

@@
 
-#Background
+# Background
 
@@
 
-No sensor location has complete data for the period between between 1/1/2014 and `r paste(day(today()), month(today()), year(today()), sep = "/")`.
-
-# Literature Review
+The period of data which is being used for analysis is between 01/01/2014 to 22/10/2017 (date of this paper).
+This is due to the poor quality of the dataset pre-2014 with many missing values with few sensors being installed during at the start of pedestrian counts recording (in 2009).
+
+Even with a truncated period for the data, no sensor location has complete data for the period between between 01/01/2014 and 31/12/2016 (the training period to be used for predictive models in the paper).
+As a result, imputation of the data is required for meaningful analysis of pedestrian behaviour.
+
+### Literature Review
 <!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->
@@
 The improvement in the imputation model is significant, where the predicted values lie closer to the dashed line.
-The simple model shows bias caused by the false zero-counts, causing underestimation.
-")}
+The simple model shows bias caused by the false zero-counts, causing underestimation.")}
 

### 2017-10-22 thesis_draft/01-chap1.Rmd
updated intro

@@
 Imputation methods are explored due to the lack of complete data in the pedestrian dataset.
+Without imputation on the data, some sensors lack information about pedestrian activity.
+This makes statistical inference and model estimation difficult to compare between locations.
 
@@
 Efficient methods for solving these issues is also a priority due to the size of the data.
+As such, the imputation method used should be computationally feasible.
 
@@
 
-The dataset being analysed can be obtained from the official City of Melbourne's open data[@com_data].
+The dataset being analysed can be obtained from the official City of Melbourne's online open data platform.
 The pedestrian data is in the form of hourly pedestrian counts for `r length(unique(ped_data$Sensor_ID))` sensor locations.
@@
 It is available in `.csv` format, allowing for easy data import into R.
-We can also access the pedestrian data using the the `rwalkr` package [@_wang_rwalkr].
+Access ti the pedestrian data using the the `rwalkr` package.
 This package allows R to import the data from `data.melbourne.vic.gov.au` which is updated monthly, or from the data source which is used by `pedestrian.melbourne.vic.gov.au` which is updated daily.
@@
 
+The data source used in this paper is accessed with the `rwalkr::walk_melb()` function, which imports data from `pedestrian.melbourne.vic.gov.au`.
+This data source was chosen due to missing values being explicit (all possible rows in the data are given, even if data is missing).
+It is also updated more frequently.
+
 The period of data which is being used for analysis is between 01/01/2014 to 22/10/2017 (date of this paper).
@@
 
+The sensor data is collected by sensors located throughout the Melbourne CBD under awnings and street poles.
+The sensors record the number of people passing through a counting zone each hour, sending this data to a server to be saved.
+The date of installation varies, as new sensors are constantly being installed over time.
+
 ### Literature Review
-<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->
-
-<!-- http://ir.lib.uwo.ca/cgi/viewcontent.cgi?article=1250&context=csce2016   -->
-<!-- - traffic counters with 40% to 60% missing values -->
-<!-- - heuristic methods -->
-<!--   - using "good historical values" or historical average values resulted -->
-<!--     in MARE (mean absolute relative error) reaching up to 80% -->
-<!--   - using moving average values worked better than other heuristic methods -->
-<!--   - all heuristic methods inhenrently cannot reflect sudden fluctuations during abnormal periods -->
-<!-- - pattern matching methods -->
-<!--   - find "candidate" volume pattern to compare to "study curve" and use the candidate which best matches the study curve -->
-<!-- - ARIMA -->
-<!--   - did not work well with multiple seasonalities -->
-<!-- - AI methods -->
-<!--   - genetic algorithms (GAs) and artificial neural networks (ANNs) -->
-
-<!-- Method used in the study was kNN with an interaction using a weather model, which they found to work best with large variations in traffic during winter -->
-<!-- (severe weather conditions affecting traffic volume)   -->
 
@@
 proposed using non-parametric estimation methods with k-Nearest Neighbours (kNN) for imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.
+Similarly, most other applications of imputation methods on traffic data used non-parametric regressions for traffic flow forecasting (Zhong, J., Ling, S.
+2015).
+
+Another important part of analysis which needed to be conducted when performing data imputation is assessment of robustness (Zhong, M.
+et al.
+2005).
+Zhong et al.
+found that modeles using additional observations as input and more sophisticated prediction techniques were able to make better imputations on a consistent basis.
 

### 2017-10-22 thesis_draft/00-acknowledge.Rmd
acknowledgements

---
knit: "bookdown::render_book"
---

# Acknowledgements {-}

I would like to thank my supervisor, Professor Dianne Cook, for assisting with my research and offering great advice. I would also like to thank Earo Wang for her support, as well as creating the extremely helpful `rwalkr` package.

Thank you to Professor Xueyan Zhao for coordinating a great Econometrics Honours course! And lastly, all my fellow Honours Class of 2017 for making the course so enjoyable!

## Software used

R and RStudio were used for all analysis in this paper. The following R packages were used: `tidyverse`, containing the packages: `ggplot2, dplyr, tidyr, readr, purr, tibble`, `lubridate`, `rwalkr`, `knitr`, `plyr`, `foreach`, `doSNOW`, `pander`, `gridExtra`, `ggmap`, `viridis`, `broom`, `bookdown`

### 2017-10-22 thesis_draft/02-chap2.Rmd
acknowledgements

@@
 
+Implementation of the imputation algorithm was made such that it is efficient, reproducible and general enough to allow for changes to model parameterisation and new data with from new sensor installations.
+
+Using the imputed data, a predictive model to be used for point estimates for a given time was built and packaged in a function in R, `ped_predict()`, to facilitate predictions in an easy and tidy format.
+The model has been saved so that predictions can be made without the end user needing to estimate the model prior to prediction.
+
 \nocite{*}

### 2017-10-22 thesis_draft/00-abstract.Rmd
spell check

@@
 
-The City of Melbourne provides an open data platform to access council datasets, with the intention "to increase transparency, improve public services and support new economic and social initiatives".
-This paper will focus primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.
-In particular, we want to be able to model the pedestrian traffic at different locations in the CBD, and forecast the traffic.
-In addition, we aim to have an interactive, dynamic data visualisation of the model to help us to understand how Melbourne operates.
+The City of Melbourne provides access pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD.
+Through data analysis and exploration, the aim is analyse the pedestrian behaviour at different locations in the CBD in order to predict pedestrian counts.
+Issues in the dataset, with missing values and false zero counts, are investigated and solutions proposed.
+An improved imputation method is proposed, using different models based on the proportion of missing values.
+For sensor locations with a small proportion of missing values used generalised linear models using only time and date based variables as predictors, while sensors with a large proportion of missings used neighbouring sensor counts as predictors.
+Using this imputed data, predictive models can be trained more effectively, producing better estimates.
+A predictive model for on-the-fly predictions was wrapped into a function to facilitate point estimates in an easy and tidy format.
 

### 2017-10-22 thesis_draft/01-chap1.Rmd
spell check

@@
 ori_ts2 <- ori_ts
+
+ped_loc <- read.csv("../data/Pedestrian_sensor_locations.csv")
 ```
@@
 
+```{r, fig.cap=paste("Plot of hourly counts for Flagstaff Station, Southbank and State Library sensors during the training period of 2014-2016.
+It is observed that Southbank has a period without any data (pre installation), while Flagstaff Station appears to have complete data.
+State Library appears complete, although the first half of 2015 shows suspeciously low hourly counts for a long period.
+")}
+pd <- ggplot(dfa %>% filter(Date_Time < "2017-01-01 00:00:00",
+                            Date_Time > "2013-12-31 00:00:00")) +
+        scale_y_continuous(limits = c(0, 7000)) +
+        xlab(NULL) +
+        theme_minimal()
+
+p1 <-  pd + geom_line(aes(x = Date_Time, y = `Flagstaff Station`))
+p2 <-  pd + geom_line(aes(x = Date_Time, y = `Southbank`))
+p3 <-  pd + geom_line(aes(x = Date_Time, y = `State Library`)) +
+            xlab("Time and Date")
+
+grid.arrange(p2, p1, p3, nrow = 3, ncol = 1)
+```
+
 Even with a truncated period for the data, no sensor location has complete data for the period between between 01/01/2014 and 31/12/2016 (the training period to be used for predictive models in the paper).
@@
 The date of installation varies, as new sensors are constantly being installed over time.
+
+```{r, fig.cap = paste("Map of all the pedestrian count sensors currently installed around the Melbourne CBD.
+The proximity between sensors is very close at some locations.
+This indicates potential for imputation using nearby sensor counts as predictors.")}
+melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
+                               mean(range(ped_loc$Latitude))),
+                                zoom=14)
+ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
+                         data = ped_loc, colour = "red",
+                         size = 4, alpha = 0.7, shape = 17) +
+              xlab(NULL) + ylab(NULL) +
+              theme(plot.title = element_text(hjust = 0.5),
+                    plot.caption = element_text(hjust = 0.5),
+                    axis.text.x = element_blank(),
+                    axis.text.y = element_blank(),
+                    axis.ticks = element_blank())
+```
 
@@
 proposed using non-parametric estimation methods with k-Nearest Neighbours (kNN) for imputation, although this is unlikely to work well with the lack of data at certain sensor locations in the Melbourne CBD Pedestrian data.
+This is because kNN regression works best when there is a lot of historical data.
+The lack of historical data at newer sensor installations, where less than 12 months of training data is available, makes kNN unfeasible.
 Similarly, most other applications of imputation methods on traffic data used non-parametric regressions for traffic flow forecasting (Zhong, J., Ling, S.
-2015).
+2015) involving kNN.
 
@@
 Zhong et al.
-found that modeles using additional observations as input and more sophisticated prediction techniques were able to make better imputations on a consistent basis.
+found that models using additional observations as input and more sophisticated prediction techniques were able to make better imputations on a consistent basis.
 
@@
 
-```{r naprop_hist, fig.width=6, fig.height=3, fig.cap=paste("Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
+```{r, fig.width=6, fig.height=3, fig.cap=paste("Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts.
 The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data.
@@
 This can be interpreted as the type of day.
-This can be done as the daily patterns on the Mideweek weekdays are similar to each other.
+This can be done as the daily patterns on the Midweek weekdays are similar to each other.
 The intuition for the use of this variable is for the reduction in the number of parameters to estimate ($23 \times 2 = 46$ less parameters).
 Another consequence of this is more robust estimates for the midweek days, being less sensitive to outliers on a particular day.
-Because this model will be estimated for sensor locations with few missing values,  enough data is available to add in month as an additive effect tol help capture the annual seasonality.
+Because this model will be estimated for sensor locations with few missing values,  enough data is available to add in month as an additive effect to help capture the annual seasonality.
 This is not possible at locations with a large proportion of missing values, particularly at newer installations where less than 12 months of data is available in the training period specified.
@@
 Any large shocks to the counts at the neighbouring sensors will have an effect on the predicted counts.
+This method is a form of pattern matching, where the sensor counts being imputed can be considered the study curve, and the neighbouring sensor counts are candidates for matching.
 
@@
 This was not an issue when a for loop was used.
-Additional code for removing unnecessary elements of the GLM model object before storing to the list of models, reduucing the memory usage.
+Additional code for removing unnecessary elements of the GLM model object before storing to the list of models, reducing the memory usage.
+Significant improvements in computing time is made with the use of parallel computing.
+Parallel computing is implemented with the use of `foreach()` and `doSNOW` rather than using the base R `for()` function.
+IT was found that estimation time was reduced by 50% when using a cluster size of two parallel nodes.
+Additional nodes (more than two) did not result in a significantly reduced computing time, so only two were used.
 
@@
 
-A seperate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data) is calculated. This is to help identify overfitting of the training data.
+A separate MARE for predictions in-sample (in the training set) and the predictions out of sample (test set of 2017 data) is calculated. This is to help identify overfitting of the training data.
 
@@
 
-More importantly, the performance of the model's predictions of the out of sample observations (2017 data) is evavaluated as cross validation.
+More importantly, the performance of the model's predictions of the out of sample observations (2017 data) is evaluated as cross validation.
 Due to potential false zero counts in the test data sourced from `rwalkr::walk_melb()`, correction of false zero-counts are required to be able to properly evaluate the goodness of fit of out-of-sample observations.
@@
 As these sensor locations were both classified as locations with a large proportion of missing values, it would appear that the relationship between these locations and their neighbouring sensors has changed.
-This is because the in-saple MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_tr["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
+This is because the in-sample MARE at these sensor locations were quite acceptable at `r round(MARE_tr["Spencer St-Collins St (North)"]*100, 2)`%, `r round(MARE_tr["Spencer St-Collins St (South)"]*100, 2)`% and `r round(MARE_tr["Flinders St-Swanston St (West)"]*100, 2)`% respectively.
 
@@
 
-The plot of the average counts by time of day and type of day and data set shows that this location saw an increase in counts on weekdays which did not occur at the neghbouring sensors.
+The plot of the average counts by time of day and type of day and data set shows that this location saw an increase in counts on weekdays which did not occur at the neighbouring sensors.
 It was also seen that the pattern is slightly different during weekdays, with a higher afternoon/evening peak.
@@
 
-Another method to perform cross-validation of the robustness of the model is to remove data at a sensor, and evaulate the predictions made.
+Another method to perform cross-validation of the robustness of the model is to remove data at a sensor, and evaluate the predictions made.
 This will allow us to evaluate the imputation model's performance of out of sample but within training period observations, which is the primary concern with them imputation process.

### 2017-10-22 thesis_draft/02-chap2.Rmd
spell check

@@
 
-Using a GLM predicts well compared to estimating with other models such as AR($p$) or other time series models when considering the tradeoff between model complexity and prediction accuracy.
+Using a GLM predicts well compared to estimating with other models such as AR($p$) or other time series models when considering the trade off between model complexity and prediction accuracy.
 Due to multiple seasonalities, AR($p$) models would need to be considered on a hour of the week basis at each sensor location.
 Prediction with time series models would also require of historical data.
-For the objective of this predictive model, the additional imputs required by end users to predict pedestrian counts when making out of sample predictions is not desirable.
+For the objective of this predictive model, the additional inputs required by end users to predict pedestrian counts when making out of sample predictions is not desirable.
 
@@
 
+Further research to improve the imputation algorithm can be made, such as more alternative models for different proportions of missing values.
+Cross validation of different missing proportion threshold values could also improve the prediction accuracy of the imputation models.
+
 \nocite{*}

### 2017-10-23 thesis_draft/03-refs.Rmd
bib added as separate rmd, manually typed :(

---
knit: "bookdown::render_book"
---

# References {-}

City of Melbourne. (2017) City of Melbourne - Open Data Portal, data.melbourne.vic.gov.au

City of Melbourne. (2017) City Of Melbourne: Pedestrian Counting System, pedestrian.melbourne.vic.gov.au

David Kahle and Hadley Wickham. (2016) ggmap: Spatial visualization with ggplot2

Simon Garnier. (2017) The viridis color palettes

Lionel Henry and Hadley Wickham. (2017) purrr: Functional programming tools

Kirill Müller and Hadley Wickham. (2017) tibble: Simple data frames

Hyuk-Jae Roh, Satish Sharma, and Prasanta K. Sahu. (2016) Imputation of missing classified traffic data during winter season

Vitalie Spinu, Garrett Grolemund, and Hadley Wickham. (2016) lubridate: Make dealing with dates a little easier

Mario Pineda-Krch. (2010) Great-circle distance calculations in R, R-bloggers

Earo Wang. (2017) rwalkr: API to Melbourne Pedestrian Data

Hadley Wickham and Winston Chang. (2016) ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics

Hadley Wickham and Romain Francois and Lionel Henry and Kirill Müller. (2017) dplyr: A grammar of data manipulation

Jing-ting Zhong and Shuai Ling. (2015) Key factors of k-nearest neighbours nonparametric regression in short-time traffic flow forecasting, Proceedings of the 21st International Conference on Industrial Engineering and Engineering Management 2014, 9–12

Ming Zhong, Satish Sharma, and Zhaobin Liu. (2005) Assessing robustness of imputation models based on data from different jurisdictions: Examples of Alberta and Saskatchewan, Canada, Transportation Research Record: Journal of the Transportation Research Board 1917, 116–126

Nina Zumel. (2014) Trimming the fat from glm() models in R, Win-Vector LLC

### 2017-10-23 thesis_draft/01-chap1.Rmd
pdf with png

@@
                       message = FALSE, warning = FALSE,
-                      fig.pos = 'h')
+                      fig.pos = 'h', dev = 'png')
 library(tidyverse)
@@
 
-The main issue which is encountered with traffic data in general is the presence of missing values. For the Melbourne CBD pedestrian data, data is missing as a result of the
+The main issue which is encountered with traffic data in general is the presence of missing values.
+For the Melbourne CBD pedestrian data, data is missing as a result of sensors being installed over time, as well as technical issues.
+As a consequence, data imputation, the estimation of missing data, is required.
 
@@
 
-# Motivation
+## Motivation
+
 Modelling pedestrian activity in the Melbourne CBD will provide insights to the social and consumer behaviour of the people within the city.
@@
 
-The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
+The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a quasipoisson error distribution, where specify the model as:
 $$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
